---
title: "NB02 3: Vannmiljø QC"
author: "Sam Welch"
date: last-modified
format: html
execute:
  freeze: true   # parked 2026-07-29, see _quarto.yml render list
---

I am also confused over notebook numbering. We'll sort it out at some point. 

# FIXME: Doesn't work rn because we're having an issue with the validation function. W/e.

This appendix documents the data quality validation results for Vannmiljø data converted to eData format.
```{r}
#| label: setup
#| include: false

library(targets)
library(pointblank)
library(dplyr)
library(gt)

tar_load(vm_edata_validation)
```

## Validation Overview

The following sections show validation results for each eData table created from Vannmiljø data.

### Campaign Table
```{r}
#| label: validation-campaign

get_agent_report(interrogate(vm_edata_validation$campaign))
```

### Reference Table
```{r}
#| label: validation-reference

get_agent_report(interrogate(vm_edata_validation$reference))
```

### Parameters Table
```{r}
#| label: validation-parameters

get_agent_report(interrogate(vm_edata_validation$parameters))
```

### Sites Table
```{r}
#| label: validation-sites

get_agent_report(interrogate(vm_edata_validation$sites))
```

### Samples Table
```{r}
#| label: validation-samples

get_agent_report(interrogate(vm_edata_validation$samples))
```

### Biota Table
```{r}
#| label: validation-biota

get_agent_report(interrogate(vm_edata_validation$biota))
```

### Measurements Table
```{r}
#| label: validation-measurements

get_agent_report(interrogate(vm_edata_validation$measurements))
```

## Validation Summary
```{r}
#| label: validation-summary

# Create summary table
# validation_summary <- data.frame(
#   Table = names(vm_edata_validation),
#   Passed = sapply(vm_edata_validation, x_passed),
#   Failed = sapply(vm_edata_validation, x_failed),
#   Pass_Rate = sapply(vm_edata_validation, function(x) {
#     total <- x_passed(x) + x_failed(x)
#     if (total == 0) {
#       return(NA)
#     }
#     round(100 * x_passed(x) / total, 1)
#   })
# ) |>
#   gt() |>
#   tab_header(
#     title = "eData Validation Summary",
#     subtitle = "Results across all Vannmiljø tables"
#   ) |>
#   fmt_number(columns = Pass_Rate, decimals = 1, pattern = "{x}%") |>
#   data_color(
#     columns = Pass_Rate,
#     colors = scales::col_numeric(
#       palette = c("red", "yellow", "green"),
#       domain = c(0, 100)
#     )
#   )

# validation_summary
```