---
title: "Notebook 08 - Emissions Data"
author: "Sam Welch"
date: last-modified
format: html
execute:
  freeze: true   # parked 2026-07-29, see _quarto.yml render list
---

*Currently a holding space for data on sources.*. Cache = true, fyi.

```{r}
#| label: setup
suppressMessages(
  {
    library(tibble)
    library(dplyr)
    library(ggplot2)
    library(here)
    library(devtools)
    library(sf)
    library(summarytools)
    library(DT)
    library(readxl)

    i_am("README.md")

    load_all(quiet = TRUE)
  }
)
```

# Copper Emissions & Source Data

## Closed Mines in Norway

Dataset available as a pdf here: https://www.vannportalen.no/miljotiltak/nedlagte-gruver/. Doesn't specify sources.

## Submarine Tailing Disposal in/off Norway

Note: STD locations (named) extracted from Kvassnes & Iversen 2013. Coordinates estimated by LLM, confidence self-reported by LLM. @kvassnesWasteSitesMines2013

```{r}
#| label: tbl-std-data
mine_tailings_coords() |> DT::datatable()
```

## Apple Production Areas in Norway

According to https://static02.nmbu.no/mina/studier/moppgaver/2018-Paulsen.pdf, 7 municipalities represent the main fruit districts in Norway. We also have sales of copper oxide for 2020-24 from https://mattilsynet-xp7prod.enonic.cloud/\_/attachment/inline/8311cd70-8877-4712-98cc-f7bc1dba46ac:0f491609e72f1d5f0dca5ad1030a4b979e8b4695/Omsetningsstatistikk%20for%20plantevernmidler%202020%20-%202024%20med%20uthevinger.pdf (table 2a, p4, thanks to Roger Holten).

@mattilsynetOmsetningsstatistikkPlantevernmidler202020242025 - does the citation work now?

N.B: Re Ronny Berdinsen, Mattilsynet (2025.12.15)

Hei

Tallene ble korrigerte pga en feilrapportering i 2022. Riktige tall er:

-	2022: 4332 liter (9482liter - 5150 liter)
-	2023: 2400 liter

Merk også at statistikken ikke viser plantevernmidler brukt pr år, men importert mengde. Det er ikke gitt at det som er importert i eksempelvis 2022 er brukt av bonden samme år.


```{r}
#| label: tbl-apple-emissions
# number_of_sites <- nrow(municipality_centroids())
# sales_data_split <- copper_oxide_sales() |>
#   mutate(
#     amount_kg = amount_kg / number_of_sites,
#     average_kg = average_kg / number_of_sites
#   )

# cross_join(municipality_centroids(), sales_data_split) |>
#   DT::datatable()

```

Although not explicitly stated, we assume that:

-   Copper products used as fungicides are exclusively applied in these areas and enter the environment via soil and freshwater
-   Apple farmers use the arthimetic mean of sold copper oxide for 2020 - 2024 every year. all applied copper oxide enters the environment.
-   The total used copper is split equally between the "main fruit districts" named in Paulsen (2018).

## Aquaculture Locations in Norway

Although I don't immediately have data for copper use in aquaculture in Norway (although this data probably exist?), the locations of licensed aquaculture sites are freely available from Fiskeridirektoratet as shapefiles (@norwegiandirectorateoffisheriesAquacultureLocationsNorway2025).

Actually the source ended up being https://portal.fiskeridir.no/portal/apps/webappviewer/index.html?id=bed4faeea84e4d6dbc548d43cc134c96&layer=Akvakulturregisteret, which is annoyingly detail-free.


```{r}
#| label: load-aquaculture-data
aquaculture <- sf::read_sf(here(
  "inst/extdata/shapefiles/Aquaculture_Locations/Fiskeridir_Akvakulturregistere.shx"
)) # that was easy

# View summary in browser
# aquaculture |> sf::st_drop_geometry() |> dfSummary()

# Whatever, the data's a mess but we'll treat it as a starting point. We can do a simple map in any case.
```

This data's a mess but we'll treat it as a starting point. We can do a simple map in any case, using the same hex-bin approach as elsewhere.


```{r}
#| label: fig-sites-hex-map
#| fig-cap: All Fiskerdirektoriet licensed sites. Doesn't include number of animals per site, because they're reported in lots of different units and I haven't worked out how to convert them yet.
#| fig-width: 12
# Plot
norway_map <- map_data("world", region = "Norway(?!:Svalbard)")
ggplot() +
  geom_polygon(
    data = norway_map,
    aes(x = long, y = lat, group = group),
    fill = "lightgray",
    colour = "white"
  ) +
  geom_hex(
    data = aquaculture,
    aes(x = lon, y = lat),
    bins = 100,
    linewidth = 0
  ) +
  scale_fill_viridis_b() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() +
  theme(axis.text = element_blank(), axis.title = element_blank())
```

And an accompanying table, although here we aggregate by kommune, rather than binned based on spatial proximity.

```{r}
#| label: tbl-most-aquaculture
aquaculture |>
  group_by(kommune, fylke) |>
  reframe(n = n()) |>
  arrange(desc(n)) |>
  datatable()
```

Technically geotraces isn't emissions either, but nevermind. Unfortunately they provide their data in a weird geochemistry/oceanography format (ODV). https://geotraces.webodv.awi.de/IDP2025%3Eseawater%3EGEOTRACES_IDP2025_Seawater/service/DataExtraction


```{r}
#| label: geotraces
library(readr)
library(oce)
library(here)
geotraces <- read.ctd.odv(
  here(
    "inst/extdata/exposure/Geotraces/IDP2025_seawater_GEOTRACES_IDP2025_Seawater_w1efqRDC.txt"
  )
)

show(geotraces)

library(maps)

world_map <- map_data("world")
library(sf)
geotraces_tibble <- as_tibble(geotraces@data, .name_repair = "universal") |>
  mutate(lon = longitude, lat = latitude) |>
  st_as_sf(coords = c("lon", "lat")) |>
  mutate(Cu_D_nmol_kg = `Cu_D_CONC..nmol.kg.`)


geotraces_tibble_grouped <- geotraces_tibble |>
  group_by(longitude, latitude) |>
  reframe(mean_conc = mean(Cu_D_nmol_kg, na.rm = TRUE)) |>
  filter(!is.na(mean_conc))
#  |>
# st_as_sf(coords = c("longitude", "latitude"))

# desnity mapping: https://www.andrewheiss.com/blog/2023/07/28/gradient-map-fills-r-sf/
# actually, this might not be it because we don't want density, we want to visualise the mean concentration
# what I want is a statistically valid way to assume concentration gradients from point data

# claude suggests kriging:
# 2. Kriging (Gaussian Process interpolation)
# More statistically principled — models spatial autocorrelation explicitly via a variogram. Gives you uncertainty estimates too, which is valuable for saying "concentration here is X ± Y". This is probably closest to what you want for making defensible statements.

# This might be overkill

ggplot() +
  geom_polygon(
    data = world_map,
    aes(x = long, y = lat, group = group),
    fill = "lightgray",
    colour = "white"
  ) +
  geom_hex(
    data = geotraces_tibble_grouped,
    aes(x = longitude, y = latitude),
    bins = 50
  ) +
  scale_fill_viridis_c() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() +
  theme(axis.text = element_blank(), axis.title = element_blank())
```
