---
title: "Notebook XX - Aquaculture Copper Emissions (Hammerfest)"
subtitle: "Matching the Miljodirektoratet aquaculture Cu-to-water export to licensed farm locations"
date: last-modified
format: html
execute:
  freeze: true   # parked / temporary, not in the _quarto.yml render list
  echo: true     # working notebook: the name-matching logic is the point
  warning: false
  message: false
---

## What this is

`data/raw/Copper-Aquaculture-Emissions.xlsx` turned up as a per-facility,
per-year record of **copper discharged to water** by Norwegian aquaculture sites
(Miljodirektoratet / Norske utslipp style export, sheet `Utslipp`). It is
national, 2002-2023, one row per facility x operator x year, every value in kg
and every value carrying basis code **`B` (*Beregnet*, i.e. modelled)** rather
than measured. Copper from a marine farm is overwhelmingly net-pen antifouling
coating, so a modelled figure here is expected.

The file has **no location number** (`loknr`), only a free-text `Anleggsnavn` of
the form `"<site>, matfisk (<operator>)"`. To place the emissions we match that
name against the Fiskeridirektoratet aquaculture register
(`inst/extdata/shapefiles/Aquaculture_Locations/`), which does carry `loknr`,
coordinates, species and licence status.

**Scope: Hammerfest kommune only** (post-2020 merged kommune, so the old
Kvalsund sites are included; the file has already recoded them to `Hammerfest`).

```{r}
#| label: setup
library(readxl)
library(sf)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(flextable)
library(here)

here::i_am("README.md")
pkgload::load_all(here::here(), quiet = TRUE)

ft <- function(d, cap = NULL) {
  flextable(d) |>
    theme_vanilla() |>
    bold(part = "header") |>
    fontsize(size = 9, part = "all") |>
    padding(padding = 2, part = "all") |>
    set_caption(cap) |>
    set_table_properties(layout = "autofit", width = 1)
}

# Site-name key: lower-case, drop the ", matfisk (...)" / " akva (...)" tail,
# squish whitespace. Diacritics are kept -- both sources are Norwegian and
# agree on them.
norm_site <- function(x) {
  x |>
    str_to_lower() |>
    str_replace_all("\\s*[,(].*$", "") |>
    str_replace_all("\\s+akva$", "") |>
    str_replace_all("[[:space:]]+", " ") |>
    str_trim()
}
```

## Load and key both sources

```{r}
#| label: load-emissions
emissions <- read_excel(
  here::here("data/raw/Copper-Aquaculture-Emissions.xlsx"),
  sheet = "Utslipp",
  skip = 1,
  .name_repair = "unique_quiet"
) |>
  rename(
    anlegg = 1,
    fylke = 2,
    kommune = 3,
    year = 4,
    cu_kg = 5,
    basis = 6,
    unit = 7
  ) |>
  mutate(
    year = as.integer(year),
    cu_kg = as.numeric(cu_kg),
    site_key = norm_site(anlegg),
    operator = str_match(anlegg, "\\(([^)]*)\\)\\s*$")[, 2]
  )

em_ha <- emissions |>
  filter(str_to_upper(kommune) == "HAMMERFEST") |>
  group_by(site_key) |>
  summarise(
    operators = paste(sort(unique(operator)), collapse = " | "),
    year_from = min(year),
    year_to = max(year),
    n_year_rows = n(),
    cu_total_kg = round(sum(cu_kg, na.rm = TRUE), 1),
    cu_per_year = round(mean(cu_kg, na.rm = TRUE), 2),
    .groups = "drop"
  )
```

```{r}
#| label: load-register
register <- sf::read_sf(here::here(
  "inst/extdata/shapefiles/Aquaculture_Locations/Fiskeridir_Akvakulturregistere.shp"
)) |>
  sf::st_drop_geometry() |>
  as_tibble() |>
  filter(str_to_upper(kommune) == "HAMMERFEST") |>
  mutate(site_key = norm_site(navn)) |>
  group_by(site_key) |>
  summarise(
    loknr = paste(sort(unique(loknr)), collapse = "/"),
    navn = navn[1],
    status = paste(sort(unique(status_lok)), collapse = "/"),
    lon = mean(lon),
    lat = mean(lat),
    species = paste(sort(unique(til_arter)), collapse = " | "),
    capacity = paste0(kapasitet_[1], " ", kapasitet1[1]),
    .groups = "drop"
  )

# Combined Hammerfest-area AEP bounding box, read live from the manifest.
aep_box <- readr::read_csv(
  here::here("data/clean/aep/aep_manifest.csv"),
  show_col_types = FALSE
) |>
  summarise(
    lon_min = min(lon_min),
    lon_max = max(lon_max),
    lat_min = min(lat_min),
    lat_max = max(lat_max)
  )
```

## Match

```{r}
#| label: match
matched <- em_ha |>
  left_join(register, by = "site_key") |>
  mutate(
    in_aep_box = !is.na(lat) &
      lon >= aep_box$lon_min &
      lon <= aep_box$lon_max &
      lat >= aep_box$lat_min &
      lat <= aep_box$lat_max
  )

n_sites <- nrow(matched)
n_matched <- sum(!is.na(matched$loknr))
n_in_box <- sum(matched$in_aep_box, na.rm = TRUE)
```

`r n_matched` of `r n_sites` Hammerfest-kommune emission sites match a
currently-licensed location by name (1:1 -- no two current Hammerfest licences
share a normalised name). `r n_in_box` of the matched sites fall inside the
combined AEP bounding box (`r sprintf("%.2f-%.2f E, %.2f-%.2f N", aep_box$lon_min, aep_box$lon_max, aep_box$lat_min, aep_box$lat_max)`);
the rest are elsewhere in the kommune (Sorroysund, Revsbotn, the Kvalsund side).

### Matched: emission site to register `loknr`

```{r}
#| label: tbl-matched
#| tbl-cap: "Hammerfest-kommune aquaculture sites reporting copper to water, matched to the Fiskeridirektoratet register. Cu is the sum / mean of annual modelled discharge over the reporting years shown. Coordinates, `loknr` and species come from the register."
matched |>
  filter(!is.na(loknr)) |>
  transmute(
    Site = str_to_title(site_key),
    loknr,
    Lon = round(lon, 4),
    Lat = round(lat, 4),
    `In AEP box` = if_else(in_aep_box, "yes", "no"),
    `Reg. status` = status,
    Years = paste0(year_from, "-", year_to),
    `n yr` = n_year_rows,
    `Cu total (kg)` = cu_total_kg,
    `Cu (kg/yr)` = cu_per_year,
    Species = species
  ) |>
  arrange(desc(`Cu total (kg)`)) |>
  ft()
```

### Emissions but no current licence

Historic sites -- decommissioned or licence surrendered before the register
snapshot, so no coordinates are available from it. Operator strings are the
pre-2015 holders (Polarlaks, Stolt Sea Farm, early Cermaq).

```{r}
#| label: tbl-historic
#| tbl-cap: "Hammerfest-kommune sites with reported copper discharge but no current aquaculture licence."
matched |>
  filter(is.na(loknr)) |>
  transmute(
    Site = str_to_title(site_key),
    Operators = operators,
    Years = paste0(year_from, "-", year_to),
    `n yr` = n_year_rows,
    `Cu total (kg)` = cu_total_kg
  ) |>
  arrange(desc(`Cu total (kg)`)) |>
  ft()
```

### Licensed sites in the AEP box with no emissions record

```{r}
#| label: tbl-nomatch-reverse
#| tbl-cap: "Currently-licensed aquaculture locations inside the AEP bounding box that do not appear in the copper emissions file."
register |>
  filter(
    !is.na(lon),
    lon >= aep_box$lon_min,
    lon <= aep_box$lon_max,
    lat >= aep_box$lat_min,
    lat <= aep_box$lat_max,
    !site_key %in% em_ha$site_key
  ) |>
  transmute(
    loknr,
    Site = navn,
    Lon = round(lon, 4),
    Lat = round(lat, 4),
    Species = species,
    Capacity = capacity
  ) |>
  ft()
```

## Totals

```{r}
#| label: totals
tibble::tribble(
  ~Scope                                  , ~`Cu to water 2002-2023 (kg)`                                           ,
  "All Hammerfest-kommune emission sites" , round(sum(matched$cu_total_kg, na.rm = TRUE), 0)                        ,
  "Matched to a current licence"          , round(sum(matched$cu_total_kg[!is.na(matched$loknr)], na.rm = TRUE), 0) ,
  "Inside the AEP bounding box"           , round(sum(matched$cu_total_kg[matched$in_aep_box], na.rm = TRUE), 0)    ,
  "Historic sites (no current licence)"   , round(sum(matched$cu_total_kg[is.na(matched$loknr)], na.rm = TRUE), 0)
) |>
  ft()
```

```{r}
#| label: fig-cu-by-site
#| fig-cap: "Total modelled copper to water 2002-2023 per Hammerfest-kommune aquaculture site, split by whether the site matches a current licence. Written to images/dev/ for review outside the rendered notebook."
#| fig-width: 8
#| fig-height: 6
plot_df <- matched |>
  mutate(
    site_lab = str_to_title(site_key),
    group = if_else(is.na(loknr), "historic (no licence)", "current licence")
  )

p <- ggplot(
  plot_df,
  aes(
    x = cu_total_kg,
    y = reorder(site_lab, cu_total_kg),
    fill = group
  )
) +
  geom_col() +
  scale_fill_manual(
    values = c(
      "current licence" = "#2B5F75",
      "historic (no licence)" = "#B5876B"
    )
  ) +
  labs(
    x = "Cu to water 2002-2023 (kg, modelled)",
    y = NULL,
    fill = NULL,
    title = "Aquaculture copper discharge, Hammerfest kommune"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom", panel.grid.major.y = element_blank())

dir.create(here::here("images/dev"), showWarnings = FALSE, recursive = TRUE)
ggsave(
  here::here("images/dev/aquaculture-cu-by-site.png"),
  p,
  width = 8,
  height = 6,
  dpi = 150,
  bg = "white"
)
p
```

## AEP-001 figure: facility map and emission trends

Two-panel figure for the methods section, scoped to the **AEP-001 bounding box
only** (23.40-24.12 E, 70.44-70.78 N, read live from `aep_manifest.csv`), not the
combined Hammerfest-area box used earlier in this notebook. Both panels show the
same set: farms inside the box that carry *both* a register location and a
modelled copper-to-water series. "Still reporting" means a modelled value in 2022
or 2023; "Stopped" means the last modelled year is 2021 or earlier. All values
are modelled (basis `B`), kg Cu to water per year.

```{r}
#| label: aep001-prep
# A001 box only.
a001 <- readr::read_csv(
  here::here("data/clean/aep/aep_manifest.csv"),
  show_col_types = FALSE
) |>
  filter(aep_id == "A001")

# Per-facility, per-year modelled Cu to water, Hammerfest kommune. A site-year
# can carry several operator rows, so sum.
em_year <- emissions |>
  filter(str_to_upper(kommune) == "HAMMERFEST") |>
  group_by(site_key, year) |>
  summarise(cu_kg = sum(cu_kg, na.rm = TRUE), .groups = "drop")

active_cutoff <- 2022L

# Long table: one row per (facility, year) for facilities in the A001 box with
# register coordinates.
fac <- em_year |>
  group_by(site_key) |>
  mutate(last_year = max(year)) |>
  ungroup() |>
  inner_join(
    distinct(register, site_key, navn, lon, lat, species, capacity),
    by = "site_key"
  ) |>
  filter(
    lon >= a001$lon_min, lon <= a001$lon_max,
    lat >= a001$lat_min, lat <= a001$lat_max
  ) |>
  mutate(
    status = if_else(
      last_year >= active_cutoff, "Still reporting", "Stopped"
    ),
    site_lab = str_to_title(site_key),
    # `capacity` is the register string, e.g. "5400 TN": TN = tonnes of
    # maximum allowed biomass (MTB). All eight in-box farms are licensed for
    # the same species mix (Laks / Regnbueørret / Ørret).
    cap_t = readr::parse_number(capacity),
    species_lbl = dplyr::case_when(
      species == "Laks, Regnbueørret, Ørret" ~ "salmon & trout",
      TRUE ~ species
    ),
    # Facet-strip label: site name, then species and licensed capacity.
    facility = paste0(
      site_lab, "\n", species_lbl, ", ", scales::comma(cap_t), " t MTB"
    )
  )

# Facet order: biggest cumulative discharge first, then the 8-farm sum last.
fac_order <- fac |>
  group_by(facility) |>
  summarise(tot = sum(cu_kg), .groups = "drop") |>
  arrange(desc(tot)) |>
  pull(facility)

total_lab <- "Sum of all 8 farms"
fac_levels <- c(fac_order, total_lab)
fac <- mutate(fac, facility = factor(facility, levels = fac_levels))

# 9th panel: yearly total modelled Cu across the 8 in-box farms.
fac_total <- fac |>
  group_by(year) |>
  summarise(cu_kg = sum(cu_kg), .groups = "drop") |>
  mutate(facility = factor(total_lab, levels = fac_levels))

# Period mean of the summed series, for a reference line in the sum panel.
total_mean <- tibble::tibble(
  facility = factor(total_lab, levels = fac_levels),
  cu_kg = mean(fac_total$cu_kg)
)

fac_pts <- fac |>
  distinct(site_key, site_lab, facility, navn, lon, lat, status, last_year)

n_fac <- nrow(fac_pts)
```

`r n_fac` farms in the AEP-001 box have both a register location and a modelled
emissions series (`r sum(fac_pts$status == "Still reporting")` still reporting in
2022-2023, `r sum(fac_pts$status == "Stopped")` stopped earlier).

```{r}
#| label: fig-aep001-aquaculture
#| fig-cap: "Aquaculture copper to water inside the AEP-001 bounding box. (a) Licensed farm locations that carry a modelled copper-to-water record, coloured by whether they were still reporting in 2022-2023 or had stopped earlier; red box is the AEP-001 extent (coastline is Natural Earth 10m, too coarse to resolve every bay). (b) Modelled copper discharge per farm by year, 2002-2023; strip labels carry species and licensed capacity (t maximum allowed biomass). One panel per farm plus a final grey panel giving the yearly sum across all eight, with a dashed period-mean line; every panel has an independent y axis, and the sum panel's is roughly an order of magnitude larger. All values modelled (basis B). Coordinates: Fiskeridirektoratet aquaculture register; emissions: Miljodirektoratet / Norske utslipp."
#| fig-width: 13
#| fig-height: 7
status_cols <- c("Still reporting" = "#2B5F75", "Stopped" = "#B5876B")

## Panel a: farm locations -------------------------------------------------
pad <- 0.15
span_lon <- a001$lon_max - a001$lon_min
span_lat <- a001$lat_max - a001$lat_min
view <- c(
  xmin = a001$lon_min - span_lon * pad,
  xmax = a001$lon_max + span_lon * pad,
  ymin = a001$lat_min - span_lat * pad,
  ymax = a001$lat_max + span_lat * pad
)
centre_lat <- mean(c(a001$lat_min, a001$lat_max))
eqc <- sf::st_crs(paste0("+proj=eqc +lat_ts=", centre_lat, " +datum=WGS84"))

land <- rnaturalearth::ne_countries(
  scale = 10, country = "Norway", returnclass = "sf"
)
view_sfc <- sf::st_as_sfc(sf::st_bbox(view, crs = sf::st_crs(4326)))
land_crop <- suppressWarnings(sf::st_crop(land, view_sfc))
box_sfc <- sf::st_as_sfc(sf::st_bbox(
  c(
    xmin = a001$lon_min, ymin = a001$lat_min,
    xmax = a001$lon_max, ymax = a001$lat_max
  ),
  crs = sf::st_crs(4326)
))
pts_sf <- sf::st_as_sf(
  fac_pts, coords = c("lon", "lat"), crs = sf::st_crs(4326)
)

pa <- ggplot() +
  geom_sf(
    data = land_crop, fill = "grey86", colour = "grey70", linewidth = 0.2
  ) +
  geom_sf(data = box_sfc, fill = NA, colour = "#b23b3b", linewidth = 0.7) +
  # No legend key from either map layer; panel b's points are the single
  # collected legend for the whole figure.
  geom_sf(data = pts_sf, aes(colour = status), size = 2.4, show.legend = FALSE) +
  ggrepel::geom_text_repel(
    data = fac_pts,
    aes(x = lon, y = lat, label = site_lab, colour = status),
    size = 2.8, fontface = "bold", seed = 1, min.segment.length = 0,
    box.padding = 0.4, max.overlaps = Inf, show.legend = FALSE
  ) +
  scale_colour_manual(values = status_cols, name = NULL) +
  coord_sf(
    crs = eqc, default_crs = sf::st_crs(4326),
    xlim = c(view["xmin"], view["xmax"]),
    ylim = c(view["ymin"], view["ymax"]), expand = FALSE
  ) +
  theme_void(base_size = 11) +
  theme(
    panel.background = element_rect(fill = "#eaf2f7", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "bottom"
  )

## Panel b: per-farm trends, plus the 8-farm yearly sum ----------------
pb <- ggplot(mapping = aes(year, cu_kg)) +
  # Lines contribute no legend key, so `guides = "collect"` merges panel b's
  # point key with panel a's into a single legend.
  geom_line(
    data = fac, aes(colour = status, group = site_key),
    linewidth = 0.6, show.legend = FALSE
  ) +
  geom_point(data = fac, aes(colour = status), size = 1.4) +
  # Summed series drawn neutral grey so it does not read as a status, with its
  # period mean as a dashed reference line (sum panel only).
  geom_hline(
    data = total_mean, aes(yintercept = cu_kg),
    colour = "grey35", linetype = "dashed", linewidth = 0.4
  ) +
  geom_text(
    data = total_mean,
    aes(x = 2002, y = cu_kg, label = paste0("mean ", round(cu_kg, 1), " kg/yr")),
    hjust = 0, vjust = -0.5, size = 2.6, colour = "grey35"
  ) +
  geom_line(data = fac_total, colour = "grey20", linewidth = 0.6) +
  geom_point(data = fac_total, colour = "grey20", size = 1.4) +
  facet_wrap(~facility, ncol = 3, scales = "free_y") +
  scale_colour_manual(values = status_cols, name = NULL) +
  scale_x_continuous(breaks = seq(2005, 2020, by = 5), limits = c(2002, 2023)) +
  scale_y_continuous(
    limits = c(0, NA), expand = expansion(mult = c(0, 0.1))
  ) +
  labs(x = NULL, y = "Modelled Cu to water (kg/yr)") +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold", size = 9)
  )

fig <- patchwork::wrap_plots(pa, pb, ncol = 2, widths = c(1, 2)) +
  patchwork::plot_layout(guides = "collect") +
  patchwork::plot_annotation(tag_levels = "a", tag_suffix = ")") &
  theme(
    legend.position = "bottom",
    plot.tag = element_text(face = "bold")
  )

# Manuscript figure for _02-methods.qmd (Aquaculture section). Not a target;
# committed like fig01/fig03/fig04 and embedded by static path.
ggsave(
  here::here("figures/fig12-aquaculture-emissions.png"),
  fig,
  width = 13, height = 7, dpi = 200, bg = "white"
)
fig
```

## Caveats

- **Name match only.** The emissions export carries no `loknr`; the join is on
  a normalised site name within Hammerfest kommune. It is 1:1 here, but a future
  register snapshot with two same-named Hammerfest sites would break that.
- **Coordinates come entirely from the register side.** The seven historic
  sites have none; they could be geolocated from Norske utslipp facility pages
  or an older Fiskeridirektoratet extract.
- **All discharge values are modelled** (`B`), consistent with copper being
  back-calculated from antifouling use. Relevant to the Evidence score for any
  aquaculture KTR edge.
- **Operator strings change across years** with company renames (Polarlaks ->
  Cermaq -> "Cermaq Norway Salmon AS" etc.); sites are collapsed on name, not
  operator.
- Not wired into `_targets.R` or the `_quarto.yml` render list. Build by hand
  with `quarto render docs/NBXX-aquaculture.qmd`.
