---
title: "Notebook XX - Copper Sources: PRTR Releases and REACH Declarations"
subtitle: "What the reported-emissions data can and cannot support as AEP source nodes"
execute:
  freeze: true
---

```{r}
#| label: setup
suppressMessages({
  library(dplyr)
  library(ggplot2)
  library(scales)
  library(flextable)
  library(here)
})
here::i_am("README.md")
suppressMessages(pkgload::load_all(here::here(), quiet = TRUE))

prtr <- read_prtr_long(here("inst/extdata/emissions"))
reach <- read_reach_declarations(here("inst/extdata/emissions/REACH_copper_prtd.xlsx"))

sector_pal <- c(
  "Land-based industry" = "#2B5F75",
  "Waste water treatment" = "#5A9367",
  "Offshore oil and gas" = "#B5876B",
  "Landfills" = "#C9B458"
)

fmt <- function(x, digits = 3) {
  formatC(x, format = "g", digits = digits, big.mark = ",")
}

tbl <- function(d, caption = NULL) {
  flextable(d) |>
    colformat_double(digits = 2) |>
    autofit() |>
    set_caption(caption)
}
```

## Why this notebook exists

Sam, 2026-08-06: *"If we're going to do source nodes we need to go to the REACH
PRTR data and work with that rather than adding too much putative stuff that
won't match up."* Agreed. This is that data, plotted, so the source nodes can be
decided from something rather than asserted.

**Read the two datasets as different kinds of number and never add them
together.**

| Dataset | What it measures | Unit | Geography |
|---|---|---|---|
| `norske_utslipp_*.xlsx` | **Releases** to air, water, subsurface | kg/yr | Facility, with fylke and kommune |
| `REACH_copper_prtd.xlsx` | Copper **in commerce**: imported + produced − exported | tonnes/yr | National only |

REACH sits upstream of any emission. A tonne declared is not a tonne released,
and the ratio between them is the thing nobody in this project has estimated.

::: {.callout-important}
## Two traps, both hit while building this

**Sum within a year, then average over years.** The first aggregate here was the
mean of individual facility-year values, labelled as the sector release. For
land-based industry to water that is 149 kg/yr against a true national annual
total of 12,848: a factor of 25, in the direction that makes a source look
negligible. `summarise_prtr_releases()` now returns both, named differently.

**A blank cell is "not reported", not zero.** Facilities report only the media
they are required to. Averaging an unreported medium in as a measured zero drags
every sector mean down by however many facilities happen not to report it.
:::

## What is in the files

```{r}
#| label: tbl-coverage
#| tbl-cap: "Coverage of the PRTR facility files."
prtr |>
  summarise(
    `Facilities` = n_distinct(facility),
    `Facility-years` = n(),
    `First year` = min(year),
    `Last year` = max(year),
    `Media` = paste(sort(unique(medium)), collapse = ", "),
    .by = source_category
  ) |>
  rename(`Source category` = source_category) |>
  arrange(desc(`Facility-years`)) |>
  tbl()
```

## National releases

```{r}
#| label: fig-national-totals
#| fig-cap: "Annual national copper release by sector and medium. Points are years; the line is the reported total across all facilities in that year. Note the log y axis: the sectors span three orders of magnitude."
#| fig-height: 6
annual <- prtr_annual_totals(prtr)

ggplot(annual, aes(year, total_kg, colour = source_category)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.1) +
  scale_colour_manual(values = sector_pal, name = NULL) +
  scale_y_log10(labels = label_number(big.mark = ",")) +
  facet_wrap(~medium, ncol = 1, scales = "free_y") +
  labs(
    x = NULL, y = "Reported release (kg/yr)",
    caption = "Norwegian PRTR (norske utslipp). Log y axis."
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", panel.grid.minor = element_blank())
```

::: {.callout-warning}
## A falling line is not necessarily falling emissions

Reporting completeness changes over time. A sector's total can drop because
fewer facilities reported, not because less copper was released. The panel below
puts facility count beside the total so the two can be told apart before anyone
writes "copper releases have declined" in the manuscript.

**Look at the last point on the land-based industry lines.** It falls roughly a
hundredfold in a single year across a hundred-odd facilities. That is a
submission deadline, not an environmental event: the most recent year is
routinely partial because facilities report on a lag. The same thing happens at
the *start* of the landfills series, where the reporting obligation was phasing
in.

Those years are **excluded** from every magnitude in this notebook, not merely
flagged. A partial year enters a mean as a low-emission year, so leaving it in
understates the source. This was found by plotting the series; it was invisible
in the aggregate table.
:::

```{r}
#| label: tbl-incomplete
#| tbl-cap: "Years excluded as incomplete reporting. A year is flagged only where the group typically has at least five reporters, this year's reporter count fell below half the usual, and the total fell below a quarter of the usual. All three are required, so a genuinely clean year is not mistaken for a missing one."
prtr_complete_years(prtr) |>
  filter(!complete) |>
  transmute(
    `Source category` = source_category,
    Medium = medium,
    Year = year,
    `Reported total (kg)` = total_kg,
    `Facilities reporting` = n_facilities
  ) |>
  tbl()
```

::: {.callout-note}
## The rule is gated on group size, and that gate matters

The first version of this detector used only "is the total far below the
median?", and applied to the four Hammerfest facilities it dropped **6 of 13
years**. With that few reporters an individual facility *is* the series, so a low
year is a low year rather than a missing report. `prtr_complete_years()` now
refuses to judge completeness below five typical reporters, which is why nothing
is dropped from the Repparfjorden figures below.
:::

```{r}
#| label: fig-reporting-completeness
#| fig-cap: "Number of facilities reporting, by sector and year. Compare against the totals above before reading any trend as a real change in release."
#| fig-height: 5
ggplot(annual, aes(year, n_facilities, colour = source_category)) +
  geom_step(linewidth = 0.7) +
  scale_colour_manual(values = sector_pal, name = NULL) +
  facet_wrap(~medium, ncol = 1, scales = "free_y") +
  labs(x = NULL, y = "Facilities reporting") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", panel.grid.minor = element_blank())
```

```{r}
#| label: tbl-national
#| tbl-cap: "National releases. `total_kg_yr` is the annual national total averaged over reported years, and is the figure an AEP source node should carry. `mean_kg_yr` describes a typical facility and is much smaller."
national <- summarise_prtr_releases(prtr, c("source_category", "medium"))

national |>
  transmute(
    `Source category` = source_category,
    Medium = medium,
    `Total kg/yr` = total_kg_yr,
    `sd (annual)` = sd_total_kg_yr,
    Facilities = n_facilities,
    Years = n_years,
    Period = paste0(year_min, "-", year_max),
    `Mean kg/yr per facility-year` = mean_kg_yr
  ) |>
  tbl()
```

## The distribution behind those totals

```{r}
#| label: fig-facility-spread
#| fig-cap: "Distribution of individual facility-year releases, by sector. Binned density with a boxplot; the log axis spans nine orders of magnitude. This is why the mean facility-year and the sector total are so far apart."
#| fig-height: 5
prtr |>
  filter(value_kg > 0) |>
  ggplot(aes(value_kg, source_category)) +
  geom_violin(fill = "grey85", colour = "grey55", linewidth = 0.3) +
  geom_boxplot(width = 0.14, outlier.size = 0.4, fill = "white") +
  # label_log(), NOT label_number(). Across nine orders of magnitude
  # label_number() produces "0.000 / 0.010 / 100.000" and three labels for the
  # whole axis, which is worse than no axis at all.
  scale_x_log10(breaks = 10^(-4:5), labels = label_log(), minor_breaks = NULL) +
  facet_wrap(~medium, ncol = 1, scales = "free_y") +
  labs(
    x = "Reported release (kg/yr, log scale)", y = NULL,
    caption = "Zero and unreported values excluded. One row per facility per year."
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())
```

**A handful of facilities dominate every sector.** That matters for the AEP in a
specific way: a source node's magnitude is a national total, but the exposure it
drives is local to wherever those few facilities are. The national node is
honest about total load and says nothing about where it lands.

```{r}
#| label: tbl-top-facilities
#| tbl-cap: "The twelve largest single facilities by mean annual release to water."
prtr |>
  filter(medium == "Water") |>
  summarise(
    `Mean kg/yr` = mean(value_kg),
    Years = n_distinct(year),
    .by = c(facility, fylke, source_category)
  ) |>
  arrange(desc(`Mean kg/yr`)) |>
  head(12) |>
  rename(Facility = facility, Fylke = fylke, Sector = source_category) |>
  tbl()
```

## Repparfjorden (A002)

The A002 bounding box sits inside **Hammerfest kommune**, which absorbed
Kvalsund in the 2020 municipal reform. Both names are matched, or the pre-2020
rows are silently lost.

```{r}
#| label: tbl-hammerfest
#| tbl-cap: "Every PRTR facility reporting copper in Hammerfest or Kvalsund kommune."
hf <- filter_prtr_kommune(prtr)

hf |>
  summarise(
    `Mean kg/yr` = mean(value_kg),
    `sd` = sd(value_kg),
    Years = n_distinct(year),
    Period = paste0(min(year), "-", max(year)),
    .by = c(facility, source_category, medium)
  ) |>
  arrange(desc(`Mean kg/yr`)) |>
  rename(Facility = facility, Sector = source_category, Medium = medium) |>
  tbl()
```

```{r}
#| label: fig-hammerfest-series
#| fig-cap: "Reported copper release by Hammerfest facilities over time. Points are individual reported years; gaps are years with no report, not years with no release."
#| fig-height: 5
# A REPORTED ZERO IS NOT PLOTTABLE ON A LOG AXIS and vanishes silently, taking
# its line segment with it. Counted and stated rather than dropped quietly.
n_zero <- sum(hf$value_kg <= 0)

# COLOUR BY FACILITY, NOT BY SECTOR. With five facilities sharing two sector
# colours, the connecting lines read as though they join different facilities.
# Names are trimmed because the register spells them out in full, parenthetical
# operator and all.
short <- function(x) {
  stringr::str_trunc(
    stringr::str_to_sentence(sub(" [(].*[)]$", "", x)), 38
  )
}

hf_plot <- hf |>
  filter(value_kg > 0) |>
  mutate(facility_short = short(facility))

ggplot(hf_plot, aes(year, value_kg, colour = facility_short, shape = medium)) +
  geom_line(aes(group = interaction(facility_short, medium)), linewidth = 0.5, alpha = 0.7) +
  geom_point(size = 2.4) +
  scale_colour_brewer(palette = "Dark2", name = NULL) +
  scale_shape_manual(values = c(Air = 17, Water = 16), name = NULL) +
  scale_y_log10(breaks = 10^(-3:2), labels = label_log(), minor_breaks = NULL) +
  guides(colour = guide_legend(ncol = 1, order = 1)) +
  labs(
    x = NULL, y = "Reported release (kg/yr, log scale)",
    caption = if (n_zero > 0) {
      paste(
        n_zero, "reported zero value(s) omitted: not plottable on a log axis.",
        "A reported zero is a declared nil return, not a missing report."
      )
    } else {
      NULL
    }
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right", panel.grid.minor = element_blank())
```

::: {.callout-important}
## The reported releases are not the dominant historic source, and the diagram must say so

```{r}
#| label: hammerfest-total
#| output: asis
hf_water <- summarise_prtr_releases(
  filter(hf, medium == "Water"), "medium"
)
cat(sprintf(
  "All reported copper to water in Hammerfest kommune is **%s kg/yr** across %d facilities (%d-%d).\n",
  fmt(hf_water$total_kg_yr), hf_water$n_facilities,
  hf_water$year_min, hf_water$year_max
))
```

Repparfjorden received on the order of a **million tonnes of tailings** over
seven years of submarine tailings disposal in the 1970s (Kvassnes & Iversen
2013). A source node built from PRTR alone would say industry contributes
essentially nothing to this fjord. That is true of *reported* emissions and
badly false of the system.

Hence **N008, submarine tailing disposal**, as a separate `external` node.
Two things about it still need resolving, and both are unit questions rather
than data questions:

1. **A million tonnes of what?** Tailings mass, not copper mass. A source node's
   magnitude should be copper. Repparfjorden ore ran roughly 0.7% Cu and
   tailings retain some fraction of that, so the copper figure is likely three
   orders of magnitude smaller than the tailings figure. Transcribing the
   tailings tonnage into `external_value` and labelling it `tonnes` will read as
   a copper release.
2. **A seven-year total is not an annual rate.** Every PRTR node here is kg/yr.
   A total and a rate on the same diagram will be compared as though they were
   the same quantity.
:::

## REACH declarations

```{r}
#| label: fig-reach
#| fig-cap: "Declared net copper quantity by year: imported + produced - exported, for declared chemicals. This is copper in commerce, not copper released."
#| fig-height: 4
ry <- reach_complete_years(reach)

ggplot(ry, aes(year, tonnes, fill = complete)) +
  geom_col(width = 0.65) +
  scale_fill_manual(
    values = c(`TRUE` = "#2B5F75", `FALSE` = "#B5876B"),
    labels = c(`TRUE` = "Plausibly complete", `FALSE` = "Suspect: far below median"),
    name = NULL
  ) +
  scale_y_continuous(labels = label_number(big.mark = ",", scale = 1e-3, suffix = " kt")) +
  labs(x = NULL, y = "Declared net copper (tonnes/yr)") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", panel.grid.minor = element_blank())
```

::: {.callout-warning}
## Do not average across the break

```{r}
#| label: reach-note
#| output: asis
ok <- ry |> filter(complete)
cat(sprintf(
  paste(
    "Years flagged plausibly complete (%s) average **%s tonnes/yr** (sd %s).",
    "The remaining years (%s) sit at %s tonnes, a %sx step rather than a trend.\n"
  ),
  paste(ok$year, collapse = ", "),
  fmt(mean(ok$tonnes)), fmt(sd(ok$tonnes)),
  paste(ry$year[!ry$complete], collapse = ", "),
  paste(fmt(ry$tonnes[!ry$complete]), collapse = " and "),
  fmt(mean(ok$tonnes) / mean(ry$tonnes[!ry$complete]), 2)
))
```

A sixfold drop between consecutive years is far more likely to be incomplete
recent reporting than a real collapse in Norwegian copper imports. **Check
whether the recent years are fully reported before using any mean of this
series.** `reach_complete_years()` detects the cut from the data rather than
hard-coding a year, so it moves if the file is refreshed.
:::

```{r}
#| label: tbl-reach-sector
#| tbl-cap: "Declared net copper by sector, plausibly complete years only."
reach |>
  filter(year %in% ry$year[ry$complete]) |>
  summarise(
    `Mean tonnes/yr` = mean(netto_tonn),
    `sd` = sd(netto_tonn),
    Years = n_distinct(year),
    .by = sector
  ) |>
  arrange(desc(`Mean tonnes/yr`)) |>
  mutate(sector = coalesce(sector, "Unclassified")) |>
  rename(Sector = sector) |>
  tbl()
```

## What this supports as source nodes

`scripts/scaffold_prtr_nodes.R` appends these as `external` nodes with
`external_value`, `external_sd`, `external_n` and `external_unit` filled and
every EPEQ score left blank. Scores and coordinates are judgement and are not
scaffolded.

```{r}
#| label: tbl-proposed
#| tbl-cap: "Proposed source nodes. Magnitudes are annual totals in kg/yr."
bind_rows(
  national |>
    transmute(
      AEP = "A001",
      Label = paste0(source_category, " releases to ", tolower(medium)),
      `kg/yr` = total_kg_yr, sd = sd_total_kg_yr, Years = n_years
    ),
  summarise_prtr_releases(hf, "medium") |>
    transmute(
      AEP = "A002",
      Label = paste0("Reported point sources to ", tolower(medium)),
      `kg/yr` = total_kg_yr, sd = sd_total_kg_yr, Years = n_years
    )
) |>
  tbl()
```

**Still missing, and not fillable from this data:**

- **Aquaculture antifouling application (N007).** Not a reported PRTR release, so
  nothing here can fill it. It is currently `node_type = empirical` with no
  members, which is why it warns on every build and renders "no measured data".
  Either it becomes `external` with a figure derived from the aquaculture
  register, or it is dropped.
- **The ratio between declared and released copper.** REACH says roughly 60 kt
  enters Norwegian commerce annually; PRTR says roughly 22 t is reported
  released to water and air. That is a factor of about 2,700, and the gap is
  stock accumulation, unreported diffuse release, and export in products. It is
  the most interesting number in this notebook and it is not one this dataset
  can settle.
