---
title: "EXPECT Copper AEP"
author: "Sam Welch"
date: "2025.11.25"
lightbox: true
format: 
  html:
    grid: 
      body-width: 1500px
      sidebar-width: 0px
    page-layout: full
    code-fold: true
    max-width: 4500px
    fig-format: svg
    toc: true
execute:
  debug: true
preview: 
  port: 4200
---

```{r}
#| label: targets-setup
#| warning: false
#| output: false
library(STOPeData)
library(DT)
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
library(viridis)
library(stringr)
library(ggridges)
library(ggh4x)
library(tidygraph)
library(ggraph)
library(ggiraph)
library(leaflet)
library(sf)
library(shiny)
library(dplyr)
library(bslib)
library(ggtext)

source("_targets.R")
tar_make()


# load in any data we need from the targets workflow
literature_merged_data <- tar_read(load_literature_pqt)
literature_reference_data <- tar_read(reference_data)
literature_campaign_data <- tar_read(campaign_data)
literature_sites_data <- tar_read(sites_data)
literature_qc <- tar_read(data_quality_report)
wgs84_geo <- tar_read(wgs84_geography)

species_names <- arrow::read_parquet("data/clean/ecotox_2025_06_12_species.parquet")
species_lookup <- species_names |>
  with(setNames(common_name, latin_name))

copper_toxicity_thresholds <- tar_read(copper_toxicity_thresholds)
# Define threshold colors based on the report ----
threshold_colours <- c(
  "Background (I)" = "#0070C0", # Blue
  "Good (II)" = "#00B050", # Green
  "Moderate (III)" = "#FFC000", # Yellow
  "Poor (IV)" = "#FF6600", # Orange
  "Very Poor (V)" = "#FF0000" # Red
)

compartment_colours <- c(
  # Aquatic ----
  "Freshwater" = "#4A90E2", # clear blue
  "Marine/Salt Water" = "#1E5A8E", # deep ocean blue
  "Brackish/Transitional Water" = "#6BA5C8", # mid-blue (between fresh/marine)
  "Groundwater" = "#2C5F7B", # darker, subdued blue
  "Wastewater" = "#8B7355", # murky brown
  "Liquid Growth Medium" = "#9FD8CB", # pale greenish-blue
  "Rainwater" = "#B3D9FF", # light sky blue
  "Stormwater" = "#607D8B", # grey-blue
  "Leachate" = "#6B5344", # dark muddy brown
  "Aquatic Sediment" = "#EBCF1B", # your existing yellow
  "Porewater" = "#7FA0B8", # muted blue-grey
  "Sludge" = "#67AB33", # your existing green

  # Atmospheric ----
  "Indoor Air" = "#E8F4F8", # very pale blue-grey
  "Outdoor Air" = "#87CEEB", # sky blue

  # Terrestrial ----
  "Terrestrial Biological Residue" = "#8B6F47", # organic brown
  "Soil H Horizon (Peat)" = "#3E2723", # very dark brown
  "Soil O Horizon (Organic)" = "#5D4037", # dark organic brown
  "Soil A Horizon (Topsoil)" = "#795548", # medium brown
  "Soil E Horizon (Mineral)" = "#A1887F", # light greyish-brown
  "Soil S Horizon (Mineral)" = "#BCAAA4", # pale brown-grey
  "Soil C Horizon (Parent Material)" = "#D7CCC8", # very pale brown
  "Soil R Horizon (Bedrock)" = "#9E9E9E", # grey

  # Biota ----
  "Biota, Terrestrial" = "#558B2F", # forest green
  "Biota, Aquatic" = "#00695C", # teal
  "Biota, Atmospheric" = "#B39DDB", # light purple (birds/insects)
  "Biota, Other" = "#9C27B0" # distinct purple
)

# Helper function for sample sizes
calculate_sample_sizes <- function(data) {
  data |>
    group_by(
      REFERENCE_ID,
      OCEAN_COUNTRY,
      SITE_GEOGRAPHIC_FEATURE,
      ENVIRON_COMPARTMENT_SUB
    ) |>
    summarise(total_n = sum(MEASURED_N, na.rm = TRUE), .groups = "drop")
}


```

## Dataset Overview

### To Do

-   [ ] Add probable contamination sources to each paper
-   [ ] Make sure we handle missing characters properly
-   [ ] Reach out to Dag or Jens to ask if MD have any copper emissions data
-   [ ] Talk to Marianne from reference group (or anyone else on the reference group)
-   [ ] Get more data in 
-   [ ] Make network diagram visualise uncertainty, etc.
-   [ ] Functional and risk assessment trophic levels may not be perfectly aligned 
    -   [ ] might be better off using ERA grouping - talk to Li about it
-   [ ] Fix icon colouration in map

## Data Quality Overview {#sec-data-quality}

```{r}
#| label: quality-summary
#| echo: false
#| output: asis

#' Generate a text summary of data quality issues
#'
#' @param qc_report Output from check_data_quality()
#' @return A character string with markdown-formatted summary
format_quality_summary <- function(qc_report) {
  s <- qc_report$summary

  glue::glue(
    "
Dataset contains **{s$total_rows}** rows from **{s$total_references}** references across **{s$total_sites}** sites.

**Issues identified:**

- **Measurements**: {s$n_rows_missing_measurements} rows ({s$n_refs_missing_measurements} references) missing all measurement data (value, LOQ, and LOD)
- **Sample size**: {s$n_rows_missing_n} rows ({s$n_refs_missing_n} references) missing or zero sample size
- **Methods**: {s$n_rows_missing_methods} rows ({s$n_refs_missing_methods} references) missing method information
- **Uncertainty**: {s$n_rows_missing_uncertainty} rows ({s$n_refs_missing_uncertainty} references) missing uncertainty type
- **Site data**: {s$n_sites_missing_data} sites ({s$n_refs_missing_sites} references) missing location/geographic data
- **Biota data**: {s$n_rows_missing_biota} rows ({s$n_refs_missing_biota} references) missing species/tissue information
"
  )
}

cat(format_quality_summary(literature_qc))
```

::: {.callout-note collapse="true"}
## Missing Data

### Missing Measurements

Rows where measured value, LOQ, and LOD are all missing.

Known issues:

-   2025PelkonenEnvironmentalImpactOf does not report LOD/LOQ values in main paper or ESI.

```{r}
#| label: missing-measurements
#| echo: false

if (nrow(literature_qc$missing_measurements) > 0) {
  DT::datatable(
    literature_qc$missing_measurements |>
      dplyr::select(-sample_ids),
    options = list(pageLength = 10),
    caption = "References with missing measurement data"
  )
} else {
  cat("*No missing measurements found.*")
}
```

### Missing Sample Size

Rows where MEASURED_N is missing or zero.

```{r}
#| label: missing-n
#| echo: false

if (nrow(literature_qc$missing_n) > 0) {
  DT::datatable(
    literature_qc$missing_n |>
      dplyr::select(-sample_ids),
    options = list(pageLength = 10),
    caption = "References with missing sample size"
  )
} else {
  cat("*No missing sample sizes found.*")
}
```

### Missing Methods

Rows missing any of: analytical protocol, extraction protocol, fractionation protocol, or sampling protocol.

```{r}
#| label: missing-methods
#| echo: false

#| label: missing-methods
#| echo: false

if (nrow(literature_qc$missing_methods) > 0) {
  display_df <- literature_qc$missing_methods |>
    dplyr::mutate(
      analytical_issue = dplyr::case_when(
        missing_analytical ~ "Missing",
        .default = NA_character_
      ),
      extraction_issue = dplyr::case_when(
        missing_extraction ~ "Missing",
        .default = NA_character_
      ),
      fractionation_issue = dplyr::case_when(
        missing_fractionation ~ "Missing",
        .default = NA_character_
      ),
      sampling_issue = dplyr::case_when(
        missing_sampling ~ "Missing",
        .default = NA_character_
      )
    ) |>
    dplyr::select(
      REFERENCE_ID,
      n_rows,
      analytical_issue,
      extraction_issue,
      fractionation_issue,
      sampling_issue
    )

  DT::datatable(
    display_df,
    options = list(pageLength = 10),
    caption = "References with missing method information"
  )
} else {
  cat("*No missing methods found.*")
}
```

### Missing Uncertainty

Rows where UNCERTAINTY_TYPE or UNCERTAINTY\_\*\_STANDARD are missing.

```{r}
#| label: missing-uncertainty
#| echo: false

if (nrow(literature_qc$missing_uncertainty) > 0) {
  display_df <- literature_qc$missing_uncertainty |>
    dplyr::mutate(
      uncertainty_type_issue = dplyr::case_when(
        type_missing ~ "Missing",
        type_not_reported ~ "Not reported",
        type_not_relevant ~ "Not relevant",
        .default = NA_character_
      ),
      upper_bound_issue = dplyr::case_when(
        upper_missing ~ "Missing",
        upper_zero ~ "Zero",
        upper_below_value ~ "< Average",
        .default = NA_character_
      ),
      lower_bound_issue = dplyr::case_when(
        lower_missing ~ "Missing",
        lower_zero ~ "Zero",
        lower_above_value ~ "> Average",
        .default = NA_character_
      )
    ) |>
    dplyr::select(
      REFERENCE_ID,
      n_rows,
      uncertainty_type_issue,
      upper_bound_issue,
      lower_bound_issue
    )

  DT::datatable(
    display_df,
    options = list(pageLength = 10),
    caption = "References with missing or problematic uncertainty"
  )
} else {
  cat("*No uncertainty issues found.*")
}


# tar_read(measurements_data) |>
#   filter(str_detect(REFERENCE_ID, "Olsvik")) |>
#   select(
#     MEASURED_VALUE,
#     UNCERTAINTY_UPPER,
#     UNCERTAINTY_LOWER,
#     MEASURED_UNIT,
#     SAMPLE_ID
#   ) |>
#   view()

# tar_read(literature_joined) |>
#   filter(str_detect(REFERENCE_ID, "Olsvik")) |>
#   select(
#     MEASURED_VALUE,
#     UNCERTAINTY_UPPER,
#     UNCERTAINTY_LOWER,
#     MEASURED_UNIT,
#     SAMPLE_ID
#   ) |>
#   view()

# tar_read(literature_clean) |>
#   filter(str_detect(REFERENCE_ID, "Olsvik")) |>
#   select(
#     MEASURED_VALUE,
#     UNCERTAINTY_UPPER,
#     UNCERTAINTY_LOWER,
#     MEASURED_UNIT,
#     SAMPLE_ID
#   ) |>
#   view()

# tar_read(literature_clean) |>
#   filter(str_detect(REFERENCE_ID, "Olsvik")) |>
#   select(
#     MEASURED_VALUE,
#     UNCERTAINTY_UPPER,
#     UNCERTAINTY_LOWER,
#     MEASURED_UNIT,
#     SAMPLE_ID
#   ) |>
#   standardise_measured_units(
#     value_columns = c(
#       "MEASURED_VALUE",
#       "UNCERTAINTY_UPPER",
#       "UNCERTAINTY_LOWER"
#     ),
#     unit_column = "MEASURED_UNIT"
#   ) |>
#   view()
```

### Missing Site Data

Sites missing coordinates, geographic features, country, or ocean information.

```{r}
#| label: missing-sites
#| echo: false

if (nrow(literature_qc$missing_sites) > 0) {
  DT::datatable(
    literature_qc$missing_sites,
    options = list(pageLength = 10),
    caption = "Sites with missing geographic data"
  )
} else {
  cat("*No missing site data found.*")
}
```

### Missing Biota Data

Biota samples missing species, tissue, lifestage, gender, or species group.

```{r}
#| label: missing-biota
#| echo: false

if (nrow(literature_qc$missing_biota) > 0) {
  DT::datatable(
    literature_qc$missing_biota |>
      dplyr::select(-sample_ids),
    options = list(pageLength = 10),
    caption = "References with missing biota information"
  )
} else {
  cat("*No missing biota data found.*")
}
```
:::

## Literature Summary

@tbl-paper-summaries provides a comprehensive summary of all papers included in this analysis, showing the number of measurements extracted from each study and their associated sampling campaigns.

```{r}
#| label: tbl-paper-summaries
#| tbl-cap: "Summary of analysed papers showing publication details, associated campaigns, and number of copper measurements extracted from each study."
#| warning: false
literature_merged_data |>
  dplyr::left_join(literature_reference_data) |>
  dplyr::left_join(
    literature_campaign_data |> select(CAMPAIGN_NAME_SHORT, CAMPAIGN_COMMENT)
  ) |>
  dplyr::group_by(
    TITLE,
    YEAR,
    AUTHOR,
    CAMPAIGN_COMMENT
  ) |>
  dplyr::reframe(SUM_N = sum(MEASURED_N)) |>
  datatable(options = list(pageLength = 5, lengthMenu = c(5, 10, 15, 20)))
```

### Compartment Coverage

```{r}
#| label: fig-paper-subcompartments
#| fig-cap: "Number of environmental sub-compartments sampled per paper. Each bar represents a unique publication, with height indicating the diversity of compartments (e.g., surface water, sediment, soil) analysed within that study."
#| warning: false
#| fig-width: 10
#| fig-height: 8

# Prepare data for plotting ----
plot_data <- literature_merged_data |>
  group_by(REFERENCE_ID, ENVIRON_COMPARTMENT_SUB) |>
  distinct() |>
  reframe(count = sum(MEASURED_N, na.rm = TRUE)) |>
  # Join year data for y-axis ordering
  left_join(
    literature_merged_data |> distinct(REFERENCE_ID, YEAR),
    by = "REFERENCE_ID"
  ) |>
  filter(!is.na(ENVIRON_COMPARTMENT_SUB))

# Calculate text color threshold ----
threshold_value <- mean(plot_data$count, na.rm = TRUE)

# Create plot ----
ggplot(
  plot_data,
  aes(
    y = REFERENCE_ID,
    x = ENVIRON_COMPARTMENT_SUB,
    fill = count
  )
) +
  geom_tile() +
  geom_text(aes(
    label = count,
    color = if_else(count > threshold_value, "black", "white")
  )) +
  scale_fill_viridis(name = "Sample Size (n)") +
  scale_color_identity() +
  theme_aep_plots() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    axis.title.y = element_blank(),
    legend.position = "none"
  ) +
  labs(x = "Environmental Sub-Compartment")
```

### Species Coverage

@fig-paper-species illustrates the taxonomic diversity captured across studies, highlighting which papers focused on biota measurements and the range of species groups investigated.

```{r}
#| label: fig-paper-species
#| fig-cap: "Sample size per species per paper across the dataset."
#| warning: false
#| fig-width: 16
#| fig-height: 8

# Prepare data for plotting ----
plot_data <- literature_merged_data |>
  group_by(REFERENCE_ID, SAMPLE_SPECIES) |>
  distinct() |>
  reframe(count = sum(MEASURED_N, na.rm = TRUE), SPECIES_GROUP) |>
  # Join year data for y-axis ordering
  left_join(
    literature_merged_data |> distinct(REFERENCE_ID, YEAR),
    by = "REFERENCE_ID"
  ) |>
  filter(!is.na(SAMPLE_SPECIES)) |>  distinct()

# Create formatted labels
plot_data <- plot_data |>
  mutate(
    formatted_species = glue::glue(
      "**{species_lookup[SAMPLE_SPECIES]}**<br>*({SAMPLE_SPECIES})*"
    )
  )


# Calculate text color threshold ----
threshold_value <- 2000 # set manually, as the 2050 mussels tend to overwhelm the others

# Create plot ----

ggplot(
  plot_data,
  aes(
    y = REFERENCE_ID,  # Use the formatted label
    x = formatted_species,
    fill = count
  )
) +
  geom_tile() +
  geom_text(aes(
    label = count,
    color = if_else(count > threshold_value, "black", "white")
  )) +
  scale_fill_viridis(name = "Sample Size (n)") +
  scale_color_identity() +
  # theme_aep_plots() + # breaks markdown
  theme(
    axis.text.x = element_markdown(angle = 90, hjust = 1),  # Enable markdown rendering
    axis.title.y = element_blank(),
    legend.position = "none"
  ) +
  labs(x = "Sampled Species") +
  facet_wrap(
    facets = vars(SPECIES_GROUP),
    space = "free_x",
    scales = "free_x",
    nrow = 1
  )
```

## Geographic Distribution of Sampling Sites

@fig-sampling-map shows the spatial distribution of all sampling locations included in the analysis. This map helps visualise the geographic scope of copper monitoring in Arctic regions and identify areas with high or low sampling intensity.

```{r}
#| label: fig-sampling-map
#| warning: false

# Turn off spherical geometry ----
sf::sf_use_s2(FALSE)

# Prepare spatial data ----
data_sf <- literature_merged_data |>
  filter(
    !is.na(LATITUDE),
    !is.na(LONGITUDE),
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  st_as_sf(
    coords = c("LONGITUDE", "LATITUDE"),
    crs = st_crs("WGS84"),
    remove = FALSE)

  # Prepare apple data ----
  number_of_sites <- nrow(municipality_centroids())
  sales_data_split <- copper_oxide_sales() |> mutate(amount_kg = amount_kg / number_of_sites,
  average_kg = average_kg / number_of_sites)

  apple_data <- cross_join(municipality_centroids(), sales_data_split) |>
  group_by(name, x, y) |>
  summarise(
    avg_copper_kg = mean(amount_kg, na.rm = TRUE),
    .groups = "drop"
  ) |>
  st_as_sf(
    coords = c("x", "y"),
    crs = st_crs("WGS84"),
    remove = FALSE
  )

environ_compartments_vocabulary <- environ_compartments_vocabulary() |>
  setdiff(c("Not relevant", "Not reported", "Other"))

# Create colour palette for main compartments ----
compartment_pal <- colorFactor(
  palette = palette.colors(n = 4, palette = "Tableau"),
  domain = environ_compartments_vocabulary
)

# Create colour palette for measured values ----
value_range <- range(data_sf$MEASURED_VALUE_STANDARD, na.rm = TRUE)
value_pal <- colorNumeric(
  palette = "viridis",
  domain = value_range,
  na.color = "gray"
)

# Create sub-compartment letter codes ----
subcomp_codes <- c(
  # Aquatic
  "Freshwater" = "F",
  "Marine/Salt Water" = "M",
  "Brackish/Transitional Water" = "B",
  "Groundwater" = "G",
  "Wastewater" = "WW",
  "Liquid Growth Medium" = "LGM",
  "Rainwater" = "R",
  "Stormwater" = "SW",
  "Leachate" = "L",
  "Aquatic Sediment" = "AS",
  "Porewater" = "P",
  "Sludge" = "Sl",
  # Atmospheric
  "Indoor Air" = "IA",
  "Outdoor Air" = "OA",
  # Terrestrial
  "Terrestrial Biological Residue" = "TBR",
  "Soil H Horizon (Peat)" = "H",
  "Soil O Horizon (Organic)" = "O",
  "Soil A Horizon (Topsoil)" = "A",
  "Soil E Horizon (Mineral)" = "E",
  "Soil S Horizon (Mineral)" = "S",
  "Soil C Horizon (Parent Material)" = "C",
  "Soil R Horizon (Bedrock)" = "R",
  # Biota
  "Biota, Terrestrial" = "BT",
  "Biota, Aquatic" = "BA",
  "Biota, Atmospheric" = "BAm",
  "Biota, Other" = "BO"
)

# Add codes to data ----
data_sf <- data_sf |>
  mutate(
    SUBCOMP_CODE = subcomp_codes[ENVIRON_COMPARTMENT_SUB],
    SUBCOMP_CODE = ifelse(is.na(SUBCOMP_CODE), "?", SUBCOMP_CODE)
  )

# Define map colours ----
marine_colors <- list(
  default = "lightblue",
  highlight = "darkblue"
)

country_colours <- list(
  default = "lightgreen",
  highlight = "darkgreen"
)

# Calculate polygon centroids for labels ----
marine_centroids <- wgs84_geo$marine_polys |>
  filter(highlight_name) |>
  st_centroid()

country_centroids <- wgs84_geo$countries |>
  filter(
    highlight_name,
    !name %in% c("Greenland", "Iceland", "Norway")
  ) |>
  st_centroid()

# Create study connection lines ----
study_lines <- data_sf |>
  st_drop_geometry() |>
  group_by(REFERENCE_ID) |>
  filter(n() > 1) |>
  summarise(
    coords = list(cbind(LONGITUDE, LATITUDE)),
    .groups = "drop"
  )

# Convert to linestring sf object ----
study_lines_sf <- study_lines |>
  rowwise() |>
  mutate(
    geometry = list(st_linestring(coords))
  ) |>
  ungroup() |>
  select(-coords) |>
  st_sf(crs = st_crs("WGS84"))

# Create map ----
map <- leaflet() |>
  addProviderTiles(providers$CartoDB.Positron) |>

  # Add ocean polygons (no popups) ----
  addPolygons(
    data = wgs84_geo$marine_polys |> filter(highlight_name),
    fillColor = ~ ifelse(
      highlight_name == TRUE,
      marine_colors[["highlight"]],
      marine_colors[["default"]]
    ),
    fillOpacity = ~ ifelse(highlight_name == TRUE, 0.3, 0),
    color = "black",
    weight = 1,
    group = "Geography"
  ) |>

  # Add country polygons (no popups) ----
  addPolygons(
    data = wgs84_geo$countries |> filter(highlight_name),
    fillColor = ~ ifelse(
      highlight_name == TRUE,
      country_colours[["highlight"]],
      country_colours[["default"]]
    ),
    fillOpacity = ~ ifelse(highlight_name == TRUE, 0.3, 0),
    weight = 1,
    color = "black",
    group = "Geography"
  ) |>

  # Add permanent labels for oceans ----
  addLabelOnlyMarkers(
    data = marine_centroids,
    label = ~name,
    labelOptions = labelOptions(
      noHide = TRUE,
      direction = "center",
      textOnly = TRUE,
      style = list(
        "color" = "darkblue",
        "font-weight" = "bold",
        "font-size" = "14px",
        "text-shadow" = "1px 1px 2px white, -1px -1px 2px white"
      )
    ),
    group = "Geography"
  ) |>

  # Add permanent labels for countries ----
  addLabelOnlyMarkers(
    data = country_centroids,
    label = ~name,
    labelOptions = labelOptions(
      noHide = TRUE,
      direction = "center",
      textOnly = TRUE,
      style = list(
        "color" = "darkgreen",
        "font-weight" = "bold",
        "font-size" = "12px",
        "text-shadow" = "1px 1px 2px white, -1px -1px 2px white"
      )
    ),
    group = "Geography"
  ) |>

  # Add study connection lines ----
  addPolylines(
    data = study_lines_sf,
    color = "purple",
    weight = 2,
    opacity = 0.5,
    dashArray = "5, 5",
    label = ~ paste0("Study: ", REFERENCE_ID),
    labelOptions = labelOptions(
      style = list("font-weight" = "normal")
    ),
    group = "Study Connections"
  ) |>

  # Add mine tailings markers ----
  addMarkers(
    data = mine_tailings_coords(),
    lng = ~LONGITUDE,
    lat = ~LATITUDE,
    icon = list(
      iconUrl = "https://upload.wikimedia.org/wikipedia/commons/8/8c/Mining_symbol.svg",
      iconSize = c(15, 15)
    ),
    popup = ~ paste0(
      "<b>Mine:</b> ",
      MINE_NAME,
      "<br>",
      "<b>Location:</b> ",
      SITE_NAME,
      "<br>",
      "<b>Ore Type:</b> ",
      ORE_TYPE,
      "<br>",
      "<b>Status:</b> ",
      ifelse(ACTIVE_2013, "Active (2013)", "Terminated"),
      "<br>",
      "<b>Details:</b> ",
      KEY_DETAILS,
      "<br>",
      "<b>Coordinate Confidence:</b> ",
      CONFIDENCE,
      "/5"
    ),
    group = "Mine Tailings"
  ) |> 

  # apple markers... for apples
  addMarkers(
    data = apple_data,
    lng = ~x,
    lat = ~y,
    icon = list(
      iconUrl = "https://www.svgrepo.com/show/42619/apple-fruit.svg",
      iconSize = c(15, 15)
    ),
    popup = ~ paste0(
      "<b>Municipality:</b> ",
      name,
      "<br>",
      "<b>Average Copper Applied:</b> ",
      round(avg_copper_kg, 1),
      " kg/year",
      "<br>",
      "<b>Period:</b> 2020-2024"
    ),
    group = "Apple Production"
  ) |> 

  # add case study sites
    addCircles(
      data = expect_case_studies(),
    lng = ~lon,
    lat = ~lat,
    radius = 100000,  # 100km radius
    color = "hotpink",
    fillColor = "hotpink",
    fillOpacity = 0.3,
    opacity = 0.6,
    weight = 2,
    popup = ~paste0(
      "<strong>", site_name, "</strong><br/>",
      "<em>Lat: ", lat, ", Lon: ", lon, "</em><br/><br/>",
      summary
    ),
    group =       "Potential Case Studies"

  )

# Add layers for each main compartment ----
for (comp in environ_compartments_vocabulary) {
  data_subset <- data_sf |>
    filter(ENVIRON_COMPARTMENT == comp)

  if (nrow(data_subset) > 0) {
    # Build popup HTML with conditional species info
    popup_html <- if (comp == "Biota") {
      ~ paste0(
        "<b>Reference:</b> ",
        REFERENCE_ID,
        "<br>",
        "<b>Site:</b> ",
        SITE_CODE,
        "<br>",
        "<b>Date:</b> ",
        SAMPLING_DATE,
        "<br>",
        "<b>Compartment:</b> ",
        ENVIRON_COMPARTMENT,
        "<br>",
        "<b>Sub-compartment:</b> ",
        ENVIRON_COMPARTMENT_SUB,
        " (",
        SUBCOMP_CODE,
        ")",
        "<br>",
        "<b>Species:</b> ",
        SAMPLE_SPECIES,
        "<br>",
        "<b>Measured Value:</b> ",
        round(MEASURED_VALUE_STANDARD, 3),
        " ",
        MEASURED_UNIT_STANDARD
      )
    } else {
      ~ paste0(
        "<b>Reference:</b> ",
        REFERENCE_ID,
        "<br>",
        "<b>Site:</b> ",
        SITE_CODE,
        "<br>",
        "<b>Date:</b> ",
        SAMPLING_DATE,
        "<br>",
        "<b>Compartment:</b> ",
        ENVIRON_COMPARTMENT,
        "<br>",
        "<b>Sub-compartment:</b> ",
        ENVIRON_COMPARTMENT_SUB,
        " (",
        SUBCOMP_CODE,
        ")",
        "<br>",
        "<b>Measured Value:</b> ",
        round(MEASURED_VALUE_STANDARD, 3),
        " ",
        MEASURED_UNIT_STANDARD
      )
    }

    map <- map |>
      addCircleMarkers(
        data = data_subset,
        fillColor = ~ value_pal(MEASURED_VALUE_STANDARD),
        color = compartment_pal(comp),
        weight = 2,
        fillOpacity = 0.7,
        radius = 6,
        label = ~SUBCOMP_CODE,
        labelOptions = labelOptions(
          noHide = TRUE,
          direction = "center",
          textOnly = TRUE,
          style = list(
            "color" = compartment_pal(comp),
            "font-weight" = "normal",
            "font-size" = "8px"
          )
        ),
        popup = popup_html,
        group = "Samples"
      )
  }
}

# Add site labels as separate toggleable layer ----
map <- map |>
  addLabelOnlyMarkers(
    data = data_sf,
    label = ~SITE_CODE,
    labelOptions = labelOptions(
      noHide = FALSE,
      direction = "top",
      textOnly = TRUE,
      style = list(
        "color" = "black",
        "font-size" = "9px",
        "background" = "rgba(255, 255, 255, 0.7)",
        "padding" = "2px"
      )
    ),
    group = "Site Labels"
  ) |>

  # Add layer controls ----
  addLayersControl(
    baseGroups = character(0),
    overlayGroups = c(
      "Geography",
      "Study Connections",
      "Mine Tailings",
      "Apple Production",
      "Samples",
      "Site Labels",
      "Potential Case Studies"
    ),
    options = layersControlOptions(collapsed = FALSE)
  ) |>

  # Add legends ----
  addLegend(
    position = "bottomright",
    pal = compartment_pal,
    values = environ_compartments_vocabulary,
    title = "Compartment<br>(border colour)",
    opacity = 1
  ) |>
  addLegend(
    position = "bottomleft",
    pal = value_pal,
    values = data_sf$MEASURED_VALUE_STANDARD,
    title = "Measured Value<br>(fill colour)",
    opacity = 0.8,
    labFormat = labelFormat(digits = 1)
  )

# Display map ----
bslib::card(full_screen = TRUE, map, style = "text-align: left; z-index: 999999;") # yes, of course, that makes perfect sense
```


## Copper Concentrations by Environmental Compartment

Get a basic overview of the orders of magnitude for different sub-compartments (@fig-basic-subcompartments), so we can plot them on scales that make sense. Note that I haven't included uncertainty or sample sizes.

```{r}
#| label: fig-basic-subcompartments

literature_merged_data |> 
  filter(!is.na(MEASURED_UNIT_STANDARD)) |> 
  ggplot(aes(x = MEASURED_VALUE_STANDARD, y = ENVIRON_COMPARTMENT_SUB, color = ENVIRON_COMPARTMENT_SUB)) +
  geom_point() +
  facet_nested(rows = vars(ENVIRON_COMPARTMENT, MEASURED_UNIT_STANDARD), 
  scales = "free_y", 
  space = "free_y", 
  nest_line = element_line()
  ) +
  theme_minimal() +
    scale_color_manual(
    values = compartment_colours,
    breaks = names(compartment_colours)
  ) +
    theme(legend.position = "none") +
    labs(x = "", y = "")
   
```

The following sections present copper concentration distributions for each major environmental compartment. Plots are based on reported means uncertainty bounds (SD, range, 95% CI, etc.) and overlaid with relevant regulatory thresholds where available.

### Aquatic Compartment

#### Freshwater

@fig-boxplot-freshwater shows we have a fair amount of data for water. It's not strictly useful to include saltwater, freshwater and inbetween on here, so I think it makes sense to split it - but it won't make a massive difference to the final product.

Some of these units raise some big questions. Why is brackish/transitional using dry? And why are there some aquatic biota using mg/L?

```{r}
#| label: fig-boxplot-freshwater
#| fig-cap: "Copper concentrations in water column samples by reference"
#| fig-width: 10
# Calculate sample sizes
sample_sizes_water <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB == "Freshwater",
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  calculate_sample_sizes()

# Get first facet for labels
first_facet <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB == "Freshwater",
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  distinct(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE) |>
  slice(1)

# Create plot
literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB == "Freshwater",
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  ggplot(aes(
    x = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    y = REFERENCE_ID,
    colour = ENVIRON_COMPARTMENT_SUB
  )) +
  # ... (all your other geoms remain the same) ...
  geom_crossbar(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "95% Confidence Interval"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    )
  ) +
  geom_linerange(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    linewidth = 0.8
  ) +
  geom_point(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    size = 2
  ) +
  geom_boxplot(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "1st-3rd Quartile"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      lower = UNCERTAINTY_LOWER_STANDARD,
      middle = MEASURED_OR_IMPUTED_VALUE_STANDARD,
      upper = UNCERTAINTY_UPPER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    stat = "identity"
  ) +
  geom_point(
    alpha = 0.5,
    data = \(x) {
      filter(
        x,
        is_not_reported(UNCERTAINTY_TYPE) | is_not_relevant(UNCERTAINTY_TYPE)
      )
    }
  ) +
  geom_text(
    data = sample_sizes_water,
    aes(x = -Inf, y = REFERENCE_ID, label = paste0("n=", total_n)),
    hjust = -1, vjust = -1,
    inherit.aes = FALSE,
    size = 3
  ) +
  geom_vline(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(ENVIRON_COMPARTMENT_SUB %in% sub_comps, THRESHOLD_TYPE == "Classification boundary")
    },
    aes(xintercept = THRESHOLD_VALUE, color = THRESHOLD_CLASS),
    linetype = "dashed",
    linewidth = 0.8,
    alpha = 0.6
  ) +
  # Threshold labels - only in first facet
  geom_text(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(ENVIRON_COMPARTMENT_SUB %in% sub_comps) |>
        cross_join(first_facet)
    },
    aes(x = THRESHOLD_VALUE, y = Inf, label = THRESHOLD_CLASS, color = THRESHOLD_CLASS),
    vjust = 1.5,
    angle = 90,
    hjust = 1,
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_color_manual(
    values = c(threshold_colours, compartment_colours),
    breaks = names(compartment_colours)
  ) +
  labs(
    x = "Concentration (mg/L)",
    y = "Reference",
    colour = "Sub-compartment"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.box = "vertical"
  ) +
  facet_nested(
    rows = vars(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE),
    scales = "free_y",
    space = "free_y",
    shrink = TRUE,
    labeller = label_wrap_gen(width = 15),
    nest_line = element_line()
  ) +
      coord_cartesian(
    xlim = range(
      literature_merged_data$MEASURED_OR_IMPUTED_VALUE_STANDARD[
        literature_merged_data$ENVIRON_COMPARTMENT_SUB == "Freshwater"
      ],
      na.rm = TRUE
    ) |> round(),
    clip = "off"  # allows threshold labels to extend beyond plot area if needed
  ) 
```

#### Salt and Transitional Water

@fig-boxplot-saltwater 

```{r}
#| label: fig-boxplot-saltwater
#| fig-cap: "Copper concentrations in water column samples by reference"
#| fig-width: 10
# Calculate sample sizes
chosen_compartments <- c("Marine/Salt Water", "Brackish/Transitional Water")

sample_sizes_water <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  calculate_sample_sizes()

# Get first facet for labels
first_facet <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  distinct(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE) |>
  slice(1)

# Create plot
literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  ggplot(aes(
    x = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    y = REFERENCE_ID,
    colour = ENVIRON_COMPARTMENT_SUB
  )) +
  # ... (all your other geoms remain the same) ...
  geom_crossbar(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "95% Confidence Interval"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    )
  ) +
  geom_linerange(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    linewidth = 0.8
  ) +
  geom_point(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    size = 2
  ) +
  geom_boxplot(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "1st-3rd Quartile"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      lower = UNCERTAINTY_LOWER_STANDARD,
      middle = MEASURED_OR_IMPUTED_VALUE_STANDARD,
      upper = UNCERTAINTY_UPPER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    stat = "identity"
  ) +
  geom_point(
    alpha = 0.5,
    data = \(x) {
      filter(
        x,
        is_not_reported(UNCERTAINTY_TYPE) | is_not_relevant(UNCERTAINTY_TYPE)
      )
    }
  ) +
  geom_text(
    data = sample_sizes_water,
    aes(x = -Inf, y = REFERENCE_ID, label = paste0("n=", total_n)),
    hjust = -1, vjust = -1,
    inherit.aes = FALSE,
    size = 3
  ) +
  geom_vline(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(ENVIRON_COMPARTMENT_SUB %in% sub_comps, THRESHOLD_TYPE == "Classification boundary")
    },
    aes(xintercept = THRESHOLD_VALUE, color = THRESHOLD_CLASS),
    linetype = "dashed",
    linewidth = 0.8,
    alpha = 0.6
  ) +
  # Threshold labels - only in first facet
  geom_text(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(ENVIRON_COMPARTMENT_SUB %in% sub_comps) |>
        cross_join(first_facet)
    },
    aes(x = THRESHOLD_VALUE, y = Inf, label = THRESHOLD_CLASS, color = THRESHOLD_CLASS),
    vjust = 1.5,
    angle = 90,
    hjust = 1,
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_color_manual(
    values = c(threshold_colours, compartment_colours),
    breaks = names(compartment_colours)
  ) +
  labs(
    x = "Concentration (mg/L)",
    y = "Reference",
    colour = "Sub-compartment"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.box = "vertical"
  ) +
  facet_nested(
    rows = vars(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE),
    scales = "free_y",
    space = "free_y",
    shrink = TRUE,
    labeller = label_wrap_gen(width = 15),
    nest_line = element_line()
  ) +
      coord_cartesian(
    xlim = range(
      literature_merged_data$MEASURED_OR_IMPUTED_VALUE_STANDARD[
        literature_merged_data$ENVIRON_COMPARTMENT_SUB %in% chosen_compartments
      ],
      na.rm = TRUE
    ) |> round(),
    clip = "off"  # allows threshold labels to extend beyond plot area if needed
  ) 
```

#### Aquatic Sediment

@fig-boxplot-sediment

```{r}
#| label: fig-boxplot-sediment
#| fig-cap: "Copper concentrations in aquatic sediment samples by reference"
#| fig-width: 10
# Calculate sample sizes
chosen_compartments <- c("Aquatic Sediment")

sample_sizes_water <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  calculate_sample_sizes()

# Get first facet for labels
first_facet <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  distinct(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE) |>
  slice(1)

# Create plot
literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  ggplot(aes(
    x = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    y = REFERENCE_ID,
    colour = ENVIRON_COMPARTMENT_SUB
  )) +
  # ... (all your other geoms remain the same) ...
  geom_crossbar(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "95% Confidence Interval"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    )
  ) +
  geom_linerange(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    linewidth = 0.8
  ) +
  geom_point(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    size = 2
  ) +
  geom_boxplot(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "1st-3rd Quartile"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      lower = UNCERTAINTY_LOWER_STANDARD,
      middle = MEASURED_OR_IMPUTED_VALUE_STANDARD,
      upper = UNCERTAINTY_UPPER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    stat = "identity"
  ) +
  geom_point(
    alpha = 0.5,
    data = \(x) {
      filter(
        x,
        is_not_reported(UNCERTAINTY_TYPE) | is_not_relevant(UNCERTAINTY_TYPE)
      )
    }
  ) +
  geom_text(
    data = sample_sizes_water,
    aes(x = -Inf, y = REFERENCE_ID, label = paste0("n=", total_n)),
    hjust = -1, vjust = -1,
    inherit.aes = FALSE,
    size = 3
  ) +
  geom_vline(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(ENVIRON_COMPARTMENT_SUB %in% sub_comps, THRESHOLD_TYPE == "Classification boundary")
    },
    aes(xintercept = THRESHOLD_VALUE, color = THRESHOLD_CLASS),
    linetype = "dashed",
    linewidth = 0.8,
    alpha = 0.6
  ) +
  # Threshold labels - only in first facet
  geom_text(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(ENVIRON_COMPARTMENT_SUB %in% sub_comps) |>
        cross_join(first_facet)
    },
    aes(x = THRESHOLD_VALUE, y = Inf, label = THRESHOLD_CLASS, color = THRESHOLD_CLASS),
    vjust = 1.5,
    angle = 90,
    hjust = 1,
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_color_manual(
    values = c(threshold_colours, compartment_colours),
    breaks = names(compartment_colours)
  ) +
  labs(
    x = "Concentration (mg/L)",
    y = "Reference",
    colour = "Sub-compartment"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.box = "vertical"
  ) +
  facet_nested(
    rows = vars(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE),
    scales = "free_y",
    space = "free_y",
    shrink = TRUE,
    labeller = label_wrap_gen(width = 15),
    nest_line = element_line()
  ) +
      coord_cartesian(
    xlim = range(
      literature_merged_data$MEASURED_OR_IMPUTED_VALUE_STANDARD[
        literature_merged_data$ENVIRON_COMPARTMENT_SUB %in% chosen_compartments
      ],
      na.rm = TRUE
    ) |> round(),
    clip = "off"  # allows threshold labels to extend beyond plot area if needed
  ) 
```

### Terrestrial Compartment

Currently, data on copper in the terrestrial compartment is limited to one study (@fig-boxplot-terrestrial). 


```{r}
#| label: fig-boxplot-terrestrial
#| fig-cap: "Copper concentrations in terrestrial samples by reference"
#| fig-width: 10
# Calculate sample sizes
chosen_compartments <- c("Soil O Horizon (Organic)")

sample_sizes_water <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  calculate_sample_sizes()

# Get first facet for labels
first_facet <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  distinct(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE) |>
  slice(1)

# Create plot
literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  ggplot(aes(
    x = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    y = REFERENCE_ID,
    colour = ENVIRON_COMPARTMENT_SUB
  )) +
  # ... (all your other geoms remain the same) ...
  geom_crossbar(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "95% Confidence Interval"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    )
  ) +
  geom_linerange(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    linewidth = 0.8
  ) +
  geom_point(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    size = 2
  ) +
  geom_boxplot(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "1st-3rd Quartile"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      lower = UNCERTAINTY_LOWER_STANDARD,
      middle = MEASURED_OR_IMPUTED_VALUE_STANDARD,
      upper = UNCERTAINTY_UPPER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    stat = "identity"
  ) +
  geom_point(
    alpha = 0.5,
    data = \(x) {
      filter(
        x,
        is_not_reported(UNCERTAINTY_TYPE) | is_not_relevant(UNCERTAINTY_TYPE)
      )
    }
  ) +
  geom_text(
    data = sample_sizes_water,
    aes(x = -Inf, y = REFERENCE_ID, label = paste0("n=", total_n)),
    hjust = -1, vjust = -1,
    inherit.aes = FALSE,
    size = 3
  ) +
  geom_vline(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(ENVIRON_COMPARTMENT_SUB %in% sub_comps, THRESHOLD_TYPE == "Classification boundary")
    },
    aes(xintercept = THRESHOLD_VALUE, color = THRESHOLD_CLASS),
    linetype = "dashed",
    linewidth = 0.8,
    alpha = 0.6
  ) +
  # Threshold labels - only in first facet
  geom_text(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(ENVIRON_COMPARTMENT_SUB %in% sub_comps) |>
        cross_join(first_facet)
    },
    aes(x = THRESHOLD_VALUE, y = Inf, label = THRESHOLD_CLASS, color = THRESHOLD_CLASS),
    vjust = 1.5,
    angle = 90,
    hjust = 1,
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_color_manual(
    values = c(threshold_colours, compartment_colours),
    breaks = names(compartment_colours)
  ) +
  labs(
    x = "Concentration (mg/L)",
    y = "Reference",
    colour = "Sub-compartment"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.box = "vertical"
  ) +
  facet_nested(
    rows = vars(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE),
    scales = "free_y",
    space = "free_y",
    shrink = TRUE,
    labeller = label_wrap_gen(width = 15),
    nest_line = element_line()
  ) +
      coord_cartesian(
    xlim = range(
      literature_merged_data$MEASURED_OR_IMPUTED_VALUE_STANDARD[
        literature_merged_data$ENVIRON_COMPARTMENT_SUB %in% chosen_compartments
      ],
      na.rm = TRUE
    ) |> round(),
    clip = "off"  # allows threshold labels to extend beyond plot area if needed
  ) 
```

### Biota Compartment

#### Aquatic Biota

@fig-points-biota presents copper body burdens across different species groups. Note: There appear to be issues with the distribution calculations for biota that require further investigation.

```{r}
#| label: fig-points-biota
#| fig-cap: "Density distributions of copper concentrations in biota, separated by species group. Copper is an essential micronutrient, so baseline concentrations vary widely among taxa. High concentrations may indicate either regulatory adaptation or toxic exposure. Note: Distribution calculations require verification."
#| warning: false
#| fig-width: 10


literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT == "Biota"
  ) |>
  ggplot(aes(
    x = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    y = SAMPLE_SPECIES,
    colour = SPECIES_GROUP
  )) +
  geom_pointrange(aes(
    xmin = UNCERTAINTY_UPPER_STANDARD,
    xmax = UNCERTAINTY_UPPER_STANDARD
  )) +
  facet_wrap(
    facet = vars(OCEAN_COUNTRY),
    ncol = 1,
    scales = "free_y",
    space = "free_y"
  )

```

## Temporal Trends

### Abiotic Compartments

@fig-temporal-abiotic examines temporal trends in copper concentrations across abiotic environmental compartments. The current dataset has limited temporal coverage, but these trends may become more apparent as additional data from monitoring databases (e.g., Vannmiljø) are incorporated.

```{r}
#| label: fig-temporal-abiotic
#| fig-cap: "Temporal trends in copper concentrations across abiotic environmental sub-compartments. Points represent mean values with error bars showing standard deviations. Limited temporal coverage currently restricts trend interpretation, but patterns may emerge with additional historical monitoring data."
#| warning: false
#| fig-width: 10
#| fig-height: 12

p <- literature_merged_data |>
  filter(
    !is.na(MEASURED_UNIT_STANDARD),
    UNCERTAINTY_TYPE == "Standard Deviation",
    ENVIRON_COMPARTMENT != "Biota"
  ) |>
  group_by(ENVIRON_COMPARTMENT_SUB, SAMPLE_SPECIES) |>
  ggplot(aes(
    x = YEAR,
    y = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    color = ENVIRON_COMPARTMENT_SUB
  )) +
  geom_point() +
  geom_errorbar(
    aes(ymin = UNCERTAINTY_LOWER_STANDARD, ymax = UNCERTAINTY_UPPER_STANDARD),
    width = .2,
  ) +
  facet_wrap(
    facets = vars(ENVIRON_COMPARTMENT_SUB),
    ncol = 1,
    scales = "free_y"
  ) +
  labs(
    x = "Year",
    y = "Copper Concentration (standardised units)",
    color = "Sub-compartment"
  ) +
  theme_minimal()

plotly::ggplotly(p)
```

### Biota

@fig-temporal-biota shows temporal patterns in biotic copper concentrations across species groups. Interpretation should consider both temporal changes in exposure and potential differences in species composition across sampling years.

```{r}
#| label: fig-temporal-biota
#| fig-cap: "Temporal trends in copper body burdens across different species groups. Points are coloured by species to help distinguish measurement series. Error bars represent reported uncertainty ranges. Vertical facets separate species groups to account for baseline differences in copper accumulation patterns."
#| warning: false
#| fig-width: 10
#| fig-height: 12

p <- literature_merged_data |>
  filter(
    !is.na(MEASURED_UNIT_STANDARD),
    ENVIRON_COMPARTMENT == "Biota"
  ) |>
  ggplot(aes(
    x = YEAR,
    y = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    color = SAMPLE_SPECIES
  )) +
  geom_point() +
  geom_errorbar(
    aes(ymin = UNCERTAINTY_LOWER_STANDARD, ymax = UNCERTAINTY_UPPER_STANDARD),
    width = .2,
  ) +
  facet_wrap(
    facets = vars(SPECIES_GROUP),
    ncol = 1,
    scales = "free_y"
  ) +
  labs(
    x = "Year",
    y = "Copper Concentration (standardised units)",
    color = "Species"
  ) +
  theme_minimal()

plotly::ggplotly(p)
```

# Network Diagram

```{r}
#| echo: false
# Network Diagram ----
#| fig-width: 10
#| fig-height: 12

unique_compartments_and_sites <- literature_merged_data |>
  select(
    SITE_GEOGRAPHIC_FEATURE,
    SITE_GEOGRAPHIC_FEATURE_SUB,
    ENVIRON_COMPARTMENT,
    ENVIRON_COMPARTMENT_SUB,
    MEASURED_N
  ) |>
  distinct() |>
  filter(if_all(everything(), ~ !is.na(.))) |> 
  group_by(SITE_GEOGRAPHIC_FEATURE, SITE_GEOGRAPHIC_FEATURE_SUB, ENVIRON_COMPARTMENT, ENVIRON_COMPARTMENT_SUB) |> 
  reframe(sum_n = sum(MEASURED_N)) |> 
  mutate(compartment_name_n = glue::glue("{ENVIRON_COMPARTMENT_SUB}\n{sum_n}"))


# TODO: This is very speculative and just generated by AI. I haven't yet reviewed it to see what's missing.
# Copper flow network edges ----
copper_flow_edges <- tribble(
  ~from, ~to, ~flow_type, ~plausibility, ~magnitude, ~evidence_quality, ~description,
  
  # Atmospheric deposition to surface waters ----
  "Atmospheric", "Freshwater", "deposition", 0.95, "medium", "high", "Wet and dry deposition from atmospheric particulates",
  "Atmospheric", "Brackish/Transitional Water", "deposition", 0.95, "medium", "high", "Wet and dry deposition to coastal/estuarine waters",
  "Atmospheric", "Aquatic Sediment", "deposition", 0.90, "low", "medium", "Direct deposition to exposed sediments",
  "Atmospheric", "Soil O Horizon (Organic)", "deposition", 0.95, "medium", "high", "Atmospheric deposition to terrestrial surfaces",
  
  # Terrestrial to aquatic flows ----
  "Soil O Horizon (Organic)", "Freshwater", "runoff", 0.95, "high", "high", "Surface runoff and leaching from soils",
  "Soil O Horizon (Organic)", "Groundwater", "percolation", 0.85, "medium", "medium", "Downward migration through soil profile",
  "Biota, Terrestrial", "Soil O Horizon (Organic)", "excretion_decomposition", 0.90, "low", "medium", "Excretion and decomposition of terrestrial organisms",
  
  # Groundwater to surface water ----
  "Groundwater", "Freshwater", "baseflow", 0.90, "medium", "high", "Groundwater discharge to streams/rivers",
  "Groundwater", "Brackish/Transitional Water", "discharge", 0.75, "low", "medium", "Submarine groundwater discharge to coastal waters",
  
  # Freshwater flows ----
  "Freshwater", "Brackish/Transitional Water", "advection", 0.95, "high", "high", "River discharge to estuaries",
  "Freshwater", "Aquatic Sediment", "sedimentation", 0.95, "high", "high", "Particulate settling from water column",
  "Freshwater", "Biota, Aquatic", "uptake", 0.95, "medium", "high", "Bioaccumulation by freshwater organisms",
  
  # Brackish/transitional water flows ----
  "Brackish/Transitional Water", "Aquatic Sediment", "sedimentation", 0.95, "high", "high", "Particulate settling in estuarine environments",
  "Brackish/Transitional Water", "Biota, Aquatic", "uptake", 0.95, "medium", "high", "Bioaccumulation by estuarine/coastal organisms",
  
  # Sediment processes ----
  "Aquatic Sediment", "Freshwater", "resuspension", 0.85, "medium", "high", "Sediment resuspension and porewater release",
  "Aquatic Sediment", "Brackish/Transitional Water", "resuspension", 0.85, "medium", "high", "Sediment resuspension in coastal/estuarine areas",
  "Aquatic Sediment", "Biota, Aquatic", "uptake", 0.90, "low", "medium", "Benthic organism uptake from sediments",
  "Aquatic Sediment", "Groundwater", "diagenesis", 0.70, "low", "low", "Deep burial and diagenetic processes",
  
  # Biota cycling ----
  "Biota, Aquatic", "Aquatic Sediment", "excretion_decomposition", 0.95, "medium", "high", "Fecal pellets, mortality, decomposition",
  "Biota, Aquatic", "Freshwater", "excretion", 0.90, "low", "medium", "Dissolved excretion by aquatic organisms",
  "Biota, Aquatic", "Brackish/Transitional Water", "excretion", 0.90, "low", "medium", "Dissolved excretion in coastal waters",
  "Biota, Aquatic", "Biota, Terrestrial", "trophic_transfer", 0.80, "low", "medium", "Predation by terrestrial organisms (e.g., birds, mammals)",
  
  # Wastewater flows ----
  "Wastewater", "Freshwater", "discharge", 0.90, "high", "high", "Treated or untreated wastewater discharge",
  "Wastewater", "Brackish/Transitional Water", "discharge", 0.80, "medium", "medium", "Coastal wastewater discharge",
  "Wastewater", "Sludge", "treatment", 0.95, "high", "high", "Copper partitioning to sludge during treatment",
  
  # Sludge applications ----
  "Sludge", "Soil O Horizon (Organic)", "application", 0.85, "medium", "medium", "Agricultural application of biosolids",
  "Sludge", "Freshwater", "discharge", 0.60, "low", "low", "Illegal or emergency discharge (historical)",
  
  # Terrestrial-atmospheric ----
  "Soil O Horizon (Organic)", "Atmospheric", "resuspension", 0.70, "low", "low", "Wind erosion and dust generation",
  "Biota, Terrestrial", "Atmospheric", "volatilization", 0.40, "negligible", "low", "Unlikely for copper but included for completeness"
)

# Add reverse flow indicators for bidirectional processes if needed ----
# don't work yet
# copper_flow_edges |> left_join(unique_compartments_and_sites |> select(ENVIRON_COMPARTMENT_SUB, compartment_name_n), by = "ENVIRON_COMPARTMENT_SUB")

# Summary statistics ----
# copper_flow_edges |>
#   count(flow_type, sort = TRUE)

# copper_flow_edges |>
#   count(magnitude)

# Then plot
graph <- as_tbl_graph(copper_flow_edges)

```


```{r}
#| warning: false
#| width: 20
ggraph(graph, layout = "tree") +
    geom_edge_diagonal2(aes(label = flow_type) , colour = "darkgrey", label_colour = "black",
  strength = 0.8,
  check_overlap = TRUE,
  angle_calc = "along",
                  arrow = arrow(length = unit(4, 'mm')), 
                start_cap = circle(15, 'mm'),
                end_cap = circle(15, 'mm')) +
  geom_node_point(size = 40, aes(colour = name))+
  geom_node_text(aes(label = str_wrap(name, 10)), size = 5,) +
  scale_fill_manual(values = compartment_colours) +
  theme_void() +
  theme(legend.position = "none")
```

# Appendices

## Appendix 1: Copper Sources Data

### Submarine Tailing Disposal in/off Norway

Note: STD locations (named) extracted from Kvassnes & Iversen 2013. Coordinates estimated by LLM, confidence self-reported by LLM.

```{r}
#| label: tbl-std-data
mine_tailings_coords() |> DT::datatable()

```

### Apple Production Areas in Norway

According to https://static02.nmbu.no/mina/studier/moppgaver/2018-Paulsen.pdf, 7 municipalities represent the main fruit districts in Norway. We also have sales of copper oxide for 2020-24 from https://mattilsynet-xp7prod.enonic.cloud/_/attachment/inline/8311cd70-8877-4712-98cc-f7bc1dba46ac:0f491609e72f1d5f0dca5ad1030a4b979e8b4695/Omsetningsstatistikk%20for%20plantevernmidler%202020%20-%202024%20med%20uthevinger.pdf (table 2a, p4, thanks to Roger Holten). 

```{r}
#| label: tbl-apple-emissions
number_of_sites <- nrow(municipality_centroids())
sales_data_split <- copper_oxide_sales() |> mutate(amount_kg = amount_kg / number_of_sites,
average_kg = average_kg / number_of_sites)

cross_join(municipality_centroids(), sales_data_split) |> 
   DT::datatable()

```


Although not explicitly stated, we assume that:

- Copper products used as fungicides are exclusively applied in these areas and enter the environment via soil and freshwater
- Apple farmers use the arthimetic mean of sold copper oxide for 2020 - 2024 every year. all applied copper oxide enters the environment.
- The total used copper is split equally between the "main fruit districts" named in Paulsen (2018). 

### Manufacturing Emissions of Copper in Norway

Taken from https://www.norskeutslipp.no/no/Komponenter/Utslipp/Kobber/?ComponentType=utslipp&ComponentPageID=73&SectorID=600, which includes self-reported copper emissions by land-based Norwegian industries for 1985-2024. Date is *very* spotty and incomplete, but it allows us to highlight fylke with higher emissions. Map to follow.

We assume:

- Reported concentrations are fully correct
- Everything ends up in the environment.


```{r}
#| label: fig-industrial-emissions
#| wifth: 16

source("temp_R/scratchpad_norske_utslippe.R")

top_municipalities_plot
```