---
title: "Notepad 07 - Emissions Data"
author: "Sam Welch"
date: "2025.12.17"
lightbox: true
format:
  html:
    grid:
      body-width: 1500px
      sidebar-width: 0px
    page-layout: full
    code-fold: true
    max-width: 4500px
    fig-format: svg
    toc: true
---

*Currently a holding space for data on sources.*. 


# Appendices

## Appendix 1: Copper Sources Data

### Submarine Tailing Disposal in/off Norway

Note: STD locations (named) extracted from Kvassnes & Iversen 2013. Coordinates estimated by LLM, confidence self-reported by LLM.

```{r}
#| label: tbl-std-data

devtools::load_all()
library(tibble)
library(dplyr)
library(ggplot2)
library(here)
mine_tailings_coords() |> DT::datatable()

```

### Apple Production Areas in Norway

According to https://static02.nmbu.no/mina/studier/moppgaver/2018-Paulsen.pdf, 7 municipalities represent the main fruit districts in Norway. We also have sales of copper oxide for 2020-24 from https://mattilsynet-xp7prod.enonic.cloud/\_/attachment/inline/8311cd70-8877-4712-98cc-f7bc1dba46ac:0f491609e72f1d5f0dca5ad1030a4b979e8b4695/Omsetningsstatistikk%20for%20plantevernmidler%202020%20-%202024%20med%20uthevinger.pdf (table 2a, p4, thanks to Roger Holten).

N.B: Re Ronny Berdinsen, Mattilsynet (2025.12.15)

Hei

Tallene ble korrigerte pga en feilrapportering i 2022. Riktige tall er:

-	2022: 4332 liter (9482liter - 5150 liter)
-	2023: 2400 liter

Merk også at statistikken ikke viser plantevernmidler brukt pr år, men importert mengde. Det er ikke gitt at det som er importert i eksempelvis 2022 er brukt av bonden samme år.


```{r}
#| label: tbl-apple-emissions
number_of_sites <- nrow(municipality_centroids())
sales_data_split <- copper_oxide_sales() |>
  mutate(
    amount_kg = amount_kg / number_of_sites,
    average_kg = average_kg / number_of_sites
  )

cross_join(municipality_centroids(), sales_data_split) |>
  DT::datatable()

```

Although not explicitly stated, we assume that:

-   Copper products used as fungicides are exclusively applied in these areas and enter the environment via soil and freshwater
-   Apple farmers use the arthimetic mean of sold copper oxide for 2020 - 2024 every year. all applied copper oxide enters the environment.
-   The total used copper is split equally between the "main fruit districts" named in Paulsen (2018).

### Manufacturing Emissions of Copper in Norway

Taken from https://www.norskeutslipp.no/no/Komponenter/Utslipp/Kobber/?ComponentType=utslipp&ComponentPageID=73&SectorID=600, which includes self-reported copper emissions by land-based Norwegian industries for 1985-2024. Date is *very* spotty and incomplete, but it allows us to highlight fylke with higher emissions. Map to follow.

We assume:

-   Reported concentrations are fully correct
-   Everything ends up in the environment.

```{r}
#| label: fig-industrial-emissions
#| fig-width: 16

source(here("temp_R/scratchpad_norske_utslippe.R"))

top_municipalities_plot
```

