---
title: "Appendix 04 - Units, Unit Errors, and Corrections"
# Formatting lives in docs/_metadata.yml.
---

```{r}
#| label: setup
#| include: false

library(targets)
library(here)
library(tidyverse)
library(flextable)
pkgload::load_all(quiet = TRUE)

here::i_am("index.qmd")
```

Unit errors are a frequent issue in science. In this project, we run into two main types:
1. Human error (e.g. using the wrong conversion factor for a pair of units)
2. Encoding error (e.g. special characters such as µ (mu) being corrupted and thus not being properly read further down a pipeline)

While the former (1) typically only occur when units are explicitly converted, the latter (2) can occur anywhere where files are converted between formats. Encoding errors can be introduced without warning, and cause data groups to be mis-converted or dropped without warning. Because data are processed by multiple parties at different stages of this and others' pipelines, it becomes extremely important to identify, flag, and triage unit errors if data analyses are to be trusted. 

A thorough review of the AEP's upstream and local data determined that both types of errors had been introduced at multiple points, both in source data during processing by its owner, and in this project. Consequently, we elected to scrutinise the data both manually and automatically to correct unit errors from source data owners, and prevent the introduction of new errors. 

This document serves as a reference for the authors, for LLM-based coding, and for readers, and provides a detailed overview of the unit error detection and correction process.

How this project handles units, how it finds units that arrived wrong, and how
to correct one.

Everything bellow is LLM-written and needs to be re-written.

# Unit Correction Functions

```{r}

```

The **single source of truth is `data/clean/decisions/unit_corrections.csv`**. Every
correction shown on this page is read from it live, so nothing here can drift
away from what the pipeline actually did. This document explains the file; it
does not duplicate it.

# The one fact behind most of the errors

::: callout-important
## µg/g **is** mg/kg

Micro-over-gram cancels against milli-over-kilogram. They are the same number,
not a thousand apart.
:::

Three separate 1000x faults surfaced on 2026-08-05 from that single
misconception, in three different places:

1. A hand extraction recorded `µg/kg` where the paper said `µg/g`.
2. Our own `standardise_measured_units()` divided every genuine `µg/g` row by
   1000.
3. A Vannmiljø submitter multiplied their `µg/g` values by 1000 to "convert"
   them to `mg/kg`, and said so in the comment field.

The first two were ours and are fixed. The third is in data we do not control,
which is what the corrections layer exists for.

# Three files, three different jobs

| File | Job | Changes a number? |
|---|---|---|
| `R/fct_units.R` | Conversions **this project** performs | Yes, by rule |
| `R/fct_unit_anomalies.R` | **Detection** of errors that arrived already made | No, reports only |
| `R/fct_unit_corrections.R` | The **human decision** to override them | Yes, by hand |

They are deliberately separate. Conversion is a lookup and should be automatic.
Overriding a national monitoring database is a scientific judgement, and it has
to be reviewable in a diff and defensible in the methods.

# Two traps that cost a day each

## Never write a micro sign you do not have to

`µ` reaches this project as at least four different characters: `U+03BC` (Greek
mu), `U+00B5` (micro sign), a plain `u`, and `U+FFFD` (a micro sign already
destroyed by an encoding round trip). CSV plus Windows plus Excel will mangle it
eventually, and **the failure is silent**: an unmatched unit becomes `NA`, the
row is dropped, and a whole reference can vanish without a message. Eighteen
rows of `2000JulshamnTraceElementLevels` were lost this way for months.

Write `ug` in anything hand-entered. `normalise_unit_string()` maps every
variant onto `u` before matching, and that includes the `comment_match` column
of the corrections file, so a plain `ug/g` typed into the spreadsheet will match
a real micro sign in the data.

## A foothill three orders from the main mode is a unit error

Copper in a given matrix spans maybe two to three orders in reality. Anything
sitting a clean $10^3$ away from the mode is almost always mg/kg against ug/kg,
or ug/g against ug/kg. **Check the unit before reaching for a statistical
explanation.** Two separate 1000x faults were found exactly this way, one in the
extraction and one in the pipeline.

# Where a correction happens in the pipeline

```
literature_clean_standardised   <- our conversions
        |
load_literature_pqt             <- row_id assigned here
        |
literature_corrected            <- unit_corrections.csv applied here
        |
literature_analysis_ready       <- hygiene (drop NA / <= 0)
        |
everything: summary, triage, group_ids, node cards, AEP
```

Above the hygiene step, so a correction lands before anything is dropped or
summarised. Below `literature_clean_standardised`, so our conversions and their
errors stay separate concerns.

::: callout-note
## Why not correct inside a notebook, or per AEP node

The same bad rows feed the triage panels, the summary table and
`group_decisions.csv`. Correcting only where you noticed the problem leaves all
of those showing the old numbers while the AEP shows the new ones, which is
worse than not correcting at all.
:::

A correction scales `MEASURED_VALUE_STANDARD` **and** both LOD/LOQ standard and
imputed columns, because a submitter who multiplied their values multiplied the
limits reported alongside them. `MEASURED_VALUE` is never touched: it stays as
the audit trail, so a corrected row still shows what the source actually said.

# What is currently flagged and not yet corrected

`unit_anomaly_report` runs on **corrected** data, which makes it a shrinking
to-do list rather than a static record. A group still listed here after a
correction has been written means the correction was insufficient, not that it
is missing.

```{r}
#| label: tbl-anomaly-offsets
#| tbl-cap: "Campaigns sitting a whole number of decades from their peers within the same sampling group. `log10_ratio` is the offset; `near_decade` is how close that offset is to a clean power of ten, where closer to zero is more suspicious."

anomalies <- tar_read(unit_anomaly_report)

anomalies$offsets |>
  filter(check_units) |>
  select(
    any_of(c(
      "ENVIRON_COMPARTMENT_SUB",
      "SAMPLE_SPECIES",
      "SAMPLE_TISSUE",
      "CAMPAIGN_NAME_SHORT",
      "n_rows",
      "geo_mean",
      "peer_geo_mean",
      "log10_ratio",
      "near_decade"
    ))
  ) |>
  flextable() |>
  colformat_double(digits = 2) |>
  autofit()
```

```{r}
#| label: tbl-anomaly-comments
#| tbl-cap: "Rows whose own comment describes unit arithmetic. The cheapest detector, because it is not an inference: the row says what was done to it."

anomalies$comments |>
  flextable() |>
  colformat_double(digits = 3) |>
  autofit()
```

# Corrections currently applied

```{r}
#| label: tbl-corrections
#| tbl-cap: "Live contents of `data/clean/decisions/unit_corrections.csv`."

corrections <- tar_read(unit_corrections)

if (nrow(corrections) == 0) {
  flextable(tibble(` ` = "No corrections have been written yet."))
} else {
  corrections |>
    select(-row_ids) |>
    flextable() |>
    autofit()
}
```

```{r}
#| label: tbl-correction-effect
#| tbl-cap: "What those corrections touched, per sampling group. `n_corrected` counts measurements (`sum(MEASURED_N)`); `n_rows_corrected` counts rows."

effect <- tar_read(unit_correction_report)

if (nrow(effect) == 0) {
  flextable(tibble(` ` = "Nothing corrected."))
} else {
  effect |>
    select(
      any_of(c(
        "group_id",
        "SAMPLE_SPECIES",
        "SAMPLE_TISSUE",
        "MEASURED_UNIT_STANDARD",
        "unit_correction_id",
        "unit_correction_factor",
        "n_rows_corrected",
        "n_corrected"
      ))
    ) |>
    flextable() |>
    autofit()
}
```

# How to write a correction

Five steps. Steps 2 and 4 are the ones people skip, and both are checked.

## 1. Convince yourself it is a unit error

The bar is the same as anything else that reaches the methods section. Useful
evidence, roughly in descending order of strength:

- **The row says so.** A `MEASUREMENT_COMMENT` describing unit arithmetic is as
  close to proof as this gets. See @tbl-anomaly-comments.
- **A clean factor of exactly 1000** against the other campaigns measuring the
  same thing in the same matrix. `near_decade` near zero in
  @tbl-anomaly-offsets. There is no environmental process that produces a clean
  three orders.
- **The corrected values become ordinary.** If dividing by 1000 puts a group
  back on top of every other campaign measuring the same thing, that is
  corroboration.
- **The source paper disagrees with the extraction.** Then fix the extraction in
  `data/raw/eData/`, not here. This file is for data we do not control.

::: callout-warning
## Two orders is not automatically a unit error

`check_units` fires at two orders, not three, because a partly-affected campaign
lands between the good and bad clusters and never reaches a clean decade. That
makes it a prompt to look, not a verdict. A genuinely contaminated site is a
real explanation, and `10^2.1` is more likely to be one than `10^2.98`.
:::

## 2. Work out what the selector has to be

**This is where a correction goes wrong.** The obvious selector is usually too
coarse.

The Urban Fjord case is the worked example: 18 cod muscle rows in one campaign
spanning 0.14 to 7,724 mg/kg wet. Only the high ones are wrong. A blanket
factor on the whole campaign would divide the three correct rows by 1000 too,
and create a new error underneath the old one.

Selector columns, all optional, combining with **AND**:

| Column | Selects on | Use when |
|---|---|---|
| `comment_match` | Fixed substring of `MEASUREMENT_COMMENT` | The rows say what was done to them. **Strongest selector; prefer it.** |
| `campaign_name_short` | Exactly one campaign | The fault is one submitter's |
| `group_id` | A sampling group from `group_ids.csv` | The fault is confined to one species/tissue/unit |
| `value_min` / `value_max` | `MEASURED_VALUE_STANDARD` bounds | No comment exists and you must isolate the bad cluster by magnitude |

::: callout-tip
## Write `ug`, not `µg`

`comment_match` normalises the micro sign on both sides, so `ug/g (w.w.)`
matches a comment containing a real `µ`. Do not paste a micro sign into the
spreadsheet.
:::

Value bounds are always tested against the **original** values, never against
values another correction has already scaled, so the order of rows in the CSV
never changes the result.

## 3. Fill in the row

One row per correction. Leave `row_ids` **blank**; step 4 fills it.

| Column | What goes in it |
|---|---|
| `correction_id` | `C001`, `C002`, ... Unique, never reused |
| `group_id` | Selector, or blank |
| `campaign_name_short` | Selector, or blank |
| `comment_match` | Selector, or blank |
| `value_min`, `value_max` | Selector bounds, or blank |
| `factor` | What to **multiply** by. A 1000x overstatement is `0.001` |
| `row_ids` | **Leave blank.** Written by the scaffold script |
| `reason` | Why this is a unit error, in a sentence. Required |
| `evidence` | What convinced you. Required |
| `date_added` | `YYYY-MM-DD` |

`reason` and `evidence` are required by the validator, not by convention. A
correction that documents nothing cannot be defended to a reviewer, and this
file's whole purpose is to be defensible.

## 4. Record the rows

```bash
Rscript scripts/scaffold_unit_corrections.R
```

This resolves each selector against the current data and writes the matching
`row_id`s back into the file. It prints the row count, the measurement count,
and the value range it matched, so **read that output**: it is the last check
before the numbers change.

::: callout-note
## Why the file carries both a selector and a list of row ids

Because either alone fails silently, in opposite directions.

A **selector alone** silently widens. If a re-export adds two more rows matching
the same comment, they get corrected without anyone deciding they should be.

**Row ids alone** silently narrow, and record no reasoning. A row that should be
corrected but arrived later is simply missed, and a reader cannot tell a
deliberate omission from an oversight.

Holding both, and requiring them to agree, turns a silent change of extent into
a build failure that names exactly what drifted, in both directions. That is
what makes this an audit trail rather than a record, and it is why `row_id`
exists.

The scaffold is a hand-run script and never a pipeline target, for the same
reason. If the pipeline re-resolved the ids on every build they would track the
data automatically, and the check would detect nothing.
:::

## 5. Review the diff, then rebuild

`git diff data/clean/decisions/unit_corrections.csv` before `tar_make()`. A correction
invalidates everything downstream, including the triage panels, the node cards
and the AEP.

# When the build refuses

Every failure in this layer is an error rather than a warning, deliberately. A
file that overwrites measured values must not be allowed to half-work.

| Message | What happened | What to do |
|---|---|---|
| `matches no rows` | The selector is stale or wrong | Check for a typo, or a micro sign that should be `ug` |
| `no longer matches the rows it recorded` | The data changed under a correction | Read the named gained/lost ids. If the change is right, re-run the scaffold with `--refresh` |
| `overlaps` | Two corrections claim the same row | Narrow one. Compounding order must never decide a value |
| `reason` / `evidence` `are required` | Undocumented correction | Write the sentence |
| `no selector set` | Every selector column blank | Would have matched the entire dataset |
| `factor must be finite and positive` | Bad factor | A 1000x overstatement is `0.001`, not `-1000` |

::: callout-warning
## `--refresh` is not a way to make an error go away

It re-records what a selector currently matches, which means it accepts drift.
Use it only after the build has told you what changed and you have looked at it.
:::

# Row identifiers

`row_id` is what makes a per-row reference survive a rebuild. It is assigned in
`load_literature_pqt` and inherited by everything downstream.

It is **not** a sequential counter. A counter is positional: Vannmiljø is
re-exported periodically and the eData files are edited whenever an extraction
fault is found, so one inserted row shifts every id after it, and a correction
keyed on `R01234` would then silently overwrite a different measurement.

`row_id` is `SAMPLE_ID` instead, which is already built from site, parameter,
compartment, date and subsample. Insert a row and every other id is unchanged.
Where two rows genuinely share a `SAMPLE_ID`, `SUBSAMPLE` is appended; if that
still leaves a tie the build **stops**, because a counter fallback would
reintroduce positional ids by the back door.

```{r}
#| label: tbl-row-id-collisions
#| tbl-cap: "Rows whose `SAMPLE_ID` was shared and had to be disambiguated. A non-empty table is a data-entry defect in the source extraction, not a property of the data."

collisions <- tar_read(row_id_collisions)

if (nrow(collisions) == 0) {
  flextable(tibble(` ` = "No collisions. Every SAMPLE_ID is unique."))
} else {
  collisions |>
    select(any_of(c(
      "sample_id",
      "source_file",
      "n_sharing",
      "SUBSAMPLE",
      "SAMPLE_SPECIES",
      "SAMPLE_TISSUE"
    ))) |>
    flextable() |>
    autofit()
}
```

# Standing decisions

- **Molar units are excluded.** 42 rows carry `µM` and are dropped. They *are*
  convertible (1 µM Cu is 63.55 µg/L), so this is a choice: a molar measurement
  is not obviously comparable with the total/dissolved mix already in the data,
  and 42 rows do not justify the methods paragraph that defending the conversion
  would need. Revisit only if a chosen AEP node depends on them.
- **Mixed units are refused, never averaged**, when an AEP node resolves its
  members.
- **`mg/kg` with no dry/wet basis is not silently assigned to either.** They are
  different standard units and are not interconvertible without a moisture
  content this project does not hold.

# For the methods section

If any correction survives to submission it needs a sentence, because this is
the pipeline overriding a national monitoring database. The defensible form is
roughly: values from *n* campaigns were divided by 1000 where the submitter's
own comment recorded that µg/g values had been multiplied by 1000 prior to
submission, µg/g being equivalent to mg/kg; original values are retained
unmodified.

The exclusion of molar units belongs there too, since it is a deliberate
exclusion rather than an absence.
