---
title: "Dev - Node Card Styling"
# Formatting lives in docs/_metadata.yml.
#
# Code is SHOWN here, against the project default of `echo: false`. This is a
# notebook about how the card is built, so hiding the calls that build it
# would leave only the pictures, which is the half you can get from
# figures/node_cards/ already.
execute:
  echo: true
format:
  html:
    code-fold: false
---

```{r}
#| label: setup
#| include: false

library(ggplot2)
library(patchwork)
library(ggview)
pkgload::load_all(quiet = TRUE)

here::i_am("index.qmd")
source(here::here("tests/testthat/helper-fixtures.R"))
```

Every figure here is shown exactly once, and where it exists as a file on disk
it is shown *from* that file. A whole card is built silently, written at its
real size, and then included; only the single panels, which have no artefact of
their own, are drawn by knitr. Printing a card inline as well as including the
written PNG showed each one twice, at two different sizes, which is worse than
useless in a notebook about how the thing looks.

Sizing is a chunk option throughout, never a `ggview::canvas()`: see the note
under the geometry table below.

```{r}
#| label: fixtures

node <- node_fixture()
ids <- ids_fixture()
data <- data_fixture()
members <- members_fixture(c("G001", "G002"))
card <- aep_node_report_cards(node, members, data, ids)
```

```{r}
#| label: card

card_width <- 2.4
card_height <- 1.8
card_dpi <- 300
# Named rather than inline, because the unpacked rebuild below has to pass the
# identical limits for its comparison against this card to mean anything.
card_limits <- c(0.1, 100)

# Built, not printed. The written file below is what gets shown.
p <- node_card(
  node,
  card,
  members,
  data,
  ids,
  limits = card_limits,
  dpi = card_dpi
)
```

```{r}
#| label: write
#| include: false

card_path <- here::here("figures/dummy_node_card.png")
save_ggplot(
  p,
  card_path,
  width = card_width,
  height = card_height,
  dpi = card_dpi,
  bg = NULL
)
```

```{r}
#| label: fig-card
#| echo: false
#| fig-cap: !expr sprintf("The finished card at its real size, %s x %s in at %s dpi, as written to `figures/dummy_node_card.png`. Everything below takes it apart.", card_width, card_height, card_dpi)

knitr::include_graphics(card_path)
```

## Unpacking `node_card()`

`node_card()` is three plots and a stack. It builds each panel from its own
exported function, then makes one `patchwork::wrap_plots()` call and merges a
background theme into all three. Nothing else happens in it. This section calls
those three functions separately so each piece can be seen on its own, then
reassembles them by hand and checks the rebuild against the card above.

### The geometry

`node_card_heights()` owns the vertical split, and is exported rather than
inlined so that anything illustrating the layout reads the real ratio instead
of a second, driftable copy of the same three numbers. The units are relative;
patchwork normalises them, so only the share matters.

```{r}
#| label: unpack-geometry

heights <- node_card_heights()
frac <- heights / sum(heights)

tibble::tibble(
  panel = names(heights),
  height = unname(heights),
  share = sprintf("%.1f%%", 100 * unname(frac)),
  inches = round(unname(frac) * card_height, 3)
)
```

Those inches are the constraint the whole design is fighting. The badge band
is a fifth of an inch tall, and text sizes are absolute, so a panel previewed
at any other size lies about its own legibility. Every preview below is
therefore drawn at card width by the panel's own share of card height, not
blown up to fill the page.

**That takes a chunk option, not just a `canvas()`.** Under knitr the figure
device is sized by `fig-width` / `fig-height`, which default project-wide to
12x8, and a `ggview::canvas()` does not bring it down to the card size: the
figure the `card` chunk prints inline measures 4.5x3.2 inches, and only the
written `figures/dummy_node_card.png` is a true 2.4x1.8. So each preview below
sets its device explicitly with `!expr`, reading the same `frac` as the table
above.

### The background

The level colour comes from none of the three panel functions. `node_card()`
builds one theme and merges it into all three with patchwork's `&`. That is
what lets it survive on top of `theme_void()` (header, badges) and
`triage_theme()` (strips) alike, since `&` merges into each panel's theme
rather than replacing it. Both `plot.background` and `panel.background` are
set: a colour visible in one but not the other reads as a rendering bug rather
than a fill.

```{r}
#| label: unpack-background

bg <- node_card_bg_colour(node)  # keyed on `level`, here "medium"
bg_theme <- ggplot2::theme(
  plot.background = ggplot2::element_rect(fill = bg, colour = NA),
  panel.background = ggplot2::element_rect(fill = bg, colour = NA)
)

# What `node_card()` does to each panel with `&`, done to one panel at a time
# so the previews below show what actually lands on the card. Sizing is the
# chunk's job, not this function's.
with_bg <- function(p) p + bg_theme

bg
```

### 1. Header, `node_card_header()`

An empty `ggplot()` under `theme_void()`, used as a text canvas with hand-set
limits (`x` 0 to 1, `y` -0.75 to 3.1). Three `annotate("text")` calls share the
`x = 0.5` anchor: the wrapped node label, the headline geometric mean, and the
counts line. The headline turns red and gains a `(!)` where it diverges from
the median by more than half an order, via `headline_is_suspect()`.

The node id is the exception. It is an `annotation_custom()` textGrob pinned
18x18 **pixels** from the physical top-left corner in `npc` units, which is the
only reason `dpi` is an argument to this function at all: the offset is
recomputed as `18 / dpi` inches so it lands on the same pixel whatever
resolution the card is saved at.

```{r}
#| label: fig-unpack-header
#| fig-width: !expr card_width
#| fig-height: !expr frac[["header"]] * card_height
#| fig-cap: "Header panel alone, at the 46% of card height it occupies. Node id in the corner, wrapped label, geometric mean with unit, then measurements, references and year range."

header <- node_card_header(node, card, dpi = card_dpi)
with_bg(header)
```

### 2. Badges, `node_epeq_badges()`

Four `geom_tile()` squares, filled from `epeq_score_colours()` (three levels
low to high, plus grey for unscored, which must not read as a low score), each
carrying its abbreviation and score on one line. Two letters per criterion
because Essentiality and Evidence share an initial.

Note the `text_size`: `node_card()` calls this at 3.4, not the function's own
default of 2.4, which is sized for full-size figures rather than a 2.4 inch
card. The x limits are symmetric about the tiles' centre so the band lines up
with the centred text above it.

```{r}
#| label: fig-unpack-badges
#| fig-width: !expr card_width
#| fig-height: !expr frac[["badges"]] * card_height
#| fig-cap: "Badge band alone, the whole of its 0.22 in. Fixture scores are Es 3, Pl 3, Ev 2, Qn 2, so two green and two yellow tiles; grey would mean unscored rather than low."

badges <- node_epeq_badges(node, text_size = 3.4)
with_bg(badges)
```

### 3. Strips, `node_group_strips()`

One violin per constituent group rather than one for the pooled node, because
the pooled geometric mean cannot tell you whether the lumping was sound and two
strips an order apart can. Value on x on the shared per-unit log scale, groups
on y, capped at `max_groups` with any remainder named rather than dropped
silently.

This is the only one of the three that touches the data, and so the only one
that can degrade: `triage_empty_plot()` for an external node or one with no
rows left after its restrictions, points instead of violins below the
`triage_use_points()` threshold. Group ids are drawn inside the panel, since a
y axis would inset this panel relative to the two `theme_void()` panels above
and put the whole text block off centre.

```{r}
#| label: fig-unpack-strips
#| fig-width: !expr card_width
#| fig-height: !expr frac[["strips"]] * card_height
#| fig-cap: "Strips panel alone. Two fixture groups, G001 and G002, twenty rows between them, which is under the thirty-row `triage_use_points()` threshold, so points rather than violins. The log value axis is the only axis the card keeps."

strips <- node_group_strips(
  node,
  members,
  data,
  ids,
  limits = card_limits,
  max_groups = 3
)
with_bg(strips)
```

### 4. The stack

One `wrap_plots()` in a single column with the heights from above, then
`& bg_theme`. The `&` is load-bearing: `+` would add the theme to the last
plot only, leaving the header and badges on white.

```{r}
#| label: fig-unpack-compose
#| fig-width: !expr card_width
#| fig-height: !expr card_height
#| fig-cap: "The three panels above stacked by hand. Compare with @fig-card, which `node_card()` produced; the check below says they are the same file."

composed <- patchwork::wrap_plots(
  header,
  badges,
  strips,
  ncol = 1,
  heights = node_card_heights()
) & bg_theme

composed
```

Rebuilt by hand from the three panels above, this is the same card
`node_card()` returns. Checked by rendering both and comparing the files, not
by eye, since a difference of a few pixels in one panel's height is exactly
the sort of thing eyeballing a screenshot misses.

```{r}
#| label: unpack-verify

reference <- node_card(
  node, card, members, data, ids,
  limits = card_limits, dpi = card_dpi
)

png_of <- function(x) {
  f <- tempfile(fileext = ".png")
  save_ggplot(
    x, f,
    width = card_width, height = card_height, dpi = card_dpi, bg = NULL
  )
  unname(tools::md5sum(f))
}

c(
  heights_identical = identical(
    composed$patches$layout$heights,
    reference$patches$layout$heights
  ),
  png_identical = png_of(composed) == png_of(reference)
)
```

## Stress test

```{r}
#| label: stress-fixtures

stress_node <- node_fixture(
  node_id = "N014-arctic-coastal-benthic-copper-bioaccumulation-network",
  label = "Coastal Sediment-Associated Copper"
)

stress_group_ids <- c(
  "G010-Bm-Cfj-G-mor-Liv-Mw",
  "G011-Bm-Cfj-G-mor-Mus-Mw",
  "G012-Bm-Cfj-M-edu-St-Mw",
  "G013-Bm-Cfj-C-pag-Mus-Mw",
  "G014-Bm-Cfj-P-pla-Liv-Mw"
)

stress_ids <- tibble::tibble(
  ENVIRON_COMPARTMENT = "Biota",
  ENVIRON_COMPARTMENT_SUB = "Biota, Marine",
  SPECIES_GROUP = c("Fish", "Fish", "Mollusc", "Crustacean", "Fish"),
  SAMPLE_SPECIES = c(
    "Gadus morhua", "Gadus morhua", "Mytilus edulis",
    "Cancer pagurus", "Pleuronectes platessa"
  ),
  SAMPLE_TISSUE = c("Liver", "Muscle", "Soft tissue", "Muscle", "Liver"),
  SITE_GEOGRAPHIC_FEATURE = "Coastal, fjord",
  SITE_GEOGRAPHIC_FEATURE_SUB = "Not reported",
  MEASURED_UNIT_STANDARD = "mg/kg (wet)",
  group_id = stress_group_ids
)

stress_data <- purrr::list_rbind(purrr::imap(stress_group_ids, function(gid, i) {
  row <- stress_ids[stress_ids$group_id == gid, ]
  n_rows <- 30
  tibble::tibble(
    ENVIRON_COMPARTMENT = row$ENVIRON_COMPARTMENT,
    ENVIRON_COMPARTMENT_SUB = row$ENVIRON_COMPARTMENT_SUB,
    SPECIES_GROUP = row$SPECIES_GROUP,
    SAMPLE_SPECIES = row$SAMPLE_SPECIES,
    SAMPLE_TISSUE = row$SAMPLE_TISSUE,
    SITE_GEOGRAPHIC_FEATURE = row$SITE_GEOGRAPHIC_FEATURE,
    SITE_GEOGRAPHIC_FEATURE_SUB = row$SITE_GEOGRAPHIC_FEATURE_SUB,
    MEASURED_UNIT_STANDARD = row$MEASURED_UNIT_STANDARD,
    MEASURED_VALUE_STANDARD = 10^stats::rnorm(n_rows, mean = log10(i), sd = 0.4),
    MEASURED_N = sample(10:80, n_rows, replace = TRUE),
    LATITUDE = seq(60, 78, length.out = n_rows),
    LONGITUDE = 10,
    SAMPLING_DATE = seq(as.Date("2000-01-01"), by = "year", length.out = n_rows),
    REFERENCE_ID = sample(paste0("Ref", LETTERS[1:8]), n_rows, replace = TRUE),
    CAMPAIGN_NAME_SHORT = paste0("Stress Campaign ", i)
  )
}))

stress_members <- members_fixture(stress_group_ids, stress_node$node_id[1])
stress_card <- aep_node_report_cards(stress_node, stress_members, stress_data, stress_ids)
```

```{r}
#| label: stress-card

stress_path <- here::here("figures/dummy_node_card_stress.png")

p_stress <- node_card(
  stress_node,
  stress_card,
  stress_members,
  stress_data,
  stress_ids,
  limits = range(stress_data$MEASURED_VALUE_STANDARD),
  dpi = card_dpi
)

save_ggplot(
  p_stress,
  stress_path,
  width = card_width,
  height = card_height,
  dpi = card_dpi,
  bg = NULL
)
```

```{r}
#| label: fig-stress-card
#| echo: false
#| fig-cap: "The same card under a long composite id, a long label and five constituent groups, of which `max_groups` keeps the three largest. Written to `figures/dummy_node_card_stress.png`."

knitr::include_graphics(stress_path)
```