---
title: "Notebook 02 - Vannmiljo Conversion"
author: "Sam Welch"
date: "2025.12.16"
html:
  toc-depth: 3
  toc-expand: 3
---

This notebook summarises and reports on the process of importing, cleaning, and reformatting data from the Norwegian Environmental Ministry's Vannmiljø database.

```{mermaid}
%%| label: fig-mermaid-process
%%| fig-width: 12
%%| fig-cap: |
%%|   A placeholder for a basic flow diagram of what we do here.
flowchart TB
    subgraph Input["Data Loading"]
        A1[Raw Measurements<br/>Vm_Copper_2025.12.05.xlsx<br/>138,615 hits]
        A2[Raw Sites<br/>3 Excel files<br/>split due to export limit]
        A3[Lookup Tables<br/>medium, vannkategori,<br/>methods analysis/sampling,<br/>campaigns, units, species]
    end
    
    subgraph Merge["Data Merging"]
        B1[Join Sites + Measurements<br/>on Vannlok_kode]
        B2[vm_merged<br/>Add all lookup tables]
    end
    
    subgraph Explore["Initial Exploration on vm_merged"]
        C1[Heatmaps:<br/>Campaign × Vannkategori<br/>Campaign × Medium<br/>Vannkategori × Medium<br/>Campaign × Species<br/>Units × Medium]
        C2[Sample Distribution<br/>cumulative % table<br/>90% coverage goal]
    end
    
    subgraph Filter["Apply Filters"]
        D1[vm_filtered<br/>Compartment filters<br/>Point sites only<br/>Date: 2010-01-01 to 2025-12-05<br/>Exclude Svalbard]
        D2[138,615 → 129,917 rows<br/>93.7% retained]
    end
    
    subgraph QualityCheck["Conflict Detection & Resolution"]
        E1[Detect Conflicts<br/>vannkategori vs medium<br/>compartment/subcompartment]
        E2[resolve_compartment_conflicts<br/>Biota trumps all<br/>Sediment overrides water<br/>Wildcard handling]
        E3[vm_compartment_conflicts_resolved]
        E4[Filter out unresolved<br/>remove FLAG rows]
        E5[vm_compartment_conflicts_resolved_filter]
    end
    
    subgraph Tables["Create eData Tables"]
        F1[edata_campaign]
        F2[edata_reference]
        F3[edata_parameters<br/>Copper only]
        F4[edata_sites<br/>unique sites<br/>UTM33→WGS84 transform]
        F5[edata_methods<br/>from vm_lookup_methods]
    end
    
    subgraph Samples["Sample Processing"]
        G1[edata_samples_wide<br/>from resolved+filtered<br/>generate_sample_id]
        G2[vm_samples_biota_only<br/>filter to biota<br/>add tissue, species, lifestage]
        G3[biota_samples_merged<br/>bind_rows of both]
    end
    
    subgraph BiotaSpecial["Biota Table"]
        H1[edata_biota<br/>species, tissue, gender<br/>from resolved data]
    end
    
    subgraph Measurements["Measurements Table"]
        I1[edata_measurements<br/>from biota_samples_merged<br/>values, flags, LOD/LOQ<br/>convert operators & units]
    end
    
    subgraph Export["Export & Validation"]
        J1[Assemble 7 tables<br/>into list]
        J2[export_campaign_zip<br/>Vm_2010_2025_timestamp.zip<br/>to data/raw/eData/]
        J3[Validate<br/>compartment counts<br/>outlier detection z≥3]
    end
    
    A1 --> B1
    A2 --> B1
    B1 --> B2
    A3 --> B2
    
    B2 --> C1
    B2 --> C2
    
    C1 --> D1
    C2 --> D1
    B2 --> D1
    
    D1 --> D2
    D2 --> E1
    E1 --> E2
    E2 --> E3
    E3 --> E4
    E4 --> E5
    
    E5 --> F1
    E5 --> F2
    E5 --> F3
    E5 --> F4
    A3 --> F5
    
    E5 --> G1
    E5 --> G2
    G1 --> G3
    G2 --> G3
    
    E5 --> H1
    
    G3 --> I1
    
    F1 --> J1
    F2 --> J1
    F3 --> J1
    F4 --> J1
    F5 --> J1
    G1 --> J1
    H1 --> J1
    I1 --> J1
    
    J1 --> J2
    J2 --> J3
    
    style Input fill:#e1f5ff
    style Merge fill:#fff4e1
    style Explore fill:#f0e1ff
    style Filter fill:#ffe1e1
    style QualityCheck fill:#e1ffe1
    style Tables fill:#fff0f0
    style Samples fill:#ffe1f0
    style BiotaSpecial fill:#f0ffe1
    style Measurements fill:#ffffe1
    style Export fill:#f5f5f5
```

# To Do:
-   [ ] Investigate some very strange detection and quantification limits
-   [ ] Add probable contamination source
-   [ ] More sanity checking of extracted values
-   [ ] Move boring code to targets
-   [ ] Work out how to do CREED for this data, then implement.
## Optional Improvements
-   [ ] Improve heatmaps using https://github.com/raivokolde/pheatmap
-   [ ] Make real Sankey diagrams

# Libraries & Setup

Import required packages and set working directory to project root.

```{r}
#| label: setup-libraries
#| warning: false
#| message: false

knitr::opts_chunk$set(dev = "svglite") # otherwise some SVGs cause problems in firefox?

library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(readr)
library(readxl)
library(glue)
library(ggplot2)
library(sf)
library(DT)
library(summarytools)
library(STOPeData)
library(eDataDRF)
library(here)
library(knitr)
library(devtools)
library(tibble)
library(pointblank)

validation_config <- list(
  # For lookups: warn if ANY fail (we want to know), but don't stop (we'll investigate)
  lookup = action_levels(warn_at = 0.0001, stop_at = NULL), # Never stops, always warns

  # For critical stuff later: stop at 5% failure
  standard = action_levels(warn_at = 0.01, stop_at = 0.05)
)

load_all()

i_am("README.md") |> suppressMessages()
```

## Configuration

See @tbl-config-filters and @tbl-config-metadata for a summary of the parameters of this extraction from Vannmiljø.

```{r}
#| label: setup-configuration
#| warning: false

config <- list(
  # File paths
  files = list(
    # TODO: Make imports more consistent and timestamp everything
    vm_copper = here("data/raw/vannmiljo/Vm_Copper_2025.12.05.xlsx"),
    vm_sites_1 = here("data/raw/vannmiljo/Vm_Copper_Sites_2025.12.05-1.xlsx"),
    vm_sites_2 = here("data/raw/vannmiljo/Vm_Copper_Sites_2025.12.05-2.xlsx"),
    vm_sites_3 = here("data/raw/vannmiljo/Vm_Copper_Sites_2025.12.05-3.xlsx"),
    lookup_medium = here("data/clean/Vm_medium_lookup_matrix_filled.csv"),
    lookup_vannkategori = here("data/clean/vm_sites_codes_lookup.csv"),
    lookup_methods = here("data/clean/vm_methods_lookup_filled.csv"),
    methods_analysis = here(
      "data/raw/vannmiljo/Vannmiljø_Analysemetode_2025-12-15.xlsx"
    ),
    methods_sampling = here(
      "data/raw/vannmiljo/Vannmiljø_Prøvetakingsmetode_2025-12-15.xlsx"
    ),
    lookup_campaigns = here("data/clean/Vm_lookup_campaigns.csv"),
    lookup_units = here(
      "data/raw/vannmiljo/Vannmiljø_Enhet_2025-12-30.xlsx"
    ),
    lookup_species = here("data/clean/Vm_species_lookup.csv") # not actually used rn
  ),

  # Filter criteria
  # * is used as a wildcard in compartment lookups
  filters = list(
    date_start = as.Date("2010-01-01"),
    date_end = as.Date("2025-12-05"),
    exclude_sites = "Svalbard, ENSB-Kilde 2",
    compartment = c("Aquatic", "Biota", "*"),
    compartment_sub = c(
      "Freshwater",
      "Aquatic Sediment",
      "Marine/Salt Water",
      "Brackish/Transitional Water",
      "Biota, Aquatic",
      "*"
    )
  ),

  # Campaign metadata
  metadata = list(
    campaign_name = "Vannmiljø Copper Monitoring 2010-2025",
    campaign_short = "Vm_2010_2025",
    reference_id = "VannmiljøCopper2010-2025",
    entered_by = "Sam Welch",
    organisation = "Miljødirektoratet"
  )
)

```

```{r}
#| label: tbl-config-filters

config$filters |>
  enframe(name = "Parameter", value = "Value") |>
  kable()
```

```{r}
#| label: tbl-config-metadata

config$metadata |>
  enframe(name = "Parameter", value = "Value") |>
  kable()
```

## Helper Functions

```{r}
#| label: setup-helper-functions
# Convert Vannmiljø operator codes to eData flags
vm_convert_operator <- function(col) {
  case_match(
    col,
    "=" ~ "",
    "<" ~ "< LOQ",
    ">" ~ "What is going on here. Why higher than?",
    "ND" ~ "< LOD"
  )
}

vm_convert_unit <- function(col) {
  case_match(
    col,
    "µg/l" ~ "µg/L",
    "mg/kg t.v." ~ "mg/kg (dry)",
    "mg/kg v.v." ~ "mg/kg (wet)",
    .default = "UNIT CONVERSION ERROR"
  )
}
```

# Load Raw Data

## Measurements Data

Downloaded 2025.12.05, data for 2025.01.01 - 2025.12.05, all kommune, all media, all campaigns

Search for Kobber & Kobberpyrition: 138615 hits

NOTE: Data may extend beyond 2025 date range despite request parameters

```{r}
#| label: load-raw-measurements

# Downloaded 2025.12.05, data for 2025.01.01 - 2025.12.05, all kommune, all media, all campaigns
# Search for Kobber & Kobberpyrition: 138615 hits
# NOTE: Data may extend beyond 2025 date range despite request parameters

if (!exists("vm_raw_copper")) {
  vm_raw_copper <- read_excel(
    path = config$files$vm_copper,
    sheet = 1,
    guess_max = 138615 # Avoid automatic type conversion warnings
  )
}
```

`summarytools::dfSummary()` generates an initial summary of the dataset here. It's not perfect but it's one line of code, and generally very helpful.

::: {.callout-note collapse="true" title="Initial Dataset Summary"}

```{r}
#| label: Initial data inspection
print(
  dfSummary(
    vm_raw_copper,
    varnumbers = FALSE,
    valid.col = FALSE,
    graph.magnif = 0.82
  ),
  method = 'render'
)
```

:::

## Sites Data

Vannmiljø export limit requires splitting site codes into multiple files

Procedure:

1.  Extract unique site codes: `vm_raw_copper\$Vannlok_kode |> unique() |> writeClipboard()`
2.  Replace newlines with commas: Ctrl+H find "\n" replace with ", "
3.  Split into chunks (Vannmiljø fails silently above unknown threshold)
4.  Download as separate Excel files

```{r}
#| label: load-raw-sites
#| echo: false

vm_raw_sites <- read_excel(
  config$files$vm_sites_1,
  guess_max = 10000
) |>
  add_row(read_excel(
    config$files$vm_sites_2,
    guess_max = 10000
  )) |>
  add_row(read_excel(
    config$files$vm_sites_3,
    guess_max = 10000
  ))

message(glue("Loaded {nrow(vm_raw_sites)} site records"))
```

## Lookup Tables

This part is very messy - we generated a bunch of lookup tables (semi-reproducibly) and filled them by hand/with Claude. I don't have a lot of confidence in any of them.

Some of the lookups are actually direct from Vannmiljø though - they're the xlsx ones. They're also bad, but for different reasons!

```{r}
#| label: load-lookups
#| echo: false

# Medium ID lookup
# Maps Vannmiljø Medium_id to eData compartment/species/tissue fields
vm_lookup_medium_raw <- read_csv(
  config$files$lookup_medium,
  guess_max = 100, # this may not be the best idea in the world
  show_col_types = FALSE # TODO: Come back to this
)

vm_lookup_medium <- vm_lookup_medium_raw |>
  rename_with(
    ~ paste0(., "_medium"),
    c(
      ENVIRON_COMPARTMENT,
      ENVIRON_COMPARTMENT_SUB,
      SPECIES_GROUP,
      SAMPLE_SPECIES,
      SPECIES_GENDER,
      SAMPLE_TISSUE,
      SITE_GEOGRAPHIC_FEATURE,
      SITE_GEOGRAPHIC_FEATURE_SUB
    )
  )

# Vannkategori lookup
# Maps Vannmiljø Vannkategori codes to eData site/compartment fields
vm_lookup_vannkategori_raw <- read_csv(
  config$files$lookup_vannkategori,
  show_col_types = FALSE
)

vm_lookup_vannkategori <- vm_lookup_vannkategori_raw |>
  rename_with(
    ~ paste0(., "_vkat"),
    c(
      ENVIRON_COMPARTMENT,
      ENVIRON_COMPARTMENT_SUB,
      SITE_GEOGRAPHIC_FEATURE,
      SITE_GEOGRAPHIC_FEATURE_SUB,
      n
    )
  )

# Methods lookup
vm_lookup_analysis <- read_excel(config$files$methods_analysis)
vm_lookup_sampling <- read_excel(config$files$methods_sampling)
vm_lookup_methods <- read_csv(
  config$files$lookup_methods,
  show_col_types = FALSE
) |>
  mutate(
    CAMPAIGN_NAME = config$metadata$campaign_short,
    PROTOCOL_ID = generate_protocol_id(
      PROTOCOL_CATEGORY,
      PROTOCOL_NAME,
      1, # FIXME: We don't have unique numbers for this rn.
      config$metadata$campaign_short
    )
  )
vm_lookup_campaigns <- read_csv(
  config$files$lookup_campaigns,
  show_col_types = FALSE
)

vm_lookup_units <- read_excel(config$files$lookup_units)

vm_lookup_species <- read_csv(
  config$files$lookup_species,
  show_col_types = FALSE
)

lookups <- grepv("vm_lookup", ls())

message(glue(
  "Loaded {length(lookups)} lookup tables: {paste(lookups, collapse = ', ')}"
))
```

# Data Processing

## Merge Datasets

We join sites and measurements by `"Vannlok_kode" = "Vannlokalitetskode"`, then our compartment/etc. lookups. We check for and remove:
- Non-null values: Vannlokalitetsnavn, Medium_id, Registrerings_id,`UTM33 Ost (X)`, `UTM33 Nord (Y)`
- Medium_id, Enhet_id, Aktivitet_id, Vannkategori, VitenskapligNavn not in our lookups

Note that we only use four lookups at this stage: `vm_lookup_medium`, `vm_lookup_vannkategori`, `vm_lookup_campaigns`, and `vm_lookup_units`.

```{r}
#| label: processing-merge-datasets
#| echo: false

# Join sites to measurements
vm_merged_sites <- vm_raw_copper |>
  left_join(
    vm_raw_sites,
    by = c("Vannlok_kode" = "Vannlokalitetskode")
  )

message(glue(
  "Joined vm_raw_copper and vm_raw_sites by Vannlokalitetskode. Output: vm_merged_sites ({nrow(vm_merged_sites)} x {ncol(vm_merged_sites)})"
))

# Add lookup tables
vm_merged_lookups <- vm_merged_sites |>
  left_join(
    vm_lookup_medium |>
      rename(MediumID_Name = Name, MediumID_Description = Description),
    by = c(Medium_id = "MediumID")
  ) |>
  left_join(
    vm_lookup_vannkategori |>
      rename(
        Vannkategori_Name = Name,
        Vannkategori_Description = Description
      ),
    by = c(Vannkategori = "VannkategoriID")
  ) |>
  left_join(
    vm_lookup_campaigns,
    by = c(Aktivitet_id = "ActivityID")
  ) |>
  left_join(
    vm_lookup_units |> rename(Unit_Name = Name, Unit_Description = Description),
    by = c(Enhet_id = "UnitID")
  ) |>
  left_join(
    vm_lookup_species,
    by = "VitenskapligNavn"
  ) |>
  # pointblank-based VALID II validation
  col_vals_not_null(
    columns = vars(
      Vannlokalitetsnavn,
      Medium_id,
      Registrerings_id,
      `UTM33 Ost (X)`,
      `UTM33 Nord (Y)`
    )
  ) |>
  col_vals_in_set(
    columns = vars(Medium_id),
    set = vm_lookup_medium$MediumID,
    label = "All Medium_ids have lookup matches",
    actions = validation_config$lookup
  ) |>
  col_vals_in_set(
    columns = vars(Enhet_id),
    set = vm_lookup_units$UnitID,
    label = "All Enhet_ids have lookup matches",
    actions = validation_config$lookup
  ) |>
  col_vals_in_set(
    columns = vars(Aktivitet_id),
    set = vm_lookup_campaigns$ActivityID,
    label = "All Aktivitet_ids have lookup matches",
    actions = validation_config$lookup
  ) |>
  col_vals_in_set(
    columns = vars(Vannkategori),
    set = vm_lookup_vannkategori$VannkategoriID,
    label = "All Vannkategori have lookup matches",
    actions = validation_config$lookup
  )


message(
  glue(
    "Left joined vm_merged_sites to lookup tables: vm_lookup_medium, vm_lookup_vannkategori, vm_lookup_campaigns, vm_lookup_units.
    vm_merged_sites ({nrow(vm_merged_sites)} x {ncol(vm_merged_sites)}) 
    + vm_lookup_medium 
    + vm_lookup_vannkategori 
    + vm_lookup_campaigns 
    + vm_lookup_units 
    + vm_lookup_species
    ---> vm_merged_lookups ({nrow(vm_merged_lookups)} x {ncol(vm_merged_lookups)})"
  )
)
```

## Pre-Transformation Exploration

We don't yet have an oversight over the distribution of site types, biota, tissue, and media in the dataset. This will inform what and where we filter. Unfortunately not all Vannkategori codes have a definition assigned to them, so I've attempted to guess.

```{r}
#| label: fig-heatmap-campaign-vannkategory
#| fig-cap: "Distribution of Vannmiljø copper samples by monitoring campaign and reported water category, all years"
#| fig-width: 12
#| fig-height: 10
#| warning: false

# Create a summary of counts by the two variables
vm_exploration_campaign_vannkategory <- vm_merged_lookups |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN = case_when(
      is.na(Name_EN) ~ "Other",
      TRUE ~ Name_EN
    )
  ) |>
  group_by(CAMPAIGN_DESCRIPTION_EN, Vannkategori_Description) |>
  summarise(n_samples = n(), .groups = "drop")

# Calculate text color threshold
threshold_value <- mean(
  log10(vm_exploration_campaign_vannkategory$n_samples),
  na.rm = TRUE
)

# Calculate total samples per campaign for ordering (y-axis)
campaign_order <- vm_exploration_campaign_vannkategory |>
  group_by(CAMPAIGN_DESCRIPTION_EN) |>
  summarise(total = sum(n_samples)) |>
  arrange(total) |> # puts highest values at top
  pull(CAMPAIGN_DESCRIPTION_EN)

# Calculate total samples per water category for ordering (x-axis)
vannkategori_order <- vm_exploration_campaign_vannkategory |>
  group_by(Vannkategori_Description) |>
  summarise(total = sum(n_samples)) |>
  arrange(desc(total)) |> # puts highest values on left
  pull(Vannkategori_Description)

# Wrap text, then create complete grid with all combinations
vm_exploration_complete <- vm_exploration_campaign_vannkategory |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN_wrapped = str_wrap(CAMPAIGN_DESCRIPTION_EN, 30),
    Vannkategori_Description_wrapped = str_wrap(Vannkategori_Description, 20)
  ) |>
  complete(
    CAMPAIGN_DESCRIPTION_EN_wrapped = str_wrap(campaign_order, 30),
    Vannkategori_Description_wrapped = str_wrap(vannkategori_order, 20),
    fill = list(n_samples = 0)
  ) |>
  mutate(
    # Factor with wrapped levels in the correct order
    CAMPAIGN_DESCRIPTION_EN_wrapped = factor(
      CAMPAIGN_DESCRIPTION_EN_wrapped,
      levels = str_wrap(campaign_order, 30)
    ),
    Vannkategori_Description_wrapped = factor(
      Vannkategori_Description_wrapped,
      levels = str_wrap(vannkategori_order, 20)
    )
  )

vm_exploration_complete |>
  ggplot(aes(
    x = Vannkategori_Description_wrapped,
    y = CAMPAIGN_DESCRIPTION_EN_wrapped,
    fill = n_samples
  )) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_text(
    aes(
      label = if_else(n_samples > 0, scales::comma(n_samples), ""),
      color = if_else(
        n_samples > 0 & log10(n_samples) > threshold_value,
        "black",
        "white"
      )
    ),
    size = 3
  ) +
  scale_fill_viridis_c(trans = "log10", na.value = "grey90") +
  scale_color_identity() +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(
    x = "",
    y = ""
  ) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 7),
    axis.text.y = element_text(size = 7),
    plot.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "none"
  )
```

@fig-heatmap-campaign-vannkategory clearly shows the the majority of copper contamination river and lake data from Vannmiljø is monitoring of mine-affected river systems. In general, sites expected to be contaminated (from mining, aquaculture, or other industries) are very well represented in the dataset. This is not unexpected, but needs to be kept in mind when interpreting the results. 

Rivers, lakes, and coastal waters are the majority of the dataset, which is appropriate for a primarily water-focused database.

```{r}
#| label: fig-heatmap-medium-campaign
#| fig-cap: "Distribution of Vannmiljø copper samples by monitoring campaign and reported sampling medium, all years"
#| fig-width: 12
#| fig-height: 10
#| warning: false

# Create a summary of counts by the two variables
vm_exploration_campaign_x_medium <- vm_merged_lookups |>
  mutate(
    Name_EN = case_when(
      is.na(Name_EN) ~ "Other",
      TRUE ~ Name_EN
    )
  ) |>
  group_by(Name_EN, MediumID_Name) |>
  summarise(n_samples = n(), .groups = "drop")

threshold_value_medium <- mean(
  log10(vm_exploration_campaign_x_medium$n_samples),
  na.rm = TRUE
)

campaign_order_medium <- vm_exploration_campaign_x_medium |>
  group_by(Name_EN) |>
  summarise(total = sum(n_samples)) |>
  arrange(total) |>
  pull(Name_EN)

medium_order <- vm_exploration_campaign_x_medium |>
  group_by(MediumID_Name) |>
  summarise(total = sum(n_samples)) |>
  arrange(desc(total)) |>
  pull(MediumID_Name)

vm_exploration_campaign_x_medium_complete <- vm_exploration_campaign_x_medium |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN_wrapped = str_wrap(Name_EN, 30),
    MediumID_Name_wrapped = str_wrap(MediumID_Name, 20)
  ) |>
  complete(
    CAMPAIGN_DESCRIPTION_EN_wrapped = str_wrap(campaign_order_medium, 30),
    MediumID_Name_wrapped = str_wrap(medium_order, 20),
    fill = list(n_samples = 0)
  ) |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN_wrapped = factor(
      CAMPAIGN_DESCRIPTION_EN_wrapped,
      levels = str_wrap(campaign_order_medium, 30)
    ),
    MediumID_Name_wrapped = factor(
      MediumID_Name_wrapped,
      levels = str_wrap(medium_order, 20)
    )
  )

vm_exploration_campaign_x_medium_complete |>
  ggplot(aes(
    x = MediumID_Name_wrapped,
    y = CAMPAIGN_DESCRIPTION_EN_wrapped,
    fill = n_samples
  )) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_text(
    aes(
      label = if_else(n_samples > 0, scales::comma(n_samples), ""),
      color = if_else(
        n_samples > 0 & log10(n_samples) > threshold_value_medium,
        "black",
        "white"
      )
    ),
    size = 3
  ) +
  scale_fill_viridis_c(trans = "log10", na.value = "grey90") +
  scale_color_identity() +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(x = "", y = "") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 7),
    axis.text.y = element_text(size = 7),
    plot.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "none"
  )
```

We also check reported water category by sampling medium, for the sake of completeness (@fig-heatmap-vannkategori-medium). From my interpretation, there are some combinations that make little sense (e.g. water category = air, sampling medium = soil), However, these are a minority. 

```{r}
#| label: fig-heatmap-vannkategori-medium
#| fig-cap: "Distribution of Vannmiljø copper samples by water category and reported sampling medium, all years"
#| fig-width: 12
#| fig-height: 3
#| warning: false

# Create a summary of counts by the two variables
vm_exploration_vkat_x_medium <- vm_merged_lookups |>
  group_by(Vannkategori_Description, MediumID_Name) |>
  summarise(n_samples = n(), .groups = "drop")

threshold_value_vkat_medium <- mean(
  log10(vm_exploration_vkat_x_medium$n_samples),
  na.rm = TRUE
)

vannkategori_order_vkat <- vm_exploration_vkat_x_medium |>
  group_by(Vannkategori_Description) |>
  summarise(total = sum(n_samples)) |>
  arrange(total) |>
  pull(Vannkategori_Description)

medium_order_vkat <- vm_exploration_vkat_x_medium |>
  group_by(MediumID_Name) |>
  summarise(total = sum(n_samples)) |>
  arrange(desc(total)) |>
  pull(MediumID_Name)

vm_exploration_vkat_x_medium_complete <- vm_exploration_vkat_x_medium |>
  mutate(
    Vannkategori_Description_wrapped = str_wrap(Vannkategori_Description, 20),
    MediumID_Name_wrapped = str_wrap(MediumID_Name, 20)
  ) |>
  complete(
    Vannkategori_Description_wrapped = str_wrap(vannkategori_order_vkat, 20),
    MediumID_Name_wrapped = str_wrap(medium_order_vkat, 20),
    fill = list(n_samples = 0)
  ) |>
  mutate(
    Vannkategori_Description_wrapped = factor(
      Vannkategori_Description_wrapped,
      levels = str_wrap(vannkategori_order_vkat, 20)
    ),
    MediumID_Name_wrapped = factor(
      MediumID_Name_wrapped,
      levels = str_wrap(medium_order_vkat, 20)
    )
  )

vm_exploration_vkat_x_medium_complete |>
  ggplot(aes(
    x = MediumID_Name_wrapped,
    y = Vannkategori_Description_wrapped,
    fill = n_samples
  )) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_text(
    aes(
      label = if_else(n_samples > 0, scales::comma(n_samples), ""),
      color = if_else(
        n_samples > 0 & log10(n_samples) > threshold_value_vkat_medium,
        "black",
        "white"
      )
    ),
    size = 3
  ) +
  scale_fill_viridis_c(trans = "log10", na.value = "grey90") +
  scale_color_identity() +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(x = "", y = "") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 7),
    axis.text.y = element_text(size = 7),
    plot.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "none"
  )
```

@tbl-sample-distribution gives the cumulative proportion of samples corresponding to each combination of reported water category and medium. This allows us to make strategic decisions about which groups to prioritise conversion into our eData format. I have elected to use 90% as an arbitrary cut-off point - that is, once I have 90% coverage of water categories and sampling media (before date-based sampling), I will consider that sufficient data.

```{r}
#| label: tbl-sample-distribution

# Calculate sample distribution with cumulative percentages
sample_distribution <- vm_exploration_vkat_x_medium |>
  arrange(desc(n_samples)) |>
  mutate(
    pct_of_total = 100 * n_samples / sum(n_samples),
    cumulative_n = cumsum(n_samples),
    cumulative_pct = 100 * cumulative_n / sum(n_samples)
  ) |>
  select(
    Vannkategori_Description,
    MediumID_Name,
    n_samples,
    pct_of_total,
    cumulative_n,
    cumulative_pct
  )

datatable(
  sample_distribution,
  colnames = c(
    "Water Category",
    "Medium",
    "N Samples",
    "% of Total",
    "Cumulative N",
    "Cumulative %"
  ),
  options = list(
    pageLength = 10,
    order = list(list(3, 'desc'))
  )
) |>
  formatRound(columns = c("pct_of_total", "cumulative_pct"), digits = 1) |>
  formatCurrency(
    columns = c("n_samples", "cumulative_n"),
    currency = "",
    digits = 0
  )
```


We also consider sample size of @fig-heatmap-campaign-biota by campaign. This includes many largely irrelevant (not-aquatic) species, but the vast majority are standard aquatic monitoring species (mussels, cod, edible crabs, etc.).

```{r}
#| label: fig-heatmap-campaign-biota
#| fig-cap: "Distribution of Vannmiljø copper samples by monitoring campaign and species, all years"
#| fig-width: 12
#| fig-height: 20
#| warning: false

# Create a summary of counts by the two variables (filter to non-NA species)
vm_exploration_species <- vm_merged_lookups |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN = case_when(
      is.na(Name_EN) ~ "Other",
      TRUE ~ Name_EN
    )
  ) |>
  filter(!is.na(VitenskapligNavn), VitenskapligNavn != "Artsuavhengig") |>
  group_by(CAMPAIGN_DESCRIPTION_EN, VitenskapligNavn) |>
  summarise(n_samples = n(), .groups = "drop")

# Calculate text color threshold
threshold_value_species <- mean(
  log10(vm_exploration_species$n_samples),
  na.rm = TRUE
)

campaign_order_species_new <- vm_exploration_species |>
  group_by(CAMPAIGN_DESCRIPTION_EN) |>
  summarise(total = sum(n_samples)) |>
  arrange(desc(total)) |>
  pull(CAMPAIGN_DESCRIPTION_EN)

species_order_new <- vm_exploration_species |>
  group_by(VitenskapligNavn) |>
  summarise(total = sum(n_samples)) |>
  arrange(total) |>
  pull(VitenskapligNavn)

vm_exploration_species_complete <- vm_exploration_species |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN_wrapped = str_wrap(CAMPAIGN_DESCRIPTION_EN, 30),
    VitenskapligNavn_wrapped = str_wrap(VitenskapligNavn, 20)
  ) |>
  complete(
    CAMPAIGN_DESCRIPTION_EN_wrapped = str_wrap(campaign_order_species_new, 30),
    VitenskapligNavn_wrapped = str_wrap(species_order_new, 20),
    fill = list(n_samples = 0)
  ) |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN_wrapped = factor(
      CAMPAIGN_DESCRIPTION_EN_wrapped,
      levels = str_wrap(campaign_order_species_new, 30)
    ),
    VitenskapligNavn_wrapped = factor(
      VitenskapligNavn_wrapped,
      levels = str_wrap(species_order_new, 20)
    )
  )

vm_exploration_species_complete |>
  ggplot(aes(
    x = VitenskapligNavn_wrapped,
    y = CAMPAIGN_DESCRIPTION_EN_wrapped,
    fill = n_samples
  )) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_text(
    aes(
      label = if_else(n_samples > 0, scales::comma(n_samples), ""),
      color = if_else(
        n_samples > 0 & log10(n_samples) > threshold_value_species,
        "black",
        "white"
      )
    ),
    size = 3
  ) +
  scale_fill_viridis_c(trans = "log10", na.value = "grey90") +
  scale_color_identity() +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(x = "", y = "") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 7),
    axis.text.y = element_text(size = 7),
    plot.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "none"
  ) +
  coord_flip()

```

Finally, we check units by sampling medium @fig-heatmap-units-medium. We would expect mass/volume units to be used for water and air samples, and mass/mass (dry or wet) units to be used for biota, soil, sediment, and other largely solid matrices. Fortunately this is borne out by the data. 

```{r}
#| label: fig-heatmap-units-medium
#| fig-cap: "Distribution of Vannmiljø copper samples by monitoring campaign and species, all years"
#| fig-width: 12
#| fig-height: 3
#| warning: false

vm_exploration_units_x_medium <- vm_merged_lookups |>
  group_by(Unit_Name, MediumID_Name) |>
  summarise(n_samples = n(), .groups = "drop")
threshold_value_units_x_medium <- mean(
  log10(vm_exploration_units_x_medium$n_samples),
  na.rm = TRUE
)

unit_order <- vm_exploration_units_x_medium |>
  group_by(Unit_Name) |>
  summarise(total = sum(n_samples)) |>
  arrange(total) |>
  pull(Unit_Name)

medium_order_units <- vm_exploration_units_x_medium |>
  group_by(MediumID_Name) |>
  summarise(total = sum(n_samples)) |>
  arrange(desc(total)) |>
  pull(MediumID_Name)

vm_exploration_units_x_medium_complete <- vm_exploration_units_x_medium |>
  mutate(
    Unit_Name_wrapped = str_wrap(Unit_Name, 20),
    MediumID_Name_wrapped = str_wrap(MediumID_Name, 20)
  ) |>
  complete(
    Unit_Name_wrapped = str_wrap(unit_order, 20),
    MediumID_Name_wrapped = str_wrap(medium_order_units, 20),
    fill = list(n_samples = 0)
  ) |>
  mutate(
    Unit_Name_wrapped = factor(
      Unit_Name_wrapped,
      levels = str_wrap(unit_order, 20)
    ),
    MediumID_Name_wrapped = factor(
      MediumID_Name_wrapped,
      levels = str_wrap(medium_order_units, 20)
    )
  )


vm_exploration_units_x_medium_complete |>
  ggplot(aes(
    x = MediumID_Name_wrapped,
    y = Unit_Name_wrapped,
    fill = n_samples
  )) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_text(
    aes(
      label = if_else(n_samples > 0, scales::comma(n_samples), ""),
      color = if_else(
        n_samples > 0 & log10(n_samples) > threshold_value_units_x_medium,
        "black",
        "white"
      )
    ),
    size = 3
  ) +
  scale_fill_viridis_c(trans = "log10", na.value = "grey90") +
  scale_color_identity() +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(x = "", y = "") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 10),
    axis.text.y = element_text(size = 10),
    plot.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "none",
  )

```

Let's check out methods (@tbl-methods-by-year). Anything with an eData Classification column has been matched with the lookup. Vannmiljø methods without this are either not able to be matched (e.g. UKJENT) or aren't included in our filtered dataset (see next section).

```{r}
#| label: tbl-methods-by-year
#| tbl-cap: "Vannmiljø sampling and analysis method IDs reported by number of samples (rows), last year used, and corresponding eData Classification (if relevant)."

vm_exploration_methods_x_year <- vm_merged_lookups |>
  select(Tid_provetak, Provetakmetode_id, Analysemetode_id, Medium_id) |>
  pivot_longer(
    cols = c(Provetakmetode_id, Analysemetode_id),
    names_to = "Method_Type",
    values_to = "Method_Name"
  ) |>
  mutate(
    Year = ymd_hms(Tid_provetak) |> year(),
    Method_Type = case_match(
      Method_Type,
      "Provetakmetode_id" ~ "Sampling",
      "Analysemetode_id" ~ "Analysis"
    ),
    Base_Method = str_extract(
      Method_Name,
      pattern = "([A-Z\\-]* [0-9]{4,5}?(-[0-9]))"
    )
  )

vm_exploration_methods_x_year_table <- vm_exploration_methods_x_year |>
  group_by(Method_Name) |>
  reframe(
    Rows = n(),
    Method_Type = paste(Method_Type),
    Last_Year_Used = last(Year),
    Base_Method
  ) |>
  distinct() |>
  arrange(desc(Rows)) |>
  left_join(
    vm_lookup_methods |> select(ISO_ID, PROTOCOL_NAME),
    by = c("Method_Name" = "ISO_ID")
  )

datatable(
  vm_exploration_methods_x_year_table,
  colnames = c(
    "n",
    "Vm ID",
    "Rows",
    "Type",
    "Last Used",
    "Base Code",
    "eData Classification"
  )
)

```

## Apply Filters

As discussed above, to avoid spending excessive time on low-incidence edge cases we filter the data at this stage to the most relevant compartments and dates. The results of this summarisation are calculated below.

Included:

-   See config tables in [Configuration](##Configuration) for specifics
-   Point sites only (no polygons)
-   Mainland Norway only (excluded Svalbard)

Excluded:

-   Svalbard sites (n=1)

```{r}
#| label: processing-apply-filters-and-summarise

# Track exclusions step by step
exclusion_summary <- tibble(
  Filter = character(),
  Rows_Excluded = integer(),
  Rows_Remaining = integer(),
  Excluded_Values = character()
)

# Start with merged data
current_data <- vm_merged_lookups |>
  mutate(SAMPLING_DATE = as.Date(Tid_provetak))

initial_count <- nrow(current_data)

# Step 1: Compartment filters
excluded_compartments <- current_data |>
  filter(
    !(ENVIRON_COMPARTMENT_vkat %in%
      config$filters$compartment |
      ENVIRON_COMPARTMENT_medium %in% config$filters$compartment) |
      !(ENVIRON_COMPARTMENT_SUB_vkat %in%
        config$filters$compartment_sub |
        ENVIRON_COMPARTMENT_SUB_medium %in% config$filters$compartment_sub)
  ) |>
  group_by(ENVIRON_COMPARTMENT_SUB_vkat, ENVIRON_COMPARTMENT_SUB_medium) |>
  reframe(n = n()) |>
  mutate(
    label = glue(
      "{ENVIRON_COMPARTMENT_SUB_vkat}/{ENVIRON_COMPARTMENT_SUB_medium} (n={n})"
    )
  ) |>
  pull(label) |>
  paste(collapse = "; ")

after_compartment <- current_data |>
  filter(
    ENVIRON_COMPARTMENT_vkat %in%
      config$filters$compartment |
      ENVIRON_COMPARTMENT_medium %in% config$filters$compartment,
    ENVIRON_COMPARTMENT_SUB_vkat %in%
      config$filters$compartment_sub |
      ENVIRON_COMPARTMENT_SUB_medium %in% config$filters$compartment_sub
  )

exclusion_summary <- exclusion_summary |>
  add_row(
    Filter = "Wrong compartment/ subcompartment",
    Rows_Excluded = initial_count - nrow(after_compartment),
    Rows_Remaining = nrow(after_compartment),
    Excluded_Values = excluded_compartments
  )

# Step 2: Site type filter
excluded_sites <- after_compartment |>
  filter(
    Objekttype != "point" |
      Vannlokalitetsnavn == config$filters$exclude_sites
  ) |>
  group_by(Objekttype, Vannlokalitetsnavn) |>
  reframe(n = n()) |>
  mutate(label = glue("{Objekttype}: {Vannlokalitetsnavn} (n={n})")) |>
  pull(label) |>
  paste(collapse = "; ")

after_site_type <- after_compartment |>
  filter(
    Objekttype == "point",
    Vannlokalitetsnavn != config$filters$exclude_sites
  )

exclusion_summary <- exclusion_summary |>
  add_row(
    Filter = "Not point sites or excluded sites",
    Rows_Excluded = nrow(after_compartment) - nrow(after_site_type),
    Rows_Remaining = nrow(after_site_type),
    Excluded_Values = excluded_sites
  )

# Step 3: Date range filter
excluded_dates <- after_site_type |>
  filter(SAMPLING_DATE < config$filters$date_start) |>
  group_by(year = year(SAMPLING_DATE)) |>
  reframe(n = n()) |>
  mutate(label = glue("{year} (n={n})")) |>
  pull(label) |>
  paste(collapse = "; ")

after_date <- after_site_type |>
  filter(SAMPLING_DATE >= config$filters$date_start)

exclusion_summary <- exclusion_summary |>
  add_row(
    Filter = "Outside date range",
    Rows_Excluded = nrow(after_site_type) - nrow(after_date),
    Rows_Remaining = nrow(after_date),
    Excluded_Values = excluded_dates
  )

# Add total row
exclusion_summary <- exclusion_summary |>
  add_row(
    Filter = "**Total**",
    Rows_Excluded = initial_count - nrow(after_date),
    Rows_Remaining = nrow(after_date),
    Excluded_Values = ""
  ) |>
  mutate(
    Percent_Excluded = round(100 * Rows_Excluded / initial_count, 1),
    Percent_Remaining = round(100 * Rows_Remaining / initial_count, 1)
  )

# Store final filtered data
vm_filtered <- after_date

# Validation
vm_filtered_validation <- vm_filtered |>
  mutate(
    .compartment_valid = ENVIRON_COMPARTMENT_vkat %in%
      config$filters$compartment |
      ENVIRON_COMPARTMENT_medium %in% config$filters$compartment,
    .sub_compartment_valid = ENVIRON_COMPARTMENT_SUB_vkat %in%
      config$filters$compartment_sub |
      ENVIRON_COMPARTMENT_SUB_medium %in% config$filters$compartment_sub
  ) |>
  col_vals_equal(
    columns = c(.compartment_valid, .sub_compartment_valid),
    value = TRUE,
    label = "At least one compartment source is within specified filter criteria",
    actions = validation_config$standard
  ) |>
  col_vals_in_set(
    Objekttype,
    c("point"),
    label = "All sites are of type point (not polygon)",
    actions = validation_config$standard
  ) |>
  col_vals_not_in_set(
    Vannlokalitetsnavn,
    config$filters$exclude_sites,
    label = "No sites are in set of excluded sites",
    actions = validation_config$standard
  )

message(glue(
  "vm_merged_lookups ({nrow(vm_merged_lookups)} x {ncol(vm_merged_lookups)}) -> vm_filtered ({nrow(vm_filtered)} x {ncol(vm_filtered)}, {round(100*nrow(vm_filtered)/nrow(vm_merged_lookups), 1)}% of starting data)"
))

```

For completeness, we also report on excluded groups (@tbl-processing-excluded-data)

- [ ] Table HTML is broken here.

```{r}
#| label: fig-filter-visualisation
#| fig-cap: Vannmiljø copper sample size at different steps of filtration
#| fig-width: 12

# Create filtering summary for visualization
filter_flow <- exclusion_summary |>
  filter(Filter != "**Total**") |>
  mutate(
    stage = c("Compartments", "Point sites", "Date range"),
    pct = Percent_Remaining,
    label = glue("{scales::comma(Rows_Remaining)}\n({pct}%)")
  ) |>
  add_row(
    Filter = "Raw",
    stage = "Raw",
    Rows_Remaining = initial_count,
    pct = 100,
    label = glue("{scales::comma(initial_count)}\n(100%)"),
    .before = 1
  )

ggplot(filter_flow, aes(x = stage, y = Rows_Remaining)) +
  geom_col(aes(fill = stage)) +
  geom_text(aes(label = label), vjust = -0.5, size = 3) +
  scale_fill_brewer(name = "Filtering Stage", palette = "Set1") +
  scale_x_discrete(limits = filter_flow$stage) +
  scale_y_continuous(
    labels = scales::comma,
    expand = expansion(mult = c(0, 0.1))
  ) +
  labs(
    x = NULL,
    y = "Number of Records"
  ) +
  theme_minimal()
```

A Sankey diagram would be better, but @fig-filter-visualisation will do for now.

```{r}
#| label: tbl-processing-excluded-data
#| tbl-cap: "Summary of excluded data by filter criterion"

datatable(
  exclusion_summary,
  colnames = c(
    "Filter Criterion",
    "Rows Excluded",
    "Rows Remaining",
    "Excluded Values",
    "% Excluded",
    "% Remaining"
  ),
  rownames = FALSE,
  escape = c("Excluded Values"),
  options = list(
    columnDefs = list(
      list(width = "300px", targets = 3)
    )
  )
) |>
  formatCurrency(
    columns = c("Rows_Excluded", "Rows_Remaining"),
    currency = "",
    digits = 0
  )
```

At time of writing, we've got 94% of compartment data, which is well above our goal of 90%.

## Quality Checks

We check for conflicts between our two lookups here - because we look up compartments, sites, biota, etc. based on both site `Vannlokalkode` and measurments' `MediumID`, it's possible that we might end up with conflicts and/or cases where neither of these give us the relevant data and we have to guess from context. Right now we filter down to only freshwater cases, hence the fact that we don't get hits below.

Note: It's not necessarily a problem that something turns up as a conflict - we fix a lot of these in the next steps.

```{r}
#| label: tbl-compartment-conflicts
#| tbl-cap: "Conflicts between vannkategori and medium lookups for environmental compartments"

# Check for conflicts between lookup sources
conflict_check <- vm_filtered |>
  group_by(
    ENVIRON_COMPARTMENT_vkat,
    ENVIRON_COMPARTMENT_medium,
    ENVIRON_COMPARTMENT_SUB_vkat,
    ENVIRON_COMPARTMENT_SUB_medium
  ) |>
  reframe(count = n()) |>
  mutate(
    compartment_conflict = ENVIRON_COMPARTMENT_vkat !=
      ENVIRON_COMPARTMENT_medium,
    subcompartment_conflict = ENVIRON_COMPARTMENT_SUB_vkat !=
      ENVIRON_COMPARTMENT_SUB_medium,
    Conflict = case_when(
      compartment_conflict & subcompartment_conflict ~ "Main, Sub",
      compartment_conflict ~ "Main",
      subcompartment_conflict ~ "Sub",
      .default = NA_character_
    )
  ) |>
  filter(!is.na(Conflict)) |>
  arrange(desc(count))

if (nrow(conflict_check) > 0) {
  warning(
    glue(
      "{nrow(conflict_check)} conflicts detected between vannkategori and medium lookups.
      Result: conflict_check ({nrow(conflict_check)} x {ncol(conflict_check)})"
    )
  )
}

datatable(
  conflict_check |>
    select(
      ENVIRON_COMPARTMENT_vkat,
      ENVIRON_COMPARTMENT_medium,
      ENVIRON_COMPARTMENT_SUB_vkat,
      ENVIRON_COMPARTMENT_SUB_medium,
      count,
      Conflict
    ),
  colnames = c(
    "Compartment (Vannkategory)",
    "Compartment (MediumName)",
    "Subcompartment (Vannkategory)",
    "Subcompartment (MediumName)",
    "Rows",
    "Conflict Type"
  )
)
```

@tbl-compartment-conflicts lists all combinations of compartments and subcompartments, and whether any conflict is detected between the two compartments. I feel a need to be explicit and reproducible about what trumps what where, so see below.

-   Subcompartment (Vkat) == Freshwater & Subcompartment (Medium) == Aquatic Sediment -\> Aquatic Sediment
-   If one column returns biota, we use that (even if it's a fish e.g. living in the sea...)
-   If only one (sub)compartment value is a wildcard (*), use the other. If both are wildcards, flag the row for exclusion. 

```{r}
#| label: resolve-compartment-conflicts

# TODO: This maybe should be a table, although the logic is a bit complicated and twisty
resolve_compartment_conflicts <- function(df) {
  df <- df |>
    mutate(
      # SET COMPARTMENT
      ENVIRON_COMPARTMENT_resolved = case_when(
        # Explicit rule: Biota trumps all
        ENVIRON_COMPARTMENT_vkat == "Biota" |
          ENVIRON_COMPARTMENT_medium == "Biota" ~ "Biota",

        # if ENVIRON_COMPARTMENT_vkat is wildcard then use ENVIRON_COMPARTMENT_medium
        !has_data_issue(ENVIRON_COMPARTMENT_vkat) &
          !has_data_issue(ENVIRON_COMPARTMENT_medium) &
          ENVIRON_COMPARTMENT_vkat == "*" &
          ENVIRON_COMPARTMENT_medium != "*" ~ ENVIRON_COMPARTMENT_medium,

        # if ENVIRON_COMPARTMENT_medium is wildcard then use ENVIRON_COMPARTMENT_vkat
        !has_data_issue(ENVIRON_COMPARTMENT_vkat) &
          !has_data_issue(ENVIRON_COMPARTMENT_medium) &
          ENVIRON_COMPARTMENT_vkat != "*" &
          ENVIRON_COMPARTMENT_medium == "*" ~ ENVIRON_COMPARTMENT_vkat,

        # if there's no conflict, use the first value  - unless both are *
        ENVIRON_COMPARTMENT_vkat == ENVIRON_COMPARTMENT_medium &
          ENVIRON_COMPARTMENT_medium != "*" ~ ENVIRON_COMPARTMENT_medium,

        # If none of these rules resolve the conflict, flag the row
        .default = "FLAG: Compartment conflict."
      ),

      # SET SUBCOMPARTMENT
      ENVIRON_COMPARTMENT_SUB_resolved = case_when(
        # Handle Biota compartment first (all biota rules together)
        ENVIRON_COMPARTMENT_resolved == "Biota" ~ case_when(
          # Prefer explicit "Biota, Aquatic" from either source
          ENVIRON_COMPARTMENT_SUB_vkat == "Biota, Aquatic" |
            ENVIRON_COMPARTMENT_SUB_medium ==
              "Biota, Aquatic" ~ "Biota, Aquatic",

          # Use biota-specific lookup if available
          !is.na(ENVIRON_COMPARTMENT_SUB_biota) &
            ENVIRON_COMPARTMENT_SUB_biota !=
              "*" ~ ENVIRON_COMPARTMENT_SUB_biota,

          .default = "FLAG: Compartment conflict."
        ),

        # Non-biota compartments: simpler logic
        # Sediment trumps water
        ENVIRON_COMPARTMENT_SUB_medium ==
          "Aquatic Sediment" ~ "Aquatic Sediment",

        # Wildcard handling
        ENVIRON_COMPARTMENT_SUB_vkat == "*" &
          ENVIRON_COMPARTMENT_SUB_medium !=
            "*" ~ ENVIRON_COMPARTMENT_SUB_medium,
        ENVIRON_COMPARTMENT_SUB_medium == "*" &
          ENVIRON_COMPARTMENT_SUB_vkat != "*" ~ ENVIRON_COMPARTMENT_SUB_vkat,

        # No conflict
        ENVIRON_COMPARTMENT_SUB_vkat == ENVIRON_COMPARTMENT_SUB_medium &
          ENVIRON_COMPARTMENT_SUB_medium != "*" ~ ENVIRON_COMPARTMENT_SUB_vkat,

        .default = "FLAG: Compartment conflict."
      )
    )

  # Count potential conflicts (where both sources differ and neither is NA)
  # Why is this generating negative results?
  n_potential_compartment <- df |>
    filter(
      !has_data_issue(ENVIRON_COMPARTMENT_vkat),
      !has_data_issue(ENVIRON_COMPARTMENT_medium),
      ENVIRON_COMPARTMENT_vkat != ENVIRON_COMPARTMENT_medium
    ) |>
    nrow()

  n_potential_sub <- df |>
    filter(
      !has_data_issue(ENVIRON_COMPARTMENT_SUB_vkat),
      !has_data_issue(ENVIRON_COMPARTMENT_SUB_medium),
      ENVIRON_COMPARTMENT_SUB_vkat != ENVIRON_COMPARTMENT_SUB_medium
    ) |>
    nrow()

  # Count resolved
  n_resolved <- df |>
    tally(ENVIRON_COMPARTMENT_resolved != "FLAG: Compartment conflict.") -
    n_potential_compartment
  n_resolved_sub <- df |>
    tally(ENVIRON_COMPARTMENT_SUB_resolved != "FLAG: Compartment conflict.") -
    n_potential_sub

  # Count unresolved
  n_unresolved <- df |>
    tally(ENVIRON_COMPARTMENT_resolved == "FLAG: Compartment conflict.")
  n_unresolved_sub <- df |>
    tally(ENVIRON_COMPARTMENT_SUB_resolved == "FLAG: Compartment conflict.")

  # todo: this is a little confusing, because n_resolved currently includes all the problem-free rows too
  message(glue(
    "Compartments: {n_resolved} resolved, {n_unresolved} flagged for removal (of {n_potential_compartment} conflicts)"
  ))
  message(glue(
    "Subcompartments: {n_resolved_sub} resolved, {n_unresolved_sub} flagged for removal (of {n_potential_sub} conflicts)"
  ))

  if (n_unresolved + n_unresolved_sub > 0) {
    warning(
      "Some (sub)compartments still aren't resolved and have been flagged for removal."
    )
  }

  df
}

vm_compartment_conflicts_resolved <- resolve_compartment_conflicts(
  vm_filtered
)

test <- vm_filtered |>
  filter(
    ENVIRON_COMPARTMENT_medium == "Biota",
    !str_detect(
      ENVIRON_COMPARTMENT_SUB_vkat,
      "Biota"
    )
  ) |>
  group_by(
    ENVIRON_COMPARTMENT_medium,
    ENVIRON_COMPARTMENT_vkat,
    ENVIRON_COMPARTMENT_SUB_medium,
    ENVIRON_COMPARTMENT_SUB_vkat,
    ENVIRON_COMPARTMENT_SUB_biota
  ) |>
  reframe(n())

vm_compartment_conflicts_resolved |>
  # inspect data
  group_by(
    ENVIRON_COMPARTMENT_resolved,
    ENVIRON_COMPARTMENT_SUB_resolved
  ) |>
  reframe(Rows = n(), Mean = mean(Verdi))

```

In some cases we're unable to resolve conflicts with recourse to the simple rules above. In these cases, we simply remove the rows, and create `vm_compartment_conflicts_resolved_removed`. 

```{r}
#| label: remove-compartment-conflicts
vm_compartment_conflicts_resolved_removed <- vm_compartment_conflicts_resolved |>
  filter(
    ENVIRON_COMPARTMENT_resolved != "FLAG: Compartment conflict.",
    ENVIRON_COMPARTMENT_SUB_resolved != "FLAG: Compartment conflict."
  )

vm_compartment_conflicts_resolved_removed |>
  # inspect data
  group_by(
    ENVIRON_COMPARTMENT_resolved,
    ENVIRON_COMPARTMENT_SUB_resolved
  ) |>
  reframe(Rows = n(), Mean = mean(Verdi))

# Now validate with pointblank
vm_compartment_conflicts_resolved_removed |>
  col_vals_in_set(
    ENVIRON_COMPARTMENT_resolved,
    eDataDRF::environ_compartments_vocabulary()
  ) |>
  col_vals_in_set(
    ENVIRON_COMPARTMENT_SUB_resolved,
    purrr::flatten(eDataDRF::environ_compartments_sub_vocabulary())
  )

# Todo: Could this be a conditional code chunk?
# vm_compartment_conflicts_resolved_removed |>
#   filter(
#     !(ENVIRON_COMPARTMENT_resolved %in%
#       eDataDRF::environ_compartments_vocabulary())
#   ) |>
#   group_by(ENVIRON_COMPARTMENT_resolved) |>
#   reframe(n = n())

# vm_compartment_conflicts_resolved_removed |>
#   filter(
#     !(ENVIRON_COMPARTMENT_SUB_resolved %in%
#       purrr::flatten(eDataDRF::environ_compartments_sub_vocabulary()))
#   ) |>
#   group_by(ENVIRON_COMPARTMENT_SUB_resolved) |>
#   reframe(n = n())

message(glue(
  "vm_filtered ({nrow(vm_filtered)} x {ncol(vm_filtered)}) -> vm_compartment_conflicts_resolved_removed ({nrow(vm_compartment_conflicts_resolved_removed)} x {ncol(vm_compartment_conflicts_resolved_removed)})"
))

```

And let's take a look at anyting we failed to resolve: @tbl-unresolved-conflicts. At time of writing we have 1916 rows with unresolved conflicts, but this is a very small number and I don't (currently) judge it a good use of time to chase down these edge cases.

```{r}
#| label: tbl-unresolved-conflicts
#| tbl-cap: "Remaining unresolved conflicts after applying resolution rules"

# Check for unresolved conflicts
unresolved_conflicts_summary <- vm_compartment_conflicts_resolved |>
  # Keep only rows where conflicts weren't resolved
  filter(
    ENVIRON_COMPARTMENT_SUB_resolved == "FLAG: Compartment conflict." &
      ENVIRON_COMPARTMENT_resolved == "FLAG: Compartment conflict."
  ) |>
  group_by(
    ENVIRON_COMPARTMENT_vkat,
    ENVIRON_COMPARTMENT_medium,
    ENVIRON_COMPARTMENT_SUB_vkat,
    ENVIRON_COMPARTMENT_SUB_medium
  ) |>
  reframe(count = n()) |>
  mutate(
    compartment_conflict = ENVIRON_COMPARTMENT_vkat !=
      ENVIRON_COMPARTMENT_medium,
    subcompartment_conflict = ENVIRON_COMPARTMENT_SUB_vkat !=
      ENVIRON_COMPARTMENT_SUB_medium,
    Conflict = case_when(
      compartment_conflict & subcompartment_conflict ~ "Main, Sub",
      compartment_conflict ~ "Main",
      subcompartment_conflict ~ "Sub",
      .default = NA_character_
    )
  ) |>
  arrange(desc(count))

datatable(
  unresolved_conflicts_summary |>
    select(
      ENVIRON_COMPARTMENT_vkat,
      ENVIRON_COMPARTMENT_medium,
      ENVIRON_COMPARTMENT_SUB_vkat,
      ENVIRON_COMPARTMENT_SUB_medium,
      count,
      Conflict
    ),
  colnames = c(
    "Compartment (Vannkategory)",
    "Compartment (MediumName)",
    "Subcompartment (Vannkategory)",
    "Subcompartment (MediumName)",
    "Rows",
    "Conflict Type"
  )
)
```

`vm_compartment_conflicts_resolved_removed` represents our dataset after filtering and resolution/removal of compartment lookup conflicts. However, we need to repeat the exercise for geographic features, which we also use two lookups for.

@tbl-geographic-conflicts also lists all combinations of conflicts. We haven't applied simple, rules-based resolution yet, however. 

```{r}
#| label: tbl-geographic-conflicts
#| tbl-cap: "Conflicts between vannkategori and medium lookups for geographic features"

# Check geographic feature consistency
vm_geo_conflicts_flagged <- vm_compartment_conflicts_resolved_removed |>
  mutate(
    # Only flag as conflict if both values are present and different
    # FIXME: We use different logic here and in the below block :(
    geo_conflict = !has_data_issue(SITE_GEOGRAPHIC_FEATURE_vkat) &
      !has_data_issue(SITE_GEOGRAPHIC_FEATURE_medium) &
      SITE_GEOGRAPHIC_FEATURE_vkat != "*" &
      SITE_GEOGRAPHIC_FEATURE_medium != "*" &
      SITE_GEOGRAPHIC_FEATURE_vkat != SITE_GEOGRAPHIC_FEATURE_medium,

    geo_sub_conflict = !has_data_issue(SITE_GEOGRAPHIC_FEATURE_SUB_vkat) &
      !has_data_issue(SITE_GEOGRAPHIC_FEATURE_SUB_medium) &
      SITE_GEOGRAPHIC_FEATURE_SUB_vkat != "*" &
      SITE_GEOGRAPHIC_FEATURE_SUB_medium != "*" &
      SITE_GEOGRAPHIC_FEATURE_SUB_vkat != SITE_GEOGRAPHIC_FEATURE_SUB_medium,

    Conflict = case_when(
      geo_conflict & geo_sub_conflict ~ "Main, Sub",
      geo_conflict ~ "Main",
      geo_sub_conflict ~ "Sub",
      .default = NA_character_
    )
  )

vm_geo_conflicts_flagged |> group_by(Conflict) |> reframe(n())

vm_geo_conflicts_summary <- vm_geo_conflicts_flagged |>
  group_by(
    SITE_GEOGRAPHIC_FEATURE_vkat,
    SITE_GEOGRAPHIC_FEATURE_medium,
    SITE_GEOGRAPHIC_FEATURE_SUB_vkat,
    SITE_GEOGRAPHIC_FEATURE_SUB_medium
  ) |>
  reframe(count = n(), Conflict = first(Conflict)) |>
  filter(!is.na(Conflict)) |>
  arrange(desc(count))

if (nrow(vm_geo_conflicts_summary) > 0) {
  warning(glue(
    "Geographic feature conflicts: {nrow(vm_geo_conflicts_summary)} combinations ({sum(vm_geo_conflicts_summary$count)} rows)"
  ))
}

datatable(
  vm_geo_conflicts_summary |>
    select(
      SITE_GEOGRAPHIC_FEATURE_vkat,
      SITE_GEOGRAPHIC_FEATURE_medium,
      SITE_GEOGRAPHIC_FEATURE_SUB_vkat,
      SITE_GEOGRAPHIC_FEATURE_SUB_medium,
      count,
      Conflict
    ),
  colnames = c(
    "Feature (Vannkategory)",
    "Feature (MediumName)",
    "Subfeature (Vannkategory)",
    "Subfeature (MediumName)",
    "Rows",
    "Conflict Type"
  )
)
```

If the table above is empty (I don't update this text every time I update the pipeline), it hopefully means that the simple rules (and the fact that we've *already* removed conflicting compartments, and there may be correlation between the two) have worked. 

My experience has been that `SITE_GEOGRAPHIC_FEATURE_SUB` is a lot harder to get information on, both in this conversion but also in papers. As it's less important, I've elected to just mark it as "Not reported" if it's not easy to solve with simple rules. This is unlikely to affect the eventual analysis.

```{r}
#| label: resolve-feature-conflicts

resolve_geographic_conflicts <- function(df) {
  df <- df |>
    mutate(
      # SET GEOGRAPHIC FEATURE
      SITE_GEOGRAPHIC_FEATURE_resolved = case_when(
        # if SITE_GEOGRAPHIC_FEATURE_vkat is wildcard then use SITE_GEOGRAPHIC_FEATURE_medium
        !has_data_issue(SITE_GEOGRAPHIC_FEATURE_vkat) &
          !has_data_issue(SITE_GEOGRAPHIC_FEATURE_medium) &
          SITE_GEOGRAPHIC_FEATURE_vkat == "*" &
          SITE_GEOGRAPHIC_FEATURE_medium !=
            "*" ~ SITE_GEOGRAPHIC_FEATURE_medium,

        # if SITE_GEOGRAPHIC_FEATURE_medium is wildcard then use SITE_GEOGRAPHIC_FEATURE_vkat
        !has_data_issue(SITE_GEOGRAPHIC_FEATURE_vkat) &
          !has_data_issue(SITE_GEOGRAPHIC_FEATURE_medium) &
          SITE_GEOGRAPHIC_FEATURE_vkat != "*" &
          SITE_GEOGRAPHIC_FEATURE_medium == "*" ~ SITE_GEOGRAPHIC_FEATURE_vkat,

        # if there's no conflict (unless both are *), use the first value
        SITE_GEOGRAPHIC_FEATURE_vkat == SITE_GEOGRAPHIC_FEATURE_medium &
          SITE_GEOGRAPHIC_FEATURE_medium !=
            "*" ~ SITE_GEOGRAPHIC_FEATURE_medium,

        # If none of these rules resolve the conflict, flag the row
        .default = "FLAG: Geographic conflict."
      ),

      # SET GEOGRAPHIC FEATURE SUB
      SITE_GEOGRAPHIC_FEATURE_SUB_resolved = case_when(
        # if SITE_GEOGRAPHIC_FEATURE_SUB_vkat is wildcard then use SITE_GEOGRAPHIC_FEATURE_SUB_medium
        !has_data_issue(SITE_GEOGRAPHIC_FEATURE_SUB_vkat) &
          !has_data_issue(SITE_GEOGRAPHIC_FEATURE_SUB_medium) &
          SITE_GEOGRAPHIC_FEATURE_SUB_vkat == "*" &
          SITE_GEOGRAPHIC_FEATURE_SUB_medium !=
            "*" ~ SITE_GEOGRAPHIC_FEATURE_SUB_medium,

        # if SITE_GEOGRAPHIC_FEATURE_SUB_medium is wildcard then use SITE_GEOGRAPHIC_FEATURE_SUB_vkat
        !has_data_issue(SITE_GEOGRAPHIC_FEATURE_SUB_vkat) &
          !has_data_issue(SITE_GEOGRAPHIC_FEATURE_SUB_medium) &
          SITE_GEOGRAPHIC_FEATURE_SUB_vkat != "*" &
          SITE_GEOGRAPHIC_FEATURE_SUB_medium ==
            "*" ~ SITE_GEOGRAPHIC_FEATURE_SUB_vkat,

        # if there's no conflict (unless both are *), use the first value
        SITE_GEOGRAPHIC_FEATURE_SUB_vkat == SITE_GEOGRAPHIC_FEATURE_SUB_medium &
          SITE_GEOGRAPHIC_FEATURE_SUB_medium !=
            "*" ~ SITE_GEOGRAPHIC_FEATURE_SUB_vkat,

        # Sub-features are less important and not that frequently reported anyway.
        .default = "Not reported"
      )
    )

  # Count potential conflicts
  n_potential_feature <- df |>
    filter(
      !has_data_issue(SITE_GEOGRAPHIC_FEATURE_vkat),
      !has_data_issue(SITE_GEOGRAPHIC_FEATURE_medium),
      SITE_GEOGRAPHIC_FEATURE_vkat != SITE_GEOGRAPHIC_FEATURE_medium
    ) |>
    nrow()

  n_potential_sub <- df |>
    filter(
      !has_data_issue(SITE_GEOGRAPHIC_FEATURE_SUB_vkat),
      !has_data_issue(SITE_GEOGRAPHIC_FEATURE_SUB_medium),
      SITE_GEOGRAPHIC_FEATURE_SUB_vkat != SITE_GEOGRAPHIC_FEATURE_SUB_medium
    ) |>
    nrow()

  # Count unresolved
  n_unresolved <- df |>
    tally(SITE_GEOGRAPHIC_FEATURE_resolved == "FLAG: Geographic conflict.")
  n_unresolved_sub <- df |>
    tally(SITE_GEOGRAPHIC_FEATURE_SUB_resolved == "FLAG: Geographic conflict.")

  message(glue(
    "Geographic features: {n_potential_feature - n_unresolved} resolved, {n_unresolved} unresolved (of {n_potential_feature} conflicts)"
  ))
  message(glue(
    "Geographic sub-features: {n_potential_sub - n_unresolved_sub} resolved, {n_unresolved_sub} unresolved (of {n_potential_sub} conflicts)"
  ))

  if (n_unresolved + n_unresolved_sub > 0) {
    warning("Some geographic features still aren't resolved.")
  }

  df
}


# Attempt to resolve geographic conflicts, flag cases where this isn't possible
vm_compartment_geo_conflicts_resolved <- resolve_geographic_conflicts(
  vm_compartment_conflicts_resolved_removed
)

vm_compartment_geo_conflicts_resolved |>
  group_by(
    SITE_GEOGRAPHIC_FEATURE_resolved,
    SITE_GEOGRAPHIC_FEATURE_SUB_resolved
  ) |>
  reframe(n = n()) |>
  arrange(desc(n))

# what's going on with biota?
vm_compartment_geo_conflicts_resolved |>
  filter(ENVIRON_COMPARTMENT_resolved == "Biota") |>
  group_by(
    SITE_GEOGRAPHIC_FEATURE_resolved,
    SITE_GEOGRAPHIC_FEATURE_SUB_resolved
  ) |>
  reframe(n = n()) |>
  arrange(desc(n))
# All biota apparently have a conflict...
# # Not anymore!

vm_compartment_geo_conflicts_resolved |>
  filter(ENVIRON_COMPARTMENT_resolved == "Biota") |>
  group_by(
    ENVIRON_COMPARTMENT_resolved,
    ENVIRON_COMPARTMENT_SUB_resolved,
    SITE_GEOGRAPHIC_FEATURE_vkat,
    SITE_GEOGRAPHIC_FEATURE_medium,
    SITE_GEOGRAPHIC_FEATURE_SUB_vkat,
    SITE_GEOGRAPHIC_FEATURE_SUB_medium
  ) |>
  reframe(n = n()) |>
  arrange(desc(n))

# then filter out flagged cases
vm_compartment_geo_conflicts_resolved_removed <- vm_compartment_geo_conflicts_resolved |>
  filter(
    SITE_GEOGRAPHIC_FEATURE_resolved != "FLAG: Geographic conflict.",
    SITE_GEOGRAPHIC_FEATURE_SUB_resolved != "FLAG: Geographic conflict."
  )

message(glue(
  "Removed {nrow(vm_geo_conflicts_flagged) - nrow(vm_compartment_geo_conflicts_resolved_removed)} rows with unresolved geographic conflicts"
))

message(glue(
  "vm_geo_conflicts_flagged ({nrow(vm_geo_conflicts_flagged)} x {ncol(vm_geo_conflicts_flagged)}) -> vm_compartment_geo_conflicts_removed ({nrow(vm_compartment_geo_conflicts_resolved_removed)} x {ncol(vm_compartment_geo_conflicts_resolved_removed)})"
))

vm_compartment_geo_conflicts_resolved_removed |>
  # inspect data
  group_by(
    ENVIRON_COMPARTMENT_resolved,
    ENVIRON_COMPARTMENT_SUB_resolved,
    SITE_GEOGRAPHIC_FEATURE_resolved,
    SITE_GEOGRAPHIC_FEATURE_SUB_resolved
  ) |>
  reframe(Rows = n(), Mean = mean(Verdi)) |>
  # does it have biota in it? it should.
  col_vals_make_set(
    ENVIRON_COMPARTMENT_resolved,
    c("Biota", "Aquatic"),
    label = "Check vm_compartment_geo_conflicts_resolved_removed has Biota & Aquatic"
  )

```

We also have the issue that our higher level of site resolution means we need to split some of Vannmiljø's sites. Specifically, they may sample from the same longitude and latitude (one vannmiljø site) but sample different parts of the feature (e.g. sediment and the water column; two vannmiljø media/matrices). Because in our format this is two separate sites, we need to split any sites where this happens.

We do this in a fairly basic way - we take the existing `Vannlok_kode` (a string of digits, e.g. `02.61-29530) and, if we split it, append an incrementing number on the end (`02.61-29530` would become `02.61-29530-01` and `02.61-29530-02`). 

Unfortunately this is quite slow. We 

```{r}
#| label: split-sites

# split sites where different types of samples have been taken
vm_compartment_sites_split <- vm_compartment_geo_conflicts_resolved_removed |>
  group_by(Vannlok_kode) |>
  mutate(
    n_geo_combos = n_distinct(paste(
      SITE_GEOGRAPHIC_FEATURE_resolved,
      SITE_GEOGRAPHIC_FEATURE_SUB_resolved
    )),
    geo_combo = paste(
      SITE_GEOGRAPHIC_FEATURE_resolved,
      SITE_GEOGRAPHIC_FEATURE_SUB_resolved
    ),
    geo_suffix = if_else(
      n_geo_combos > 1,
      match(geo_combo, unique(geo_combo)),
      NA_integer_
    ),
    Vannlok_kode_split = case_when(
      !is.na(geo_suffix) ~ paste0(
        Vannlok_kode,
        "-",
        sprintf("%02d", geo_suffix)
      ),
      TRUE ~ Vannlok_kode
    )
  ) |>
  ungroup()

# Report on splitting
split_summary <- vm_compartment_sites_split |>
  distinct(Vannlok_kode, n_geo_combos) |>
  count(n_geo_combos, name = "n_sites") |>
  mutate(
    total_split_sites = n_geo_combos * n_sites,
    Status = if_else(n_geo_combos == 1, "Unsplit", "Split")
  )

n_original_sites <- n_distinct(vm_compartment_sites_split$Vannlok_kode)
n_split_sites <- n_distinct(vm_compartment_sites_split$Vannlok_kode_split)
n_affected <- split_summary |>
  filter(n_geo_combos > 1) |>
  pull(n_sites) |>
  sum()

message(glue(
  "Site splitting summary:\n",
  "  Original sites: {n_original_sites}\n",
  "  Sites requiring splitting: {n_affected}\n",
  "  Final unique site codes: {n_split_sites}"
))

# Clean up temporary columns
vm_compartment_sites_split <- vm_compartment_sites_split |>
  select(-n_geo_combos, -geo_combo, -geo_suffix)

message(glue(
  "vm_compartment_geo_conflicts_resolved_removed ({nrow(vm_compartment_geo_conflicts_resolved_removed)} x {ncol(vm_compartment_geo_conflicts_resolved_removed)}) -> vm_compartment_sites_split ({nrow(vm_compartment_sites_split)} x {ncol(vm_compartment_sites_split)})"
))

vm_compartment_sites_split |>
  # then check that most of what was there before is still there
  group_by(
    ENVIRON_COMPARTMENT_resolved,
    ENVIRON_COMPARTMENT_SUB_resolved,
    SITE_GEOGRAPHIC_FEATURE_resolved,
    SITE_GEOGRAPHIC_FEATURE_SUB_resolved
  ) |>
  reframe(Rows = n(), Mean = mean(Verdi)) |>
  # does it have biota in it? it should.
  col_vals_make_set(ENVIRON_COMPARTMENT_resolved, c("Biota", "Aquatic"))

```

Let's update the not-Sankey now that we've removed some more rows (@fig-conflict-resolution-visualisation).

```{r}
#| label: fig-conflict-resolution-visualisation
#| fig-cap: Vannmiljø copper sample size at different steps of filtration and conflict removal.
#| fig-width: 12

# Create filtering summary
conflict_flow <- filter_flow |>
  add_row(
    Filter = "Compartment Resolution",
    Rows_Remaining = nrow(vm_compartment_conflicts_resolved),
    pct = round(100 * Rows_Remaining / first(filter_flow$Rows_Remaining), 1),
    label = glue("{scales::comma(Rows_Remaining)}\n({pct}%)")
  ) |>
  add_row(
    Filter = "Geographic Conflicts Removed",
    Rows_Remaining = nrow(vm_compartment_sites_split),
    pct = round(100 * Rows_Remaining / first(filter_flow$Rows_Remaining), 1),
    label = glue("{scales::comma(Rows_Remaining)}\n({pct}%)")
  )

# Simple bar plot
ggplot(conflict_flow, aes(x = Filter, y = Rows_Remaining)) +
  geom_col(aes(fill = Filter)) +
  geom_text(aes(label = label), vjust = -0.5, size = 3) +
  scale_fill_brewer(palette = "Set1") +
  scale_x_discrete(limits = conflict_flow$Filter) +
  scale_y_continuous(
    labels = scales::comma,
    expand = expansion(mult = c(0, 0.1))
  ) +
  labs(
    x = NULL,
    y = "Number of Records"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

Finally, we need to run `pointblank` validation on our resolved/removed geographical feature data.

```{r}
#| label: conflict-resolution-pointblank
# first of all, does our table have data in it?
if (nrow(vm_compartment_sites_split) * ncol(vm_compartment_sites_split) != 0) {
  vm_compartment_sites_split |>
    # is there a real dataset here?
    # check that controlled vocab is respected
    col_vals_in_set(
      ENVIRON_COMPARTMENT_resolved,
      eDataDRF::environ_compartments_vocabulary(),
      label = "All ENVIRON_COMPARTMENT_resolved are in eDataDRF::environ_compartments_vocabulary()",
      actions = validation_config$standard
    ) |>
    col_vals_in_set(
      ENVIRON_COMPARTMENT_SUB_resolved,
      purrr::flatten(eDataDRF::environ_compartments_sub_vocabulary()),
      label = "All ENVIRON_COMPARTMENT_SUB_resolved are in eDataDRF::environ_compartments_sub_vocabulary()",
      actions = validation_config$standard
    ) |>
    col_vals_in_set(
      SITE_GEOGRAPHIC_FEATURE_resolved,
      eDataDRF::geographic_features_vocabulary(),
      label = "All SITE_GEOGRAPHIC_FEATURE_resolved are in eDataDRF::geographic_features_vocabulary()",
      actions = validation_config$standard
    ) |>
    col_vals_in_set(
      SITE_GEOGRAPHIC_FEATURE_SUB_resolved,
      eDataDRF::geographic_features_sub_vocabulary(),
      label = "All SITE_GEOGRAPHIC_FEATURE_SUB_resolved are in eDataDRF::geographic_features_sub_vocabulary()",
      actions = validation_config$standard
    ) |>
    # then check that most of what was there before is still there
    # e.g. in a prior version we accidentally filtered out all biota compartments
    group_by(ENVIRON_COMPARTMENT_resolved, ENVIRON_COMPARTMENT_SUB_resolved) |>
    reframe(n())
} else {
  stop("Error: vm_compartment_sites_split does not contain data.")
}

```

  # Create eData Tables
  
  ## Campaign
  
```{r}
#| label: edata-generate_campaign_table

edata_campaign <- initialise_campaign_tibble() |>
  add_row(
    CAMPAIGN_NAME_SHORT = config$metadata$campaign_short,
    CAMPAIGN_NAME = config$metadata$campaign_name,
    CAMPAIGN_START_DATE = config$filters$date_start,
    CAMPAIGN_END_DATE = config$filters$date_end,
    RELIABILITY_SCORE = NA_character_,
    RELIABILITY_EVAL_SYS = NA_character_,
    CONFIDENTIALITY_EXPIRY_DATE = as.Date(NA),
    ORGANISATION = config$metadata$organisation,
    ENTERED_BY = config$metadata$entered_by,
    ENTERED_DATE = as.Date(Sys.Date()),
    CAMPAIGN_COMMENT = "Copper and copper pyrithione measurements from Vannmiljø database covering all Norwegian municipalities and media types"
  )

message("Created campaign table")
```

## Reference

```{r}
#| label: edata-reference

edata_reference <- initialise_references_tibble() |>
  add_row(
    REFERENCE_ID = config$metadata$reference_id,
    REFERENCE_TYPE = "Database",
    DATA_SOURCE = "Vannmiljø",
    AUTHOR = config$metadata$organisation,
    TITLE = "Vannmiljø Database - Copper and Copper Pyrithione Data",
    YEAR = 2025L,
    ACCESS_DATE = config$filters$date_end,
    PERIODICAL_JOURNAL = NA_character_,
    VOLUME = NA_integer_,
    ISSUE = NA_integer_,
    PUBLISHER = NA_character_,
    INSTITUTION = config$metadata$organisation,
    DOI = NA_character_,
    URL = "https://www.vannmiljo.no/",
    ISBN_ISSN = NA_character_,
    EDITION = NA_character_,
    DOCUMENT_NUMBER = NA_character_,
    REF_COMMENT = glue(
      "Downloaded {config$filters$date_end}, data for {config$filters$date_start} to {config$filters$date_end}, ",
      "all kommune, all media, all campaigns. Search for Kobber & Kobberpyrition. 138615 hits."
    )
  )

message("Created reference table")
```

## Parameters

```{r}
#| label: edata-parameters

# No copper pyrithione reported for 2025, only copper
edata_parameters <- initialise_parameters_tibble() |>
  add_row(
    PARAMETER_TYPE = NA_character_,
    PARAMETER_TYPE_SUB = NA_character_,
    MEASURED_TYPE = NA_character_,
    PARAMETER_NAME = "Copper",
    PARAMETER_NAME_SUB = NA_character_,
    INCHIKEY_SD = NA_character_,
    PUBCHEM_CID = NA_integer_,
    CAS_RN = NA_character_,
    ENTERED_BY = glue("{config$metadata$entered_by} from Vannmiljø"),
    PARAMETER_COMMENT = NA_character_
  )

message("Created parameters table")
```

## Sites

```{r}
#| label: edata-sites-extraction

# Extract unique sites with relevant metadata
vm_sites_unique <- vm_compartment_sites_split |>
  select(
    Vannlok_kode_split,
    Vannlokalitetsnavn,
    Beskrivelse,
    `UTM33 Ost (X)`,
    `UTM33 Nord (Y)`,
    `Knytt til påvirkning`,
    SITE_GEOGRAPHIC_FEATURE_resolved,
    SITE_GEOGRAPHIC_FEATURE_SUB_resolved
  ) |>
  distinct() |>
  # Clean up emission source
  mutate(
    Emission_Source = case_when(
      `Knytt til påvirkning` %in% c("Industri", "INDUSTRI") ~ "Industrial",
      `Knytt til påvirkning` %in% c("Akvakultur", "AKVAKULTUR") ~ "Aquaculture",
      TRUE ~ NA_character_
    )
  )

message(glue("Extracted {nrow(vm_sites_unique)} unique sites"))

# Format to eData structure
edata_sites_temp <- vm_sites_unique |>
  mutate(
    SITE_CODE = glue("Vannmiljø_{Vannlok_kode_split}"),
    SITE_NAME = glue("Vannmiljø Station {Vannlokalitetsnavn}"),
    SITE_GEOGRAPHIC_FEATURE = SITE_GEOGRAPHIC_FEATURE_resolved,
    SITE_GEOGRAPHIC_FEATURE_SUB = SITE_GEOGRAPHIC_FEATURE_SUB_resolved,
    COUNTRY_ISO = "Norway",
    OCEAN_IHO = "Not relevant",
    ENTERED_BY = glue("{config$metadata$entered_by} (Vm Conversion)"),
    ENTERED_DATE = as.character(today()),
    ALTITUDE_VALUE = 0, # Altitude not relevant for water monitoring
    ALTITUDE_UNIT = "m",
    # Combine description and emission source where available
    SITE_COMMENT = case_when(
      !is.na(Beskrivelse) &
        Beskrivelse != "" &
        !is.na(Emission_Source) &
        Emission_Source != "" ~
        glue(
          "Vm Original Comment: {Beskrivelse}. Vm Emission Source: {Emission_Source}"
        ),
      !is.na(Beskrivelse) & Beskrivelse != "" ~
        glue("Vm Original Comment: {Beskrivelse}"),
      !is.na(Emission_Source) & Emission_Source != "" ~
        glue("Vm Emission Source: {Emission_Source}"),
      .default = NA_character_
    )
  )

# Reproject coordinates from UTM33 to WGS84
tryCatch(
  {
    sites_sf <- edata_sites_temp |>
      st_as_sf(coords = c("UTM33 Ost (X)", "UTM33 Nord (Y)"), crs = 25833) |>
      st_transform(4326) |>
      mutate(
        LATITUDE = st_coordinates(geometry)[, 2],
        LONGITUDE = st_coordinates(geometry)[, 1],
        SITE_COORDINATE_SYSTEM = "WGS84"
      )

    sites_coords <- sites_sf |>
      st_coordinates() |>
      as.data.frame() |>
      bind_cols(st_drop_geometry(sites_sf))

    message(glue(
      "Successfully reprojected {nrow(sites_sf)} sites from UTM33 to WGS84"
    ))
  },
  error = function(e) {
    stop(glue(
      "Failed to reproject coordinates from UTM33 to WGS84: {e$message}"
    ))
  }
)

```

```{r}
#| label: edata-create-sites-table

# Finalise sites table
tryCatch(
  {
    # Check for duplicates before joining
    sites_for_join <- sites_sf |>
      st_drop_geometry() |>
      select(SITE_CODE, LATITUDE, LONGITUDE, SITE_COORDINATE_SYSTEM) |>
      distinct()

    duplicate_sites <- sites_for_join |>
      group_by(SITE_CODE) |>
      filter(n() > 1) |>
      arrange(SITE_CODE)

    if (nrow(duplicate_sites) > 0) {
      warning(glue(
        "Found {nrow(duplicate_sites)} duplicate SITE_CODE entries in sites_sf. ",
        "Displaying first 10 conflicts:"
      ))
      print(head(duplicate_sites, 10))
    }

    edata_sites <- edata_sites_temp |>
      left_join(
        sites_for_join,
        by = "SITE_CODE",
        relationship = "many-to-many" # Temporarily allow to see what happens
      ) |>
      select(
        SITE_CODE,
        SITE_NAME,
        SITE_GEOGRAPHIC_FEATURE,
        SITE_GEOGRAPHIC_FEATURE_SUB,
        LATITUDE,
        LONGITUDE,
        SITE_COORDINATE_SYSTEM,
        COUNTRY_ISO,
        OCEAN_IHO,
        ENTERED_BY,
        ENTERED_DATE,
        SITE_COMMENT
      )

    # Check result for duplicates
    result_duplicates <- edata_sites |>
      distinct() |>
      group_by(SITE_CODE) |>
      filter(n() > 1)

    if (nrow(result_duplicates) > 0) {
      stop(glue(
        "Many-to-many join created {nrow(result_duplicates)} duplicate rows. ",
        "Check vm_sites_unique for duplicate coordinates per site."
      ))
    }
  },
  warning = function(w) {
    stop(glue(
      "Unexpected many-to-many relationship when joining sites table:\n{w}"
    ))
  }
)

# Validate against eData schema
edata_sites <- initialise_sites_tibble() |> add_row(edata_sites)

message(glue("Created sites table: {nrow(edata_sites)} sites"))
```

Visually inspecting transformed site distribution (@fig-sites-sanity-map) allows us to check that a) sites are roughly where they should be and b) the correct CRS transformation was used.

```{r}
#| label: fig-sites-sanity-map
#| fig-cap: Vannmiljø sampling sites binned into 1000 hexagons, coloured by number of sites in area.
#| fig-width: 12

# Sanity check: plot sites on map
world_map <- map_data("world")
ggplot() +
  geom_polygon(
    data = world_map,
    aes(x = long, y = lat, group = group),
    fill = "lightgray",
    colour = "white"
  ) +
  geom_hex(
    data = sites_coords,
    aes(x = X, y = Y),
    bins = 1000
  ) +
  scale_fill_viridis_c() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() +
  theme(axis.text = element_blank(), axis.title = element_blank())

```

## Methods

Because Vannmiljø publish their methods data in a format similar to ours, the majority of processing has already been done. We did the rest of the processing in Excel when assemling the lookup table, so no additional work is required here. However, this will change if we expand the filter criteria to include more compartments or older samples, and the lookup will need to be updated manually. 

```{r}
#| label: edata-methods

# Merge and categorize
edata_methods <- vm_lookup_methods

message(glue("Created methods table: {nrow(edata_methods)} protocols"))
```

## Samples

In the eData workflow, we generate a samples table by combining sites, parameters, dates and compartments. This isn't how we work with Vannmiljø, but it's useful to have it anyway because we need it to generate the biota table - which stores species, tissue, etc. information.

```{r}
#| label: edata-samples

# this function is copied from STOPeData::mod_samples_fct.R
# I guess we forgot to export it. Good candidate to be moved to eDataDRF.
# TODO: It's also quite slow, so we could probably make it better.
#' Generate Sample ID with Components ----
#' @param site_code Site code (vectorized)
#' @param parameter_name Parameter name (vectorized)
#' @param environ_compartment Environmental compartment (vectorized)
#' @param environ_compartment_sub Environmental sub-compartment (vectorized)
#' @param date Sampling date (vectorized)
#' @param subsample subsample
#' @importFrom stringr str_to_title str_remove_all
#' @import eDataDRF
#' @noRd
generate_sample_id_with_components <- function(
  site_code,
  parameter_name,
  environ_compartment,
  environ_compartment_sub,
  date,
  subsample = 1
) {
  # Create abbreviated versions for ID (vectorized)
  param_abbrev <- substr(gsub("[^A-Za-z0-9]", "", parameter_name), 1, 8)
  comp_abbrev <- substr(
    gsub("[^A-Za-z0-9]", "", environ_compartment_sub),
    1,
    12
  )
  date_abbrev <- gsub("-", "-", date)

  base_id <- glue("{site_code}-{param_abbrev}-{comp_abbrev}-{date_abbrev}")

  # Vectorized replicate
  # Subsamples will generally be text, so let's abbreviate them a bit

  # this is modified from the eDataDRF version and hopefully faster
  subsample_suffix <- stringr::str_trunc(subsample, 20, "right", ellipsis = "")
  paste0(base_id, "-R-", subsample_suffix)
}


# Create samples table

edata_samples_wide <- vm_compartment_sites_split |>
  filter(
    ENVIRON_COMPARTMENT_resolved != "FLAG: Compartment conflict.",
    ENVIRON_COMPARTMENT_SUB_resolved != "FLAG: Compartment conflict."
  ) |>
  mutate(
    # Core identifiers
    SITE_CODE = glue("Vannmiljø_{Vannlok_kode}"),
    SITE_NAME = glue("Vannmiljø Station {Vannlokalitetsnavn}"),
    PARAMETER_NAME = "Copper",
    PARAMETER_TYPE = "Not reported",

    # Compartment information
    ENVIRON_COMPARTMENT = ENVIRON_COMPARTMENT_resolved,
    ENVIRON_COMPARTMENT_SUB = ENVIRON_COMPARTMENT_SUB_resolved,
    MEASURED_CATEGORY = NA_character_,

    SAMPLING_DATE = as.character(SAMPLING_DATE),

    # Sample information
    SUBSAMPLE = "NA",
    SAMPLE_ID = generate_sample_id_with_components(
      SITE_CODE,
      PARAMETER_NAME,
      ENVIRON_COMPARTMENT,
      ENVIRON_COMPARTMENT_SUB,
      SAMPLING_DATE,
      SUBSAMPLE
    )
  )

# Validate against eData schema
edata_samples <- initialise_samples_tibble() |>
  add_row(
    edata_samples_wide |>
      select(any_of(names(initialise_samples_tibble())))
  )


message(glue(
  "Created samples table: {nrow(edata_samples)} samples"
))
```

## Biota

Biota stuff. Some genders are implicit from the tissue type reported (e.g. gonads), but I don't think I want to worry about that here. Our assumptions:

- All species reported in Vannmiljø are correct
- No attempt was made to infer gender or lifestage

Nevertheless, we need to do a bit of extra work with scientific names, species groups, and tissue before proceeding. Currently:

- Laksesmolt needs to be recoded as Juvenile Salmo salmar

That's it.

```{r}
#| label: tbl-raw-biota

# Create biota table

edata_biota <- vm_compartment_sites_split |>
  filter(ENVIRON_COMPARTMENT_resolved == "Biota")

edata_species <- unique(edata_biota$VitenskapligNavn) |> length()
message(glue(
  "{nrow(edata_biota)} biota samples found ({edata_species} unique species)"
))

biota_frequency <- count(
  edata_biota,
  VitenskapligNavn,
  ENVIRON_COMPARTMENT_SUB_resolved
) |>
  arrange(desc(n)) |>
  mutate(
    SAMPLE_SPECIES = case_when(
      VitenskapligNavn == "Laksesmolt" ~ "Salmo salar",
      TRUE ~ VitenskapligNavn
    ),
    SAMPLE_SPECIES_LIFESTAGE = case_when(
      VitenskapligNavn == "Laksesmolt" ~ "Juvenile",
      TRUE ~ "Not reported"
    )
  )

# Report data quality issues
n_missing_species_group <- biota_frequency |>
  filter(is.na(SAMPLE_SPECIES)) |>
  pull(n) |>
  sum()

n_missing_sub_compartment <- biota_frequency |>
  filter(ENVIRON_COMPARTMENT_SUB_resolved == "*") |>
  pull(n) |>
  sum()


message(glue(
  "Biota data quality:,
  - {n_missing_species_group} samples with missing species group,
  - {n_missing_sub_compartment} samples with missing subcompartment"
))


datatable(biota_frequency)

# there's certainly a more elegant way to do this than twice
edata_biota_merged <- edata_biota |>
  mutate(
    SAMPLE_SPECIES = case_when(
      VitenskapligNavn == "Laksesmolt" ~ "Salmo salar",
      TRUE ~ VitenskapligNavn
    ),
    SAMPLE_SPECIES_LIFESTAGE = case_when(
      VitenskapligNavn == "Laksesmolt" ~ "Juvenile",
      TRUE ~ "Not reported"
    ),
    # Let's guess compartment from species
    # all the species not sampled from something with a Vannkategori are terrestrial, so this isn't complicated for the
    # current dataset
    ENVIRON_COMPARTMENT_SUB_resolved = case_when(
      ENVIRON_COMPARTMENT_SUB_resolved !=
        "*" ~ ENVIRON_COMPARTMENT_SUB_resolved,
      SAMPLE_SPECIES %in%
        c(
          "Capreolus capreolus",
          "Felis catus",
          "Hylocomium splendens",
          "Larus argentatus",
          "Lichen",
          "Phoca vitulina",
          "Rattus norvegicus",
          "Somateria mollissima",
          "Turdus pilaris",
          "Vulpes vulpes"
        ) ~ "Biota, terrestrial",
    )
  )
```

We also need to convert tissue types (stored as MediumID in Vannmiljø) into our appropriate equivalent (@tbl-raw-tissue).

```{r}
#| label: tbl-raw-tissue
#| tbl-caption: Table of Vannmiljø MediumIDs corresponding to biota tissues, with sample size (n) and corresponding eData SAMPLE_TISSUE value.

# Create tissue table
# TODO: Should become a junction table (?) in the fullness of time
map_tissue_type <- function(medium_id_name) {
  case_match(
    medium_id_name,
    "Biota bløtdeler" ~ "Total soft tissues",
    "Biota gjeller" ~ "Gills",
    "Biota helkropp" ~ "Whole body",
    "Biota lever" ~ "Liver",
    "Biota muskelvev" ~ "Muscle tissue",
    "Biota plantevev" ~ "Plant tissue",
    "Biota egg" ~ "Egg",
    "Biota blod" ~ "Blood",
    "Biota skuddspiss" ~ "Shoot tip",
    "Biota fettvev" ~ "Adipose tissue",
    "Biota galle" ~ "Bile",
    .default = "Unknown Tissue"
  )
}


# TODO: Update to use geo conflicts resolved thingy as appropriate
tissue_frequency <- count(edata_biota, MediumID_Name) |>
  arrange(desc(n)) |>
  mutate(
    SAMPLE_TISSUE = map_tissue_type(MediumID_Name)
  )

# Report data quality issues
n_unknown_tissue <- tissue_frequency |>
  filter(SAMPLE_TISSUE == "Unknown Tissue") |>
  pull(n) |>
  sum()

message(glue(
  "Biota data quality: {n_unknown_tissue} samples with unknown tissue type"
))

datatable(tissue_frequency)

edata_biota_tissue_merged <- edata_biota_merged |>
  mutate(
    SAMPLE_TISSUE = map_tissue_type(MediumID_Name)
  )

```


```{r}
#| label: eData-biota

vm_samples_biota_only <- edata_biota_tissue_merged |>
  mutate(
    # Core identifiers
    SITE_CODE = glue("Vannmiljø_{Vannlok_kode}"),
    # SITE_NAME = glue("Vannmiljø Station {Vannlokalitetsnavn}"), added later
    PARAMETER_NAME = "Copper",
    # PARAMETER_TYPE = "Not reported", added later

    # Compartment information
    ENVIRON_COMPARTMENT = ENVIRON_COMPARTMENT_resolved,
    ENVIRON_COMPARTMENT_SUB = ENVIRON_COMPARTMENT_SUB_resolved,
    MEASURED_CATEGORY = NA_character_,

    SAMPLING_DATE = as.character(SAMPLING_DATE),

    # Sample information
    SUBSAMPLE = "NA",
    SAMPLE_ID = generate_sample_id_with_components(
      SITE_CODE,
      PARAMETER_NAME,
      ENVIRON_COMPARTMENT,
      ENVIRON_COMPARTMENT_SUB,
      SAMPLING_DATE,
      SUBSAMPLE
    ),

    # biota stuff
    SPECIES_GROUP = species_group,
    SAMPLE_SPECIES,
    SAMPLE_TISSUE,
    SAMPLE_SPECIES_LIFESTAGE,
    SAMPLE_SPECIES_GENDER = "Not reported",
    BIOTA_COMMENT = "Raw species names from Vannmiljø. May be erroneous.",
  )

# Validate against eData schema
edata_biota <- initialise_biota_tibble() |>
  add_row(
    vm_samples_biota_only |>
      select(any_of(names(initialise_biota_tibble())))
  )

message(glue(
  "Created biota table: {nrow(edata_biota)} rows"
))
```

Once we've handled biota separately we can merge them with the rest of the samples to create the intermediate table used to create the measurements table. This is a slightly strange way of doing things, but it works.

```{r}
#| label: merge-biota-samples

biota_samples_merged <- edata_samples_wide |> bind_rows(vm_samples_biota_only)

```

## Measurements

```{r}
#| label: edata-measurements

# Create measurements table

# TODO: Update to use geo conflicts resolved thingy as appropriate
edata_measurements <- biota_samples_merged |>
  mutate(
    # Core identifiers
    SITE_CODE,
    PARAMETER_NAME,
    SAMPLING_DATE = as.character(SAMPLING_DATE),
    CAMPAIGN_NAME_SHORT = config$metadata$campaign_short,
    REFERENCE_ID = config$metadata$reference_id,

    # Compartment information
    ENVIRON_COMPARTMENT,
    ENVIRON_COMPARTMENT_SUB,

    # Parameter classification
    PARAMETER_TYPE = "Stressor",
    MEASURED_TYPE = "Concentration",

    # Sample information
    SUBSAMPLE,
    SAMPLE_ID,

    # Measurement values
    MEASURED_FLAG = vm_convert_operator(Operator),
    MEASURED_VALUE = Verdi,
    MEASURED_UNIT = vm_convert_unit(Unit_Name),
    MEASURED_N = Ant_verdier,

    # Uncertainty (not reported in Vannmiljø)
    UNCERTAINTY_TYPE = "Not reported",
    UNCERTAINTY_UPPER = NA_real_,
    UNCERTAINTY_LOWER = NA_real_,

    # Detection limits
    LOQ_VALUE = Kvantifiseringsgrense,
    LOQ_UNIT = vm_convert_unit(Unit_Name),
    LOD_VALUE = Deteksjonsgrense,
    LOD_UNIT = vm_convert_unit(Unit_Name),

    # TODO: Fix methods reference
    # Protocol references (FIXME: Need proper protocol ID mapping)
    SAMPLING_PROTOCOL = "1",
    EXTRACTION_PROTOCOL = "2",
    FRACTIONATION_PROTOCOL = "3",
    ANALYTICAL_PROTOCOL = "4",

    # Comments
    MEASUREMENT_COMMENT = Kommentar,

    .keep = "none"
  )

# Validate against eData schema
edata_measurements <- initialise_measurements_tibble() |>
  add_row(edata_measurements)

message(glue(
  "Created measurements table: {nrow(edata_measurements)} measurements"
))
```

# Export

## Create Export

We export the resulting data to a zip file in our `data/raw/eData` folder where we keep the literature extractions.

```{r}
#| label: export-create-zip

# Assemble datasets for export
export_datasets <- list(
  edata_sites = edata_sites,
  edata_campaign = edata_campaign,
  edata_reference = edata_reference,
  edata_parameters = edata_parameters,
  edata_samples = edata_samples,
  edata_biota = edata_biota,
  edata_measurements = edata_measurements
)

# Export as ZIP
if (
  !file.exists(here(
    "data/raw/eData",
    paste0(config$metadata$campaign_short, ".zip")
  ))
) {
  timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")

  export_path <- export_campaign_zip(
    dataset_list = export_datasets,
    campaign_name = config$metadata$campaign_short,
    output_path = here(
      "data",
      "raw",
      "eData",
      glue("{config$metadata$campaign_short}_{timestamp}.zip")
    )
  )
  message(glue("Export complete: {export_path}"))
} else {
  message("Export file already exists, skipping")
}

```

## Validate Export

Do some quick checks to see if the data we're exporting looks like we expect it to. In an ideal world we would be able to automatically check and validate against constrolled vocabulary, but we're not quite there yet

### Compartments

Are our compartments valid and reasonable? @tbl-validate-export-compartments suggests they are not.

- Some Biota have * as a subcategory
- Some biota have Freshwater or Marine/Salt Water as a subcategory

Why didn't we flag these earlier?

```{r}
#| label: tbl-validate-export-compartments
count(edata_measurements, ENVIRON_COMPARTMENT, ENVIRON_COMPARTMENT_SUB) |>
  datatable()

```

@fig-vm-scatterplot is a basic plot of measured concentrations by year and unit. Without going into detailed analysis of campaigns, etc. the results are a little difficult to interprety, but it's reasonable to call some of the shapes we see down there suspicious.

```{r}
#| label: fig-vm-scatterplot
#| fig-cap: A basic scatterplot for visualising measured concentrations by year.
#| fig-width: 12

edata_measurements |>
  ggplot(
    aes(
      y = MEASURED_VALUE,
      x = SAMPLING_DATE |> as.Date(),
      colour = ENVIRON_COMPARTMENT_SUB
    )
  ) +
  geom_point(shape = 1) +
  facet_wrap(
    facets = vars(ENVIRON_COMPARTMENT_SUB),
    nrow = 3,
    scales = "free_y"
  ) +
  labs(x = "Year", y = "Measured Value")
```

We can pull out this data and take a closer look. We'll use the z-score method and assume anything more than 3 SD from the mean is worth investigation.

- [ ] Todo: find a citation for use of z-scores
- [ ] Implement using pointblank

```{r}
#| label: tbl-vm-outliers

edata_measurements |>
  group_by(MEASURED_UNIT, ENVIRON_COMPARTMENT_SUB) |>
  mutate(z_score = scale(MEASURED_VALUE)) |> # Calculate z-score
  filter(z_score >= 3)
```
