---
title: "Notebook 02 - Vannmiljo Conversion"
author: "Sam Welch"
date: last-modified
html:
  code-overflow: wrap
---

This notebook summarises and reports on the process of importing, cleaning, and reformatting data from the Norwegian Environmental Ministry's [Vannmiljø ](https://vannmiljo.miljodirektoratet.no/)database.

Vannmijlø ("Water environment", henceforth abbreviated to Vm) is the Norwegian Environmental Department's (*Miljødirektoratet*, MD)'s clearinghouse for data on the ecological and chemical quality of bodies monitored under the Norwegian implementation of the Water Framework Directive [@lovdataForskriftOmRammer2007]. In addition, it has become a repository for data on the monitoring of other compartments including air, soil, and manufactured products (toys, artificial turf, fabrics, etc.). Vm is a data rich and extremely valuable resource for understanding patterns of contamination across the country. In respect of the fact that Vm provides a rich data resource to any interested parties 24/7, it is an excellent database.

Efforts to modernise the database are currently underway at MD, but like many long-serving databases the quality, validation and extraction of the data product provided is not always up to modern best practice. In particular, documentation on the database is extremely limited, and no unique resource identifier is available. In its own words

"Vannmiljø is the environmental authorities' professional system for registering and analysing the conditions in water. Vannmiljø thus plays a central role in planning and executing all monitoring activities required by the Water Framework Directive. Data will, however, be available for all types of case handling where information about condition and development in water environmental quality is requested." (translated from the Norwegian).

-   Vm has a basic api: https://vannmiljoapi.miljodirektoratet.no/swagger/ui/index
-   INSPIRE: https://inspire-geoportal.ec.europa.eu/srv/api/records/98dd016f-b905-463b-a633-6be2a33fcec2
-   Codes: https://vannmiljokoder.miljodirektoratet.no/
-   Some background: https://www.vannportalen.no/verktoy/vannmiljo2/hva-er-vannmiljo/

# Data Extraction

## Database Summary

-   for our purposes, vannmiljø is a table of medium quality measurements (typically but not always water)
-   and a table of sites
-   linked by a foreign key sdite

## Search Terms

| Norwegian Term | English Translation | Selected Options | Comments |
|------------------|-------------------|------------------|------------------|
| Nedlastingsdato | Download Date | 2025-12-05 |  |
| Datoperiode | Date Period | 2025-01-01 to 2025-12-05 |  |
| Kommune | Municipality | All municipalities |  |
| Medium | Medium/Matrix | All media |  |
| Kampanje | Campaign | All campaigns |  |
| Stressorer | Stressors/Parameters | Kobber (Copper) & Kobberpyrition (Copper pyrithione) |  |

## Export

Data were extracted from Vannmiljø's site. The following options were selected:





# To Do:

-   [ ] Add probable contamination source
-   [ ] More sanity checking of extracted values
-   [ ] Work out how to do CREED for this data, then implement.

## Optional Improvements

-   [ ] Improve heatmaps using https://github.com/raivokolde/pheatmap
-   [ ] Make real Sankey diagrams

# Libraries & Setup

Import required packages and set working directory to project root.

```{r}
#| label: setup-libraries
#| warning: false
#| message: false

knitr::opts_chunk$set(dev = "svglite") # otherwise some SVGs cause problems in firefox?

suppressMessages
({
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(lubridate)
  library(readr)
  library(readxl)
  library(glue)
  library(ggplot2)
  library(sf)
  library(DT)
  library(summarytools)
  library(eDataDRF)
  library(here)
  library(knitr)
  library(devtools)
  library(tibble)
  library(pointblank)
  library(targets)
})

suppressMessages(
load_all(quiet = TRUE))

# aesthetics
theme_set(theme_minimal(base_size = 12))

i_am("README.md") |> suppressMessages()
```

## Configuration

See @tbl-vm_config-filters and @tbl-vm_config-metadata for a summary of the parameters of this extraction from Vannmiljø.

```{r}
#| label: setup-configuration
#| warning: false

# this needs to be rationalised at some point
vm_config <- list(
  # # File paths
  # files = list(
  #   # TODO: Make imports more consistent and timestamp everything
  #   vm_copper = here("data/raw/vannmiljo/Vm_Copper_2025.12.05.xlsx"),
  #   vm_sites_1 = here("data/raw/vannmiljo/Vm_Copper_Sites_2025.12.05-1.xlsx"),
  #   vm_sites_2 = here("data/raw/vannmiljo/Vm_Copper_Sites_2025.12.05-2.xlsx"),
  #   vm_sites_3 = here("data/raw/vannmiljo/Vm_Copper_Sites_2025.12.05-3.xlsx"),
  #   lookup_medium = here("data/clean/Vm_medium_lookup_matrix_filled.csv"),
  #   lookup_vannkategori = here("data/clean/vm_sites_codes_lookup.csv"),
  #   lookup_methods = here("data/clean/vm_methods_lookup_filled.csv"),
  #   methods_analysis = here(
  #     "data/raw/vannmiljo/Vannmiljø_Analysemetode_2025-12-15.xlsx"
  #   ),
  #   methods_sampling = here(
  #     "data/raw/vannmiljo/Vannmiljø_Prøvetakingsmetode_2025-12-15.xlsx"
  #   ),
  #   lookup_campaigns = here("data/clean/Vm_lookup_campaigns.csv"),
  #   lookup_units = here(
  #     "data/raw/vannmiljo/Vannmiljø_Enhet_2025-12-30.xlsx"
  #   ),
  #   lookup_species = here("data/clean/Vm_species_lookup.csv") # not actually used rn
  # ),

  # Filter criteria
  # * is used as a wildcard in compartment lookups
  filters = list(
    date_start = as.Date("2010-01-01"),
    date_end = as.Date("2025-12-05"),
    exclude_sites = "Svalbard, ENSB-Kilde 2",
    compartment = c("Aquatic", "Biota", "*"),
    compartment_sub = c(
      "Freshwater",
      "Aquatic Sediment",
      "Marine/Salt Water",
      "Brackish/Transitional Water",
      "Biota, Aquatic",
      "*"
    )
  ),

  # Campaign metadata
  metadata = list(
    campaign_name = "Vannmiljø Copper Monitoring 2010-2025",
    campaign_short = "Vm_2010_2025",
    reference_id = "VannmiljøCopper2010-2025",
    entered_by = "Sam Welch",
    organisation = "Miljødirektoratet"
  )
)


```

Load in raw vannmiljø data and lookups (which are a mixture of raw data from vannmiljø and data I've processed myself) from the `targets` workflow.

```{r}
#| label: load-raw-data

tar_load(c(
  vm_raw_copper,
  vm_raw_sites,
  vm_lookup_medium,
  vm_lookup_vannkategori,
  vm_lookup_analysis,
  vm_lookup_sampling,
  vm_lookup_methods,
  vm_lookup_campaigns,
  vm_lookup_units,
  vm_lookup_species,
  vm_join_sites_measurements,
  vm_compartment_geo_conflicts_resolved,
  vm_join_sites_measurements_lookup,
  vm_data_with_dates,
  vm_filtered_compartments,
  vm_filtered_sites,
  vm_filtered_dates,
  vm_compartment_conflicts_resolved,
  vm_compartment_conflicts_resolved_removed,
  vm_compartment_geo_conflicts_resolved_removed,
  vm_sites_split,
  vm_sites_split_clean, # Last Vannmiljø processing step
  vm_edata_campaign, # eData campaign metadata
  vm_edata_reference, # eData reference metadata
  vm_edata_parameters, # eData parameters (copper)
  vm_edata_sites, # eData sites with reprojected coordinates
  vm_edata_samples, # eData samples
  vm_edata_biota, # eData biota-specific information
  vm_edata_measurements # eData measurements with values and limits
))
```

```{r}
#| label: tbl-vm_config-filters

vm_config$filters |>
  enframe(name = "Parameter", value = "Value") |>
  kable()
```

```{r}
#| label: tbl-vm_config-metadata

vm_config$metadata |>
  enframe(name = "Parameter", value = "Value") |>
  kable()
```

## Helper Functions

These are also in fct_vm_processing.R but I include them here to remind myself I need a better documented solution.

```{r}
#| label: setup-helper-functions
# Convert Vannmiljø operator codes to eData flags
vm_convert_operator <- function(col) {
  case_match(
    col,
    "=" ~ "",
    "<" ~ "< LOQ",
    ">" ~ "What is going on here. Why higher than?",
    "ND" ~ "< LOD"
  )
}

vm_convert_unit <- function(col) {
  case_match(
    col,
    "µg/l" ~ "µg/L",
    "mg/kg t.v." ~ "mg/kg (dry)",
    "mg/kg v.v." ~ "mg/kg (wet)",
    .default = "UNIT CONVERSION ERROR"
  )
}
```

# Load Raw Data

## Measurements Data

Downloaded 2025.12.05, data for 2025.01.01 - 2025.12.05, all kommune, all media, all campaigns

Search for Kobber & Kobberpyrition: 138615 hits

NOTE: Data may extend beyond 2025 date range despite request parameters

`summarytools::dfSummary()` generates an initial summary of the dataset here. It's not perfect but it's one line of code, and generally very helpful.

::: {.callout-note collapse="true" title="Initial Dataset Summary"}
```{r}
#| label: Initial data inspection
print(
  dfSummary(
    vm_raw_copper,
    varnumbers = FALSE,
    valid.col = FALSE,
    graph.magnif = 0.82
  ),
  method = 'render'
)
```
:::

## Sites Data

Vannmiljø export limit requires splitting site codes into multiple files

Procedure:

1.  Extract unique site codes: `vm_raw_copper\$Vannlok_kode |> unique() |> writeClipboard()`
2.  Replace newlines with commas: Ctrl+H find "\n" replace with ", "
3.  Split into chunks (Vannmiljø fails silently above unknown threshold)
4.  Download as separate Excel files

## Lookup Tables

This part is very messy - we generated a bunch of lookup tables (semi-reproducibly) and filled them by hand/with Claude. I don't have a lot of confidence in any of them.

Some of the lookups are actually direct from Vannmiljø though - they're the xlsx ones. They're also bad, but for different reasons!

```{r}
#| label: lookups

lookups <- grepv("vm_lookup", ls())

message(glue(
  "Loaded {length(lookups)} lookup tables: {paste(lookups, collapse = ', ')}"
))

```

# Data Processing

## Join Datasets

We join sites and measurements by `"Vannlok_kode" = "Vannlokalitetskode"`, then our compartment/etc. lookups. We check for and remove:

-   Non-null values: Vannlokalitetsnavn, Medium_id, Registrerings_id,`UTM33 Ost (X)`, `UTM33 Nord (Y)`
-   Medium_id, Enhet_id, Aktivitet_id, Vannkategori, VitenskapligNavn not in our lookups

Note that we only use four lookups at this stage: `vm_lookup_medium`, `vm_lookup_vannkategori`, `vm_lookup_campaigns`, and `vm_lookup_units`.

```{r}
#| label: processing-merge-datasets
#| echo: false

message(glue(
  "Joined vm_raw_copper and vm_raw_sites by Vannlokalitetskode. Output: vm_join_sites_measurements ({nrow(vm_join_sites_measurements)} x {ncol(vm_join_sites_measurements)})"
))

message(
  glue(
    "Left joined vm_join_sites_measurements to lookup tables: vm_lookup_medium, vm_lookup_vannkategori, vm_lookup_campaigns, vm_lookup_units.
    vm_join_sites_measurements ({nrow(vm_join_sites_measurements)} x {ncol(vm_join_sites_measurements)}) 
    + vm_lookup_medium 
    + vm_lookup_vannkategori 
    + vm_lookup_campaigns 
    + vm_lookup_units 
    + vm_lookup_species
    ---> vm_join_sites_measurements_lookup ({nrow(vm_join_sites_measurements_lookup)} x {ncol(vm_join_sites_measurements_lookup)})"
  )
)
```

## Pre-Transformation Exploration

We don't yet have an oversight over the distribution of site types, biota, tissue, and media in the dataset. This will inform what and where we filter. Unfortunately not all Vannkategori codes have a definition assigned to them, so I've attempted to guess.

```{r}
#| label: fig-heatmap-campaign-vannkategory
#| fig-cap: "Distribution of Vannmiljø copper samples by monitoring campaign and reported water category, all years"
#| fig-width: 12
#| fig-height: 14
#| warning: false

wrap_x <- 30
wrap_y <- 50

# Create a summary of counts by the two variables
vm_exploration_campaign_vannkategory <- vm_join_sites_measurements_lookup |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN = case_when(
      is.na(Name_EN) ~ "Other",
      TRUE ~ Name_EN
    )
  ) |>
  group_by(CAMPAIGN_DESCRIPTION_EN, Vannkategori_Description) |>
  summarise(n_samples = n(), .groups = "drop")

# Calculate text color threshold
threshold_value <- mean(
  log10(vm_exploration_campaign_vannkategory$n_samples),
  na.rm = TRUE
)

# Calculate total samples per campaign for ordering (y-axis)
campaign_order <- vm_exploration_campaign_vannkategory |>
  group_by(CAMPAIGN_DESCRIPTION_EN) |>
  summarise(total = sum(n_samples)) |>
  arrange(total) |> # puts highest values at top
  pull(CAMPAIGN_DESCRIPTION_EN)

# Calculate total samples per water category for ordering (x-axis)
vannkategori_order <- vm_exploration_campaign_vannkategory |>
  group_by(Vannkategori_Description) |>
  summarise(total = sum(n_samples)) |>
  arrange(desc(total)) |> # puts highest values on left
  pull(Vannkategori_Description)

# Wrap text, then create complete grid with all combinations
vm_exploration_complete <- vm_exploration_campaign_vannkategory |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN_wrapped = str_wrap(CAMPAIGN_DESCRIPTION_EN, wrap_y),
    Vannkategori_Description_wrapped = str_wrap(
      Vannkategori_Description,
      wrap_x
    )
  ) |>
  complete(
    CAMPAIGN_DESCRIPTION_EN_wrapped = str_wrap(campaign_order, wrap_y),
    Vannkategori_Description_wrapped = str_wrap(vannkategori_order, wrap_x),
    fill = list(n_samples = 0)
  ) |>
  mutate(
    # Factor with wrapped levels in the correct order
    CAMPAIGN_DESCRIPTION_EN_wrapped = factor(
      CAMPAIGN_DESCRIPTION_EN_wrapped,
      levels = str_wrap(campaign_order, wrap_y)
    ),
    Vannkategori_Description_wrapped = factor(
      Vannkategori_Description_wrapped,
      levels = str_wrap(vannkategori_order, wrap_x)
    )
  )

vm_exploration_complete |>
  ggplot(aes(
    x = Vannkategori_Description_wrapped,
    y = CAMPAIGN_DESCRIPTION_EN_wrapped,
    fill = n_samples
  )) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_text(
    aes(
      label = if_else(n_samples > 0, scales::comma(n_samples), ""),
      color = if_else(
        n_samples > 0 & log10(n_samples) > threshold_value,
        "black",
        "white"
      )
    ),
    size = 3
  ) +
  scale_fill_viridis_c(trans = "log10", na.value = "grey90") +
  scale_color_identity() +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(
    x = "",
    y = ""
  ) +
  theme(
    axis.text.x = element_text(),
    axis.text.y = element_text(),
    plot.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "none"
  )
```

@fig-heatmap-campaign-vannkategory clearly shows the the majority of copper contamination river and lake data from Vannmiljø is monitoring of mine-affected river systems. In general, sites expected to be contaminated (from mining, aquaculture, or other industries) are very well represented in the dataset. This is not unexpected, but needs to be kept in mind when interpreting the results.

Rivers, lakes, and coastal waters are the majority of the dataset, which is appropriate for a primarily water-focused database.

```{r}
#| label: fig-heatmap-medium-campaign
#| fig-cap: "Distribution of Vannmiljø copper samples by monitoring campaign and reported sampling medium, all years"
#| fig-width: 12
#| fig-height: 14
#| warning: false

# Create a summary of counts by the two variables
vm_exploration_campaign_x_medium <- vm_join_sites_measurements_lookup |>
  mutate(
    Name_EN = case_when(
      is.na(Name_EN) ~ "Other",
      TRUE ~ Name_EN
    )
  ) |>
  group_by(Name_EN, MediumID_Name) |>
  summarise(n_samples = n(), .groups = "drop")

threshold_value_medium <- mean(
  log10(vm_exploration_campaign_x_medium$n_samples),
  na.rm = TRUE
)

campaign_order_medium <- vm_exploration_campaign_x_medium |>
  group_by(Name_EN) |>
  summarise(total = sum(n_samples)) |>
  arrange(total) |>
  pull(Name_EN)

medium_order <- vm_exploration_campaign_x_medium |>
  group_by(MediumID_Name) |>
  summarise(total = sum(n_samples)) |>
  arrange(desc(total)) |>
  pull(MediumID_Name)

vm_exploration_campaign_x_medium_complete <- vm_exploration_campaign_x_medium |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN_wrapped = str_wrap(Name_EN, 30),
    MediumID_Name_wrapped = str_wrap(MediumID_Name, wrap_y)
  ) |>
  complete(
    CAMPAIGN_DESCRIPTION_EN_wrapped = str_wrap(campaign_order_medium, 30),
    MediumID_Name_wrapped = str_wrap(medium_order, 20),
    fill = list(n_samples = 0)
  ) |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN_wrapped = factor(
      CAMPAIGN_DESCRIPTION_EN_wrapped,
      levels = str_wrap(campaign_order_medium, 30)
    ),
    MediumID_Name_wrapped = factor(
      MediumID_Name_wrapped,
      levels = str_wrap(medium_order, 20)
    )
  )

vm_exploration_campaign_x_medium_complete |>
  ggplot(aes(
    x = MediumID_Name_wrapped,
    y = CAMPAIGN_DESCRIPTION_EN_wrapped,
    fill = n_samples
  )) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_text(
    aes(
      label = if_else(n_samples > 0, scales::comma(n_samples), ""),
      color = if_else(
        n_samples > 0 & log10(n_samples) > threshold_value_medium,
        "black",
        "white"
      )
    ),
    size = 3
  ) +
  scale_fill_viridis_c(trans = "log10", na.value = "grey90") +
  scale_color_identity() +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(x = "", y = "") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 7),
    axis.text.y = element_text(size = 7),
    plot.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "none"
  )
```

We also check reported water category by sampling medium, for the sake of completeness (@fig-heatmap-vannkategori-medium). From my interpretation, there are some combinations that make little sense (e.g. water category = air, sampling medium = soil), However, these are a minority.

```{r}
#| label: fig-heatmap-vannkategori-medium
#| fig-cap: "Distribution of Vannmiljø copper samples by water category and reported sampling medium, all years"
#| fig-width: 12
#| fig-height: 3
#| warning: false

# Create a summary of counts by the two variables
vm_exploration_vkat_x_medium <- vm_join_sites_measurements_lookup |>
  group_by(Vannkategori_Description, MediumID_Name) |>
  summarise(n_samples = n(), .groups = "drop")

threshold_value_vkat_medium <- mean(
  log10(vm_exploration_vkat_x_medium$n_samples),
  na.rm = TRUE
)

vannkategori_order_vkat <- vm_exploration_vkat_x_medium |>
  group_by(Vannkategori_Description) |>
  summarise(total = sum(n_samples)) |>
  arrange(total) |>
  pull(Vannkategori_Description)

medium_order_vkat <- vm_exploration_vkat_x_medium |>
  group_by(MediumID_Name) |>
  summarise(total = sum(n_samples)) |>
  arrange(desc(total)) |>
  pull(MediumID_Name)

vm_exploration_vkat_x_medium_complete <- vm_exploration_vkat_x_medium |>
  mutate(
    Vannkategori_Description_wrapped = str_wrap(Vannkategori_Description, 20),
    MediumID_Name_wrapped = str_wrap(MediumID_Name, 20)
  ) |>
  complete(
    Vannkategori_Description_wrapped = str_wrap(vannkategori_order_vkat, 20),
    MediumID_Name_wrapped = str_wrap(medium_order_vkat, 20),
    fill = list(n_samples = 0)
  ) |>
  mutate(
    Vannkategori_Description_wrapped = factor(
      Vannkategori_Description_wrapped,
      levels = str_wrap(vannkategori_order_vkat, 20)
    ),
    MediumID_Name_wrapped = factor(
      MediumID_Name_wrapped,
      levels = str_wrap(medium_order_vkat, 20)
    )
  )

vm_exploration_vkat_x_medium_complete |>
  ggplot(aes(
    x = MediumID_Name_wrapped,
    y = Vannkategori_Description_wrapped,
    fill = n_samples
  )) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_text(
    aes(
      label = if_else(n_samples > 0, scales::comma(n_samples), ""),
      color = if_else(
        n_samples > 0 & log10(n_samples) > threshold_value_vkat_medium,
        "black",
        "white"
      )
    ),
    size = 3
  ) +
  scale_fill_viridis_c(trans = "log10", na.value = "grey90") +
  scale_color_identity() +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(x = "", y = "") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 7),
    axis.text.y = element_text(size = 7),
    plot.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "none"
  )
```

@tbl-sample-distribution gives the cumulative proportion of samples corresponding to each combination of reported water category and medium. This allows us to make strategic decisions about which groups to prioritise conversion into our eData format. I have elected to use 90% as an arbitrary cut-off point - that is, once I have 90% coverage of water categories and sampling media (before date-based sampling), I will consider that sufficient data.

```{r}
#| label: tbl-sample-distribution

# Calculate sample distribution with cumulative percentages
sample_distribution <- vm_exploration_vkat_x_medium |>
  arrange(desc(n_samples)) |>
  mutate(
    pct_of_total = 100 * n_samples / sum(n_samples),
    cumulative_n = cumsum(n_samples),
    cumulative_pct = 100 * cumulative_n / sum(n_samples)
  ) |>
  select(
    Vannkategori_Description,
    MediumID_Name,
    n_samples,
    pct_of_total,
    cumulative_n,
    cumulative_pct
  )

datatable(
  sample_distribution,
  colnames = c(
    "Water Category",
    "Medium",
    "N Samples",
    "% of Total",
    "Cumulative N",
    "Cumulative %"
  ),
  options = list(
    pageLength = 10,
    order = list(list(3, 'desc'))
  )
) |>
  formatRound(columns = c("pct_of_total", "cumulative_pct"), digits = 1) |>
  formatCurrency(
    columns = c("n_samples", "cumulative_n"),
    currency = "",
    digits = 0
  )
```

We also consider sample size of @fig-heatmap-campaign-biota by campaign. This includes many largely irrelevant (not-aquatic) species, but the vast majority are standard aquatic monitoring species (mussels, cod, edible crabs, etc.).

```{r}
#| label: fig-heatmap-campaign-biota
#| fig-cap: "Distribution of Vannmiljø copper samples by monitoring campaign and species, all years"
#| fig-width: 12
#| fig-height: 25
#| warning: false

# todo: facet this by species group, it's too tall otherwise
# Create a summary of counts by the two variables (filter to non-NA species)
vm_exploration_species <- vm_join_sites_measurements_lookup |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN = case_when(
      is.na(Name_EN) ~ "Other",
      TRUE ~ Name_EN
    )
  ) |>
  filter(!is.na(VitenskapligNavn), VitenskapligNavn != "Artsuavhengig") |>
  group_by(CAMPAIGN_DESCRIPTION_EN, VitenskapligNavn) |>
  summarise(n_samples = n(), .groups = "drop")

# Calculate text color threshold
threshold_value_species <- mean(
  log10(vm_exploration_species$n_samples),
  na.rm = TRUE
)

campaign_order_species_new <- vm_exploration_species |>
  group_by(CAMPAIGN_DESCRIPTION_EN) |>
  summarise(total = sum(n_samples)) |>
  arrange(desc(total)) |>
  pull(CAMPAIGN_DESCRIPTION_EN)

species_order_new <- vm_exploration_species |>
  group_by(VitenskapligNavn) |>
  summarise(total = sum(n_samples)) |>
  arrange(total) |>
  pull(VitenskapligNavn)

vm_exploration_species_complete <- vm_exploration_species |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN_wrapped = str_wrap(CAMPAIGN_DESCRIPTION_EN, 30),
    VitenskapligNavn_wrapped = str_wrap(VitenskapligNavn, 20)
  ) |>
  complete(
    CAMPAIGN_DESCRIPTION_EN_wrapped = str_wrap(campaign_order_species_new, 30),
    VitenskapligNavn_wrapped = str_wrap(species_order_new, 20),
    fill = list(n_samples = 0)
  ) |>
  mutate(
    CAMPAIGN_DESCRIPTION_EN_wrapped = factor(
      CAMPAIGN_DESCRIPTION_EN_wrapped,
      levels = str_wrap(campaign_order_species_new, 30)
    ),
    VitenskapligNavn_wrapped = factor(
      VitenskapligNavn_wrapped,
      levels = str_wrap(species_order_new, 20)
    )
  )

vm_exploration_species_complete |>
  ggplot(aes(
    x = VitenskapligNavn_wrapped,
    y = CAMPAIGN_DESCRIPTION_EN_wrapped,
    fill = n_samples
  )) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_text(
    aes(
      label = if_else(n_samples > 0, scales::comma(n_samples), ""),
      color = if_else(
        n_samples > 0 & log10(n_samples) > threshold_value_species,
        "black",
        "white"
      )
    ),
    size = 3
  ) +
  scale_fill_viridis_c(trans = "log10", na.value = "grey90") +
  scale_color_identity() +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(x = "", y = "") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 7),
    axis.text.y = element_text(size = 7),
    plot.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "none"
  ) +
  coord_flip()

```

Finally, we check units by sampling medium @fig-heatmap-units-medium. We would expect mass/volume units to be used for water and air samples, and mass/mass (dry or wet) units to be used for biota, soil, sediment, and other largely solid matrices. Fortunately this is borne out by the data.

```{r}
#| label: fig-heatmap-units-medium
#| fig-cap: "Units used in Vannmiljø copper samples by MediumName"
#| fig-width: 12
#| fig-height: 3
#| warning: false

vm_exploration_units_x_medium <- vm_join_sites_measurements_lookup |>
  group_by(Unit_Name, MediumID_Name) |>
  summarise(n_samples = n(), .groups = "drop")
threshold_value_units_x_medium <- mean(
  log10(vm_exploration_units_x_medium$n_samples),
  na.rm = TRUE
)

unit_order <- vm_exploration_units_x_medium |>
  group_by(Unit_Name) |>
  summarise(total = sum(n_samples)) |>
  arrange(total) |>
  pull(Unit_Name)

medium_order_units <- vm_exploration_units_x_medium |>
  group_by(MediumID_Name) |>
  summarise(total = sum(n_samples)) |>
  arrange(desc(total)) |>
  pull(MediumID_Name)

vm_exploration_units_x_medium_complete <- vm_exploration_units_x_medium |>
  mutate(
    Unit_Name_wrapped = str_wrap(Unit_Name, 20),
    MediumID_Name_wrapped = str_wrap(MediumID_Name, 20)
  ) |>
  complete(
    Unit_Name_wrapped = str_wrap(unit_order, 20),
    MediumID_Name_wrapped = str_wrap(medium_order_units, 20),
    fill = list(n_samples = 0)
  ) |>
  mutate(
    Unit_Name_wrapped = factor(
      Unit_Name_wrapped,
      levels = str_wrap(unit_order, 20)
    ),
    MediumID_Name_wrapped = factor(
      MediumID_Name_wrapped,
      levels = str_wrap(medium_order_units, 20)
    )
  )

vm_exploration_units_x_medium_complete |>
  ggplot(aes(
    x = MediumID_Name_wrapped,
    y = Unit_Name_wrapped,
    fill = n_samples
  )) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_text(
    aes(
      label = if_else(n_samples > 0, scales::comma(n_samples), ""),
      color = if_else(
        n_samples > 0 & log10(n_samples) > threshold_value_units_x_medium,
        "black",
        "white"
      )
    ),
    size = 3
  ) +
  scale_fill_viridis_c(trans = "log10", na.value = "grey90") +
  scale_color_identity() +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(x = "", y = "") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 10),
    axis.text.y = element_text(size = 10),
    plot.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "none",
  )

```

Let's check out methods (@tbl-methods-by-year). Anything with an eData Classification column has been matched with the lookup. Vannmiljø methods without this are either not able to be matched (e.g. UKJENT) or aren't included in our filtered dataset (see next section).

```{r}
#| label: tbl-methods-by-year
#| tbl-cap: "Vannmiljø sampling and analysis method IDs reported by number of samples (rows), last year used, and corresponding eData Classification (if relevant)."

vm_exploration_methods_x_year <- vm_join_sites_measurements_lookup |>
  select(Tid_provetak, Provetakmetode_id, Analysemetode_id, Medium_id) |>
  pivot_longer(
    cols = c(Provetakmetode_id, Analysemetode_id),
    names_to = "Method_Type",
    values_to = "Method_Name"
  ) |>
  mutate(
    Year = ymd_hms(Tid_provetak) |> year(),
    Method_Type = case_match(
      Method_Type,
      "Provetakmetode_id" ~ "Sampling",
      "Analysemetode_id" ~ "Analysis"
    ),
    Base_Method = str_extract(
      Method_Name,
      pattern = "([A-Z\\-]* [0-9]{4,5}?(-[0-9]))"
    )
  )

vm_exploration_methods_x_year_table <- vm_exploration_methods_x_year |>
  group_by(Method_Name) |>
  reframe(
    Rows = n(),
    Method_Type = paste(Method_Type),
    Last_Year_Used = last(Year),
    Base_Method
  ) |>
  distinct() |>
  arrange(desc(Rows)) |>
  left_join(
    vm_lookup_methods |> select(ISO_ID, PROTOCOL_NAME),
    by = c("Method_Name" = "ISO_ID")
  )

datatable(
  vm_exploration_methods_x_year_table,
  colnames = c(
    "n",
    "Vm ID",
    "Rows",
    "Type",
    "Last Used",
    "Base Code",
    "eData Classification"
  )
)

```

## Apply Filters

As discussed above, to avoid spending excessive time on low-incidence edge cases we filter the data at this stage to the most relevant compartments and dates. The results of this summarisation are calculated below.

Included:

-   See vm_config tables in [Configuration](##Configuration) for specifics
-   Point sites only (no polygons)
-   Mainland Norway only (excluded Svalbard)

Excluded:

-   Svalbard sites (n=1)

```{r}
#| label: processing-apply-filters-and-summarise
# Load the intermediate targets ----

# Build exclusion summary ----
exclusion_summary <- tibble(
  Filter = character(),
  Rows_Excluded = integer(),
  Rows_Remaining = integer(),
  Excluded_Values = character()
)

# initial_count <- nrow(vm_data_with_dates)

# # Step 1: Compartment exclusions ----
# excluded_compartments <- vm_data_with_dates |>
#   anti_join(
#     vm_filtered_compartments,
#     by = join_by(across(everything()))
#   ) |>
#   group_by(ENVIRON_COMPARTMENT_SUB_vkat, ENVIRON_COMPARTMENT_SUB_medium) |>
#   reframe(n = n()) |>
#   mutate(
#     label = glue(
#       "{ENVIRON_COMPARTMENT_SUB_vkat}/{ENVIRON_COMPARTMENT_SUB_medium} (n={n})"
#     )
#   ) |>
#   pull(label) |>
#   paste(collapse = "; ")

# exclusion_summary <- exclusion_summary |>
#   add_row(
#     Filter = "Wrong compartment/subcompartment",
#     Rows_Excluded = initial_count - nrow(vm_filtered_compartments),
#     Rows_Remaining = nrow(vm_filtered_compartments),
#     Excluded_Values = excluded_compartments
#   )

# # Step 2: Site type exclusions ----
# excluded_sites <- vm_filtered_compartments |>
#   anti_join(
#     vm_filtered_sites,
#     by = join_by(across(everything()))
#   ) |>
#   group_by(Objekttype, Vannlokalitetsnavn) |>
#   reframe(n = n()) |>
#   mutate(label = glue("{Objekttype}: {Vannlokalitetsnavn} (n={n})")) |>
#   pull(label) |>
#   paste(collapse = "; ")

# exclusion_summary <- exclusion_summary |>
#   add_row(
#     Filter = "Not point sites or excluded sites",
#     Rows_Excluded = nrow(vm_filtered_compartments) - nrow(vm_filtered_sites),
#     Rows_Remaining = nrow(vm_filtered_sites),
#     Excluded_Values = excluded_sites
#   )

# # Step 3: Date range exclusions ----
# excluded_dates <- vm_filtered_sites |>
#   anti_join(
#     vm_filtered_dates,
#     by = join_by(across(everything()))
#   ) |>
#   group_by(year = year(SAMPLING_DATE)) |>
#   reframe(n = n()) |>
#   mutate(label = glue("{year} (n={n})")) |>
#   pull(label) |>
#   paste(collapse = "; ")

# exclusion_summary <- exclusion_summary |>
#   add_row(
#     Filter = "Outside date range",
#     Rows_Excluded = nrow(vm_filtered_sites) - nrow(vm_filtered_dates),
#     Rows_Remaining = nrow(vm_filtered_dates),
#     Excluded_Values = excluded_dates
#   )

# # Add total row ----
# exclusion_summary <- exclusion_summary |>
#   add_row(
#     Filter = "<b>Total</b>",
#     Rows_Excluded = initial_count - nrow(vm_filtered_dates),
#     Rows_Remaining = nrow(vm_filtered_dates),
#     Excluded_Values = ""
#   ) |>
#   mutate(
#     Percent_Excluded = round(100 * Rows_Excluded / initial_count, 1),
#     Percent_Remaining = round(100 * Rows_Remaining / initial_count, 1)
#   )

# # Display the summary ----
# exclusion_summary

message(glue(
  "vm_join_sites_measurements_lookup ({nrow(vm_join_sites_measurements_lookup)} x {ncol(vm_join_sites_measurements_lookup)}) -> vm_filtered_dates ({nrow(vm_filtered_dates)} x {ncol(vm_filtered_dates)}, {round(100*nrow(vm_filtered_dates)/nrow(vm_join_sites_measurements_lookup), 1)}% of starting data)"
))

```

For completeness, we also report on excluded groups (@tbl-processing-excluded-data)

-   [ ] Table HTML is broken here.

```{r}
#| label: fig-filter-visualisation
#| fig-cap: Vannmiljø copper sample size at different steps of filtration
#| fig-width: 12

# Create filtering summary for visualization
# filter_flow <- exclusion_summary |>
#   filter(Filter != "<b>Total</b>") |>
#   mutate(
#     stage = c("Compartments", "Point sites", "Date range"),
#     pct = Percent_Remaining,
#     label = glue("{scales::comma(Rows_Remaining)}\n({pct}%)")
#   ) |>
#   add_row(
#     Filter = "Raw",
#     stage = "Raw",
#     Rows_Remaining = initial_count,
#     pct = 100,
#     label = glue("{scales::comma(initial_count)}\n(100%)"),
#     .before = 1
#   )

# ggplot(filter_flow, aes(x = stage, y = Rows_Remaining)) +
#   geom_col(aes(fill = stage)) +
#   geom_text(aes(label = label), vjust = -0.5, size = 3) +
#   scale_fill_brewer(name = "Filtering Stage", palette = "Set1") +
#   scale_x_discrete(limits = filter_flow$stage) +
#   scale_y_continuous(
#     labels = scales::comma,
#     expand = expansion(mult = c(0, 0.1))
#   ) +
#   labs(
#     x = NULL,
#     y = "Number of Records"
#   ) +
#   theme(legend.position = "none")
```

A Sankey diagram would be better, but @fig-filter-visualisation will do for now. @tbl-processing-excluded-data provides a summary of what was excluded, and why.

```{r}
#| label: tbl-processing-excluded-data
#| tbl-cap: "Summary of excluded data by filter criterion"

datatable(
  exclusion_summary,
  colnames = c(
    "Filter Criterion",
    "Rows Excluded",
    "Rows Remaining",
    "Excluded Values",
    "% Excluded",
    "% Remaining"
  ),
  rownames = FALSE,
  escape = c("Excluded Values"),
  options = list(
    columnDefs = list(
      list(width = "300px", targets = 3)
    )
  )
) |>
  formatCurrency(
    columns = c("Rows_Excluded", "Rows_Remaining"),
    currency = "",
    digits = 0
  )
```

At time of writing, we've got 94% of compartment data, which is well above our goal of 90%.

## Quality Checks

We check for conflicts between our two lookups here - because we look up compartments, sites, biota, etc. based on both site `Vannlokalkode` and measurments' `MediumID`, it's possible that we might end up with conflicts and/or cases where neither of these give us the relevant data and we have to guess from context. Right now we filter down to only freshwater cases, hence the fact that we don't get hits below.

Note: It's not necessarily a problem that something turns up as a conflict - we fix a lot of these in the next steps.

```{r}
#| label: tbl-compartment-conflicts
#| tbl-cap: "Conflicts between vannkategori and medium lookups for environmental compartments"

# Check for conflicts between lookup sources
conflict_check <- vm_filtered_dates |>
  group_by(
    ENVIRON_COMPARTMENT_vkat,
    ENVIRON_COMPARTMENT_medium,
    ENVIRON_COMPARTMENT_SUB_vkat,
    ENVIRON_COMPARTMENT_SUB_medium
  ) |>
  reframe(count = n()) |>
  mutate(
    compartment_conflict = ENVIRON_COMPARTMENT_vkat !=
      ENVIRON_COMPARTMENT_medium,
    subcompartment_conflict = ENVIRON_COMPARTMENT_SUB_vkat !=
      ENVIRON_COMPARTMENT_SUB_medium,
    Conflict = case_when(
      compartment_conflict & subcompartment_conflict ~ "Main, Sub",
      compartment_conflict ~ "Main",
      subcompartment_conflict ~ "Sub",
      .default = NA_character_
    )
  ) |>
  filter(!is.na(Conflict)) |>
  arrange(desc(count))

if (nrow(conflict_check) > 0) {
  warning(
    glue(
      "{nrow(conflict_check)} conflicts detected between vannkategori and medium lookups.
      Result: conflict_check ({nrow(conflict_check)} x {ncol(conflict_check)})"
    )
  )
}

datatable(
  conflict_check |>
    select(
      ENVIRON_COMPARTMENT_vkat,
      ENVIRON_COMPARTMENT_medium,
      ENVIRON_COMPARTMENT_SUB_vkat,
      ENVIRON_COMPARTMENT_SUB_medium,
      count,
      Conflict
    ),
  colnames = c(
    "Compartment (Vannkategory)",
    "Compartment (MediumName)",
    "Subcompartment (Vannkategory)",
    "Subcompartment (MediumName)",
    "Rows",
    "Conflict Type"
  )
)
```

@tbl-compartment-conflicts lists all combinations of compartments and subcompartments, and whether any conflict is detected between the two compartments. I feel a need to be explicit and reproducible about what trumps what where, so see below.

-   Subcompartment (Vkat) == Freshwater & Subcompartment (Medium) == Aquatic Sediment then Aquatic Sediment
-   If one column returns biota, we use that (even if it's a fish e.g. living in the sea...)
-   If only one (sub)compartment value is a wildcard (\*), use the other. If both are wildcards, flag the row for exclusion.
-   If we know a row is biota (because it has a species name) and we don't have any other information on its sampling site (e.g. a `Vannkategori`), we guess from the species

```{r}
#| label: resolve-compartment-conflicts

test <- vm_filtered_dates |>
  filter(
    ENVIRON_COMPARTMENT_medium == "Biota",
    !str_detect(
      ENVIRON_COMPARTMENT_SUB_vkat,
      "Biota"
    )
  ) |>
  group_by(
    ENVIRON_COMPARTMENT_medium,
    ENVIRON_COMPARTMENT_vkat,
    ENVIRON_COMPARTMENT_SUB_medium,
    ENVIRON_COMPARTMENT_SUB_vkat,
    ENVIRON_COMPARTMENT_SUB_biota
  ) |>
  reframe(n())

vm_compartment_conflicts_resolved |>
  # inspect data
  group_by(
    ENVIRON_COMPARTMENT_resolved,
    ENVIRON_COMPARTMENT_SUB_resolved
  ) |>
  reframe(Rows = n(), Mean = mean(Verdi))

```

In some cases we're unable to resolve conflicts with recourse to the simple rules above. In these cases, we simply remove the rows, and create `vm_compartment_conflicts_resolved_removed`.

```{r}
#| label: remove-compartment-conflicts

vm_compartment_conflicts_resolved_removed |>
  # inspect data
  group_by(
    ENVIRON_COMPARTMENT_resolved,
    ENVIRON_COMPARTMENT_SUB_resolved
  ) |>
  reframe(Rows = n(), Mean = mean(Verdi))

# Now validate with pointblank
vm_compartment_conflicts_resolved_removed |>
  col_vals_in_set(
    ENVIRON_COMPARTMENT_resolved,
    eDataDRF::environ_compartments_vocabulary()
  ) |>
  col_vals_in_set(
    ENVIRON_COMPARTMENT_SUB_resolved,
    purrr::flatten(eDataDRF::environ_compartments_sub_vocabulary())
  )

# Todo: Could this be a conditional code chunk?
# vm_compartment_conflicts_resolved_removed |>
#   filter(
#     !(ENVIRON_COMPARTMENT_resolved %in%
#       eDataDRF::environ_compartments_vocabulary())
#   ) |>
#   group_by(ENVIRON_COMPARTMENT_resolved) |>
#   reframe(n = n())

# vm_compartment_conflicts_resolved_removed |>
#   filter(
#     !(ENVIRON_COMPARTMENT_SUB_resolved %in%
#       purrr::flatten(eDataDRF::environ_compartments_sub_vocabulary()))
#   ) |>
#   group_by(ENVIRON_COMPARTMENT_SUB_resolved) |>
#   reframe(n = n())

message(glue(
  "vm_filtered_dates ({nrow(vm_filtered_dates)} x {ncol(vm_filtered_dates)}) -> vm_compartment_conflicts_resolved_removed ({nrow(vm_compartment_conflicts_resolved_removed)} x {ncol(vm_compartment_conflicts_resolved_removed)})"
))

```

And let's take a look at anyting we failed to resolve: @tbl-unresolved-conflicts. At time of writing we have 1916 rows with unresolved conflicts, but this is a very small number and I don't (currently) judge it a good use of time to chase down these edge cases.

```{r}
#| label: tbl-unresolved-conflicts
#| tbl-cap: "Remaining unresolved conflicts after applying resolution rules"

# Check for unresolved conflicts
unresolved_conflicts_summary <- vm_compartment_conflicts_resolved |>
  # Keep only rows where conflicts weren't resolved
  filter(
    ENVIRON_COMPARTMENT_SUB_resolved == "FLAG: Compartment conflict." &
      ENVIRON_COMPARTMENT_resolved == "FLAG: Compartment conflict."
  ) |>
  group_by(
    ENVIRON_COMPARTMENT_vkat,
    ENVIRON_COMPARTMENT_medium,
    ENVIRON_COMPARTMENT_SUB_vkat,
    ENVIRON_COMPARTMENT_SUB_medium
  ) |>
  reframe(count = n()) |>
  mutate(
    compartment_conflict = ENVIRON_COMPARTMENT_vkat !=
      ENVIRON_COMPARTMENT_medium,
    subcompartment_conflict = ENVIRON_COMPARTMENT_SUB_vkat !=
      ENVIRON_COMPARTMENT_SUB_medium,
    Conflict = case_when(
      compartment_conflict & subcompartment_conflict ~ "Main, Sub",
      compartment_conflict ~ "Main",
      subcompartment_conflict ~ "Sub",
      .default = NA_character_
    )
  ) |>
  arrange(desc(count))

datatable(
  unresolved_conflicts_summary |>
    select(
      ENVIRON_COMPARTMENT_vkat,
      ENVIRON_COMPARTMENT_medium,
      ENVIRON_COMPARTMENT_SUB_vkat,
      ENVIRON_COMPARTMENT_SUB_medium,
      count,
      Conflict
    ),
  colnames = c(
    "Compartment (Vannkategory)",
    "Compartment (MediumName)",
    "Subcompartment (Vannkategory)",
    "Subcompartment (MediumName)",
    "Rows",
    "Conflict Type"
  )
)
```

`vm_compartment_conflicts_resolved_removed` represents our dataset after filtering and resolution/removal of compartment lookup conflicts. However, we need to repeat the exercise for geographic features, which we also use two lookups for.

@tbl-geographic-conflicts also lists all combinations of conflicts. We haven't applied simple, rules-based resolution yet, however.

```{r}
#| label: tbl-geographic-conflicts
#| tbl-cap: "Conflicts between vannkategori and medium lookups for geographic features"

# Check geographic feature consistency
vm_geo_conflicts_flagged <- vm_compartment_conflicts_resolved_removed |>
  mutate(
    # Only flag as conflict if both values are present and different
    # FIXME: We use different logic here and in the below block :(
    geo_conflict = !has_data_issue(SITE_GEOGRAPHIC_FEATURE_vkat) &
      !has_data_issue(SITE_GEOGRAPHIC_FEATURE_medium) &
      SITE_GEOGRAPHIC_FEATURE_vkat != "*" &
      SITE_GEOGRAPHIC_FEATURE_medium != "*" &
      SITE_GEOGRAPHIC_FEATURE_vkat != SITE_GEOGRAPHIC_FEATURE_medium,

    geo_sub_conflict = !has_data_issue(SITE_GEOGRAPHIC_FEATURE_SUB_vkat) &
      !has_data_issue(SITE_GEOGRAPHIC_FEATURE_SUB_medium) &
      SITE_GEOGRAPHIC_FEATURE_SUB_vkat != "*" &
      SITE_GEOGRAPHIC_FEATURE_SUB_medium != "*" &
      SITE_GEOGRAPHIC_FEATURE_SUB_vkat != SITE_GEOGRAPHIC_FEATURE_SUB_medium,

    Conflict = case_when(
      geo_conflict & geo_sub_conflict ~ "Main, Sub",
      geo_conflict ~ "Main",
      geo_sub_conflict ~ "Sub",
      .default = NA_character_
    )
  )

vm_geo_conflicts_flagged |> group_by(Conflict) |> reframe(n())

vm_geo_conflicts_summary <- vm_geo_conflicts_flagged |>
  group_by(
    SITE_GEOGRAPHIC_FEATURE_vkat,
    SITE_GEOGRAPHIC_FEATURE_medium,
    SITE_GEOGRAPHIC_FEATURE_SUB_vkat,
    SITE_GEOGRAPHIC_FEATURE_SUB_medium
  ) |>
  reframe(count = n(), Conflict = first(Conflict)) |>
  filter(!is.na(Conflict)) |>
  arrange(desc(count))

if (nrow(vm_geo_conflicts_summary) > 0) {
  warning(glue(
    "Geographic feature conflicts: {nrow(vm_geo_conflicts_summary)} combinations ({sum(vm_geo_conflicts_summary$count)} rows)"
  ))
}

datatable(
  vm_geo_conflicts_summary |>
    select(
      SITE_GEOGRAPHIC_FEATURE_vkat,
      SITE_GEOGRAPHIC_FEATURE_medium,
      SITE_GEOGRAPHIC_FEATURE_SUB_vkat,
      SITE_GEOGRAPHIC_FEATURE_SUB_medium,
      count,
      Conflict
    ),
  colnames = c(
    "Feature (Vannkategory)",
    "Feature (MediumName)",
    "Subfeature (Vannkategory)",
    "Subfeature (MediumName)",
    "Rows",
    "Conflict Type"
  )
)
```

If the table above is empty (I don't update this text every time I update the pipeline), it hopefully means that the simple rules (and the fact that we've *already* removed conflicting compartments, and there may be correlation between the two) have worked.

My experience has been that `SITE_GEOGRAPHIC_FEATURE_SUB` is a lot harder to get information on, both in this conversion but also in papers. As it's less important, I've elected to just mark it as "Not reported" if it's not easy to solve with simple rules. This is unlikely to affect the eventual analysis.

Unlike environmental compartments we don't apply any special logic here. We just pick whatever's not a wildcard, and if both disagree, we drop the row.

```{r}
#| label: resolve-feature-conflicts

vm_compartment_geo_conflicts_resolved |>
  group_by(
    SITE_GEOGRAPHIC_FEATURE_resolved,
    SITE_GEOGRAPHIC_FEATURE_SUB_resolved
  ) |>
  reframe(n = n()) |>
  arrange(desc(n))


vm_compartment_geo_conflicts_resolved |>
  filter(ENVIRON_COMPARTMENT_resolved == "Biota") |>
  group_by(
    ENVIRON_COMPARTMENT_resolved,
    ENVIRON_COMPARTMENT_SUB_resolved,
    SITE_GEOGRAPHIC_FEATURE_vkat,
    SITE_GEOGRAPHIC_FEATURE_medium,
    SITE_GEOGRAPHIC_FEATURE_SUB_vkat,
    SITE_GEOGRAPHIC_FEATURE_SUB_medium
  ) |>
  reframe(n = n()) |>
  arrange(desc(n))


message(glue(
  "Removed {nrow(vm_geo_conflicts_flagged) - nrow(vm_compartment_geo_conflicts_resolved_removed)} rows with unresolved geographic conflicts"
))

message(glue(
  "vm_geo_conflicts_flagged ({nrow(vm_geo_conflicts_flagged)} x {ncol(vm_geo_conflicts_flagged)}) -> vm_compartment_geo_conflicts_removed ({nrow(vm_compartment_geo_conflicts_resolved_removed)} x {ncol(vm_compartment_geo_conflicts_resolved_removed)})"
))

vm_compartment_geo_conflicts_resolved_removed |>
  # inspect data
  group_by(
    ENVIRON_COMPARTMENT_resolved,
    ENVIRON_COMPARTMENT_SUB_resolved,
    SITE_GEOGRAPHIC_FEATURE_resolved,
    SITE_GEOGRAPHIC_FEATURE_SUB_resolved
  ) |>
  reframe(Rows = n(), Mean = mean(Verdi)) |>
  # does it have biota in it? it should.
  col_vals_make_set(
    ENVIRON_COMPARTMENT_resolved,
    c("Biota", "Aquatic"),
    label = "Check vm_compartment_geo_conflicts_resolved_removed has Biota & Aquatic"
  )

```

We also have the issue that our higher level of site resolution means we need to split some of Vannmiljø's sites. Specifically, they may sample from the same longitude and latitude (one vannmiljø site) but sample different parts of the feature (e.g. sediment and the water column; two vannmiljø media/matrices). Because in our format this is two separate sites, we need to split any sites where this happens.

We do this in a fairly basic way - we take the existing `Vannlok_kode` (a string of digits, e.g. `02.61-29530`) and, if we split it, append an incrementing number on the end (`02.61-29530` would become `02.61-29530-01` and `02.61-29530-02`).

Unfortunately this is quite slow.

```{r}
#| label: split-sites

# split sites where different types of samples have been taken

# Report on splitting
split_summary <- vm_sites_split |>
  distinct(Vannlok_kode, n_geo_combos) |>
  count(n_geo_combos, name = "n_sites") |>
  mutate(
    total_split_sites = n_geo_combos * n_sites,
    Status = if_else(n_geo_combos == 1, "Unsplit", "Split")
  )

n_original_sites <- n_distinct(vm_sites_split$Vannlok_kode)
n_split_sites <- n_distinct(vm_sites_split$Vannlok_kode_split)
n_affected <- split_summary |>
  filter(n_geo_combos > 1) |>
  pull(n_sites) |>
  sum()

message(glue(
  "Site splitting summary:\n",
  "  Original sites: {n_original_sites}\n",
  "  Sites requiring splitting: {n_affected}\n",
  "  Final unique site codes: {n_split_sites}"
))


message(glue(
  "vm_compartment_geo_conflicts_resolved_removed ({nrow(vm_compartment_geo_conflicts_resolved_removed)} x {ncol(vm_compartment_geo_conflicts_resolved_removed)}) -> vm_sites_split ({nrow(vm_sites_split)} x {ncol(vm_sites_split)})"
))

vm_sites_split |>
  # then check that most of what was there before is still there
  group_by(
    ENVIRON_COMPARTMENT_resolved,
    ENVIRON_COMPARTMENT_SUB_resolved,
    SITE_GEOGRAPHIC_FEATURE_resolved,
    SITE_GEOGRAPHIC_FEATURE_SUB_resolved
  ) |>
  reframe(Rows = n(), Mean = mean(Verdi))
# does it have biota in it? it should.
# col_vals_make_set(ENVIRON_COMPARTMENT_resolved, c("Biota", "Aquatic"))

```

Let's update the not-Sankey now that we've removed some more rows (@fig-conflict-resolution-visualisation). Note the columns "Wrong compartment" and "Compartment Conflicts Removed" - these steps are pre- and post-transformation, which is why the samples are different.

```{r}
#| label: fig-conflict-resolution-visualisation
#| fig-cap: Vannmiljø copper sample size at different steps of filtration and conflict removal.
#| fig-width: 12

# # Create filtering summary
# conflict_flow <- filter_flow |>
#   add_row(
#     Filter = "Compartment Conflicts Removed",
#     Rows_Remaining = nrow(vm_compartment_conflicts_resolved),
#     pct = round(100 * Rows_Remaining / first(filter_flow$Rows_Remaining), 1),
#     label = glue("{scales::comma(Rows_Remaining)}\n({pct}%)")
#   ) |>
#   add_row(
#     Filter = "Geographic Conflicts Removed",
#     Rows_Remaining = nrow(vm_compartment_sites_split),
#     pct = round(100 * Rows_Remaining / first(filter_flow$Rows_Remaining), 1),
#     label = glue("{scales::comma(Rows_Remaining)}\n({pct}%)")
#   )

# # Simple bar plot
# ggplot(
#   conflict_flow,
#   aes(x = Filter |> str_wrap(width = 20), y = Rows_Remaining)
# ) +
#   geom_col(aes(fill = Filter)) +
#   geom_text(aes(label = label), vjust = -0.5, size = 3) +
#   scale_fill_brewer(palette = "Set1") +
#   scale_x_discrete(limits = conflict_flow$Filter |> str_wrap(width = 20)) +
#   scale_y_continuous(
#     labels = scales::comma,
#     expand = expansion(mult = c(0, 0.1))
#   ) +
#   labs(
#     x = NULL,
#     y = "Number of Records"
#   ) +
#   theme_minimal() +
#   theme(legend.position = "none")
```

Finally, we need to run `pointblank` validation on our resolved/removed geographical feature data.

```{r}
#| label: conflict-resolution-pointblank
# # first of all, does our table have data in it?
# if (nrow(vm_sites_split_clean) * ncol(vm_sites_split_clean) != 0) {
#   vm_sites_split_clean |>
#     # is there a real dataset here?
#     # check that controlled vocab is respected
#     col_vals_in_set(
#       ENVIRON_COMPARTMENT_resolved,
#       eDataDRF::environ_compartments_vocabulary(),
#       label = "All ENVIRON_COMPARTMENT_resolved are in eDataDRF::environ_compartments_vocabulary()",
#       actions = validation_config$standard
#     ) |>
#     col_vals_in_set(
#       ENVIRON_COMPARTMENT_SUB_resolved,
#       purrr::flatten(eDataDRF::environ_compartments_sub_vocabulary()),
#       label = "All ENVIRON_COMPARTMENT_SUB_resolved are in eDataDRF::environ_compartments_sub_vocabulary()",
#       actions = validation_config$standard
#     ) |>
#     col_vals_in_set(
#       SITE_GEOGRAPHIC_FEATURE_resolved,
#       eDataDRF::geographic_features_vocabulary(),
#       label = "All SITE_GEOGRAPHIC_FEATURE_resolved are in eDataDRF::geographic_features_vocabulary()",
#       actions = validation_config$standard
#     ) |>
#     col_vals_in_set(
#       SITE_GEOGRAPHIC_FEATURE_SUB_resolved,
#       eDataDRF::geographic_features_sub_vocabulary(),
#       label = "All SITE_GEOGRAPHIC_FEATURE_SUB_resolved are in eDataDRF::geographic_features_sub_vocabulary()",
#       actions = validation_config$standard
#     ) |>
#     # then check that most of what was there before is still there
#     # e.g. in a prior version we accidentally filtered out all biota compartments
#     group_by(ENVIRON_COMPARTMENT_resolved, ENVIRON_COMPARTMENT_SUB_resolved) |>
#     reframe(n())
# } else {
#   stop("Error: vm_sites_split_clean does not contain data.")
# }

```

\# Create eData Tables

\## Campaign

```{r}
#| label: edata-generate_campaign_table

```

## Reference

```{r}
#| label: edata-reference

```

## Parameters

Note: Only setup for copper copper, should probably be extended in case we accidentally include some copper pyrithrione (which is apparently in Vannmiljø) some time

```{r}
#| label: edata-parameters

# TODO: See above

```

## Sites

```{r}
#| label: edata-sites-extraction

```

```{r}
#| label: edata-create-sites-table

```

Visually inspecting transformed site distribution (@fig-sites-sanity-map) allows us to check that a) sites are roughly where they should be and b) the correct CRS transformation was used.

I swear there wasn't a site in Africa last time I checked...

```{r}
#| label: fig-sites-sanity-map
#| fig-cap: Vannmiljø sampling sites binned into 50 hexagons, coloured by number of sites in area.
#| fig-width: 12

# Sanity check: plot sites on map
world_map <- map_data("world")
ggplot() +
  geom_polygon(
    data = world_map,
    aes(x = long, y = lat, group = group),
    fill = "lightgray",
    colour = "white"
  ) +
  geom_hex(
    data = vm_edata_sites,
    aes(x = LONGITUDE, y = LATITUDE),
    bins = 50
  ) +
  scale_fill_viridis_c() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() +
  theme(axis.text = element_blank(), axis.title = element_blank())

```

## Methods

Because Vannmiljø publish their methods data in a format similar to ours, the majority of processing has already been done. We did the rest of the processing in Excel when assemling the lookup table, so no additional work is required here. However, this will change if we expand the filter criteria to include more compartments or older samples, and the lookup will need to be updated manually.

```{r}
#| label: edata-methods


```

## Samples

In the eData workflow, we generate a samples table by combining sites, parameters, dates and compartments. This isn't how we work with Vannmiljø, but it's useful to have it anyway because we need it to generate the biota table - which stores species, tissue, etc. information.

```{r}
#| label: edata-samples

```

## Biota

Biota stuff. Some genders are implicit from the tissue type reported (e.g. gonads), but I don't think I want to worry about that here. Our assumptions:

-   All species reported in Vannmiljø are correct
-   No attempt was made to infer gender or lifestage

Nevertheless, we need to do a bit of extra work with scientific names, species groups, and tissue before proceeding. Currently:

-   Laksesmolt needs to be recoded as Juvenile Salmo salmar

That's it.

```{r}
#| label: tbl-raw-biota

```

We also need to convert tissue types (stored as MediumID in Vannmiljø) into our appropriate equivalent (@tbl-raw-tissue).

```{r}
#| label: tbl-raw-tissue
#| tbl-caption: Table of Vannmiljø MediumIDs corresponding to biota tissues, with sample size (n) and corresponding eData SAMPLE_TISSUE value.

# tissue_frequency <- count(vm_edata_biota, MediumID_Name) |>
#   arrange(desc(n)) |>
#   mutate(
#     SAMPLE_TISSUE = map_tissue_type(MediumID_Name)
#   )

# # Report data quality issues
# n_unknown_tissue <- tissue_frequency |>
#   filter(SAMPLE_TISSUE == "Unknown Tissue") |>
#   pull(n) |>
#   sum()

# message(glue(
#   "Biota data quality: {n_unknown_tissue} samples with unknown tissue type"
# ))

# datatable(tissue_frequency)

# edata_biota_tissue_merged <- edata_biota_merged |>
#   mutate(
#     SAMPLE_TISSUE = map_tissue_type(MediumID_Name)
#   )

```

```{r}
#| label: eData-biota

```

Once we've handled biota separately we can merge them with the rest of the samples to create the intermediate table used to create the measurements table. This is a slightly strange way of doing things, but it works.

```{r}
#| label: merge-biota-samples

```

## Measurements

```{r}
#| label: edata-measurements

```

# Export

## Create Export

We export the resulting data to a zip file in our `data/raw/eData` folder where we keep the literature extractions.

Disabled, there's better ways to do this...

```{r}
#| label: export-create-zip

# # Assemble datasets for export
# export_datasets <- list(
#   edata_sites = vm_edata_sites,
#   edata_campaign = vm_edata_campaign,
#   edata_reference = vm_edata_reference,
#   edata_parameters = vm_edata_parameters,
#   edata_samples = vm_edata_samples,
#   edata_biota = vm_edata_biota,
#   edata_measurements = vm_edata_measurements
# )

# # Export as ZIP
# if (
#   !file.exists(here(
#     "data/raw/eData",
#     paste0(vm_config$metadata$campaign_short, ".zip")
#   ))
# ) {
#   timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")

#   export_path <- export_campaign_zip(
#     dataset_list = export_datasets,
#     campaign_name = vm_config$metadata$campaign_short,
#     output_path = here(
#       "data",
#       "raw",
#       "eData",
#       glue("{vm_config$metadata$campaign_short}_{timestamp}.zip")
#     )
#   )
#   message(glue("Export complete: {export_path}"))
# } else {
#   message("Export file already exists, skipping")
# }

```

## Validate Export

Do some quick checks to see if the data we're exporting looks like we expect it to. In an ideal world we would be able to automatically check and validate against constrolled vocabulary, but we're not quite there yet

### Compartments

Are our compartments valid and reasonable? @tbl-validate-export-compartments gives a quick summary..

```{r}
#| label: tbl-validate-export-compartments
count(vm_edata_measurements, ENVIRON_COMPARTMENT, ENVIRON_COMPARTMENT_SUB) |>
  datatable()

```

@fig-vm-scatterplot is a basic plot of measured concentrations by year and unit. Without going into detailed analysis of campaigns, etc. the results are a little difficult to interpret, but it's reasonable to call some of the shapes we see down there suspicious.

```{r}
#| label: fig-vm-scatterplot
#| fig-cap: A basic scatterplot for visualising measured concentrations by year. Values more/less than 3 S.D from the mean per group highlighted in red.
#| fig-width: 12
#| fig-height: 12
edata_zscore <- vm_edata_measurements |>
  group_by(ENVIRON_COMPARTMENT_SUB) |>
  mutate(z_score = scale(MEASURED_VALUE)[, 1]) |>
  ungroup()

ggplot(
  data = edata_zscore,
  aes(
    y = MEASURED_VALUE,
    x = SAMPLING_DATE |> as.Date()
  )
) +
  geom_point(shape = 1, alpha = 0.5, colour = "grey") +
  geom_point(
    data = edata_zscore |> filter(abs(z_score) >= 3),
    colour = "red",
    shape = 1
  ) +
  facet_wrap(
    facets = vars(ENVIRON_COMPARTMENT_SUB),
    nrow = 7,
    scales = "free_y"
  ) +
  labs(
    x = "Year",
    y = "Measured Value",
    caption = "Red circles indicate |z-score| ≥ 3"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

We can pull out this data and take a closer look. We'll use the z-score method and assume anything more than 3 SD from the mean is worth investigation.

-   [ ] Todo: find a citation for use of z-scores
-   [ ] Implement using pointblank

```{r}
#| label: tbl-vm-outliers

vm_edata_measurements |>
  group_by(MEASURED_UNIT, ENVIRON_COMPARTMENT_SUB) |>
  mutate(z_score = scale(MEASURED_VALUE)) |> # Calculate z-score
  filter(z_score >= 3)
```