---
title: "Notebook 03 - Visualisation"
author: "Sam Welch"
date: last-modified
format:
  html:
    fig-format: svg
---

```{r}
#| label: targets-setup
#| warning: false
#| output: false
library(DT)
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
library(viridis)
library(stringr)
library(ggridges)
library(ggh4x)
library(tidygraph)
library(ggraph)
library(ggiraph)
library(leaflet)
library(sf)
library(shiny)
library(dplyr)
library(bslib)
library(ggtext)
library(data.table)
library(eDataDRF)
library(here)
library(targets)
library(purrr)

devtools::load_all(quiet = TRUE)
```

```{r}
#| label: load-data
# load in any data we need from the targets workflow

literature_merged_data <- tar_read(load_literature_pqt)
literature_reference_data <- tar_read(reference_data)
literature_campaign_data <- tar_read(campaign_data)
literature_sites_data <- tar_read(sites_data)
literature_qc <- tar_read(data_quality_report)
wgs84_geo <- tar_read(wgs84_geography)

species_names <- eDataDRF::species_names_vocabulary()
species_lookup <- species_names |>
  with(setNames(SPECIES_COMMON_NAME, SPECIES_NAME))
```

```{r}
#| label: themes-and-helpers
copper_toxicity_thresholds <- tar_read(copper_toxicity_thresholds)

# Helper function for sample sizes
calculate_sample_sizes <- function(data) {
  data |>
    group_by(
      REFERENCE_ID,
      OCEAN_COUNTRY,
      SITE_GEOGRAPHIC_FEATURE,
      ENVIRON_COMPARTMENT_SUB
    ) |>
    summarise(total_n = sum(MEASURED_N, na.rm = TRUE), .groups = "drop")
}
```

## Dataset Overview

### To Do

-   [ ] Add probable contamination sources to each paper
-   [ ] Make sure we handle missing characters properly
-   [ ] Reach out to Dag or Jens to ask if MD have any copper emissions data
-   [ ] Talk to Marianne from reference group (or anyone else on the reference group)
-   [ ] Get more data in
-   [ ] (Maybe) renable plotly
-   [ ] Make network diagram visualise uncertainty, etc.
-   [ ] Functional and risk assessment trophic levels may not be perfectly aligned
    -   [ ] might be better off using ERA grouping - talk to Li about it
-   [ ] Fix icon colouration in map
-   [ ] Flag papers with transfer data (again)
-   [ ] Add CREED stuff
-   [ ] Add missing species names
-   [ ] Get some salinity & DOC (or at least water darkness) datasets
-   [ ] Find a good way to add sea ice to map
-   [ ] Find a good aquaculture zone and mining waste zone and characterise the sources well


### Compartment Coverage

```{r}
#| label: fig-paper-subcompartments
#| fig-cap: "Number of environmental sub-compartments sampled per paper. Each bar represents a unique publication, with height indicating the diversity of compartments (e.g., surface water, sediment, soil) analysed within that study."
#| warning: false
#| fig-width: 12
#| fig-height: 8
#| 
# Helper: parse REFERENCE_ID into a two-line y-axis label ----
format_ref_label <- function(id) {
  if (str_detect(id, "\\d{4}(-\\d{4})?")) {
    year_part <- str_extract(id, "\\d{4}(-\\d{4})?")
    remainder <- str_remove(id, "\\d{4}(-\\d{4})?")
    glue("**{remainder}**<br>{year_part}")
  } else {
    str_wrap(id, width = 40)
  }
}

# Prepare data for plotting ----
plot_data <- literature_merged_data |>
  group_by(REFERENCE_ID, ENVIRON_COMPARTMENT_SUB) |>
  distinct() |>
  reframe(count = sum(MEASURED_N, na.rm = TRUE)) |>
  left_join(
    literature_merged_data |> distinct(REFERENCE_ID, YEAR),
    by = "REFERENCE_ID"
  ) |>
  filter(!is.na(ENVIRON_COMPARTMENT_SUB), ENVIRON_COMPARTMENT_SUB != "") |>
  mutate(label_2line = map_chr(REFERENCE_ID, format_ref_label))

# Calculate text color threshold (on log10 scale) ----
threshold_value <- mean(
  log10(plot_data$count[plot_data$count > 0]),
  na.rm = TRUE
)

# Calculate total samples per reference for ordering (y-axis) ----
reference_order <- plot_data |>
  group_by(REFERENCE_ID) |>
  summarise(total = sum(count)) |>
  arrange(total) |>
  pull(REFERENCE_ID)

# Derive label order from reference_order ----
label_order <- map_chr(reference_order, format_ref_label)

# Calculate total samples per compartment for ordering (x-axis) ----
compartment_order <- plot_data |>
  group_by(ENVIRON_COMPARTMENT_SUB) |>
  summarise(total = sum(count)) |>
  arrange(desc(total)) |>
  pull(ENVIRON_COMPARTMENT_SUB)

# Create complete grid with all combinations ----
plot_data_complete <- plot_data |>
  complete(
    label_2line = label_order,
    ENVIRON_COMPARTMENT_SUB = compartment_order,
    fill = list(count = 0)
  ) |>
  mutate(
    label_2line = factor(label_2line, levels = label_order),
    ENVIRON_COMPARTMENT_SUB = factor(
      ENVIRON_COMPARTMENT_SUB,
      levels = compartment_order
    )
  )

# Create plot ----
compartments_matrix <- plot_data_complete |>
  ggplot(aes(
    x = ENVIRON_COMPARTMENT_SUB,
    y = label_2line,
    fill = count
  )) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_text(
    aes(
      label = if_else(count > 0, scales::comma(count), ""),
      color = if_else(
        count > 0 & log10(count) > threshold_value,
        "black",
        "white"
      )
    ),
    size = 3
  ) +
  scale_fill_viridis_c(
    transform = scales::transform_log10(),
    na.value = "white"
  ) +
  scale_color_identity() +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), labels = \(x) {
    str_wrap(x, 15, whitespace_only = FALSE)
  }) +
  labs(x = "", y = "") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 10),
    axis.text.y = element_markdown(lineheight = 0.85),
    axis.ticks.y = element_blank(),
    panel.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "none"
  )

ggsave(
  plot = compartments_matrix,
  filename = "images/fig_paper_subcompartments.png",
  width = 11,
  height = 10,
  units = "in",
  dpi = 300
)

compartments_matrix
```

### Species Coverage

@fig-paper-species-1, @fig-paper-species-2, @fig-paper-species-3, and @fig-paper-species-4 illustrate the taxonomic diversity captured across studies, highlighting which papers focused on biota measurements and the range of species groups investigated.
```{r}
#| label: data-prep-species-plots
#| include: false
# Prepare data for plotting ----
plot_data <- literature_merged_data |>
  group_by(REFERENCE_ID, SAMPLE_SPECIES) |>
  distinct() |>
  reframe(
    count = sum(MEASURED_N, na.rm = TRUE),
    SPECIES_GROUP,
    SPECIES_COMMON_NAME
  ) |>
  # Join year data for y-axis ordering
  filter(!is.na(SAMPLE_SPECIES)) |>
  distinct()

# Create formatted labels ----
plot_data <- plot_data |>
  mutate(
    formatted_species = glue::glue(
      "**{SPECIES_COMMON_NAME}**<br>*({SAMPLE_SPECIES})*"
    )
  )

# Define plot groups ----
# Aim for ~36-37 species per plot
plot_groups <- c(
  "Fish" = 1, # 41 species
  "Crustaceans" = 2, # 36 species
  "Molluscs" = 3, # 19 species
  "Invertebrates" = 3, # 18 species (total: 37)
  "Mammals" = 4, # 14 species
  "Worms" = 4, # 6 species
  "Birds" = 4, # 5 species
  "Moss/Hornworts" = 4, # 2 species
  "Plants" = 4, # 2 species
  "Algae" = 4, # 1 species
  "Ecosystem" = 4, # 1 species
  "Other" = 4 # 1 species (total: 32)
)

# Add plot group to data ----
plot_data <- plot_data |>
  mutate(plot_group = plot_groups[SPECIES_GROUP])

# should work automatically but actually not very good
threshold_value <- 5000

# <- mean(
#   log10(plot_data$n_samples),
#   na.rm = TRUE
# )
```

```{r}
#| label: fig-paper-species-1
#| fig-cap: "Sample size per species per paper across the dataset (Part 1: Fish)."
#| warning: false
#| fig-width: 16
#| fig-height: 5
ggplot(
  plot_data |> filter(plot_group == 1),
  aes(y = REFERENCE_ID, x = formatted_species, fill = count)
) +
  geom_tile() +
  geom_text(aes(
    label = count,
    color = if_else(count > threshold_value, "black", "white")
  )) +
  scale_fill_viridis(name = "Sample Size (n)") +
  scale_color_identity() +
  theme_aep_plots() +
  theme(
    axis.text.x.bottom = element_markdown(angle = 270, hjust = 0, vjust = 0.5),
    axis.title.y = element_blank(),
    legend.position = "none",
    base_size = 12
  ) +
  labs(x = "") +
  facet_wrap(
    facets = vars(SPECIES_GROUP),
    space = "free_x",
    scales = "free_x",
    nrow = 1
  )
```

```{r}
#| label: fig-paper-species-2
#| fig-cap: "Sample size per species per paper across the dataset (Part 2: Crustaceans)."
#| warning: false
#| fig-width: 16
#| fig-height: 8
ggplot(
  plot_data |> filter(plot_group == 2),
  aes(y = REFERENCE_ID, x = formatted_species, fill = count)
) +
  geom_tile() +
  geom_text(aes(
    label = count,
    color = if_else(count > threshold_value, "black", "white")
  )) +
  scale_fill_viridis(name = "Sample Size (n)") +
  scale_color_identity() +
  theme_aep_plots() +
  theme(
    axis.text.x.bottom = element_markdown(angle = 270, hjust = 0, vjust = 0.5),
    axis.title.y = element_blank(),
    legend.position = "none"
  ) +
  labs(x = "Sampled Species") +
  facet_wrap(
    facets = vars(SPECIES_GROUP),
    space = "free_x",
    scales = "free_x",
    nrow = 1
  )
```
```{r}
#| label: fig-paper-species-3
#| fig-cap: "Sample size per species per paper across the dataset (Part 3: Molluscs & Invertebrates)."
#| warning: false
#| fig-width: 16
#| fig-height: 8
ggplot(
  plot_data |> filter(plot_group == 3),
  aes(y = REFERENCE_ID, x = formatted_species, fill = count)
) +
  geom_tile() +
  geom_text(aes(
    label = count,
    color = if_else(count > threshold_value, "black", "white")
  )) +
  scale_fill_viridis(name = "Sample Size (n)") +
  scale_color_identity() +
  theme_aep_plots() +
  theme(
    axis.text.x.bottom = element_markdown(angle = 270, hjust = 0, vjust = 0.5),
    axis.title.y = element_blank(),
    legend.position = "none"
  ) +
  labs(x = "Sampled Species") +
  facet_wrap(
    facets = vars(SPECIES_GROUP),
    space = "free_x",
    scales = "free_x",
    nrow = 1
  )
```

```{r}
#| label: fig-paper-species-4
#| fig-cap: "Sample size per species per paper across the dataset (Part 4: Mammals, Worms, Birds & Other)."
#| warning: false
#| fig-width: 16
#| fig-height: 8
ggplot(
  plot_data |> filter(plot_group == 4),
  aes(y = REFERENCE_ID, x = formatted_species, fill = count)
) +
  geom_tile() +
  geom_text(aes(
    label = count,
    color = if_else(count > threshold_value, "black", "white")
  )) +
  scale_fill_viridis(name = "Sample Size (n)") +
  scale_color_identity() +
  theme_aep_plots() +
  theme(
    axis.text.x.bottom = element_markdown(angle = 270, hjust = 0, vjust = 0.5),
    axis.title.y = element_blank(),
    legend.position = "none"
  ) +
  labs(x = "") +
  facet_wrap(
    facets = vars(SPECIES_GROUP),
    space = "free_x",
    scales = "free_x",
    nrow = 1
  )
```

## Simple Map

@fig-sites-hex-map is a simple density map of all our sites. Needs a fair amount of aesthetic work, but gives an accurate impression of where we have sites. 

```{r}
#| label: fig-sites-hex-map
#| fig-cap: All sampling sites binned into 500 hexagons, coloured by number of sites in area.
#| fig-width: 12
library(hexbin)

sites_sample_number <- literature_merged_data |>
  group_by(SITE_CODE) |>
  reframe(LATITUDE, LONGITUDE, n = sum(MEASURED_N)) |>
  distinct()

# Create hexbin object to get bin centers
hb <- hexbin(
  sites_sample_number$LONGITUDE,
  sites_sample_number$LATITUDE,
  xbins = 100
)

# Extract hex centers and counts
hex_coords <- data.frame(
  x = hcell2xy(hb)$x,
  y = hcell2xy(hb)$y,
  count = hb@count
)

# Transform your data to a projected CRS (example: UTM or Arctic projection)
sites_sample_number_proj <- sites_sample_number |>
  filter(!is.na(LONGITUDE), !is.na(LATITUDE)) |>
  sf::st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326) |>
  sf::st_transform(crs = 3035) |> # European LAEA projection, adjust for Arctic if needed
  dplyr::mutate(
    x = sf::st_coordinates(geometry)[, 1],
    y = sf::st_coordinates(geometry)[, 2]
  )

# Plot
world_map <- map_data("world")

# Also transform your world map
world_map_proj <- sf::st_as_sf(
  world_map,
  coords = c("long", "lat"),
  crs = 4326
) |>
  sf::st_transform(crs = 3035)


ggplot() +
  geom_polygon(
    data = world_map,
    aes(x = long, y = lat, group = group),
    fill = "lightgray",
    colour = "white"
  ) +
  geom_hex(
    data = sites_sample_number,
    aes(x = LONGITUDE, y = LATITUDE),
    bins = 100,
    linewidth = 0
  ) +
  # geom_text(
  #   data = hex_coords,
  #   aes(x = x, y = y, label = count),
  #   colour = "pink",
  #   size = 3
  # ) +
  scale_fill_viridis_b(
    name = "Samples",
    breaks = c(1, 5, 10, 50, 100, 500, 1000, 2000)
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  coord_fixed(
    ratio = 2,
    xlim = c(
      get_study_area_bbox()[[1]],
      get_study_area_bbox()[[3]]
    ),
    ylim = c(
      get_study_area_bbox()[[2]],
      get_study_area_bbox()[[4]]
    )
  )
```

## Copper Concentrations by Environmental Compartment

Get a basic overview of the orders of magnitude for different sub-compartments (@fig-basic-subcompartments-dry-mass, fig-basic-subcompartments-wet-mass, fig-basic-subcompartments-concentration), so we can plot them on scales that make sense. Note that I haven't included uncertainty or sample sizes.

```{r}
#| label: fig-basic-subcompartments-dry-mass
literature_merged_data |>
  filter(
    !is.na(MEASURED_UNIT_STANDARD),
    MEASURED_UNIT_STANDARD == "mg/kg (dry)"
  ) |>
  ggplot(aes(
    x = MEASURED_VALUE_STANDARD,
    y = ENVIRON_COMPARTMENT_SUB,
    color = ENVIRON_COMPARTMENT_SUB
  )) +
  geom_point() +
  scale_x_continuous(trans = "log10") +
  facet_nested(
    rows = vars(ENVIRON_COMPARTMENT),
    scales = "free_y",
    space = "free_y",
    nest_line = element_line()
  ) +
  theme_minimal() +
  scale_color_manual(
    values = compartment_colours,
    breaks = names(compartment_colours)
  ) +
  theme(legend.position = "none") +
  labs(x = "Concentration (mg/kg dry)", y = "")
```

```{r}
#| label: fig-basic-subcompartments-concentration
literature_merged_data |>
  filter(!is.na(MEASURED_UNIT_STANDARD), MEASURED_UNIT_STANDARD == "mg/L") |>
  ggplot(aes(
    x = MEASURED_VALUE_STANDARD,
    y = ENVIRON_COMPARTMENT_SUB,
    color = ENVIRON_COMPARTMENT_SUB
  )) +
  geom_point() +
  facet_nested(
    rows = vars(ENVIRON_COMPARTMENT),
    scales = "free_y",
    space = "free_y",
    nest_line = element_line()
  ) +
  theme_minimal() +
  scale_color_manual(
    values = compartment_colours,
    breaks = names(compartment_colours)
  ) +
  theme(legend.position = "none") +
  labs(x = "Concentration (mg/L)", y = "")
```

```{r}
#| label: fig-basic-subcompartments-wet-mass
literature_merged_data |>
  filter(
    !is.na(MEASURED_UNIT_STANDARD),
    MEASURED_UNIT_STANDARD == "mg/kg (wet)"
  ) |>
  ggplot(aes(
    x = MEASURED_VALUE_STANDARD,
    y = ENVIRON_COMPARTMENT_SUB,
    color = ENVIRON_COMPARTMENT_SUB
  )) +
  geom_point() +
  facet_nested(
    rows = vars(ENVIRON_COMPARTMENT),
    scales = "free_y",
    space = "free_y",
    nest_line = element_line()
  ) +
  theme_minimal() +
  scale_color_manual(
    values = compartment_colours,
    breaks = names(compartment_colours)
  ) +
  theme(legend.position = "none") +
  labs(x = "Concentration (mg/kg wet)", y = "")
```

The following sections present copper concentration distributions for each major environmental compartment. Plots are based on reported means uncertainty bounds (SD, range, 95% CI, etc.) and overlaid with relevant regulatory thresholds where available.

### Aquatic Compartment

#### Freshwater

@fig-boxplot-freshwater shows we have a fair amount of data for water. It's not strictly useful to include saltwater, freshwater and inbetween on here, so I think it makes sense to split it - but it won't make a massive difference to the final product.

Some of these units raise some big questions. Why is brackish/transitional using dry? And why are there some aquatic biota using mg/L?

```{r}
#| label: fig-boxplot-freshwater
#| fig-cap: "Copper concentrations in water column samples by reference"
#| fig-width: 10
# Calculate sample sizes
sample_sizes_water <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB == "Freshwater",
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  calculate_sample_sizes()

# Get first facet for labels
first_facet <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB == "Freshwater",
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  distinct(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE) |>
  slice(1)

# Create plot
literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB == "Freshwater",
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  ggplot(aes(
    x = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    y = REFERENCE_ID,
    colour = ENVIRON_COMPARTMENT_SUB
  )) +
  # ... (all your other geoms remain the same) ...
  geom_crossbar(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "95% Confidence Interval"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    )
  ) +
  geom_linerange(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    linewidth = 0.8
  ) +
  geom_point(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    size = 2
  ) +
  geom_boxplot(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "1st-3rd Quartile"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      lower = UNCERTAINTY_LOWER_STANDARD,
      middle = MEASURED_OR_IMPUTED_VALUE_STANDARD,
      upper = UNCERTAINTY_UPPER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    stat = "identity"
  ) +
  geom_point(
    alpha = 0.5,
    data = \(x) {
      filter(
        x,
        is_not_reported(UNCERTAINTY_TYPE) | is_not_relevant(UNCERTAINTY_TYPE)
      )
    }
  ) +
  geom_text(
    data = sample_sizes_water,
    aes(x = -Inf, y = REFERENCE_ID, label = paste0("n=", total_n)),
    hjust = -1,
    vjust = -1,
    inherit.aes = FALSE,
    size = 3
  ) +
  geom_vline(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(
          ENVIRON_COMPARTMENT_SUB %in% sub_comps,
          THRESHOLD_TYPE == "Classification boundary"
        )
    },
    aes(xintercept = THRESHOLD_VALUE, color = THRESHOLD_CLASS),
    linetype = "dashed",
    linewidth = 0.8,
    alpha = 0.6
  ) +
  # Threshold labels - only in first facet
  geom_text(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(ENVIRON_COMPARTMENT_SUB %in% sub_comps) |>
        cross_join(first_facet)
    },
    aes(
      x = THRESHOLD_VALUE,
      y = Inf,
      label = THRESHOLD_CLASS,
      color = THRESHOLD_CLASS
    ),
    vjust = 1.5,
    angle = 90,
    hjust = 1,
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_color_manual(
    values = c(threshold_colours, compartment_colours),
    breaks = names(compartment_colours)
  ) +
  labs(
    x = "Concentration (mg/L)",
    y = "Reference",
    colour = "Sub-compartment"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.box = "vertical"
  ) +
  facet_nested(
    rows = vars(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE),
    scales = "free_y",
    space = "free_y",
    shrink = TRUE,
    labeller = label_wrap_gen(width = 15),
    nest_line = element_line()
  ) +
  coord_cartesian(
    xlim = range(
      literature_merged_data$UNCERTAINTY_UPPER_STANDARD[
        literature_merged_data$ENVIRON_COMPARTMENT_SUB == "Freshwater"
      ],
      na.rm = TRUE
    ) |>
      round(),
    clip = "off" # allows threshold labels to extend beyond plot area if needed
  )
```

#### Salt and Transitional Water

@fig-boxplot-saltwater

```{r}
#| label: fig-boxplot-saltwater
#| fig-cap: "Copper concentrations in water column samples by reference"
#| fig-width: 10
# Calculate sample sizes
chosen_compartments <- c("Marine/Salt Water", "Brackish/Transitional Water")

sample_sizes_water <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  calculate_sample_sizes()

# Get first facet for labels
first_facet <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  distinct(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE) |>
  slice(1)

# Create plot
literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  ggplot(aes(
    x = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    y = REFERENCE_ID,
    colour = ENVIRON_COMPARTMENT_SUB
  )) +
  # ... (all your other geoms remain the same) ...
  geom_crossbar(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "95% Confidence Interval"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    )
  ) +
  geom_linerange(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    linewidth = 0.8
  ) +
  geom_point(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    size = 2
  ) +
  geom_boxplot(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "1st-3rd Quartile"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      lower = UNCERTAINTY_LOWER_STANDARD,
      middle = MEASURED_OR_IMPUTED_VALUE_STANDARD,
      upper = UNCERTAINTY_UPPER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    stat = "identity"
  ) +
  geom_point(
    alpha = 0.5,
    data = \(x) {
      filter(
        x,
        is_not_reported(UNCERTAINTY_TYPE) | is_not_relevant(UNCERTAINTY_TYPE)
      )
    }
  ) +
  geom_text(
    data = sample_sizes_water,
    aes(x = -Inf, y = REFERENCE_ID, label = paste0("n=", total_n)),
    hjust = -1,
    vjust = -1,
    inherit.aes = FALSE,
    size = 3
  ) +
  geom_vline(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(
          ENVIRON_COMPARTMENT_SUB %in% sub_comps,
          THRESHOLD_TYPE == "Classification boundary"
        )
    },
    aes(xintercept = THRESHOLD_VALUE, color = THRESHOLD_CLASS),
    linetype = "dashed",
    linewidth = 0.8,
    alpha = 0.6
  ) +
  # Threshold labels - only in first facet
  geom_text(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(ENVIRON_COMPARTMENT_SUB %in% sub_comps) |>
        cross_join(first_facet)
    },
    aes(
      x = THRESHOLD_VALUE,
      y = Inf,
      label = THRESHOLD_CLASS,
      color = THRESHOLD_CLASS
    ),
    vjust = 1.5,
    angle = 90,
    hjust = 1,
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_color_manual(
    values = c(threshold_colours, compartment_colours),
    breaks = names(compartment_colours)
  ) +
  labs(
    x = "Concentration (mg/L)",
    y = "Reference",
    colour = "Sub-compartment"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.box = "vertical"
  ) +
  facet_nested(
    rows = vars(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE),
    scales = "free_y",
    space = "free_y",
    shrink = TRUE,
    labeller = label_wrap_gen(width = 15),
    nest_line = element_line()
  ) +
  coord_cartesian(
    xlim = range(
      literature_merged_data$UNCERTAINTY_UPPER_STANDARD[
        literature_merged_data$ENVIRON_COMPARTMENT_SUB %in% chosen_compartments
      ],
      na.rm = TRUE
    ) |>
      round(),
    clip = "off" # allows threshold labels to extend beyond plot area if needed
  )
```

#### Aquatic Sediment

@fig-boxplot-sediment

```{r}
#| label: fig-boxplot-sediment
#| fig-cap: "Copper concentrations in aquatic sediment samples by reference"
#| fig-width: 10
# Calculate sample sizes
chosen_compartments <- c("Aquatic Sediment")

sample_sizes_water <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  calculate_sample_sizes()

# Get first facet for labels
first_facet <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  distinct(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE) |>
  slice(1)

# Create plot
literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  ggplot(aes(
    x = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    y = REFERENCE_ID,
    colour = ENVIRON_COMPARTMENT_SUB
  )) +
  # ... (all your other geoms remain the same) ...
  geom_crossbar(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "95% Confidence Interval"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    )
  ) +
  geom_linerange(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    linewidth = 0.8
  ) +
  geom_point(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    size = 2
  ) +
  geom_boxplot(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "1st-3rd Quartile"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      lower = UNCERTAINTY_LOWER_STANDARD,
      middle = MEASURED_OR_IMPUTED_VALUE_STANDARD,
      upper = UNCERTAINTY_UPPER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    stat = "identity"
  ) +
  geom_point(
    alpha = 0.5,
    data = \(x) {
      filter(
        x,
        is_not_reported(UNCERTAINTY_TYPE) | is_not_relevant(UNCERTAINTY_TYPE)
      )
    }
  ) +
  geom_text(
    data = sample_sizes_water,
    aes(x = -Inf, y = REFERENCE_ID, label = paste0("n=", total_n)),
    hjust = -1,
    vjust = -1,
    inherit.aes = FALSE,
    size = 3
  ) +
  geom_vline(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(
          ENVIRON_COMPARTMENT_SUB %in% sub_comps,
          THRESHOLD_TYPE == "Classification boundary"
        )
    },
    aes(xintercept = THRESHOLD_VALUE, color = THRESHOLD_CLASS),
    linetype = "dashed",
    linewidth = 0.8,
    alpha = 0.6
  ) +
  # Threshold labels - only in first facet
  geom_text(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(ENVIRON_COMPARTMENT_SUB %in% sub_comps) |>
        cross_join(first_facet)
    },
    aes(
      x = THRESHOLD_VALUE,
      y = Inf,
      label = THRESHOLD_CLASS,
      color = THRESHOLD_CLASS
    ),
    vjust = 1.5,
    angle = 90,
    hjust = 1,
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_color_manual(
    values = c(threshold_colours, compartment_colours),
    breaks = names(compartment_colours)
  ) +
  labs(
    x = "Concentration (mg/L)",
    y = "Reference",
    colour = "Sub-compartment"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.box = "vertical"
  ) +
  facet_nested(
    rows = vars(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE),
    scales = "free_y",
    space = "free_y",
    shrink = TRUE,
    labeller = label_wrap_gen(width = 15),
    nest_line = element_line()
  ) +
  coord_cartesian(
    xlim = range(
      literature_merged_data$UNCERTAINTY_UPPER_STANDARD[
        literature_merged_data$ENVIRON_COMPARTMENT_SUB %in% chosen_compartments
      ],
      na.rm = TRUE
    ) |>
      round(),
    clip = "off" # allows threshold labels to extend beyond plot area if needed
  )
```

### Terrestrial Compartment

Currently, data on copper in the terrestrial compartment is limited to one study (@fig-boxplot-terrestrial).

```{r}
#| label: fig-boxplot-terrestrial
#| fig-cap: "Copper concentrations in terrestrial samples by reference"
#| fig-width: 10
# Calculate sample sizes
chosen_compartments <- c("Soil O Horizon (Organic)")

sample_sizes_water <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  calculate_sample_sizes()

# Get first facet for labels
first_facet <- literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  distinct(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE) |>
  slice(1)

# Create plot
literature_merged_data |>
  filter(
    ENVIRON_COMPARTMENT_SUB %in% chosen_compartments,
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  ggplot(aes(
    x = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    y = REFERENCE_ID,
    colour = ENVIRON_COMPARTMENT_SUB
  )) +
  # ... (all your other geoms remain the same) ...
  geom_crossbar(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "95% Confidence Interval"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    )
  ) +
  geom_linerange(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    linewidth = 0.8
  ) +
  geom_point(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "Standard Deviation"),
    size = 2
  ) +
  geom_boxplot(
    data = \(x) filter(x, UNCERTAINTY_TYPE == "1st-3rd Quartile"),
    aes(
      xmin = UNCERTAINTY_LOWER_STANDARD,
      lower = UNCERTAINTY_LOWER_STANDARD,
      middle = MEASURED_OR_IMPUTED_VALUE_STANDARD,
      upper = UNCERTAINTY_UPPER_STANDARD,
      xmax = UNCERTAINTY_UPPER_STANDARD
    ),
    stat = "identity"
  ) +
  geom_point(
    alpha = 0.5,
    data = \(x) {
      filter(
        x,
        is_not_reported(UNCERTAINTY_TYPE) | is_not_relevant(UNCERTAINTY_TYPE)
      )
    }
  ) +
  geom_text(
    data = sample_sizes_water,
    aes(x = -Inf, y = REFERENCE_ID, label = paste0("n=", total_n)),
    hjust = -1,
    vjust = -1,
    inherit.aes = FALSE,
    size = 3
  ) +
  geom_vline(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(
          ENVIRON_COMPARTMENT_SUB %in% sub_comps,
          THRESHOLD_TYPE == "Classification boundary"
        )
    },
    aes(xintercept = THRESHOLD_VALUE, color = THRESHOLD_CLASS),
    linetype = "dashed",
    linewidth = 0.8,
    alpha = 0.6
  ) +
  # Threshold labels - only in first facet
  geom_text(
    data = \(x) {
      sub_comps <- unique(x$ENVIRON_COMPARTMENT_SUB)
      copper_toxicity_thresholds |>
        filter(ENVIRON_COMPARTMENT_SUB %in% sub_comps) |>
        cross_join(first_facet)
    },
    aes(
      x = THRESHOLD_VALUE,
      y = Inf,
      label = THRESHOLD_CLASS,
      color = THRESHOLD_CLASS
    ),
    vjust = 1.5,
    angle = 90,
    hjust = 1,
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_color_manual(
    values = c(threshold_colours, compartment_colours),
    breaks = names(compartment_colours)
  ) +
  labs(
    x = "Concentration (mg/L)",
    y = "Reference",
    colour = "Sub-compartment"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.box = "vertical"
  ) +
  facet_nested(
    rows = vars(OCEAN_COUNTRY, SITE_GEOGRAPHIC_FEATURE),
    scales = "free_y",
    space = "free_y",
    shrink = TRUE,
    labeller = label_wrap_gen(width = 15),
    nest_line = element_line()
  ) +
  coord_cartesian(
    xlim = range(
      literature_merged_data$UNCERTAINTY_UPPER_STANDARD[
        literature_merged_data$ENVIRON_COMPARTMENT_SUB %in% chosen_compartments
      ],
      na.rm = TRUE
    ) |>
      round(),
    clip = "off" # allows threshold labels to extend beyond plot area if needed
  )
```

### Biota Compartment

#### Aquatic Biota

presents copper body burdens across different species groups. Note: There appear to be issues with the distribution calculations for biota that require further investigation.

```{r}
#| label: fig-points-biota
#| fig-cap: "Density distributions of copper concentrations in biota, separated by species group. Copper is an essential micronutrient, so baseline concentrations vary widely among taxa. High concentrations may indicate either regulatory adaptation or toxic exposure. Note: Distribution calculations require verification."
#| warning: false
#| fig-width: 10
#| fig-height: 10
#| include: false
# TODO: Fix pls
# literature_merged_data |>
#   filter(
#     ENVIRON_COMPARTMENT == "Biota"
#   ) |>
#   ggplot(aes(
#     x = MEASURED_OR_IMPUTED_VALUE_STANDARD,
#     y = SAMPLE_SPECIES,
#     colour = SPECIES_GROUP
#   )) +
#   geom_pointrange(aes(
#     xmin = UNCERTAINTY_UPPER_STANDARD,
#     xmax = UNCERTAINTY_UPPER_STANDARD
#   )) +
#   facet_wrap(
#     facet = vars(OCEAN_COUNTRY),
#     ncol = 1,
#     scales = "free_y",
#     space = "free_y"
#   ) |> theme_minimal()

```

## Temporal Trends

### Abiotic Compartments

 This figure examines temporal trends in copper concentrations across abiotic environmental compartments. The current dataset has limited temporal coverage, but these trends may become more apparent as additional data from monitoring databases (e.g., Vannmiljø) are incorporated.

```{r}
#| label: fig-temporal-abiotic
#| fig-cap: "Temporal trends in copper concentrations across abiotic environmental sub-compartments. Points represent mean values with error bars showing standard deviations. Limited temporal coverage currently restricts trend interpretation, but patterns may emerge with additional historical monitoring data."
#| warning: false
#| fig-width: 10
#| fig-height: 12
p <- literature_merged_data |>
  filter(
    !is.na(MEASURED_UNIT_STANDARD),
    UNCERTAINTY_TYPE == "Standard Deviation",
    ENVIRON_COMPARTMENT != "Biota"
  ) |>
  group_by(ENVIRON_COMPARTMENT_SUB, SAMPLE_SPECIES) |>
  ggplot(aes(
    x = YEAR,
    y = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    color = ENVIRON_COMPARTMENT_SUB
  )) +
  geom_point() +
  geom_errorbar(
    aes(ymin = UNCERTAINTY_LOWER_STANDARD, ymax = UNCERTAINTY_UPPER_STANDARD),
    width = .2,
  ) +
  facet_wrap(
    facets = vars(ENVIRON_COMPARTMENT_SUB),
    ncol = 1,
    scales = "free_y"
  ) +
  labs(
    x = "Year",
    y = "Copper Concentration (standardised units)",
    color = "Sub-compartment"
  ) +
  theme_minimal()
```

### Biota

@fig-temporal-biota shows temporal patterns in biotic copper concentrations across species groups. Interpretation should consider both temporal changes in exposure and potential differences in species composition across sampling years.

```{r}
#| label: fig-temporal-biota
#| fig-cap: "Temporal trends in copper body burdens across different species groups. Points are coloured by species to help distinguish measurement series. Error bars represent reported uncertainty ranges. Vertical facets separate species groups to account for baseline differences in copper accumulation patterns."
#| warning: false
#| fig-width: 10
#| fig-height: 12
p <- literature_merged_data |>
  filter(
    !is.na(MEASURED_UNIT_STANDARD),
    ENVIRON_COMPARTMENT == "Biota"
  ) |>
  ggplot(aes(
    x = YEAR,
    y = MEASURED_OR_IMPUTED_VALUE_STANDARD,
    color = SAMPLE_SPECIES
  )) +
  geom_point() +
  geom_errorbar(
    aes(ymin = UNCERTAINTY_LOWER_STANDARD, ymax = UNCERTAINTY_UPPER_STANDARD),
    width = .2,
  ) +
  # facet_wrap(
  #   facets = vars(SPECIES_GROUP),
  #   ncol = 1,
  #   scales = "free_y"
  # ) +
  labs(
    x = "Year",
    y = "Copper Concentration (standardised units)",
    color = "Species"
  ) +
  theme_minimal()

p
```

