---
title: "Notebook 01 - Quality Control"
author: "Sam Welch"
date: last-modified
---
Notebook 1 contains code for quality control of extracted data. This has been largely done with custom functions up until now but I want to try `pointblank`.

# Todo:
- [ ] Move pointblank to targets workflow

```{r}
#| label: setup

suppressMessages(
  {
    library(DT)
    library(dplyr)
    library(eDataDRF)
    library(targets)
    library(here)
    library(glue)
    library(pointblank)
    library(devtools)

    i_am("README.md")
    tar_config_set(store = here("_targets"))

    load_all(quiet = TRUE)
  }
)

# load in any data we need from the targets workflow

tar_load(data_validation)
literature_merged_data <- tar_read(load_literature_pqt)
literature_measurements_data <- tar_read(measurements_data)
literature_reference_data <- tar_read(reference_data)
literature_campaign_data <- tar_read(campaign_data)
literature_sites_data <- tar_read(sites_data)
literature_biota_data <- tar_read(biota_data)
literature_qc <- tar_read(data_quality_report)
wgs84_geo <- tar_read(wgs84_geography)

# Action levels ----
# Warn if 5% of data fails
# Notify if 7.5% of data fails
# Stop if 10% of data fails
al <- action_levels(warn_at = 0.05, stop_at = 0.1, notify_at = 0.75)
```


# Data Quality Overview

```{r}
#| label: quality-summary
#| echo: true
#| results: asis

#' Generate a text summary of data quality issues
#'
#' @param qc_report Output from check_data_quality()
#' @return A character string with markdown-formatted summary
format_quality_summary <- function(qc_report) {
  s <- qc_report$summary

  glue(
    "
Dataset contains **{s$total_rows}** rows from **{s$total_references}** references across **{s$total_sites}** sites.

**Issues identified:**

- **Measurements**: {s$n_rows_missing_measurements} rows ({s$n_refs_missing_measurements} references) missing all measurement data (value, LOQ, and LOD)
- **Sample size**: {s$n_rows_missing_n} rows ({s$n_refs_missing_n} references) missing or zero sample size
- **Methods**: {s$n_rows_missing_methods} rows ({s$n_refs_missing_methods} references) missing method information
- **Uncertainty**: {s$n_rows_missing_uncertainty} rows ({s$n_refs_missing_uncertainty} references) missing uncertainty type
- **Site data**: {s$n_sites_missing_data} sites ({s$n_refs_missing_sites} references) missing location/geographic data
- **Biota data**: {s$n_rows_missing_biota} rows ({s$n_refs_missing_biota} references) missing species/tissue information
"
  )
}

cat(format_quality_summary(literature_qc))
```

## Missing Data

### Missing Measurements

Rows where measured value, LOQ, and LOD are all missing.

Known issues:

-   2025PelkonenEnvironmentalImpactOf does not report LOD/LOQ values in main paper or ESI.

```{r}
#| label: missing-measurements
#| echo: false

if (nrow(literature_qc$missing_measurements) > 0) {
  datatable(
    literature_qc$missing_measurements,
    # select(-sample_ids),
    options = list(pageLength = 10),
    colnames = c("Source", "Timestamp", "Rows"),
    caption = "References with missing measurement data"
  )
} else {
  cat("*No missing measurements found.*")
}
```

### Missing Sample Size

Rows where MEASURED_N is missing or zero.

```{r}
#| label: missing-n
#| echo: false

if (nrow(literature_qc$missing_n) > 0) {
  datatable(
    literature_qc$missing_n,
    # select(-sample_ids),
    options = list(pageLength = 10),
    colnames = c("Source", "Timestamp", "Rows"),
    caption = "References with missing sample size"
  )
} else {
  cat("*No missing sample sizes found.*")
}
```

### Missing Methods

Rows missing any of: analytical protocol, extraction protocol, fractionation protocol, or sampling protocol.

```{r}
#| label: missing-methods
#| echo: false

if (nrow(literature_qc$missing_methods) > 0) {
  display_df <- literature_qc$missing_methods

  datatable(
    display_df,
    options = list(pageLength = 10),
    caption = "References with missing method information",
    colnames = c("Source", "Timestamp", "Rows", "Missing Protocols"),
  )
} else {
  cat("*No missing methods found.*")
}
```

### Missing Uncertainty

Rows where UNCERTAINTY_TYPE or UNCERTAINTY\_\*\_STANDARD are missing.

```{r}
#| label: missing-uncertainty
#| echo: false

if (nrow(literature_qc$missing_uncertainty) > 0) {
  display_df <- literature_qc$missing_uncertainty |>
    mutate(
      reference = sub("_Measurements.*$", "...", source_file_measurements)
    ) |>
    select(reference, n_rows, missing_elements) |>
    arrange(desc(n_rows))

  datatable(
    display_df,
    options = list(pageLength = 10),
    caption = "References with missing or problematic uncertainty",
    colnames = c("Reference", "Rows", "Uncertainty Issues")
  )
} else {
  cat("*No uncertainty issues found.*")
}

```

### Missing Site Data

Sites missing coordinates, geographic features, country, or ocean information.

```{r}
#| label: missing-sites
#| echo: false

if (nrow(literature_qc$missing_sites) > 0) {
  display_df <- literature_qc$missing_sites |>
    mutate(
      reference = sub("_Measurements.*$", "...", source_file_measurements)
    ) |>
    select(reference, n_rows, SITE_CODE, missing_elements) |>
    arrange(desc(n_rows))

  datatable(
    display_df,
    options = list(pageLength = 10),
    caption = "Sites with missing geographic data",
    colnames = c(
      "Source",
      "Timestamp",
      "Rows",
      "SITE_CODE",
      "Missing Variables"
    )
  )
} else {
  cat("*No missing site data found.*")
}
```

### Missing Biota Data

Biota samples missing species, tissue, lifestage, gender, or species group.

```{r}
#| label: missing-biota
#| echo: false

if (nrow(literature_qc$missing_biota) > 0) {
  datatable(
    literature_qc$missing_biota,
    # select(-sample_ids),
    options = list(pageLength = 10),
    caption = "References with missing biota information",
    colnames = c(
      "Source",
      "Timestamp",
      "Rows",
      "Missing Variables",
      "SAMPLE_ID"
    )
  )
} else {
  cat("*No missing biota data found.*")
}
```

### Missing Reference Data

An earlier version of the data extraction app contained a bug that meant some measurements didn't have a `REFERENCE_ID` assigned properly. These are flagged below.

```{r}
#| label: missing-reference

literature_merged_data |>
  select(REFERENCE_ID, contains("source_file_", ignore.case = FALSE)) |>
  filter(is_missing(REFERENCE_ID) | REFERENCE_ID == "Unknown Reference") |>
  distinct() |>
  datatable()
```

# Pointblank QC

`pointblank` (@iannonePointblankDataValidation2025) is a Tidyverse (?) R package that supports fast, high-quality validation of data. Ideally it should be integrated into our `targets` workflow, but for the time being life's easier if we run it here. 

## Campaign

```{r}
#| label: pointblank-campaign
#| echo: true
data_validation$campaign

```


## References

```{r}
#| label: pointblank-references
#| echo: true
data_validation$reference

```

## Sites

```{r}
#| label: pointblank-sites
#| echo: true
data_validation$sites
```

## Parameters

```{r}
#| label: pointblank-parameters
#| echo: true
data_validation$parameters
```

```{r}
#| label: pointblank-compartments
#| echo: true

literature_compartments_data <- tar_read(compartments_data)

# Create schema ----
compartments_schema <- col_schema(
  initialise_compartments_tibble() |>
    mutate(
      source_file = NA_character_,
      read_timestamp = as.POSIXlt(NA_character_)
    )
)

# Validation agent ----
agent_compartments <-
  literature_compartments_data |>
  create_agent(
    label = "Check literature_compartments_data passes format conditions.",
    actions = al
  ) |>
  # Environmental compartments
  col_vals_not_null(columns = ENVIRON_COMPARTMENT) |>
  col_vals_in_set(
    columns = ENVIRON_COMPARTMENT,
    set = environ_compartments_vocabulary()
  ) |>
  col_vals_not_null(columns = ENVIRON_COMPARTMENT_SUB) |>
  col_vals_in_set(
    columns = ENVIRON_COMPARTMENT_SUB,
    set = environ_compartments_sub_vocabulary() |> purrr::list_flatten()
  ) |>
  col_vals_not_null(columns = MEASURED_CATEGORY)

interrogate(agent_compartments)
```

## Samples

```{r}
#| label: pointblank-biota
#| echo: true
data_validation$biota
```

## Measurements

```{r}
#| label: tbl-pointblank-measurements
data_validation$measurements

```

## CREED Scores

```{r}
#| label: pointblank-creed-scores
#| echo: true

literature_creed_scores_data <- tar_read(creed_scores_data)

# Create schema ----
creed_scores_schema <- col_schema(
  initialise_CREED_scores_tibble() |>
    mutate(
      source_file = NA_character_,
      read_timestamp = as.POSIXlt(NA_character_)
    )
)

literature_creed_scores_data |>
  group_by(
    SILVER_RELEVANCE,
    SILVER_RELIABILITY,
    GOLD_RELEVANCE,
    GOLD_RELIABILITY
  ) |>
  reframe(n = n())

CREED_classifications_rb <- c(
  "Not usable",
  "Reliable with restrictions",
  "Reliable without restrictions"
)
CREED_classifications_rv <- c(
  "Not usable",
  "Relevant with restrictions",
  "Relevant without restrictions"
)

# Validation agent ----
agent_creed_scores <-
  literature_creed_scores_data |>
  create_agent(
    label = "Check literature_creed_scores_data passes format conditions.",
    actions = al
  ) |>
  # Core identifiers
  col_vals_not_null(columns = REFERENCE_ID) |>
  col_vals_not_equal(columns = REFERENCE_ID, value = "Unknown Reference") |>

  # CREED fields
  col_vals_in_set(
    columns = c(SILVER_RELIABILITY, GOLD_RELIABILITY),
    set = CREED_classifications_rb
  ) |>
  col_vals_in_set(
    columns = c(SILVER_RELEVANCE, GOLD_RELEVANCE),
    set = CREED_classifications_rv
  )

interrogate(agent_creed_scores)
```

## CREED Data

Todo: Implement!

```{r}
#| label: pointblank-creed-data
#| echo: true

# literature_creed_data <- tar_read(creed_data)

# # Create schema ----
# creed_data_schema <- col_schema(
#   initialise_CREED_data_tibble() |>
#     mutate(
#       source_file = NA_character_,
#       read_timestamp = as.POSIXlt(NA_character_)
#     )
# )

# # Validation agent ----
# agent_creed_data <-
#   literature_creed_data |>
#   create_agent(
#     label = "Check literature_creed_data passes format conditions.",
#     actions = al
#   ) |>
#   # Core identifiers
#   col_vals_not_null(columns = criterion_id) |>
#   col_vals_not_null(columns = criterion_title) |>
#   col_vals_not_null(columns = required_recommended) |>
#   col_vals_not_null(columns = relevant_data) |>
#   col_vals_not_null(columns = score)

# interrogate(agent_creed_data)
```

## Joined Table