---
title: "Notebook 04 - Maps of Sample Sites, Emission Zones, etc."
author: "Sam Welch"
date: last-modified
lightbox: true
---

# Stupid map was taking 5 minutes just to render. Now it's cached. Will have to find a better solution in future. 

```{r}
#| label: setup
#| warning: false
#| output: false
library(dplyr)
library(leaflet)
library(sf)
library(bslib)
library(eDataDRF)
library(targets)
library(devtools)
library(leaflet.extras2)

suppressMessages(
  load_all(quiet = TRUE)
) # TODO: Reduce render time by sourcing only what we specifically need?
```

```{r}
#| label: load-data
# load in any data we need from the targets workflow
literature_merged_data <- tar_read(load_literature_pqt)
wgs84_geo <- tar_read(wgs84_geography)
tar_load(copper_toxicity_thresholds)


```

```{r}
#| label: themes-and-helpers

# Helper function for sample sizes
calculate_sample_sizes <- function(data) {
  data |>
    group_by(
      REFERENCE_ID,
      OCEAN_COUNTRY,
      SITE_GEOGRAPHIC_FEATURE,
      ENVIRON_COMPARTMENT_SUB
    ) |>
    summarise(total_n = sum(MEASURED_N, na.rm = TRUE), .groups = "drop")
}


```

## Geographic Distribution of Sampling Sites

```{r}
#| label: prepare-map-data
#| warning: false

# Turn off spherical geometry ----
sf::sf_use_s2(FALSE)

# Prepare spatial data ----
data_sf <- literature_merged_data |>
  filter(
    !is.na(LATITUDE),
    !is.na(LONGITUDE),
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  st_as_sf(
    coords = c("LONGITUDE", "LATITUDE"),
    crs = st_crs("WGS84"),
    remove = FALSE
  )

# Prepare apple data ----
number_of_sites <- nrow(municipality_centroids())
sales_data_split <- copper_oxide_sales() |>
  mutate(
    amount_kg = amount_kg / number_of_sites,
    average_kg = average_kg / number_of_sites
  )

apple_data <- cross_join(municipality_centroids(), sales_data_split) |>
  group_by(name, x, y) |>
  summarise(
    avg_copper_kg = mean(amount_kg, na.rm = TRUE),
    .groups = "drop"
  ) |>
  st_as_sf(
    coords = c("x", "y"),
    crs = st_crs("WGS84"),
    remove = FALSE
  )

environ_compartments_vocabulary <- environ_compartments_vocabulary() |>
  setdiff(c("Not relevant", "Not reported", "Other"))

# Create colour palette for main compartments ----
compartment_pal <- colorFactor(
  palette = palette.colors(n = 4, palette = "Tableau"),
  domain = environ_compartments_vocabulary
)

# Create colour palette for measured values ----
value_range <- range(data_sf$MEASURED_VALUE_STANDARD, na.rm = TRUE)
value_pal <- colorNumeric(
  palette = "viridis",
  domain = value_range,
  na.color = "gray"
)

# Add codes to data ----
data_sf <- data_sf |>
  mutate(
    SUBCOMP_CODE = subcomp_codes[ENVIRON_COMPARTMENT_SUB],
    SUBCOMP_CODE = ifelse(is.na(SUBCOMP_CODE), "?", SUBCOMP_CODE)
  )

# Define map colours ----
marine_colors <- list(
  default = "lightblue",
  highlight = "darkblue"
)

country_colours <- list(
  default = "lightgreen",
  highlight = "darkgreen"
)

# Calculate polygon centroids for labels ----
marine_centroids <- wgs84_geo$marine_polys |>
  filter(highlight_name) |>
  st_centroid()

country_centroids <- wgs84_geo$countries |>
  filter(
    highlight_name,
    !name %in% c("Greenland", "Iceland", "Norway")
  ) |>
  st_centroid()

# Create study connection lines ----
study_lines <- data_sf |>
  st_drop_geometry() |>
  group_by(REFERENCE_ID) |>
  filter(n() > 1, CAMPAIGN_NAME_SHORT != "Vm_2025") |> # simply too many sites in Vm_2025 for polylines to be practical
  summarise(
    coords = list(cbind(LONGITUDE, LATITUDE)),
    .groups = "drop"
  )

# Convert to linestring sf object ----
study_lines_sf <- study_lines |>
  rowwise() |>
  mutate(
    geometry = list(st_linestring(coords))
  ) |>
  ungroup() |>
  select(-coords) |>
  st_sf(crs = st_crs("WGS84"))
```

@fig-sampling-map shows the spatial distribution of all sampling locations included in the analysis. This map helps visualise the geographic scope of copper monitoring in Arctic regions and identify areas with high or low sampling intensity.

```{r}
#| label: fig-sampling-map
#| warning: false

# First, ensure leaflet.extras2 is loaded (add to setup chunk)
# library(leaflet.extras2)

# Prepare data for hexbin - needs coordinates and a value for intensity
hexbin_data <- data_sf |>
  st_drop_geometry() |>
  group_by(LATITUDE, LONGITUDE) |>
  summarise(
    sample_size = sum(MEASURED_N, na.rm = TRUE),
    .groups = "drop"
  ) |>
  filter(!is.na(LATITUDE), !is.na(LONGITUDE))

# Create the map with hexbins
leaflet() |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addHexbin(
    lng = hexbin_data$LONGITUDE,
    lat = hexbin_data$LATITUDE,
    radius = 5,
    options = hexbinOptions(
      colorRange = c("#440154", "#21918c", "#fde725"), # viridis-ish
      radiusRange = c(1, 5),
      colorScaleExtent = c(
        min(hexbin_data$sample_size, na.rm = TRUE),
        max(hexbin_data$sample_size, na.rm = TRUE)
      )
    )
  ) |>
  # Keep your existing polygon layers if desired

  # addPolygons(
  #   data = wgs84_geo$marine_polys,
  #   fillColor = ~ ifelse(
  #     highlight_name,
  #     marine_colors$highlight,
  #     marine_colors$default
  #   ),
  #   fillOpacity = 0.2,
  #   weight = 1,
  #   color = "darkblue"
  # ) |>
  # addPolygons(
  #   data = wgs84_geo$countries,
  #   fillColor = ~ ifelse(
  #     highlight_name,
  #     country_colours$highlight,
  #     country_colours$default
  #   ),
  #   fillOpacity = 0.2,
  #   weight = 1,
  #   color = "darkgreen"
  # )
```
