---
title: "More Emissions"
format: html
---

I guess we're doing emissions again!

```{r}
suppressMessages(
  {
    library(tibble)
    library(dplyr)
    library(ggplot2)
    library(here)
    library(devtools)
    library(sf)
    library(summarytools)
    library(DT)
    library(readxl)
    library(visNetwork)
    i_am("README.md")

    # load_all(quiet = TRUE)
  }
)
```

We have a variety of datasets for sources and emissions. Or, actually, two so far.

| Name | Grouping | Reporting Unit | Notes |
|----|----|----|----|
| REACH PTRS | By product,  ,NACE category | net tonnes | Commerically sensitive |
| Norske Utslippe | By industry type, license-holder, NUTS, etc. | kg to air/water/land | Some very unconvincing spikes/troughs |

```{r}
source("R/fct_read_reach.R")
source("R/fct_read_norske_utslipp.R")
```

So what we got? - REACH reported net values (total imported - (total exported)) by sector and product type - Emissions by sector and in some cases by license-holder

Rates are fairly constant (except for some outliers), so we remove these and take means.

```{r}
reach_by_sector_en_summarised |>
  mutate(across(where(is.numeric), \(x) signif(x, 3))) |>
  DT::datatable()
```

The copper emissions data is a bit suspect - we ought to take out the big unexplained jumps before averaging. I was z-testing before but that doesn't work here, I guess?

Mapping REACH to emissions is also tricky, because the categories aren't all that similar. We could drill down to the project-level data for this?

```{r}
source_to_sector_crosswalk <- tibble::tribble(
  ~source_category  , ~nace_section , ~sector_en                                    , ~confidence , ~notes                                                 ,
  "land_industry"   , "C"           , "Manufacturing"                               , "high"      , "Core industrial emissions to air/water"               ,
  "products"        , "C"           , "Manufacturing"                               , "medium"    , "Product-related Cu emissions; could also span A, G"   ,
  "transport"       , "H"           , "Transportation and storage"                  , "high"      , "Road runoff, brake wear, exhaust"                     ,
  "water_treatment" , "E"           , "Water supply, sewerage and waste management" , "high"      , "Direct match"                                         ,
  "landfill"        , "E"           , "Water supply, sewerage and waste management" , "medium"    , "Could also reflect F (Construction) waste streams"    ,
  "oil_marine"      , "B"           , "Mining and quarrying"                        , "medium"    , "Oil & gas extraction sits in NACE B; could also be H" ,
  "households"      , NA            , NA                                            , "low"       , "Diffuse source, no clean NACE equivalent"
)
```

```{r}
copper_emissions_by_source_summarised |>
  mutate(
    across(where(is.numeric), \(x) signif(x, 3)),
    medium = recode_values(
      medium,
      "vann" ~ "Aquatic",
      "jord" ~ "Terrestrial",
      "luft" ~ "Atmospherice"
    )
  ) |>
  DT::datatable()

copper_emissions_plot
```

```{r}
# Nodes -----------------------------------------------------------------------
# TODO: This is using the emissions categories rather than the NACE stuff
source_summary <- copper_emissions_by_source_summarised |>
  group_by(source_category) |>
  summarise(
    mean_kg = mean(mean_sum_emissions_kg, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

source_nodes <- tibble(
  name = unique(copper_emissions_by_source_summarised$source_category),
  node_type = "source",
  colour = "orange"
) |>
  left_join(source_summary, by = c("name" = "source_category")) |>
  mutate(
    label = paste0(name, "\n", signif(mean_kg, 3), " kg (n=", n, ")")
  )

medium_nodes <- tibble(
  name = unique(copper_emissions_by_source_summarised$medium),
  node_type = "medium",
  colour = case_when(
    name == "vann" ~ "blue",
    name == "jord" ~ "brown",
    name == "luft" ~ "lightblue"
  )
)

organism_nodes <- tibble(
  name = c("fish", "zooplankton", "algae", "shellfish"),
  node_type = "organism",
  colour = "green"
)

nodes <- bind_rows(source_nodes, medium_nodes, organism_nodes) |>
  mutate(
    id = name,
    label = if_else(is.na(label), name, label), # source nodes already have labels
    group = node_type,
    level = case_when(
      node_type == "source" ~ 1,
      node_type == "medium" ~ 2,
      node_type == "organism" ~ 3
    )
  )

lnodes <- data.frame(
  label = c("Source", "Medium", "Organism"),
  color = c("orange", "steelblue", "forestgreen"),
  shape = "box"
)

# Links -----------------------------------------------------------------------
emission_links <- copper_emissions_by_source_summarised |>
  mutate(
    from = source_category,
    to = medium,
    value = mean_sum_emissions_kg,
    edge_type = "emission"
  ) |>
  select(from, to, value, edge_type)

# Putative medium -> organism links
# Adjust as your domain knowledge dictates
exposure_links <- tribble(
  ~from  , ~to           , ~value , ~edge_type ,
  "vann" , "fish"        , NA     , "exposure" ,
  "vann" , "zooplankton" , NA     , "exposure" ,
  "vann" , "algae"       , NA     , "exposure" ,
  "vann" , "shellfish"   , NA     , "exposure"
)

links <- bind_rows(emission_links, exposure_links) |>
  mutate(
    dashes = edge_type == "exposure",
    width = case_when(
      edge_type == "emission" ~ log1p(replace_na(value, 1)) / 3,
      TRUE ~ 1
    ),
    label = case_when(
      edge_type == "emission" ~ paste0(signif(value, 3), " kg"),
      TRUE ~ NA_character_ # no label on putative exposure edges
    )
  )

ledges <- data.frame(
  label = c("Emission (measured)", "Exposure (putative)"),
  color = c("black", "black"),
  dashes = c(FALSE, TRUE),
  arrows = "to",
  font.align = "top"
)

# Plot ------------------------------------------------------------------------
visNetwork(nodes, links, width = "100%", height = "800px") |>
  visEdges(
    arrows = "to",
    smooth = list(type = "cubicBezier", forceDirection = "horizontal"),
    font = list(size = 12, align = "middle")
  ) |>
  visGroups(groupname = "source", color = "orange", shape = "box") |>
  visNodes(
    shape = "box",
    widthConstraint = list(minimum = 120, maximum = 120),
    heightConstraint = list(minimum = 40),
    margin = list(top = 10, bottom = 10, left = 10, right = 10)
  ) |>
  visGroups(groupname = "medium", color = "steelblue", shape = "box") |>
  visGroups(groupname = "organism", color = "forestgreen", shape = "box") |>
  visOptions(highlightNearest = TRUE) |>
  visLegend(addNodes = lnodes, addEdges = ledges, useGroups = FALSE) |>
  visLayout(hierarchical = TRUE) |>
  visHierarchicalLayout(
    direction = "LR",
    levelSeparation = 500, # vertical space between levels
    nodeSpacing = 150 # horizontal space between nodes at same level
  )
```