---
title: "Notebook 08 - Emissions Data"
author: "Sam Welch"
date: last-modified
---

*Currently a holding space for data on sources.*. 


# Appendices

## Appendix 1: Copper Sources Data

### Submarine Tailing Disposal in/off Norway

Note: STD locations (named) extracted from Kvassnes & Iversen 2013. Coordinates estimated by LLM, confidence self-reported by LLM.

```{r}
#| label: tbl-std-data

devtools::load_all()
library(tibble)
library(dplyr)
library(ggplot2)
library(here)
mine_tailings_coords() |> DT::datatable()

```

### Apple Production Areas in Norway

According to https://static02.nmbu.no/mina/studier/moppgaver/2018-Paulsen.pdf, 7 municipalities represent the main fruit districts in Norway. We also have sales of copper oxide for 2020-24 from https://mattilsynet-xp7prod.enonic.cloud/\_/attachment/inline/8311cd70-8877-4712-98cc-f7bc1dba46ac:0f491609e72f1d5f0dca5ad1030a4b979e8b4695/Omsetningsstatistikk%20for%20plantevernmidler%202020%20-%202024%20med%20uthevinger.pdf (table 2a, p4, thanks to Roger Holten).

N.B: Re Ronny Berdinsen, Mattilsynet (2025.12.15)

Hei

Tallene ble korrigerte pga en feilrapportering i 2022. Riktige tall er:

-	2022: 4332 liter (9482liter - 5150 liter)
-	2023: 2400 liter

Merk også at statistikken ikke viser plantevernmidler brukt pr år, men importert mengde. Det er ikke gitt at det som er importert i eksempelvis 2022 er brukt av bonden samme år.


```{r}
#| label: tbl-apple-emissions
number_of_sites <- nrow(municipality_centroids())
sales_data_split <- copper_oxide_sales() |>
  mutate(
    amount_kg = amount_kg / number_of_sites,
    average_kg = average_kg / number_of_sites
  )

cross_join(municipality_centroids(), sales_data_split) |>
  DT::datatable()

```

Although not explicitly stated, we assume that:

-   Copper products used as fungicides are exclusively applied in these areas and enter the environment via soil and freshwater
-   Apple farmers use the arthimetic mean of sold copper oxide for 2020 - 2024 every year. all applied copper oxide enters the environment.
-   The total used copper is split equally between the "main fruit districts" named in Paulsen (2018).

### Manufacturing Emissions of Copper in Norway

Taken from https://www.norskeutslipp.no/no/Komponenter/Utslipp/Kobber/?ComponentType=utslipp&ComponentPageID=73&SectorID=600, which includes self-reported copper emissions by land-based Norwegian industries for 1985-2024. Date is *very* spotty and incomplete, but it allows us to highlight fylke with higher emissions. Map to follow.

We assume:

-   Reported concentrations are fully correct
-   Everything ends up in the environment.

```{r}
#| label: fig-industrial-emissions
#| fig-width: 16

library(readxl)
library(dplyr)
library(esquisse)

copper_industrial_emissions <- read_excel(
  here::here(
    "data",
    "raw",
    "norske_utslipp",
    "norske_utslipp_copper_land_industries_individ.xlsx"
  ),
  skip = 2
) |>
  rename(
    facility_name = Anleggsnavn,
    county = Fylke,
    municipality = Kommune,
    nace_code = `NACE Kode`,
    year = År,
    annual_emission_air = `Årlig utslipp til luft`,
    basis_air = `Grunnlagsverdi for luft`,
    annual_emission_water = `Årlig utslipp til vann`,
    basis_water = `Grunnlagsverdi for vann`,
    unit = Enhet,
    facility_id = AnleggId,
    org_number = `Org.nr.`
  )

# lots of NAs?
copper_industrial_emissions_categorized <- copper_industrial_emissions |>
  mutate(
    missing_pattern = case_when(
      is.na(annual_emission_air) &
        is.na(annual_emission_water) ~ "Both missing",
      is.na(annual_emission_air) ~ "Air missing",
      is.na(annual_emission_water) ~ "Water missing",
      TRUE ~ "Complete data"
    )
  )

# Summarize by year and missing pattern
emissions_summary <- copper_industrial_emissions_categorized |>
  group_by(year, missing_pattern) |>
  reframe(count = n())

# Create the plot
ggplot(emissions_summary, aes(x = year, y = count, fill = missing_pattern)) +
  geom_col() +
  scale_fill_manual(
    values = c(
      "Complete data" = "#2E7D32",
      "Air missing" = "#FB8C00",
      "Water missing" = "#1976D2",
      "Both missing" = "#C62828"
    )
  ) +
  labs(
    title = "Industrial Copper Emissions Data Completeness Over Time",
    x = "Year",
    y = "Number of Records",
    fill = "Data Status"
  ) +
  theme_minimal()

copper_industrial_emissions_by_year_municipality <-
  copper_industrial_emissions |>
  group_by(municipality, year) |>
  reframe(
    emissions_kg_per_year_air = sum(annual_emission_air, na.rm = TRUE),
    emissions_kg_per_year_water = sum(annual_emission_water, na.rm = TRUE)
  )

# Get top 10 municipalities by total emissions (air + water combined)
top_municipalities <- copper_industrial_emissions |>
  group_by(municipality) |>
  summarise(
    total_emissions = sum(
      annual_emission_air,
      annual_emission_water,
      na.rm = TRUE
    )
  ) |>
  slice_max(total_emissions, n = 10) |>
  pull(municipality)

# Filter and reshape data for plotting
emissions_plot_data <- copper_industrial_emissions_by_year_municipality |>
  filter(municipality %in% top_municipalities) |>
  tidyr::pivot_longer(
    cols = starts_with("emissions_kg"),
    names_to = "emission_type",
    values_to = "kg_per_year"
  ) |>
  mutate(
    emission_type = case_when(
      emission_type == "emissions_kg_per_year_air" ~ "Air",
      emission_type == "emissions_kg_per_year_water" ~ "Water"
    )
  )

# Create faceted plot
top_municipalities_plot <- ggplot(
  emissions_plot_data,
  aes(x = year, y = kg_per_year, color = emission_type)
) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~municipality, scales = "free_y", ncol = 2) +
  scale_color_manual(values = c("Air" = "#FB8C00", "Water" = "#1976D2")) +
  labs(
    title = "Top 10 Municipalities: Copper Emissions Over Time",
    x = "Year",
    y = "Emissions (kg/year)",
    color = "Emission Type"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )


top_municipalities_plot
```

