---
title: "Notebook 08 - Emissions Data"
author: "Sam Welch"
date: last-modified
bibliography: "../references.bib"
---

*Currently a holding space for data on sources.*. Cache = true, fyi.

```{r}
#| label: setup

suppressMessages(
  {
    library(tibble)
    library(dplyr)
    library(ggplot2)
    library(here)
    library(devtools)
    library(sf)
    library(summarytools)
    library(DT)

    i_am("README.md")

    load_all(quiet = TRUE)
  }
)

```

# Copper Emissions & Source Data

## Closed Mines in Norway

Dataset available as a pdf here: https://www.vannportalen.no/miljotiltak/nedlagte-gruver/. Doesn't specify sources.

## Submarine Tailing Disposal in/off Norway

Note: STD locations (named) extracted from Kvassnes & Iversen 2013. Coordinates estimated by LLM, confidence self-reported by LLM. @kvassnesWasteSitesMines2013

```{r}
#| label: tbl-std-data

# mine_tailings_coords() |> DT::datatable()

```

## Apple Production Areas in Norway

According to https://static02.nmbu.no/mina/studier/moppgaver/2018-Paulsen.pdf, 7 municipalities represent the main fruit districts in Norway. We also have sales of copper oxide for 2020-24 from https://mattilsynet-xp7prod.enonic.cloud/\_/attachment/inline/8311cd70-8877-4712-98cc-f7bc1dba46ac:0f491609e72f1d5f0dca5ad1030a4b979e8b4695/Omsetningsstatistikk%20for%20plantevernmidler%202020%20-%202024%20med%20uthevinger.pdf (table 2a, p4, thanks to Roger Holten).

@mattilsynetOmsetningsstatistikkPlantevernmidler202020242025 - does the citation work now?

N.B: Re Ronny Berdinsen, Mattilsynet (2025.12.15)

Hei

Tallene ble korrigerte pga en feilrapportering i 2022. Riktige tall er:

-	2022: 4332 liter (9482liter - 5150 liter)
-	2023: 2400 liter

Merk også at statistikken ikke viser plantevernmidler brukt pr år, men importert mengde. Det er ikke gitt at det som er importert i eksempelvis 2022 er brukt av bonden samme år.


```{r}
#| label: tbl-apple-emissions
# number_of_sites <- nrow(municipality_centroids())
# sales_data_split <- copper_oxide_sales() |>
#   mutate(
#     amount_kg = amount_kg / number_of_sites,
#     average_kg = average_kg / number_of_sites
#   )

# cross_join(municipality_centroids(), sales_data_split) |>
#   DT::datatable()

```

Although not explicitly stated, we assume that:

-   Copper products used as fungicides are exclusively applied in these areas and enter the environment via soil and freshwater
-   Apple farmers use the arthimetic mean of sold copper oxide for 2020 - 2024 every year. all applied copper oxide enters the environment.
-   The total used copper is split equally between the "main fruit districts" named in Paulsen (2018).

## Manufacturing Emissions of Copper in Norway

Taken from https://www.norskeutslipp.no/no/Komponenter/Utslipp/Kobber/?ComponentType=utslipp&ComponentPageID=73&SectorID=600, which includes self-reported copper emissions by land-based Norwegian industries for 1985-2024. Date is *very* spotty and incomplete, but it allows us to highlight fylke with higher emissions. Map to follow.

@miljodirektoratetNorskeUtslippUtslipp2025

We assume:

-   Reported concentrations are fully correct
-   Everything ends up in the environment.

```{r}
#| label: fig-industrial-emissions
#| fig-width: 16

# library(readxl)
# library(dplyr)
# library(esquisse)

# copper_industrial_emissions <- read_excel(
#   here::here(
#     "data",
#     "raw",
#     "norske_utslipp",
#     "norske_utslipp_copper_land_industries_individ.xlsx"
#   ),
#   skip = 2
# ) |>
#   rename(
#     facility_name = Anleggsnavn,
#     county = Fylke,
#     municipality = Kommune,
#     nace_code = `NACE Kode`,
#     year = År,
#     annual_emission_air = `Årlig utslipp til luft`,
#     basis_air = `Grunnlagsverdi for luft`,
#     annual_emission_water = `Årlig utslipp til vann`,
#     basis_water = `Grunnlagsverdi for vann`,
#     unit = Enhet,
#     facility_id = AnleggId,
#     org_number = `Org.nr.`
#   )

# # lots of NAs?
# copper_industrial_emissions_categorised <- copper_industrial_emissions |>
#   mutate(
#     missing_pattern = case_when(
#       is.na(annual_emission_air) &
#         is.na(annual_emission_water) ~ "Both missing",
#       is.na(annual_emission_air) ~ "Air missing",
#       is.na(annual_emission_water) ~ "Water missing",
#       TRUE ~ "Complete data"
#     )
#   )

# # Summarize by year and missing pattern
# emissions_summary <- copper_industrial_emissions_categorised |>
#   group_by(year, missing_pattern) |>
#   reframe(count = n())

# # Create the plot
# ggplot(emissions_summary, aes(x = year, y = count, fill = missing_pattern)) +
#   geom_col() +
#   scale_fill_manual(
#     values = c(
#       "Complete data" = "#2E7D32",
#       "Air missing" = "#FB8C00",
#       "Water missing" = "#1976D2",
#       "Both missing" = "#C62828"
#     )
#   ) +
#   labs(
#     title = "Industrial Copper Emissions Data Completeness Over Time",
#     x = "Year",
#     y = "Number of Records",
#     fill = "Data Status"
#   ) +
#   theme_minimal()

# copper_industrial_emissions_by_year_municipality <-
#   copper_industrial_emissions |>
#   group_by(municipality, year) |>
#   reframe(
#     emissions_kg_per_year_air = sum(annual_emission_air, na.rm = TRUE),
#     emissions_kg_per_year_water = sum(annual_emission_water, na.rm = TRUE)
#   )

# # Get top 10 municipalities by total emissions (air + water combined)
# top_municipalities <- copper_industrial_emissions |>
#   group_by(municipality) |>
#   summarise(
#     total_emissions = sum(
#       annual_emission_air,
#       annual_emission_water,
#       na.rm = TRUE
#     )
#   ) |>
#   slice_max(total_emissions, n = 10) |>
#   pull(municipality)

# # Filter and reshape data for plotting
# emissions_plot_data <- copper_industrial_emissions_by_year_municipality |>
#   filter(municipality %in% top_municipalities) |>
#   tidyr::pivot_longer(
#     cols = starts_with("emissions_kg"),
#     names_to = "emission_type",
#     values_to = "kg_per_year"
#   ) |>
#   mutate(
#     emission_type = case_when(
#       emission_type == "emissions_kg_per_year_air" ~ "Air",
#       emission_type == "emissions_kg_per_year_water" ~ "Water"
#     )
#   )

# # Create faceted plot
# top_municipalities_plot <- ggplot(
#   emissions_plot_data,
#   aes(x = year, y = kg_per_year, color = emission_type)
# ) +
#   geom_line(linewidth = 0.8) +
#   facet_wrap(~municipality, scales = "free_y", ncol = 2) +
#   scale_color_manual(values = c("Air" = "#FB8C00", "Water" = "#1976D2")) +
#   labs(
#     title = "Top 10 Municipalities: Copper Emissions Over Time",
#     x = "Year",
#     y = "Emissions (kg/year)",
#     color = "Emission Type"
#   ) +
#   theme_minimal() +
#   theme(
#     strip.text = element_text(face = "bold"),
#     legend.position = "bottom"
#   )

# top_municipalities_plot
```

## Aquaculture Locations in Norway

Although I don't immediately have data for copper use in aquaculture in Norway (although this data probably exist?), the locations of licensed aquaculture sites are freely available from Fiskeridirektoratet as shapefiles (@norwegiandirectorateoffisheriesAquacultureLocationsNorway2025).

Actually the source ended up being https://portal.fiskeridir.no/portal/apps/webappviewer/index.html?id=bed4faeea84e4d6dbc548d43cc134c96&layer=Akvakulturregisteret, which is annoyingly detail-free.


```{r}
#| label: load-aquaculture-data
aquaculture <- sf::read_sf(here(
  "inst/extdata/shapefiles/Aquaculture_Locations/Fiskeridir_Akvakulturregistere.shx"
)) # that was easy

# View summary in browser
aquaculture |> sf::st_drop_geometry() |> dfSummary()

# Whatever, the data's a mess but we'll treat it as a starting point. We can do a simple map in any case.
```

This data's a mess but we'll treat it as a starting point. We can do a simple map in any case, using the same hex-bin approach as elsewhere.


```{r}
#| label: fig-sites-hex-map
#| fig-cap: All Fiskerdirektoriet licensed sites. Doesn't include number of animals per site, because they're reported in lots of different units and I haven't worked out how to convert them yet.
#| fig-width: 12

# Plot
norway_map <- map_data("world", region = "Norway(?!:Svalbard)")
ggplot() +
  geom_polygon(
    data = norway_map,
    aes(x = long, y = lat, group = group),
    fill = "lightgray",
    colour = "white"
  ) +
  geom_hex(
    data = aquaculture,
    aes(x = lon, y = lat),
    bins = 100,
    linewidth = 0
  ) +
  scale_fill_viridis_b() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() +
  theme(axis.text = element_blank(), axis.title = element_blank())
```

And an accompanying table, although here we aggregate by kommune, rather than binned based on spatial proximity.

```{r}
#| label: tbl-most-aquaculture

aquaculture |>
  group_by(kommune, fylke) |>
  reframe(n = n()) |>
  arrange(desc(n)) |>
  datatable()
```

Technically geotraces isn't emissions either, but nevermind. Unfortunately they provide their data in a weird geochemistry/oceanography format (ODV). https://geotraces.webodv.awi.de/IDP2025%3Eseawater%3EGEOTRACES_IDP2025_Seawater/service/DataExtraction


```{r}
#| label: geotraces
library(readr)
library(oce)
geotraces <- read.ctd.odv(
  here(
    "inst/extdata/exposure/Geotraces/IDP2025_seawater_GEOTRACES_IDP2025_Seawater_w1efqRDC.txt"
  )
)

```