---
title: "Notebook 08 - Emissions Data"
author: "Sam Welch"
date: last-modified
bibliography: "../references.bib"
---

*Currently a holding space for data on sources.*. Cache = true, fyi.

```{r}
#| label: setup
suppressMessages(
  {
    library(tibble)
    library(dplyr)
    library(ggplot2)
    library(here)
    library(devtools)
    library(sf)
    library(summarytools)
    library(DT)
    library(readxl)

    i_am("README.md")

    load_all(quiet = TRUE)
  }
)
```

# Copper Emissions & Source Data

## Closed Mines in Norway

Dataset available as a pdf here: https://www.vannportalen.no/miljotiltak/nedlagte-gruver/. Doesn't specify sources.

## Submarine Tailing Disposal in/off Norway

Note: STD locations (named) extracted from Kvassnes & Iversen 2013. Coordinates estimated by LLM, confidence self-reported by LLM. @kvassnesWasteSitesMines2013

```{r}
#| label: tbl-std-data
mine_tailings_coords() |> DT::datatable()
```

## Apple Production Areas in Norway

According to https://static02.nmbu.no/mina/studier/moppgaver/2018-Paulsen.pdf, 7 municipalities represent the main fruit districts in Norway. We also have sales of copper oxide for 2020-24 from https://mattilsynet-xp7prod.enonic.cloud/\_/attachment/inline/8311cd70-8877-4712-98cc-f7bc1dba46ac:0f491609e72f1d5f0dca5ad1030a4b979e8b4695/Omsetningsstatistikk%20for%20plantevernmidler%202020%20-%202024%20med%20uthevinger.pdf (table 2a, p4, thanks to Roger Holten).

@mattilsynetOmsetningsstatistikkPlantevernmidler202020242025 - does the citation work now?

N.B: Re Ronny Berdinsen, Mattilsynet (2025.12.15)

Hei

Tallene ble korrigerte pga en feilrapportering i 2022. Riktige tall er:

-	2022: 4332 liter (9482liter - 5150 liter)
-	2023: 2400 liter

Merk også at statistikken ikke viser plantevernmidler brukt pr år, men importert mengde. Det er ikke gitt at det som er importert i eksempelvis 2022 er brukt av bonden samme år.


```{r}
#| label: tbl-apple-emissions
# number_of_sites <- nrow(municipality_centroids())
# sales_data_split <- copper_oxide_sales() |>
#   mutate(
#     amount_kg = amount_kg / number_of_sites,
#     average_kg = average_kg / number_of_sites
#   )

# cross_join(municipality_centroids(), sales_data_split) |>
#   DT::datatable()

```

Although not explicitly stated, we assume that:

-   Copper products used as fungicides are exclusively applied in these areas and enter the environment via soil and freshwater
-   Apple farmers use the arthimetic mean of sold copper oxide for 2020 - 2024 every year. all applied copper oxide enters the environment.
-   The total used copper is split equally between the "main fruit districts" named in Paulsen (2018).

## Manufacturing Emissions of Copper in Norway

Taken from https://www.norskeutslipp.no/no/Komponenter/Utslipp/Kobber/?ComponentType=utslipp&ComponentPageID=73&SectorID=600, which includes self-reported copper emissions by land-based Norwegian industries for 1985-2024. Date is *very* spotty and incomplete, but it allows us to highlight fylke with higher emissions. Map to follow.

@miljodirektoratetNorskeUtslippUtslipp2025

We assume:

-   Reported concentrations are fully correct
-   Everything ends up in the environment.

```{r}
#| label: fig-industrial-emissions
#| fig-width: 16
# library(readxl)
# library(dplyr)
# library(esquisse)

copper_industrial_emissions <- read_excel(
  here::here(
    "data",
    "raw",
    "norske_utslipp",
    "norske_utslipp_copper_land_industries_individ.xlsx"
  ),
  skip = 2
) |>
  rename(
    facility_name = Anleggsnavn,
    county = Fylke,
    municipality = Kommune,
    nace_code = `NACE Kode`,
    year = År,
    annual_emission_air = `Årlig utslipp til luft`,
    basis_air = `Grunnlagsverdi for luft`,
    annual_emission_water = `Årlig utslipp til vann`,
    basis_water = `Grunnlagsverdi for vann`,
    unit = Enhet,
    facility_id = AnleggId,
    org_number = `Org.nr.`
  )

# lots of NAs?
copper_industrial_emissions_categorised <- copper_industrial_emissions |>
  mutate(
    missing_pattern = case_when(
      is.na(annual_emission_air) &
        is.na(annual_emission_water) ~ "Both missing",
      is.na(annual_emission_air) ~ "Air missing",
      is.na(annual_emission_water) ~ "Water missing",
      TRUE ~ "Complete data"
    )
  )

# Summarize by year and missing pattern
emissions_summary <- copper_industrial_emissions_categorised |>
  group_by(year, missing_pattern) |>
  reframe(count = n())

# Create the plot
ggplot(emissions_summary, aes(x = year, y = count, fill = missing_pattern)) +
  geom_col() +
  scale_fill_manual(
    values = c(
      "Complete data" = "#2E7D32",
      "Air missing" = "#FB8C00",
      "Water missing" = "#1976D2",
      "Both missing" = "#C62828"
    )
  ) +
  labs(
    title = "Industrial Copper Emissions Data Completeness Over Time",
    x = "Year",
    y = "Number of Records",
    fill = "Data Status"
  ) +
  theme_minimal()

copper_industrial_emissions_by_year_municipality <-
  copper_industrial_emissions |>
  group_by(municipality, year) |>
  reframe(
    emissions_kg_per_year_air = sum(annual_emission_air, na.rm = TRUE),
    emissions_kg_per_year_water = sum(annual_emission_water, na.rm = TRUE)
  )

# Get top 10 municipalities by total emissions (air + water combined)
top_municipalities <- copper_industrial_emissions |>
  group_by(municipality) |>
  summarise(
    total_emissions = sum(
      annual_emission_air,
      annual_emission_water,
      na.rm = TRUE
    )
  ) |>
  slice_max(total_emissions, n = 10) |>
  pull(municipality)

# Filter and reshape data for plotting
emissions_plot_data <- copper_industrial_emissions_by_year_municipality |>
  filter(municipality %in% top_municipalities) |>
  tidyr::pivot_longer(
    cols = starts_with("emissions_kg"),
    names_to = "emission_type",
    values_to = "kg_per_year"
  ) |>
  mutate(
    emission_type = case_when(
      emission_type == "emissions_kg_per_year_air" ~ "Air",
      emission_type == "emissions_kg_per_year_water" ~ "Water"
    )
  )

# Create faceted plot
top_municipalities_plot <- ggplot(
  emissions_plot_data,
  aes(x = year, y = kg_per_year, color = emission_type)
) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~municipality, scales = "free_y", ncol = 2) +
  scale_color_manual(values = c("Air" = "#FB8C00", "Water" = "#1976D2")) +
  labs(
    title = "Top 10 Municipalities: Copper Emissions Over Time",
    x = "Year",
    y = "Emissions (kg/year)",
    color = "Emission Type"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

top_municipalities_plot
```

## Aquaculture Locations in Norway

Although I don't immediately have data for copper use in aquaculture in Norway (although this data probably exist?), the locations of licensed aquaculture sites are freely available from Fiskeridirektoratet as shapefiles (@norwegiandirectorateoffisheriesAquacultureLocationsNorway2025).

Actually the source ended up being https://portal.fiskeridir.no/portal/apps/webappviewer/index.html?id=bed4faeea84e4d6dbc548d43cc134c96&layer=Akvakulturregisteret, which is annoyingly detail-free.


```{r}
#| label: load-aquaculture-data
aquaculture <- sf::read_sf(here(
  "inst/extdata/shapefiles/Aquaculture_Locations/Fiskeridir_Akvakulturregistere.shx"
)) # that was easy

# View summary in browser
# aquaculture |> sf::st_drop_geometry() |> dfSummary()

# Whatever, the data's a mess but we'll treat it as a starting point. We can do a simple map in any case.
```

This data's a mess but we'll treat it as a starting point. We can do a simple map in any case, using the same hex-bin approach as elsewhere.


```{r}
#| label: fig-sites-hex-map
#| fig-cap: All Fiskerdirektoriet licensed sites. Doesn't include number of animals per site, because they're reported in lots of different units and I haven't worked out how to convert them yet.
#| fig-width: 12
# Plot
norway_map <- map_data("world", region = "Norway(?!:Svalbard)")
ggplot() +
  geom_polygon(
    data = norway_map,
    aes(x = long, y = lat, group = group),
    fill = "lightgray",
    colour = "white"
  ) +
  geom_hex(
    data = aquaculture,
    aes(x = lon, y = lat),
    bins = 100,
    linewidth = 0
  ) +
  scale_fill_viridis_b() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() +
  theme(axis.text = element_blank(), axis.title = element_blank())
```

And an accompanying table, although here we aggregate by kommune, rather than binned based on spatial proximity.

```{r}
#| label: tbl-most-aquaculture
aquaculture |>
  group_by(kommune, fylke) |>
  reframe(n = n()) |>
  arrange(desc(n)) |>
  datatable()
```

Technically geotraces isn't emissions either, but nevermind. Unfortunately they provide their data in a weird geochemistry/oceanography format (ODV). https://geotraces.webodv.awi.de/IDP2025%3Eseawater%3EGEOTRACES_IDP2025_Seawater/service/DataExtraction


```{r}
#| label: geotraces
library(readr)
library(oce)
library(here)
geotraces <- read.ctd.odv(
  here(
    "inst/extdata/exposure/Geotraces/IDP2025_seawater_GEOTRACES_IDP2025_Seawater_w1efqRDC.txt"
  )
)

show(geotraces)

library(maps)

world_map <- map_data("world")
library(sf)
geotraces_tibble <- as_tibble(geotraces@data, .name_repair = "universal") |>
  mutate(lon = longitude, lat = latitude) |>
  st_as_sf(coords = c("lon", "lat")) |>
  mutate(Cu_D_nmol_kg = `Cu_D_CONC..nmol.kg.`)


geotraces_tibble_grouped <- geotraces_tibble |>
  group_by(longitude, latitude) |>
  reframe(mean_conc = mean(Cu_D_nmol_kg, na.rm = TRUE)) |>
  filter(!is.na(mean_conc))
#  |>
# st_as_sf(coords = c("longitude", "latitude"))

# desnity mapping: https://www.andrewheiss.com/blog/2023/07/28/gradient-map-fills-r-sf/
# actually, this might not be it because we don't want density, we want to visualise the mean concentration
# what I want is a statistically valid way to assume concentration gradients from point data

# claude suggests kriging:
# 2. Kriging (Gaussian Process interpolation)
# More statistically principled — models spatial autocorrelation explicitly via a variogram. Gives you uncertainty estimates too, which is valuable for saying "concentration here is X ± Y". This is probably closest to what you want for making defensible statements.

# This might be overkill

ggplot() +
  geom_polygon(
    data = world_map,
    aes(x = long, y = lat, group = group),
    fill = "lightgray",
    colour = "white"
  ) +
  geom_hex(
    data = geotraces_tibble_grouped,
    aes(x = longitude, y = latitude),
    bins = 50
  ) +
  scale_fill_viridis_c() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() +
  theme(axis.text = element_blank(), axis.title = element_blank())
```

REACH/PTPR data from Vannmiljø. TODO: Check permissions before publication.


```{r}
library(tidyverse)
library(readxl)

reach_data_by_product <- read_excel(
  here("data/raw/reach/Sum kobberstoffer i produktregisteret.xlsx"),
  sheet = 4,
  skip = 1,
  col_names = c("Year", "Weight_tonnes", "Product_type"),
  col_types = c("numeric", "numeric", "text")
) |>
  replace_na(list(Product_type = "Not Reported"))

# this is incomplete
prod_name_translations <- tribble(
  ~Product_type                                                         , ~Product_type_en                    ,
  "BRENSELTILSETNINGER"                                                 , "Fuel Additives"                    ,
  "Absorpsjons/ og adsorpsjonsmaterialer"                               , "Absorption/Adsorption Materials"   ,
  "ANDRE SMØREMIDLER"                                                   , "Other Lubricants"                  ,
  "ARMERINGSMIDLER"                                                     , "Reinforcing Agents"                ,
  "Bindemidler"                                                         , "Binders"                           ,
  "Biocider"                                                            , "Biocides"                          ,
  "Borekjemikalier inkl råolje/gass"                                    , "Drilling Chemicals"                ,
  "BOREOLJER"                                                           , "Drilling Oils"                     ,
  "BUNNFELLINGSHINDRENDE MIDLER, GENERELT"                              , "Anti-Settling Agents"              ,
  "Fyllingsmidler"                                                      , "Fillers"                           ,
  "GJØDNING, GENERELT"                                                  , "Fertilisers"                       ,
  "Impregnering"                                                        , "Impregnation"                      ,
  "INSEKTSMIDDEL, INSEKTMIDLER OG ANDRE MIDLER MOT SKADEDYR PÅ PLANTER" , "Insecticides & Plant Pest Control" ,
  "Konstruksjonsmaterialer"                                             , "Construction Materials"            ,
  "Lim"                                                                 , "Adhesives"                         ,
  "Maling"                                                              , "Paint"                             ,
  "PH-REGULERENDE MIDLER, GENERELT"                                     , "pH Regulating Agents"              ,
  "Prosessregulerendemidler"                                            , "Process Control Agents"            ,
  "Rengjøring"                                                          , "Cleaning"                          ,
  "Rustbeskyttelse"                                                     , "Rust Protection"                   ,
  "SALT TIL GALVANISKE BAD"                                             , "Galvanic Bath Salts"               ,
  "Sprengstoff"                                                         , "Explosives"                        ,
  "STØPEMASSER, GENERELT"                                               , "Casting Compounds"
)

reach_data_by_industry <- read_excel(
  here("data/raw/reach/Sum kobberstoffer i produktregisteret.xlsx"),
  sheet = 3,
  skip = 1,
  col_names = c("Year", "Weight_tonnes", "Industry"),
  col_types = c("numeric", "numeric", "text")
) |>
  replace_na(list(Industry = "Not Reported"))

reach_data_by_product_grouped <- reach_data_by_product |>
  arrange(Weight_tonnes) |>
  reframe(
    Year,
    group = case_when(
      Weight_tonnes < quantile(Weight_tonnes, 0.9) ~ "Other",
      .default = Product_type
    ),
    Weight_tonnes
  ) |>
  group_by(Year, group) |>
  reframe(Weight_tonnes = sum(Weight_tonnes)) |>
  left_join(prod_name_translations, by = c(group = "Product_type"))

# ggplot(
#   data = reach_data_by_application,
#   mapping = aes(x = Year, y = Weight_tonnes, fill = Industry)
# ) +
#   geom_col(position = "stack") +
#   theme(axis.text.x = element_text(angle = 90))

ggplot(
  data = reach_data_by_product_grouped,
  mapping = aes(x = Year, y = Weight_tonnes, fill = Product_type_en)
) +
  geom_col(position = "stack") +
  theme(axis.text.x = element_text(angle = 90))
```

