---
title: "Notebook 08 - Ecology and Biota"
author: "Sam Welch"
date: last-modified
---

*A holding space for relevant data on biota diets and ADME*.

# To Do

- [ ] Cross reference with our biota data and figure out what we're missing

```{r}
#| label: setup

suppressMessages(
  {
    library(DT)
    library(dplyr)
    library(readxl)
    library(eDataDRF)
    library(targets)
    library(here)
    library(glue)
    library(ggplot2)
    library(devtools)
    library(DT)
    library(readr)
    library(taxize)

    i_am("README.md")
    tar_config_set(store = here("_targets"))

    load_all()
  }
)

literature_biota_data <- tar_read(API_biota_common_names)

set_theme(theme_minimal())

```

```{r}
#| label: literature_species

species_names <- species_names_vocabulary()

biota_data_summary <- literature_biota_data |>
  group_by(SAMPLE_SPECIES, SPECIES_COMMON_NAME) |>
  reframe(n = n()) |>
  arrange(desc(n))

# check missing common names
biota_data_summary |> filter(is.na(SPECIES_COMMON_NAME)) |> datatable()

```

In an idea world, we can find an already existing dataset with rough trophic levels (or better yet relationships) for most or all of our species. This probably doesn't exist, although FishBase seems to have pretty good data for what various species of fish eat. 

This (@mackenzieTrophicLevelsSimple2023) has TL and functional groups for ~800 Barents species

```{r}

BarentsEcologyTable <- read_excel(
  "data/raw/ecology/TL_FG_final_v2.xlsx",
  sheet = 2
) |>
  mutate(TL = na_if(TL, "NA") |> as.numeric())

BarentsEcologyReadme <- read_excel(
  "data/raw/ecology/TL_FG_final_v2.xlsx",
  sheet = 1
)

BarentsEcologyLookup <- read_excel(
  "data/raw/ecology/TL_FG_final_v2.xlsx",
  sheet = 1,
  range = "A23:B43",
  col_names = c("Code", "Functional_Group")
) |>
  filter(!is.na(Code))

BarentsEcologyClean <- BarentsEcologyTable |>
  left_join(BarentsEcologyLookup, by = c("FG" = "Code")) |>
  select(Latin, RecordName, TL, Group, Functional_Group)

BarentsEcologyClean |>
  ggplot(aes(y = TL, x = Functional_Group, colour = Functional_Group)) +
  geom_point() +
  theme(legend.position = "none")

```

```{r}
#| label: tbl-study-species-with-trophic-data

species_have_trophic <- biota_data_summary |>
  left_join(BarentsEcologyClean, by = c("SAMPLE_SPECIES" = "Latin")) |>
  filter(!is.na(Functional_Group)) |>
  nrow()

message(glue(
  "Trophic data available for {species_have_trophic} of {biota_data_summary} species."
))
```


The black-legged kittiwake is a small pelagic gull with a global population of 6–8 million pairs of which 120000 pairs breed on the west coast of Spitsbergen, Svalbard (Frederiksen et al., 2012). The kittiwakes breeding on Svalbard migrate towards Greenland and Canada in the end of October, overwintering in a widespread area in the North Atlantic between November and April (Frederiksen et al., 2012, Schultner et al., 2014). Kittiwakes have a varied diet, preying on small pelagic fish like polar cod (Boreogadus saida) and capelin (Mallotus villosus), as well as pelagic amphipods (Themisto spp.) and euphausiids (Thysanoessa spp.) (Mehlum and Gabrielsen, 1993). Other small fish species (15–20 cm), polychaetes and pteropods are also frequent in the diet of kittiwakes (Hatch, 2013). Seasonal and spatial variations in kittiwake diet occur (Anker-Nilssen et al., 2000, Lønne and Gabrielsen, 1992). During the breeding season, adult kittiwakes in Kongsfjorden feed mainly on polar cod and krill (Thysanoessa inermis) (Mehlum and Gabrielsen, 1993).

https://www.sciencedirect.com/science/article/pii/S0048969714013801?via=ihub#s0010

https://www.sciencedirect.com/science/article/pii/S0048969705008351#fig4

https://datadryad.org/dataset/doi:10.5061/dryad.4dq31c5


```{r}
#| label: phylopic-just-because
# library(rphylopic)

# # Get a single image uuid for a species
# uuid <- get_uuid(name = "Canis lupus", n = 1)
# # Get the image for that uuid
# img <- get_phylopic(uuid = uuid)
# # But multiple silhouettes can exist per species...
# uuid <- get_uuid(name = "Canis lupus", n = 5)

# img <- pick_phylopic(name = "Canis lupus", n = 5)

# # All of this functionality is available for ggplot2 as well...
# # But we use add_phylopic and geom_phylopic instead!
# library(ggplot2)
# # Get image
# uuid <- get_uuid(name = "Iris", n = 1)
# img <- get_phylopic(uuid = uuid)
# # Plot silhouettes as points!
# ggplot(iris) +
#   geom_phylopic(
#     aes(x = Sepal.Length, y = Sepal.Width),
#     img = img,
#     color = "purple",
#     height = 0.25
# )
```