---
title: "Fish"
execute:
  output: true
  echo: false
  warning: false
format: 
  html: 
    fig-width: 16
    fig-height: 20
    dpi: 300 
  docx: 
    fig-width: 16
    fig-height: 20
    reference-doc: ../custom-reference-doc.docx
    dpi: 300
---

```{r}
#| label: setup
# == Libraries ----
library(tidyverse)
library(targets)
library(hexbin)
library(ggrepel)
library(viridis)
library(patchwork)
library(ggnewscale)
library(igraph)
library(sf)
library(gridExtra)
library(grid)
library(knitr)

opts_chunk$set(dev = "svglite")

pkgload::load_all(quiet = TRUE)

# == Shared map data ----
world_map <- map_data("world")

# == Study area bbox ----
# xmin/xmax: longitude; ymin/ymax: latitude
STUDY_BBOX <- c(xmin = -65, xmax = 40, ymin = 55, ymax = 75)

# Helper: append coord_sf with fixed study area to a ggplot object.
# ratio = 2 approximates equal-area appearance at these latitudes.
study_coords <- function(p, ratio = 2) {
  coord_sf(
    xlim = c(STUDY_BBOX[["xmin"]], STUDY_BBOX[["xmax"]]),
    ylim = c(STUDY_BBOX[["ymin"]], STUDY_BBOX[["ymax"]]),
    expand = FALSE
  )
}

# == Outlier filter ----
# Z-score threshold for excluding extreme values.
# Applied locally per matrix using MEASURED_VALUE_STANDARD.
# MEASURED_VALUE is retained in the data for traceability only; never used for analysis.
Z_THRESHOLD <- 3

is_outlier <- function(x, threshold = Z_THRESHOLD) {
  abs((x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)) > threshold
}

# == Site comment cleaning patterns ----
# NOTE: \p{Lu} requires Unicode-aware regex; stringr::regex() handles this correctly.
remove_patterns <- c(
  # URLs / emails
  "(?i)https?://\\S+|www\\.\\S+|\\S+@\\S+",
  # Parenthetical / bracket content
  "\\s*\\([^)]*\\)|\\s*\\[[^]]*\\]",
  # Report codes e.g. 01.01-385, 2021/09
  "\\b\\d{1,4}(?:[\\.\\-/]\\d{1,4})+\\b",
  # Station / sample codes e.g. L1, L2, Stasjon 3
  "\\bL\\d+\\b|\\bStasjon\\s*\\d+\\b|\\bStation\\s*\\d+\\b",
  # Short numeric tokens (excludes 4-digit years)
  "\\b(?!\\d{4}\\b)\\d{1,3}\\b",
  # Report words (Norwegian / English)
  "(?i)rapport(?:\\s+nr\\.?|nr\\.?|nummer|:)?.*?\\b|(?i)report(?:\\s+no\\.?|\\s+nr\\.?|\\s+nr)?",
  "(?i)rapporter?\\b|(?i)reference(?:\\s+site)?\\b|\\bref\\.?\\b",
  # Vannmiljø / resipient noise
  "(?i)vannmilj[oø]\\b|\\bvm\\b|(?i)resipient(?:unders[kq]økelse|unde)?",
  "(?i)vm\\s*emission\\s*|(?i)miljøgifter\\b|(?i)miljøgifter\\s+i",
  # Contractor / org names
  "(?i)akvaplan-?niva|niva\\b|nina\\b|ramb[ao]ll\\b|norconsult\\b|asplan\\s+viak\\b|ngi\\b|ng\\&i\\b|APN|AS",
  "\\b\\p{Lu}[\\p{L}''\\-]+(?:\\s+\\p{Lu}[\\p{L}''\\-]+)*\\s+AS\\b",
  # Page / folio refs
  "\\bpg?\\.?\\s*\\d+\\b|\\bp\\.\\s*\\d+\\b",
  # Unit / measurement fragments
  "\\bmg/kg\\b|\\bmg/kg\\s*dry\\b|\\bmg/kg\\s*wet\\b|\\bSD\\b|\\bStandard\\s+Deviat(?:ion|ions)\\b",
  # Stray punctuation
  "[,;:\\|]{2,}|--+|\\.\\.+",
  # Literal noise strings
  "Original Comment:",
  "Emission :",
  "rsøkelse",
  "Vannmiljø Station",
  "Station"
)

clean_regex <- str_c(remove_patterns, collapse = "|")
```

As for algae, but for fish. There are a lot more or these, and there's a lot more tissue differentiation.

# Fish

```{r}
#| label: tbl-fish-variation-summary
fish <- tar_read(load_literature_pqt) |>
  filter(SPECIES_GROUP == "Fish")

fish |>
  group_by(
    SAMPLE_SPECIES,
    REFERENCE_ID,
    SITE_GEOGRAPHIC_FEATURE,
    SAMPLE_TISSUE
  ) |>
  reframe(n = n()) |>
  arrange(desc(n)) |>
  kable()
```

We have a great deal of fish data. Broadly speaking, we can group it by three factors: - Tissue (liver, gill, whole body, muscle, and in one case bile) - Site type (water ways, water bodies, coastal waters, and rarely, open ocean) - Species

It would also be useful to know the rough size, trophic level, lifestyle, etc. of these fish. We will attend to that later, though.

# Distribution of Copper Concentrations

No removal of outliers at this stage. We do need to check for unit types, though.

Two important questions:

1.  What units are used?
2.  Are they correlated with anything?

```{r}
#| label: tbl_fish_units
fish_units_exploration <- fish |>
  reframe(
    .by = c(
      'MEASURED_UNIT_STANDARD',
      "SAMPLE_SPECIES",
      "SAMPLE_TISSUE",
      "SITE_GEOGRAPHIC_FEATURE"
    ),
    mean = mean(MEASURED_VALUE_STANDARD),
    sd = sd(MEASURED_VALUE_STANDARD),
    n = sum(MEASURED_N)
  ) |>
  arrange(desc(n))

fish_units_exploration
```

For a subset of species/tissues/locations we have both dry and wet weights and can calculate some (rough) conversion factors.

```{r}
#| label: tbl_fish_units_pairwise
fish_units_lookup <- fish_units_exploration |>
  group_by(SAMPLE_SPECIES, SAMPLE_TISSUE, SITE_GEOGRAPHIC_FEATURE) |>
  filter(n() >= 2) |>
  pivot_wider(names_from = MEASURED_UNIT_STANDARD, values_from = mean) |>
  reframe(
    wet_over_dry = max(`mg/kg (wet)`, na.rm = TRUE) /
      max(`mg/kg (dry)`, na.rm = TRUE),
    n = sum(n)
  ) |>
  add_row(
    SAMPLE_SPECIES = "Generic",
    SAMPLE_TISSUE = "Generic",
    SITE_GEOGRAPHIC_FEATURE = "Generic",
    wet_over_dry = 5,
    n = NA
  )

fish_units_lookup
```

Roughly 10% of our samples (537) are dry, the remainder (3965) are wet weight. According to @cressonUnderestimationChemicalContamination2017, a conversion factor of 1:5 for dry:wet weight is typically used, but individual species vary considerably from this. Under the circumstances, we will use the calculated conversion factors for *G. morhua*, *Melanogrammus aeglefinus*, and *Pollachius pollachius*, and the default figure of 5 (despite its obvious flaws) for the rest.

```{r}
#| label: fig-fish-pointrange
threshold_data <- generate_copper_thresholds() |>
  filter(SAMPLE_SPECIES == "Gadus morhua", SAMPLE_TISSUE == "Liver") # only one background threshold; for G. morhua livers

fish_dry_normalised <- fish |>
  mutate(
    # manual, since we have so few values
    SPECIES_DRY_WET_RATIO = case_when(
      SAMPLE_SPECIES == "Gadus morhua" & SAMPLE_TISSUE == "Liver" ~ 10.9,
      SAMPLE_SPECIES == "Melanogrammus aeglefinus" &
        SAMPLE_TISSUE == "Liver" ~ 1.11,
      SAMPLE_SPECIES == "Pollachius pollachius" &
        SAMPLE_TISSUE == "Liver" ~ 1.34,
      .default = 5
    ),
    # when unit is dry, multiply by the relevant ratio, otherwise leave untouched
    MEASURED_VALUE_STANDARD_NORMALISED = case_when(
      MEASURED_UNIT_STANDARD == "mg/kg (dry)" ~ MEASURED_VALUE_STANDARD *
        SPECIES_DRY_WET_RATIO,
      .default = MEASURED_VALUE_STANDARD
    ),
    UNCERTAINTY_UPPER_STANDARD_NORMALISED = case_when(
      MEASURED_UNIT_STANDARD == "mg/kg (dry)" ~ UNCERTAINTY_UPPER_STANDARD *
        SPECIES_DRY_WET_RATIO,
      .default = UNCERTAINTY_UPPER_STANDARD
    ),
    UNCERTAINTY_LOWER_STANDARD_NORMALISED = case_when(
      MEASURED_UNIT_STANDARD == "mg/kg (dry)" ~ UNCERTAINTY_LOWER_STANDARD *
        SPECIES_DRY_WET_RATIO,
      .default = SPECIES_DRY_WET_RATIO
    ),
    MEASURED_UNIT_STANDARD_NORMALISED = "mg/kg (wet)"
  )

nrow(fish)
nrow(fish_dry_normalised)

fish_pointrange <- fish_dry_normalised |>
  mutate(
    SITE_COMMENT = SITE_COMMENT |> str_remove("^.*?\\."),
    SITE_NAME = SITE_NAME |> str_remove("Vannmiljø Station")
  ) |>
  ggplot(aes(
    y = SITE_NAME,
    x = MEASURED_VALUE_STANDARD_NORMALISED,
    colour = MEASURED_VALUE_STANDARD_NORMALISED
  )) +
  scale_colour_viridis_c() +
  geom_pointrange(aes(
    xmin = UNCERTAINTY_LOWER_STANDARD_NORMALISED,
    xmax = UNCERTAINTY_UPPER_STANDARD_NORMALISED
  )) +
  new_scale_color() +
  geom_vline(
    data = threshold_data,
    aes(xintercept = THRESHOLD_VALUE, color = TITLE_SHORT),
    linetype = "dashed",
    linewidth = 0.8,
    alpha = 0.6
  ) +
  geom_text(
    aes(
      label = str_wrap(SITE_COMMENT),
      x = UNCERTAINTY_UPPER_STANDARD_NORMALISED + 0.5
    ),
    colour = "black",
    hjust = 0
  ) +
  labs(
    y = "Site Code",
    x = "Measured Value (standardised to mg/kg (wet))"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

fish_boxplot <- fish_dry_normalised |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD_NORMALISED, y = SAMPLE_SPECIES)) +
  geom_boxplot(outlier.colour = "red") +
  scale_x_log10() +
  facet_wrap(
    facets = vars(SAMPLE_TISSUE),
    ncol = 1,
    scales = "free_y",
    space = "free_y"
  )

fish_pointrange
```

We have some improbably high values in some cases, all from Vannmiljø. Let's take a closer look:

```{r}
#| label: fig-fish-outliers
fish_outliers <- fish_dry_normalised |>
  filter(is_outlier(MEASURED_VALUE_STANDARD_NORMALISED)) # Z > 3

fish_outliers |>
  mutate(
    SITE_COMMENT = SITE_COMMENT |> str_remove("^.*?\\."),
    SITE_NAME = SITE_NAME |> str_remove("Vannmiljø Station")
  ) |>
  ggplot(aes(
    y = SITE_NAME,
    x = MEASURED_VALUE_STANDARD,
    colour = MEASURED_VALUE_STANDARD
  )) +
  scale_colour_viridis_c() +
  geom_pointrange(aes(
    xmin = UNCERTAINTY_LOWER,
    xmax = UNCERTAINTY_UPPER,
    shape = SAMPLE_TISSUE
  )) +
  new_scale_color() +
  geom_vline(
    data = threshold_data,
    aes(xintercept = THRESHOLD_VALUE, color = TITLE_SHORT),
    linetype = "dashed",
    linewidth = 0.8,
    alpha = 0.6
  ) +
  geom_text(
    aes(
      label = str_wrap(str_glue("{SAMPLE_SPECIES}, {SAMPLE_TISSUE}")),
      x = 40000
    ),
    colour = "black",
    hjust = 0
  ) +
  labs(y = "Site Code", x = "Measured Value (mg/kg (wet))") +
  facet_wrap(
    facets = vars(MEASURED_UNIT_STANDARD),
    ncol = 1,
    scales = "free_y",
    space = "free_y"
  ) +
  theme_minimal()

fish_filtered <- fish_dry_normalised |>
  filter(!is_outlier(MEASURED_VALUE_STANDARD))

nrow(fish_filtered)
```

10,000 mg of copper is 10 g copper per kg of liver. The most likely explanation here is that values were mistakenly entered with the wrong unit. However, since only a small proportion (17 rows) of data points show such irregularly high values, the most pragmatic option is to simply exclude them.

# Lumping and Splitting

We have a lot of fish data, and a lot of different species and tissues. However, data are significantly biased towards a small number of species.

```{r}
#| label: tbl-fish-species-summary
fish_species_n <- fish_filtered |>
  group_by(
    SAMPLE_SPECIES
  ) |>
  reframe(n = n()) |>
  arrange(desc(n))

fish_species_n_sum <- sum(fish_species_n$n)

fish_species_n <- fish_species_n |>
  mutate(prop = n / fish_species_n_sum)
```

From a trophic perspective we'd like to have species that represent different parts of the food web (although we wouldn't expect this to be important for copper). In practical terms, 72% of our dataset is *Gadus morhua* (Atlantic cod), while *Salmo trutta* (Brown trout), *Ctenolabrus rupestris* (Goldsinny-wrasse), and *Coregonus lavaretus* (Common whitefish) represent 3-5% of the total each. We collect some information on each species to determine their relevance to our case.

+--------------------------------------------+-------------------------+----------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------+------------------------------------------------------------------------+------------------------------------------------------------------------------+
| Species Name                               | n / proportion of total | Distribution                                                                                             | Trophic Level                                                                              | Notes                                                                  | Source                                                                       |
+============================================+=========================+==========================================================================================================+============================================================================================+========================================================================+==============================================================================+
| *Gadus morhua* (Atlantic cod)              | 3146 (72%)              | High occurrence above Arctic circle. Found from shoreline to edge of continental slope.                  | Predators of small fish and inverts, consumed by larger fish, aquatic mammals, squid       | IUCN Red List Vulnerable. Important and overfished commercial species. | <https://www.marinespecies.org/aphia.php?p=taxdetails&id=126436>\            |
|                                            |                         |                                                                                                          |                                                                                            |                                                                        | <https://www.fishbase.org/summary/Gadus_morhua.html>                         |
+--------------------------------------------+-------------------------+----------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------+------------------------------------------------------------------------+------------------------------------------------------------------------------+
| *Salmo trutta* (Brown or Sea trout)        | 204 (5%)                | Found across Norwegian & Icelandic coasts, including above Arctic circle.. Freshwater/coastal lifecycle. | Larger (#72cm) predator of small fish and inverts. Consumed by some larger fish and seals. | IUCN Least Concern. Widely eaten.                                      | <https://www.marinespecies.org/aphia.php?p=taxdetails&id=127187#attributes>\ |
|                                            |                         |                                                                                                          |                                                                                            |                                                                        | <https://www.fishbase.org/summary/Salmo_trutta.html>                         |
+--------------------------------------------+-------------------------+----------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------+------------------------------------------------------------------------+------------------------------------------------------------------------------+
| *Ctenolabrus rupestris* (Goldsinny-wrasse) | 192 (4%)                | Present in Norwegian territorial waters, including above Arctic circle.                                  | Smaller (\>=18cm) predator of invertebrates.                                               | IUCN Least Concern. Used as a cleaner fish in salmon culture.          | <https://www.marinespecies.org/aphia.php?p=taxdetails&id=126964>\            |
|                                            |                         |                                                                                                          |                                                                                            |                                                                        | <https://www.fishbase.org/summary/Ctenolabrus_rupestris.html>                |
+--------------------------------------------+-------------------------+----------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------+------------------------------------------------------------------------+------------------------------------------------------------------------------+
| *Coregonus lavaretus* (Common whitefish)   | 139 (3%)                | Freshwater sites across Norway, including above Arctic circle                                            | Medium (#40cm) predator of planktonic and benthic crustaceans.                             | IUCN Vulnerable\                                                       | <https://artsdatabanken.no/arter/takson/26176>\                              |
|                                            |                         |                                                                                                          |                                                                                            | Some ambiguity over species identification.                            | <https://www.fishbase.org/summary/Coregonus_lavaretus.html>                  |
+--------------------------------------------+-------------------------+----------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------+------------------------------------------------------------------------+------------------------------------------------------------------------------+

: Table of most common fish species in data.

Now, with that said - is there a significant difference between these species?

```{r}
#| label: fish-species-variation
fish_filtered |>
  filter(
    SAMPLE_SPECIES %in%
      c(
        "Gadus morhua",
        "Salmo trutta",
        "Coregonus lavaretus",
        "Ctenolabrus rupestris"
      )
  ) |>
  ggplot(
    mapping = aes(x = MEASURED_VALUE_STANDARD_NORMALISED, y = SAMPLE_SPECIES)
  ) +
  geom_point() +
  geom_boxplot() +
  facet_wrap(
    facets = vars(SAMPLE_TISSUE),
    ncol = 1,
    scales = "free_y",
    space = "free_y"
  )
```