---
title: "Outlier Detection"
format: html
---

Some of our datasets include a lot of outliers. For example, let's take our fish data for *Gadus morhua* (excluding any measurements in dry weight for the sake of argument, and using liver measurements because they're the most data-rich).

```{r}
#| label: setup
library(tidyverse)
library(targets)
library(hexbin)
library(ggrepel)
library(viridis)
library(patchwork)
library(ggnewscale)
library(sf)
library(gridExtra)
library(grid)
library(knitr)
library(ggtext)
library(diptest)
library(glue)

gadus_morhua <- tar_read(load_literature_pqt) |>
  filter(
    SAMPLE_SPECIES == "Gadus morhua",
    MEASURED_UNIT_STANDARD == "mg/kg (wet)",
    SAMPLE_TISSUE == "Liver"
  ) |>
  mutate(
    SAMPLING_YEAR = factor(year(SAMPLING_DATE)),
    campaign_name_pretty = case_when(
      str_detect(CAMPAIGN_NAME_SHORT, "Vm_2010_2025") ~ str_remove_all(
        CAMPAIGN_NAME_SHORT,
        "Vm_2010_2025 \\("
      ) |>
        str_remove_all("\\)$") # TODO: Prettier regex.
    ),
    .default = CAMPAIGN_NAME_SHORT
  )
```

# Methods

## Outlier detection

We use two complementary approaches to flag potential outliers in each group.

### Tukey fences (IQR method)

The Tukey fence method flags values that fall more than $1.5 \times IQR$ beyond the first or third quartile of the log-transformed values, where $IQR = Q_3 - Q_1$. Specifically, a value $x_i$ is flagged if:

$$\log_{10}(x_i) < Q_1 - 1.5 \cdot IQR \quad \text{or} \quad \log_{10}(x_i) > Q_3 + 1.5 \cdot IQR$$

We apply fences to log-transformed values because concentrations are approximately log-normally distributed.

### Robust Modified Z-Score

The Robust Modified Z-Score (RMZ) is a median-based analogue of the standard Z-score that is resistant to the influence of extreme values <!-- TODO: cite Iglewicz & Hoaglin 1993 -->. For a value $x_i$ in a group with median $\tilde{x}$, it is defined as:

$$\text{RMZ}_i = \frac{x_i - \tilde{x}}{\text{MAD}}$$

where $\text{MAD} = \text{median}(|x_i - \tilde{x}|)$ is the median absolute deviation. In R, `mad()` already scales MAD by a consistency factor ($\approx 1 / 0.6745$) to make it comparable to the standard deviation under normality <!-- TODO: cite R documentation / consistency factor source -->. Values with $|\text{RMZ}| > 3.5$ are considered outliers, following the threshold recommended by <!-- TODO: cite Iglewicz & Hoaglin 1993 -->.

Points are coloured by how many flags they receive: unflagged (black), one flag (orange), or both flags (red).

### Test for unimodality

To identify groups whose distribution may be multimodal — which can indicate a mixture of genuinely distinct populations rather than a single skewed distribution — we apply Hartigan's Dip Test <!-- TODO: cite Hartigan & Hartigan 1985 -->. The test returns a p-value for the null hypothesis of unimodality; small p-values suggest the distribution is not unimodal. We use $p < 0.05$ as our threshold.

```{r}
#| label: robust-modified-z-score
robust_modified_z_score <- function(x) {
  x_median <- median(x, na.rm = TRUE)
  # mad() in R already includes the 1/0.6745 consistency factor
  x_mad <- mad(x, na.rm = TRUE)
  (x - x_median) / x_mad
}
```

---

If we make a density plot of the total distribution we can immediately see that something is off:

```{r}
#| label: fig-distribution
no_filter <- gadus_morhua |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD)) +
  geom_density() +
  labs(title = "Values not filtered") +
  theme_minimal()

filter_5000 <- gadus_morhua |>
  filter(MEASURED_VALUE_STANDARD <= 5000) |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD)) +
  geom_density() +
  labs(title = "Values >5000 mg/kg removed") +
  theme_minimal()

filter_1000 <- gadus_morhua |>
  filter(MEASURED_VALUE_STANDARD <= 1000) |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD)) +
  geom_density() +
  labs(title = "Values >1000 mg/kg removed") +
  theme_minimal()

filter_500 <- gadus_morhua |>
  filter(MEASURED_VALUE_STANDARD <= 500) |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD)) +
  geom_density() +
  labs(title = "Values >500 mg/kg removed") +
  theme_minimal()

no_filter /
  filter_5000 /
  filter_1000 /
  filter_500 +
  plot_annotation(
    title = "Density distributions of copper concentrations in G. morhua livers when values above a certain threshold are removed."
  )
```

Our plot peaks around 0 but extends beyond the 15,000 mg/kg wet weight (that is, 15g of copper per kg of wet liver). The most extreme of these values sufficiently improbable that we can remove, and as we do so our plot begins to resemble a left-censored normal distribution. However, where do we draw the line? Furthermore, we want our workflow to be as reusable and non-specific as possible.

Expert judgement is generally sufficient to make this call, but we cannot draw on this expert judgement for each conceivable case of species, tissue, site type, etc. Vannmiljø data dominate our overall dataset, and this includes both sites monitored specifically because they are uncontaminated (e.g. reference rivers), as well as those monitored because they are known to be highly contaminated (e.g. sites affected by mining).   

In general, we would expect natural, "real" outliers to represent one or more of the following scenarios:

- A campaign has selected specific sites known to have high contamination
- An acute contamination event has occurred
- A specific site has highly localised contamination

Conversely, "artifact" outliers may be a product of:

- Analytical methodological defects or contamination
- Mistakes in data entry
- Mistakes in data processing or manipulation

Directly detecting the former is likely to be much easier than the latter. We can start by examining the distributions of concentrations by year, site, and campaign.

# Outliers by year

```{r}
#| label: outliers-year-flag
gadus_morhua_by_year <- gadus_morhua |>
  group_by(SAMPLING_YEAR) |>
  mutate(
    RMZ = robust_modified_z_score(MEASURED_VALUE_STANDARD),
    log_val = log10(MEASURED_VALUE_STANDARD),
    Q1 = quantile(log_val, 0.25),
    Q3 = quantile(log_val, 0.75),
    IQR = IQR(log_val) * 1.5,
    outlier_RMZ = abs(RMZ) > 3.5,
    outlier_IQR = log_val < (Q1 - IQR) | log_val > (Q3 + IQR),
    dot_fill = factor(outlier_RMZ + outlier_IQR),
    year_and_n = glue::glue("**{SAMPLING_YEAR}** <br> (n = {n()})")
  )

dip_results_year <- gadus_morhua_by_year |>
  group_by(SAMPLING_YEAR) |>
  reframe(
    dip_p = dip.test(MEASURED_VALUE_STANDARD)$p.value
  ) |>
  mutate(bimodal = dip_p < 0.05)
```

```{r}
#| label: fig-outliers-year
main_plot <- gadus_morhua_by_year |>
  ggplot(aes(
    y = year_and_n,
    x = MEASURED_VALUE_STANDARD
  )) +
  geom_boxplot(
    outliers = FALSE,
    whisker.colour = "darkgrey",
    staple.colour = "darkgrey",
    box.colour = "darkgrey"
  ) +
  geom_dotplot(
    dotsize = 1,
    binwidth = 0.01,
    stackdir = "centerwhole",
    binpositions = "all",
    aes(fill = dot_fill),
  ) +
  scale_fill_manual(
    values = c("0" = "black", "1" = "orange", "2" = "red"),
    labels = c("0" = "Neither", "1" = "One flag", "2" = "Both flags"),
    name = "Outlier flags"
  ) +
  scale_x_log10(
    breaks = 10^(-1:5),
    labels = scales::label_log(),
    minor_breaks = outer(1:9, 10^(-1:4)) |> as.vector()
  ) +
  labs(
    y = "Year",
    x = "Measured Value (mg/kg (wet))",
    title = "Measured copper concentration in *G. morhua* livers, published on Vannmiljø, by reported sampling year",
    subtitle = "Outlying points flagged by Tukey fences (IQR × 1.5) and Robust Modified Z-Score (threshold 3.5)"
  ) +
  theme(
    plot.title = element_markdown(),
    axis.text.y.left = element_markdown(),
    panel.background = element_blank(),
    panel.grid.major = element_line(colour = "grey90"),
    panel.grid.minor = element_line(colour = "grey95"),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.key = element_blank(),
    strip.background = element_blank(),
    axis.ticks = element_blank()
  )

dip_panel <- dip_results_year |>
  ggplot(aes(
    y = SAMPLING_YEAR,
    x = 0,
    label = format(round(dip_p, 3), nsmall = 3),
    colour = bimodal
  )) +
  geom_text(size = 3) +
  scale_colour_manual(values = c("FALSE" = "grey40", "TRUE" = "red")) +
  labs(x = "p(not unimodal)", y = NULL) +
  theme_void() +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 9)
  )

main_plot + dip_panel + plot_layout(widths = c(6, 1), guides = "collect")
```

```{r}
#| label: tbl-outliers-year-comments
gadus_morhua_by_year |>
  filter(dot_fill == 2 | dot_fill == 1) |>
  group_by(MEASUREMENT_COMMENT, campaign_name_pretty) |>
  reframe(n = n(), mean = mean(MEASURED_VALUE_STANDARD)) |>
  arrange(desc(mean))
```

# Outliers by site

```{r}
#| label: outliers-site-flag
gadus_morhua_by_site <- gadus_morhua |>
  mutate(SITE_CODE = str_remove(SITE_CODE, "Vannmiljø_")) |>
  group_by(SITE_CODE) |>
  mutate(
    RMZ = robust_modified_z_score(MEASURED_VALUE_STANDARD),
    log_val = log10(MEASURED_VALUE_STANDARD),
    Q1 = quantile(log_val, 0.25),
    Q3 = quantile(log_val, 0.75),
    IQR = IQR(log_val) * 1.5,
    outlier_RMZ = abs(RMZ) > 3.5,
    outlier_IQR = log_val < (Q1 - IQR) | log_val > (Q3 + IQR),
    dot_fill = factor(outlier_RMZ + outlier_IQR),
    site_and_n = glue::glue("**{SITE_CODE}**  (n = {n()})")
  )

# TODO: Warning here. What does it mean?
dip_results_site <- gadus_morhua_by_site |>
  group_by(SITE_CODE) |>
  reframe(
    dip_p = dip.test(MEASURED_VALUE_STANDARD)$p.value
  ) |>
  mutate(bimodal = dip_p < 0.05)
```

```{r}
#| label: fig-outliers-site
main_plot_site <- gadus_morhua_by_site |>
  ggplot(aes(
    y = site_and_n,
    x = MEASURED_VALUE_STANDARD
  )) +
  geom_boxplot(
    outliers = FALSE,
    whisker.colour = "darkgrey",
    staple.colour = "darkgrey",
    box.colour = "darkgrey"
  ) +
  geom_dotplot(
    dotsize = 0.7,
    binwidth = 0.01,
    stackdir = "centerwhole",
    binpositions = "all",
    aes(fill = dot_fill)
  ) +
  scale_fill_manual(
    values = c("0" = "black", "1" = "orange", "2" = "red"),
    labels = c("0" = "Neither", "1" = "One flag", "2" = "Both flags"),
    name = "Outlier flags"
  ) +
  scale_x_log10(
    breaks = 10^(-1:5),
    labels = scales::label_log(),
    minor_breaks = outer(1:9, 10^(-1:4)) |> as.vector()
  ) +
  facet_wrap(
    facets = vars(SITE_GEOGRAPHIC_FEATURE),
    ncol = 1,
    scales = "free_y",
    space = "free_y"
  ) +
  labs(
    y = "Site Code",
    x = "Measured Value (mg/kg (wet))",
    title = "Measured copper concentration in *G. morhua* livers, published on Vannmiljø, by reported site code",
    subtitle = "Outlying points flagged by Tukey fences (IQR × 1.5) and Robust Modified Z-Score (threshold 3.5)"
  ) +
  theme(
    plot.title = element_markdown(),
    axis.text.y.left = element_markdown(),
    panel.background = element_blank(),
    panel.grid.major = element_line(colour = "grey90"),
    panel.grid.minor = element_line(colour = "grey95"),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.key = element_blank(),
    strip.background = element_blank(),
    axis.ticks = element_blank()
  )

dip_panel_site <- dip_results_site |>
  ggplot(aes(
    y = SITE_CODE,
    x = 0,
    label = format(round(dip_p, 3), nsmall = 3),
    colour = bimodal
  )) +
  geom_text(size = 3) +
  scale_colour_manual(values = c("FALSE" = "grey40", "TRUE" = "red")) +
  labs(x = "p(not unimodal)", y = NULL) +
  theme_void() +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 9)
  )

main_plot_site +
  dip_panel_site +
  plot_layout(widths = c(6, 1), guides = "collect")
```

```{r}
#| label: tbl-outliers-site-comments
gadus_morhua_by_site |>
  filter(dot_fill == 2 | dot_fill == 1) |>
  group_by(MEASUREMENT_COMMENT, campaign_name_pretty) |>
  reframe(n = n(), mean = mean(MEASURED_VALUE_STANDARD)) |>
  arrange(desc(mean))
```

# Outliers by campaign

```{r}
#| label: outliers-campaign-flag
gadus_morhua_by_campaign <- gadus_morhua |>
  mutate(
    CAMPAIGN_NAME = str_remove(
      CAMPAIGN_NAME,
      "Vannmiljø Copper Monitoring 2010-2025"
    )
  ) |>
  group_by(CAMPAIGN_NAME) |>
  mutate(
    RMZ = robust_modified_z_score(MEASURED_VALUE_STANDARD),
    log_val = log10(MEASURED_VALUE_STANDARD),
    Q1 = quantile(log_val, 0.25),
    Q3 = quantile(log_val, 0.75),
    IQR = IQR(log_val) * 1.5,
    outlier_RMZ = abs(RMZ) > 3.5,
    outlier_IQR = log_val < (Q1 - IQR) | log_val > (Q3 + IQR),
    dot_fill = factor(outlier_RMZ + outlier_IQR),
    campaign_and_n = glue("**{CAMPAIGN_NAME}** <br> (n = {n()})") |>
      as.character()
  )

# More errors here
dip_results_campaign <- gadus_morhua_by_campaign |>
  group_by(CAMPAIGN_NAME) |>
  reframe(
    dip_p = dip.test(MEASURED_VALUE_STANDARD)$p.value
  ) |>
  mutate(bimodal = dip_p < 0.05)
```

```{r}
#| label: fig-outliers-campaign
main_plot_campaign <- gadus_morhua_by_campaign |>
  ggplot(aes(
    y = campaign_and_n,
    x = MEASURED_VALUE_STANDARD
  )) +
  geom_boxplot(
    outliers = FALSE,
    whisker.colour = "darkgrey",
    staple.colour = "darkgrey",
    box.colour = "darkgrey"
  ) +
  geom_dotplot(
    dotsize = 0.7,
    binwidth = 0.01,
    stackdir = "centerwhole",
    binpositions = "all",
    aes(fill = dot_fill)
  ) +
  scale_fill_manual(
    values = c("0" = "black", "1" = "orange", "2" = "red"),
    labels = c("0" = "Neither", "1" = "One flag", "2" = "Both flags"),
    name = "Outlier flags"
  ) +
  scale_x_log10(
    breaks = 10^(-1:5),
    labels = scales::label_log(),
    minor_breaks = outer(1:9, 10^(-1:4)) |> as.vector()
  ) +
  labs(
    y = "Campaign",
    x = "Measured Value (mg/kg (wet))",
    title = "Measured copper concentration in *G. morhua* livers, published on Vannmiljø, by monitoring campaign",
    subtitle = "Outlying points flagged by Tukey fences (IQR × 1.5) and Robust Modified Z-Score (threshold 3.5)"
  ) +
  theme(
    plot.title = element_markdown(),
    axis.text.y.left = element_markdown(),
    panel.background = element_blank(),
    panel.grid.major = element_line(colour = "grey90"),
    panel.grid.minor = element_line(colour = "grey95"),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.key = element_blank(),
    strip.background = element_blank(),
    axis.ticks = element_blank()
  )

dip_panel_campaign <- dip_results_campaign |>
  ggplot(aes(
    y = CAMPAIGN_NAME,
    x = 0,
    label = format(round(dip_p, 3), nsmall = 3),
    colour = bimodal
  )) +
  geom_text(size = 3) +
  scale_colour_manual(values = c("FALSE" = "grey40", "TRUE" = "red")) +
  labs(x = "p(not unimodal)", y = NULL) +
  theme_void() +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 9)
  )

main_plot_campaign +
  dip_panel_campaign +
  plot_layout(widths = c(6, 1), guides = "collect")
```

```{r}
#| label: tbl-outliers-campaign-comments
gadus_morhua_by_campaign |>
  filter(dot_fill == 2 | dot_fill == 1) |>
  group_by(MEASUREMENT_COMMENT, campaign_name_pretty) |>
  reframe(n = n(), mean = mean(MEASURED_VALUE_STANDARD)) |>
  arrange(desc(mean))
```

There's no immediately obvious pattern to the outliers. In the interest of scientific rigour we should investigate a number of possible patterns behind our more improbable outliers (analytical error, contamination, data entry error) and potential correlations with metadata (site, contractor, analytical technique). However, doing this for each set of outliers will quickly become prohibitively complicated. I believe (based on intuition and experience) that the most likely explanation for obvious, implausible outliers is transcription errors (either the wrong unit being used, or simply entering the wrong number), as much of the data in Vannmiljø (especially historically) is manually entered.

Under the circumstances, we need a simple, robust, defensible approach to filtering outliers that doesn't require this degree of fine attention.

