---
title: "Sample Groups"
format:
  html:
    lightbox: true
---

This is a top-but-one level notebook containing all sample groups.

```{r}
#| label: setup
#| include: false
#| warning: false

library(targets)
library(here)
library(tidyverse)
library(flextable)
pkgload::load_all(quiet = TRUE) # sample_groups_flextable() etc.

here::i_am("index.qmd")
```


```{r}
#| label: tbl-groups-summary
#| tbl-caption: tbc

# Reshaping lives in the sample_groups_table target (R/fct_sample_groups_table.R),
# so this notebook and index.qmd cannot drift apart. All 245 groups shown here;
# the manuscript shows only the large ones.
tar_read(sample_groups_table) |>
  sample_groups_flextable()
```

# Triage plots

One row of five views per group, generated by `R/fct_group_triage.R` and
written to `_triage/` as PNGs by the `triage_pilot_plots` target. Click any
plot to enlarge it.

Currently a **pilot**: 5 groups sampled at random from the `r nrow(tar_read(sample_groups_table))`-group
table, restricted to those with `n >= 100`. The sample is seeded, so it stays
the same between runs. Widen it by raising `n_sample` in the
`triage_pilot_groups` target.

::: callout-note
## Why some plots show points and others show bins

Every view switches presentation on the number of **rows** available, at a
threshold of 30 (`triage_use_points()`). Above it, values are summarised into
2D bins or hexes; below it, they are drawn as individual points. This is
deliberate: a hex map built from four observations is mostly empty cells and
conveys less than the four points themselves, while forty thousand points
overplot into a solid block and take an age to draw.

So a group can legitimately show hexes in one panel and points in another, and
two groups on this page can present the same view differently. **Aquatic /
Freshwater / River, stream, canal** is the clearest example: it clears
`n >= 100` on the summary table but has only 4 underlying rows, so its spatial
panel falls back to points.

That gap is worth watching. `n` in the summary table is `sum(MEASURED_N)`, a
count of *measurements*, whereas the plots draw one mark per *row*. A group
that aggregates many measurements into few rows will look far sparser here
than its `n` suggests. `n_rows` is reported in each heading below so the
discrepancy is visible rather than surprising.
:::

```{r}
#| label: fig-triage-plots
#| output: asis

groups <- tar_read(triage_pilot_groups)
plot_paths <- tar_read(triage_pilot_plots)

# Filename suffix -> caption. Keys match the list names in
# write_triage_plots_for_group(), which also drive the a/b/c/d/e file ordering.
captions <- c(
  a_density = "a) Overall distribution (all units)",
  b_date = "b) Concentration by date",
  c_campaign = "c) Distribution by campaign",
  d_site_type = "d) Distribution by site type",
  e_spatial = "e) Spatial distribution"
)

for (i in seq_len(nrow(groups))) {
  grp <- groups[i, , drop = FALSE]
  slug <- grp$group_slug

  cat("\n\n## ", triage_group_label(grp), "\n\n", sep = "")
  cat(
    "`n` = ",
    format(grp$n, big.mark = ","),
    " measurements across ",
    format(grp$n_rows, big.mark = ","),
    " rows.\n\n",
    sep = ""
  )

  # Paths are project-root-relative; this notebook renders from docs/
  paths <- sort(plot_paths[startsWith(basename(plot_paths), paste0(slug, "_"))])

  cat("::: {layout-ncol=5 .column-screen-inset}\n\n")
  for (p in paths) {
    key <- sub("\\.png$", "", sub(paste0("^", slug, "_"), "", basename(p)))
    cat(
      "![",
      captions[[key]],
      "](",
      file.path("..", p),
      ")",
      "{group=\"",
      slug,
      "\"}\n\n",
      sep = ""
    )
  }
  cat(":::\n\n")
}
```
