---
title: "Notebook XX - Sørfjorden Case Study"
date: last-modified
lightbox: true
format: html
execute:
  freeze: false   # parked 2026-07-29, see _quarto.yml render list
---

```{r}
#| label: setup
#| include: false
#| warning: false

library(targets)
library(here)
library(tidyverse)
library(flextable)
pkgload::load_all(quiet = TRUE) # sample_groups_flextable() etc.

here::i_am("index.qmd")
```

Sørfjorden ("The South Fjord") is the southerly arm of Hardangerfjorden on Norway's South-West coast [@thorsnaesSorfjordenHardanger2024]. In the early 1900s a number of heavy industrial plants began operation on the fjord's shores, and it became unusually contaminated with heavy metals [@ruusMonitoringEnvironmentalQuality2010]. Copper is known to have been emitted to Sørfjord as a byproduct of zinc production in the area, and unusually high copper concentrations have been detected in the tissues of fish caught in the area [@ruusMonitoringEnvironmentalQuality2010]

A003,AEP-003 - Sørfjorden,"Bounding box 59.77-60.37 N, 6.25-6.85 E, measured against load_literature_pqt the same way as A002. 760 measurements, 3 references (VannmiljøCopper2010-2025, 2003CoteurFieldContaminationOf, 2005GillanStructureOfSediment).",59.77,60.37,6.25,6.85,1900,2100,

```{r}
#| label: fig-soerfjorden-overview
#| fig-cap: Sørfjorden AEP-003 bbox (59.77-60.37N, 6.25-6.85E). a) location within the wider study area. b-d) zoomed view (padded well beyond the bbox for geographic context; data still restricted to it) showing sampling density by compartment. Hexbinned where n sites >= 30 (b), points below that (d), a "no samples" panel where a compartment is empty in this bbox (c). Adapted from scripts/build_study_area_figure.R panel a).
#| fig-width: 9
#| fig-height: 11
library(sf)
library(patchwork)

sf::sf_use_s2(FALSE)

soerfjorden_bbox <- st_bbox(
  c(xmin = 6.25, xmax = 6.85, ymin = 59.77, ymax = 60.37),
  crs = st_crs(4326)
)

# View window for panels b-d: padded well beyond the AEP bbox above. Cropping
# the map tightly to the bbox itself makes it ~2x taller than wide (a fjord is
# a narrow N-S sliver) and reads as visually meaningless. The padding is a
# viewport choice only -- site_points below still filters to the strict bbox,
# so what counts as "in this AEP" is unaffected.
view_bbox_deg <- st_bbox(
  c(
    xmin = soerfjorden_bbox[["xmin"]] - 0.35,
    xmax = soerfjorden_bbox[["xmax"]] + 0.35,
    ymin = soerfjorden_bbox[["ymin"]] - 0.12,
    ymax = soerfjorden_bbox[["ymax"]] + 0.12
  ),
  crs = st_crs(4326)
)

# UTM 32N (6-12E) rather than the eqc-projection trick used at study-area
# scale: that trick reprojects AFTER stat_bin_hex() has already binned in
# plain lon/lat degrees, so it warps the hexagons it was meant to fix (they
# came out ~2:1 height:width). Transforming points to true metres first and
# binning there is what actually gets regular hexagons, at the cost of only
# being valid near this UTM zone -- fine for one fjord, not for the
# study-area map.
utm_crs <- st_crs(32632)
view_bbox_utm <- st_bbox(st_transform(st_as_sfc(view_bbox_deg), utm_crs))

wgs84_geo <- tar_read(wgs84_geography)
literature_merged_data <- tar_read(load_literature_pqt)

# Natural Earth's rivers dataset only carries major rivers even at its finest
# (10m) scale -- it returns zero features for this fjord's catchment (checked
# 2026-08-26). Left wired up (and drawn if present) since fetching it is
# genuinely one line via the project's ne_get() cache wrapper; getting an
# Odda-scale river actually plotted needs a finer source (Kartverket N50/N250
# elvenett, or OSM).
rivers <- ne_get(
  scale = 10,
  type = "rivers_lake_centerlines",
  category = "physical",
  destdir = "data/raw/shapefiles/"
) |>
  st_crop(view_bbox_deg) |>
  suppressWarnings()

norway_bbox <- st_bbox(
  c(xmin = 4, xmax = 31.5, ymin = 57.5, ymax = 71.5),
  crs = st_crs(4326)
)

# --- a) Locator: bbox in the wider study area --------------------------

panel_locator <- ggplot() +
  geom_sf(
    data = wgs84_geo$marine_polys,
    aes(fill = I(ocean_color)),
    color = "white",
    linewidth = 0.15
  ) +
  geom_sf(
    data = wgs84_geo$countries,
    aes(
      fill = I(ifelse(
        highlight_name,
        country_colours["highlight"],
        country_colours["default"]
      ))
    ),
    color = "white",
    linewidth = 0.25
  ) +
  geom_sf(
    data = st_as_sfc(soerfjorden_bbox),
    fill = NA,
    color = "#b23b3b",
    linewidth = 0.9
  ) +
  coord_sf(
    xlim = c(norway_bbox[["xmin"]], norway_bbox[["xmax"]]),
    ylim = c(norway_bbox[["ymin"]], norway_bbox[["ymax"]]),
    expand = FALSE
  ) +
  theme_void() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(2, 2, 2, 2),
    legend.position = "none",
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.01,
      margin = margin(b = 2)
    )
  ) +
  labs(title = "a) Location") |>
    suppressWarnings()

# --- b-d) Zoomed density by compartment ---------------------------------

# Hexbin only where there's enough to bin: below n = 30 distinct sites a
# hexbin reads as more precise than it is (CLAUDE.md 4.4, "switch on n").
# Terrestrial has zero rows in this bbox (3 in the whole project, none here),
# Biota has 11 distinct sites -- both fall under that line, so they get
# points/a blank-panel message instead of a hexbin.
build_compartment_panel <- function(
  data,
  compartment,
  title,
  min_n_for_hex = 30
) {
  pts <- data |>
    filter(
      ENVIRON_COMPARTMENT == compartment,
      !is.na(LATITUDE),
      !is.na(LONGITUDE),
      !is.na(MEASURED_N),
      LATITUDE >= soerfjorden_bbox[["ymin"]],
      LATITUDE <= soerfjorden_bbox[["ymax"]],
      LONGITUDE >= soerfjorden_bbox[["xmin"]],
      LONGITUDE <= soerfjorden_bbox[["xmax"]]
    ) |>
    summarise(
      total_n = sum(MEASURED_N, na.rm = TRUE),
      .by = c(SITE_CODE, LATITUDE, LONGITUDE)
    ) |>
    st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326, remove = FALSE) |>
    st_transform(utm_crs)

  n_sites <- nrow(pts)

  base <- ggplot() +
    geom_sf(
      data = wgs84_geo$marine_polys,
      aes(fill = I(ocean_color)),
      color = "white",
      linewidth = 0.15
    ) +
    geom_sf(
      data = wgs84_geo$countries,
      aes(
        fill = I(ifelse(
          highlight_name,
          country_colours["highlight"],
          country_colours["default"]
        ))
      ),
      color = "white",
      linewidth = 0.25
    ) +
    {
      if (nrow(rivers) > 0) {
        geom_sf(data = rivers, color = "#3a7ca8", linewidth = 0.4)
      }
    } +
    ggnewscale::new_scale_fill()

  if (n_sites == 0) {
    panel <- base +
      annotate(
        "text",
        x = mean(view_bbox_utm[c("xmin", "xmax")]),
        y = mean(view_bbox_utm[c("ymin", "ymax")]),
        label = "No samples",
        size = 3,
        fontface = "italic",
        color = "grey30"
      )
  } else if (n_sites < min_n_for_hex) {
    panel <- base +
      geom_sf(data = pts, aes(size = total_n), color = "#b23b3b", alpha = 0.8) +
      scale_size_area(name = "Samples (sum n)", max_size = 5)
  } else {
    coords <- st_coordinates(pts)
    pts_df <- pts |>
      st_drop_geometry() |>
      mutate(X = coords[, 1], Y = coords[, 2])
    panel <- base +
      stat_bin_hex(
        data = pts_df,
        aes(x = X, y = Y, weight = total_n, fill = after_stat(count)),
        bins = 25,
        linewidth = 0,
        alpha = 0.8,
        inherit.aes = FALSE
      ) +
      scale_fill_viridis_c(
        name = "Samples (sum n)",
        trans = "log10",
        labels = scales::comma
      )
  }

  panel +
    # default_crs = crs (not the usual 4326): pts_df's X/Y are already in UTM
    # metres, so this tells coord_sf to treat them as such rather than
    # reproject them a second time, which is what caused the squished hexes.
    coord_sf(
      crs = utm_crs,
      default_crs = utm_crs,
      xlim = c(view_bbox_utm[["xmin"]], view_bbox_utm[["xmax"]]),
      ylim = c(view_bbox_utm[["ymin"]], view_bbox_utm[["ymax"]]),
      expand = FALSE
    ) +
    theme_void() +
    theme(
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA),
      plot.margin = margin(2, 2, 2, 2),
      legend.position = "bottom",
      legend.key.height = unit(0.25, "cm"),
      legend.key.width = unit(0.6, "cm"),
      legend.text = element_text(size = 6),
      legend.title = element_text(size = 7),
      plot.title = element_text(
        face = "bold",
        size = 10,
        hjust = 0.01,
        margin = margin(b = 2)
      )
    ) +
    labs(title = title) |>
      suppressWarnings()
}

panel_aquatic <- build_compartment_panel(
  literature_merged_data,
  "Aquatic",
  "b) Aquatic"
)
panel_terrestrial <- build_compartment_panel(
  literature_merged_data,
  "Terrestrial",
  "c) Terrestrial"
)
panel_biota <- build_compartment_panel(
  literature_merged_data,
  "Biota",
  "d) Biota"
)

panel_overview <- (panel_locator | panel_aquatic) /
  (panel_terrestrial | panel_biota)

# Written to figures/ (not .images-raw/ like the parent script) since this
# notebook's output is a real, reviewable artefact rather than a one-off
# manuscript-figure draft. See CLAUDE.md 2.3.1: Sam can't see this inline, so
# the path is the thing that matters.
ggsave(
  "figures/soerfjorden-overview.png",
  panel_overview,
  width = 9,
  height = 11,
  dpi = 300,
  bg = "white"
) |>
  suppressWarnings()

panel_overview
```

```{r}
# Table of groups filtered to bounding box
```

<!-- Embed AEP here -->