---
title: "Distributions: Crustaceans"
format: html
---

::: callout-note
This notebook was scaffolded by `scripts/generate_distribution_notebooks.R`
from the outlier-analysis targets factory (`_targets.R`,
`R/fct_outlier_groups.R`, `R/fct_outlier_detection.R`,
`R/fct_outlier_report.R`). The plots/tables below are generated, but
analysis and interpretation are added by hand, section by section.
Re-running the generator only appends sections for groups not yet present
here -- it never rewrites or removes anything you have written.

See [NBXX-Outliers.qmd](NBXX-Outliers.qmd) for the fully hand-authored
*G. morhua* liver case study this workflow was built from, including
year/site/campaign breakdowns and the reasoning behind the outlier flags.
:::

```{r}
#| label: setup
library(tidyverse)
library(targets)
library(patchwork)
library(flextable)
pkgload::load_all(quiet = TRUE)

set_flextable_defaults(
  font.size = 10,
  font.family = "Arial",
  digits = 2,
  theme_fun = theme_vanilla,
  padding = 1
)

ggplot_base_size <- 20
theme_set(theme_minimal(base_size = ggplot_base_size, base_family = "Aptos"))
```

## Calanus finmarchicus: Whole body

Crustaceans / Calanus finmarchicus / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Calanus_finmarchicus_Whole_body
grp_Crustaceans_Calanus_finmarchicus_Whole_body <- tar_read(outlier_biota_Crustaceans_Calanus_finmarchicus_Whole_body)
grp_Crustaceans_Calanus_finmarchicus_Whole_body$data <- grp_Crustaceans_Calanus_finmarchicus_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Calanus_finmarchicus_Whole_body_unit <- grp_Crustaceans_Calanus_finmarchicus_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Calanus_finmarchicus_Whole_body
#| fig-cap: "Crustaceans / Calanus finmarchicus / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Calanus_finmarchicus_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Calanus_finmarchicus_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Calanus_finmarchicus_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Calanus_finmarchicus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Calanus_finmarchicus_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Calanus_finmarchicus_Whole_body
#| fig-cap: "Crustaceans / Calanus finmarchicus / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Calanus_finmarchicus_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Calanus_finmarchicus_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Calanus_finmarchicus_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Calanus_finmarchicus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Calanus_finmarchicus_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Calanus_finmarchicus_Whole_body
#| fig-cap: "Crustaceans / Calanus finmarchicus / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Calanus_finmarchicus_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Calanus_finmarchicus_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Calanus_finmarchicus_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Calanus_finmarchicus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Calanus_finmarchicus_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Calanus_finmarchicus_Whole_body
#| fig-cap: "Crustaceans / Calanus finmarchicus / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Calanus_finmarchicus_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Calanus_finmarchicus_Whole_body_unit, ")")
)
```

## Calanus glacialis: Whole body

Crustaceans / Calanus glacialis / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Calanus_glacialis_Whole_body
grp_Crustaceans_Calanus_glacialis_Whole_body <- tar_read(outlier_biota_Crustaceans_Calanus_glacialis_Whole_body)
grp_Crustaceans_Calanus_glacialis_Whole_body$data <- grp_Crustaceans_Calanus_glacialis_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Calanus_glacialis_Whole_body_unit <- grp_Crustaceans_Calanus_glacialis_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Calanus_glacialis_Whole_body
#| fig-cap: "Crustaceans / Calanus glacialis / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Calanus_glacialis_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Calanus_glacialis_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Calanus_glacialis_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Calanus_glacialis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Calanus_glacialis_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Calanus_glacialis_Whole_body
#| fig-cap: "Crustaceans / Calanus glacialis / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Calanus_glacialis_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Calanus_glacialis_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Calanus_glacialis_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Calanus_glacialis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Calanus_glacialis_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Calanus_glacialis_Whole_body
#| fig-cap: "Crustaceans / Calanus glacialis / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Calanus_glacialis_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Calanus_glacialis_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Calanus_glacialis_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Calanus_glacialis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Calanus_glacialis_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Calanus_glacialis_Whole_body
#| fig-cap: "Crustaceans / Calanus glacialis / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Calanus_glacialis_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Calanus_glacialis_Whole_body_unit, ")")
)
```

## Calanus hyperboreus: Whole body

Crustaceans / Calanus hyperboreus / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Calanus_hyperboreus_Whole_body
grp_Crustaceans_Calanus_hyperboreus_Whole_body <- tar_read(outlier_biota_Crustaceans_Calanus_hyperboreus_Whole_body)
grp_Crustaceans_Calanus_hyperboreus_Whole_body$data <- grp_Crustaceans_Calanus_hyperboreus_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Calanus_hyperboreus_Whole_body_unit <- grp_Crustaceans_Calanus_hyperboreus_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Calanus_hyperboreus_Whole_body
#| fig-cap: "Crustaceans / Calanus hyperboreus / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Calanus_hyperboreus_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Calanus_hyperboreus_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Calanus_hyperboreus_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Calanus_hyperboreus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Calanus_hyperboreus_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Calanus_hyperboreus_Whole_body
#| fig-cap: "Crustaceans / Calanus hyperboreus / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Calanus_hyperboreus_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Calanus_hyperboreus_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Calanus_hyperboreus_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Calanus_hyperboreus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Calanus_hyperboreus_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Calanus_hyperboreus_Whole_body
#| fig-cap: "Crustaceans / Calanus hyperboreus / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Calanus_hyperboreus_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Calanus_hyperboreus_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Calanus_hyperboreus_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Calanus_hyperboreus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Calanus_hyperboreus_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Calanus_hyperboreus_Whole_body
#| fig-cap: "Crustaceans / Calanus hyperboreus / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Calanus_hyperboreus_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Calanus_hyperboreus_Whole_body_unit, ")")
)
```

## Cancer pagurus: Brown meat

Crustaceans / Cancer pagurus / Brown meat. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Cancer_pagurus_Brown_meat
grp_Crustaceans_Cancer_pagurus_Brown_meat <- tar_read(outlier_biota_Crustaceans_Cancer_pagurus_Brown_meat)
grp_Crustaceans_Cancer_pagurus_Brown_meat$data <- grp_Crustaceans_Cancer_pagurus_Brown_meat$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Cancer_pagurus_Brown_meat_unit <- grp_Crustaceans_Cancer_pagurus_Brown_meat$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Cancer_pagurus_Brown_meat
#| fig-cap: "Crustaceans / Cancer pagurus / Brown meat: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Brown_meat$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Brown_meat$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Brown_meat_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Cancer_pagurus_Brown_meat
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Brown_meat$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Cancer_pagurus_Brown_meat
#| fig-cap: "Crustaceans / Cancer pagurus / Brown meat: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Brown_meat$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Brown_meat$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Brown_meat_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Cancer_pagurus_Brown_meat
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Brown_meat$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Cancer_pagurus_Brown_meat
#| fig-cap: "Crustaceans / Cancer pagurus / Brown meat: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Brown_meat$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Brown_meat$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Brown_meat_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Cancer_pagurus_Brown_meat
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Brown_meat$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Cancer_pagurus_Brown_meat
#| fig-cap: "Crustaceans / Cancer pagurus / Brown meat: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Cancer_pagurus_Brown_meat,
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Brown_meat_unit, ")")
)
```

## Cancer pagurus: Liver

Crustaceans / Cancer pagurus / Liver. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Cancer_pagurus_Liver
grp_Crustaceans_Cancer_pagurus_Liver <- tar_read(outlier_biota_Crustaceans_Cancer_pagurus_Liver)
grp_Crustaceans_Cancer_pagurus_Liver$data <- grp_Crustaceans_Cancer_pagurus_Liver$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Cancer_pagurus_Liver_unit <- grp_Crustaceans_Cancer_pagurus_Liver$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Cancer_pagurus_Liver
#| fig-cap: "Crustaceans / Cancer pagurus / Liver: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Liver$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Liver$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Liver_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Cancer_pagurus_Liver
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Liver$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Cancer_pagurus_Liver
#| fig-cap: "Crustaceans / Cancer pagurus / Liver: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Liver$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Liver$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Liver_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Cancer_pagurus_Liver
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Liver$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Cancer_pagurus_Liver
#| fig-cap: "Crustaceans / Cancer pagurus / Liver: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Liver$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Liver$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Liver_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Cancer_pagurus_Liver
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Liver$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Cancer_pagurus_Liver
#| fig-cap: "Crustaceans / Cancer pagurus / Liver: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Cancer_pagurus_Liver,
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Liver_unit, ")")
)
```

## Cancer pagurus: Muscle

Crustaceans / Cancer pagurus / Muscle. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Cancer_pagurus_Muscle
grp_Crustaceans_Cancer_pagurus_Muscle <- tar_read(outlier_biota_Crustaceans_Cancer_pagurus_Muscle)
grp_Crustaceans_Cancer_pagurus_Muscle$data <- grp_Crustaceans_Cancer_pagurus_Muscle$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Cancer_pagurus_Muscle_unit <- grp_Crustaceans_Cancer_pagurus_Muscle$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Cancer_pagurus_Muscle
#| fig-cap: "Crustaceans / Cancer pagurus / Muscle: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Muscle$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Muscle$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Muscle_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Cancer_pagurus_Muscle
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Muscle$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Cancer_pagurus_Muscle
#| fig-cap: "Crustaceans / Cancer pagurus / Muscle: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Muscle$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Muscle$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Muscle_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Cancer_pagurus_Muscle
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Muscle$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Cancer_pagurus_Muscle
#| fig-cap: "Crustaceans / Cancer pagurus / Muscle: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Muscle$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Muscle$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Muscle_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Cancer_pagurus_Muscle
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Muscle$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Cancer_pagurus_Muscle
#| fig-cap: "Crustaceans / Cancer pagurus / Muscle: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Cancer_pagurus_Muscle,
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Muscle_unit, ")")
)
```

## Cancer pagurus: Total soft tissues

Crustaceans / Cancer pagurus / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Cancer_pagurus_Total_soft_tissues
grp_Crustaceans_Cancer_pagurus_Total_soft_tissues <- tar_read(outlier_biota_Crustaceans_Cancer_pagurus_Total_soft_tissues)
grp_Crustaceans_Cancer_pagurus_Total_soft_tissues$data <- grp_Crustaceans_Cancer_pagurus_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Cancer_pagurus_Total_soft_tissues_unit <- grp_Crustaceans_Cancer_pagurus_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Cancer_pagurus_Total_soft_tissues
#| fig-cap: "Crustaceans / Cancer pagurus / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Cancer_pagurus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Cancer_pagurus_Total_soft_tissues
#| fig-cap: "Crustaceans / Cancer pagurus / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Cancer_pagurus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Cancer_pagurus_Total_soft_tissues
#| fig-cap: "Crustaceans / Cancer pagurus / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Cancer_pagurus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Cancer_pagurus_Total_soft_tissues
#| fig-cap: "Crustaceans / Cancer pagurus / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Cancer_pagurus_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Total_soft_tissues_unit, ")")
)
```

## Cancer pagurus: Whole body

Crustaceans / Cancer pagurus / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Cancer_pagurus_Whole_body
grp_Crustaceans_Cancer_pagurus_Whole_body <- tar_read(outlier_biota_Crustaceans_Cancer_pagurus_Whole_body)
grp_Crustaceans_Cancer_pagurus_Whole_body$data <- grp_Crustaceans_Cancer_pagurus_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Cancer_pagurus_Whole_body_unit <- grp_Crustaceans_Cancer_pagurus_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Cancer_pagurus_Whole_body
#| fig-cap: "Crustaceans / Cancer pagurus / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Cancer_pagurus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Cancer_pagurus_Whole_body
#| fig-cap: "Crustaceans / Cancer pagurus / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Cancer_pagurus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Cancer_pagurus_Whole_body
#| fig-cap: "Crustaceans / Cancer pagurus / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Cancer_pagurus_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Cancer_pagurus_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Cancer_pagurus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Cancer_pagurus_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Cancer_pagurus_Whole_body
#| fig-cap: "Crustaceans / Cancer pagurus / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Cancer_pagurus_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Cancer_pagurus_Whole_body_unit, ")")
)
```

## Carcinus maenas: Total soft tissues

Crustaceans / Carcinus maenas / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Carcinus_maenas_Total_soft_tissues
grp_Crustaceans_Carcinus_maenas_Total_soft_tissues <- tar_read(outlier_biota_Crustaceans_Carcinus_maenas_Total_soft_tissues)
grp_Crustaceans_Carcinus_maenas_Total_soft_tissues$data <- grp_Crustaceans_Carcinus_maenas_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Carcinus_maenas_Total_soft_tissues_unit <- grp_Crustaceans_Carcinus_maenas_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Carcinus_maenas_Total_soft_tissues
#| fig-cap: "Crustaceans / Carcinus maenas / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Carcinus_maenas_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Carcinus_maenas_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Carcinus_maenas_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Carcinus_maenas_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Carcinus_maenas_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Carcinus_maenas_Total_soft_tissues
#| fig-cap: "Crustaceans / Carcinus maenas / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Carcinus_maenas_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Carcinus_maenas_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Carcinus_maenas_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Carcinus_maenas_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Carcinus_maenas_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Carcinus_maenas_Total_soft_tissues
#| fig-cap: "Crustaceans / Carcinus maenas / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Carcinus_maenas_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Carcinus_maenas_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Carcinus_maenas_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Carcinus_maenas_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Carcinus_maenas_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Carcinus_maenas_Total_soft_tissues
#| fig-cap: "Crustaceans / Carcinus maenas / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Carcinus_maenas_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Crustaceans_Carcinus_maenas_Total_soft_tissues_unit, ")")
)
```

## Conchoecia borealis: Whole body

Crustaceans / Conchoecia borealis / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Conchoecia_borealis_Whole_body
grp_Crustaceans_Conchoecia_borealis_Whole_body <- tar_read(outlier_biota_Crustaceans_Conchoecia_borealis_Whole_body)
grp_Crustaceans_Conchoecia_borealis_Whole_body$data <- grp_Crustaceans_Conchoecia_borealis_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Conchoecia_borealis_Whole_body_unit <- grp_Crustaceans_Conchoecia_borealis_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Conchoecia_borealis_Whole_body
#| fig-cap: "Crustaceans / Conchoecia borealis / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Conchoecia_borealis_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Conchoecia_borealis_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Conchoecia_borealis_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Conchoecia_borealis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Conchoecia_borealis_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Conchoecia_borealis_Whole_body
#| fig-cap: "Crustaceans / Conchoecia borealis / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Conchoecia_borealis_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Conchoecia_borealis_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Conchoecia_borealis_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Conchoecia_borealis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Conchoecia_borealis_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Conchoecia_borealis_Whole_body
#| fig-cap: "Crustaceans / Conchoecia borealis / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Conchoecia_borealis_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Conchoecia_borealis_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Conchoecia_borealis_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Conchoecia_borealis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Conchoecia_borealis_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Conchoecia_borealis_Whole_body
#| fig-cap: "Crustaceans / Conchoecia borealis / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Conchoecia_borealis_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Conchoecia_borealis_Whole_body_unit, ")")
)
```

## Euchaeta barbata: Whole body

Crustaceans / Euchaeta barbata / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Euchaeta_barbata_Whole_body
grp_Crustaceans_Euchaeta_barbata_Whole_body <- tar_read(outlier_biota_Crustaceans_Euchaeta_barbata_Whole_body)
grp_Crustaceans_Euchaeta_barbata_Whole_body$data <- grp_Crustaceans_Euchaeta_barbata_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Euchaeta_barbata_Whole_body_unit <- grp_Crustaceans_Euchaeta_barbata_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Euchaeta_barbata_Whole_body
#| fig-cap: "Crustaceans / Euchaeta barbata / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Euchaeta_barbata_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Euchaeta_barbata_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Euchaeta_barbata_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Euchaeta_barbata_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Euchaeta_barbata_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Euchaeta_barbata_Whole_body
#| fig-cap: "Crustaceans / Euchaeta barbata / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Euchaeta_barbata_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Euchaeta_barbata_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Euchaeta_barbata_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Euchaeta_barbata_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Euchaeta_barbata_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Euchaeta_barbata_Whole_body
#| fig-cap: "Crustaceans / Euchaeta barbata / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Euchaeta_barbata_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Euchaeta_barbata_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Euchaeta_barbata_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Euchaeta_barbata_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Euchaeta_barbata_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Euchaeta_barbata_Whole_body
#| fig-cap: "Crustaceans / Euchaeta barbata / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Euchaeta_barbata_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Euchaeta_barbata_Whole_body_unit, ")")
)
```

## Euchaeta glacialis: Whole body

Crustaceans / Euchaeta glacialis / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Euchaeta_glacialis_Whole_body
grp_Crustaceans_Euchaeta_glacialis_Whole_body <- tar_read(outlier_biota_Crustaceans_Euchaeta_glacialis_Whole_body)
grp_Crustaceans_Euchaeta_glacialis_Whole_body$data <- grp_Crustaceans_Euchaeta_glacialis_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Euchaeta_glacialis_Whole_body_unit <- grp_Crustaceans_Euchaeta_glacialis_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Euchaeta_glacialis_Whole_body
#| fig-cap: "Crustaceans / Euchaeta glacialis / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Euchaeta_glacialis_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Euchaeta_glacialis_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Euchaeta_glacialis_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Euchaeta_glacialis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Euchaeta_glacialis_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Euchaeta_glacialis_Whole_body
#| fig-cap: "Crustaceans / Euchaeta glacialis / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Euchaeta_glacialis_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Euchaeta_glacialis_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Euchaeta_glacialis_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Euchaeta_glacialis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Euchaeta_glacialis_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Euchaeta_glacialis_Whole_body
#| fig-cap: "Crustaceans / Euchaeta glacialis / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Euchaeta_glacialis_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Euchaeta_glacialis_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Euchaeta_glacialis_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Euchaeta_glacialis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Euchaeta_glacialis_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Euchaeta_glacialis_Whole_body
#| fig-cap: "Crustaceans / Euchaeta glacialis / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Euchaeta_glacialis_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Euchaeta_glacialis_Whole_body_unit, ")")
)
```

## Euchaeta norvegica: Whole body

Crustaceans / Euchaeta norvegica / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Euchaeta_norvegica_Whole_body
grp_Crustaceans_Euchaeta_norvegica_Whole_body <- tar_read(outlier_biota_Crustaceans_Euchaeta_norvegica_Whole_body)
grp_Crustaceans_Euchaeta_norvegica_Whole_body$data <- grp_Crustaceans_Euchaeta_norvegica_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Euchaeta_norvegica_Whole_body_unit <- grp_Crustaceans_Euchaeta_norvegica_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Euchaeta_norvegica_Whole_body
#| fig-cap: "Crustaceans / Euchaeta norvegica / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Euchaeta_norvegica_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Euchaeta_norvegica_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Euchaeta_norvegica_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Euchaeta_norvegica_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Euchaeta_norvegica_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Euchaeta_norvegica_Whole_body
#| fig-cap: "Crustaceans / Euchaeta norvegica / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Euchaeta_norvegica_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Euchaeta_norvegica_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Euchaeta_norvegica_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Euchaeta_norvegica_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Euchaeta_norvegica_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Euchaeta_norvegica_Whole_body
#| fig-cap: "Crustaceans / Euchaeta norvegica / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Euchaeta_norvegica_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Euchaeta_norvegica_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Euchaeta_norvegica_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Euchaeta_norvegica_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Euchaeta_norvegica_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Euchaeta_norvegica_Whole_body
#| fig-cap: "Crustaceans / Euchaeta norvegica / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Euchaeta_norvegica_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Euchaeta_norvegica_Whole_body_unit, ")")
)
```

## Eukrohnia hamata: Whole body

Crustaceans / Eukrohnia hamata / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Eukrohnia_hamata_Whole_body
grp_Crustaceans_Eukrohnia_hamata_Whole_body <- tar_read(outlier_biota_Crustaceans_Eukrohnia_hamata_Whole_body)
grp_Crustaceans_Eukrohnia_hamata_Whole_body$data <- grp_Crustaceans_Eukrohnia_hamata_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Eukrohnia_hamata_Whole_body_unit <- grp_Crustaceans_Eukrohnia_hamata_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Eukrohnia_hamata_Whole_body
#| fig-cap: "Crustaceans / Eukrohnia hamata / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Eukrohnia_hamata_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Eukrohnia_hamata_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Eukrohnia_hamata_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Eukrohnia_hamata_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Eukrohnia_hamata_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Eukrohnia_hamata_Whole_body
#| fig-cap: "Crustaceans / Eukrohnia hamata / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Eukrohnia_hamata_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Eukrohnia_hamata_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Eukrohnia_hamata_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Eukrohnia_hamata_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Eukrohnia_hamata_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Eukrohnia_hamata_Whole_body
#| fig-cap: "Crustaceans / Eukrohnia hamata / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Eukrohnia_hamata_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Eukrohnia_hamata_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Eukrohnia_hamata_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Eukrohnia_hamata_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Eukrohnia_hamata_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Eukrohnia_hamata_Whole_body
#| fig-cap: "Crustaceans / Eukrohnia hamata / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Eukrohnia_hamata_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Eukrohnia_hamata_Whole_body_unit, ")")
)
```

## Eukronia hamata: Whole body

Crustaceans / Eukronia hamata / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Eukronia_hamata_Whole_body
grp_Crustaceans_Eukronia_hamata_Whole_body <- tar_read(outlier_biota_Crustaceans_Eukronia_hamata_Whole_body)
grp_Crustaceans_Eukronia_hamata_Whole_body$data <- grp_Crustaceans_Eukronia_hamata_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Eukronia_hamata_Whole_body_unit <- grp_Crustaceans_Eukronia_hamata_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Eukronia_hamata_Whole_body
#| fig-cap: "Crustaceans / Eukronia hamata / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Eukronia_hamata_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Eukronia_hamata_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Eukronia_hamata_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Eukronia_hamata_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Eukronia_hamata_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Eukronia_hamata_Whole_body
#| fig-cap: "Crustaceans / Eukronia hamata / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Eukronia_hamata_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Eukronia_hamata_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Eukronia_hamata_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Eukronia_hamata_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Eukronia_hamata_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Eukronia_hamata_Whole_body
#| fig-cap: "Crustaceans / Eukronia hamata / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Eukronia_hamata_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Eukronia_hamata_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Eukronia_hamata_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Eukronia_hamata_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Eukronia_hamata_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Eukronia_hamata_Whole_body
#| fig-cap: "Crustaceans / Eukronia hamata / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Eukronia_hamata_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Eukronia_hamata_Whole_body_unit, ")")
)
```

## Euphausiacea: Whole body

Crustaceans / Euphausiacea / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Euphausiacea_Whole_body
grp_Crustaceans_Euphausiacea_Whole_body <- tar_read(outlier_biota_Crustaceans_Euphausiacea_Whole_body)
grp_Crustaceans_Euphausiacea_Whole_body$data <- grp_Crustaceans_Euphausiacea_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Euphausiacea_Whole_body_unit <- grp_Crustaceans_Euphausiacea_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Euphausiacea_Whole_body
#| fig-cap: "Crustaceans / Euphausiacea / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Euphausiacea_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Euphausiacea_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Euphausiacea_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Euphausiacea_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Euphausiacea_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Euphausiacea_Whole_body
#| fig-cap: "Crustaceans / Euphausiacea / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Euphausiacea_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Euphausiacea_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Euphausiacea_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Euphausiacea_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Euphausiacea_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Euphausiacea_Whole_body
#| fig-cap: "Crustaceans / Euphausiacea / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Euphausiacea_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Euphausiacea_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Euphausiacea_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Euphausiacea_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Euphausiacea_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Euphausiacea_Whole_body
#| fig-cap: "Crustaceans / Euphausiacea / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Euphausiacea_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Euphausiacea_Whole_body_unit, ")")
)
```

## Gammarus oceanicus: Whole body

Crustaceans / Gammarus oceanicus / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Gammarus_oceanicus_Whole_body
grp_Crustaceans_Gammarus_oceanicus_Whole_body <- tar_read(outlier_biota_Crustaceans_Gammarus_oceanicus_Whole_body)
grp_Crustaceans_Gammarus_oceanicus_Whole_body$data <- grp_Crustaceans_Gammarus_oceanicus_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Gammarus_oceanicus_Whole_body_unit <- grp_Crustaceans_Gammarus_oceanicus_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Gammarus_oceanicus_Whole_body
#| fig-cap: "Crustaceans / Gammarus oceanicus / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Gammarus_oceanicus_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Gammarus_oceanicus_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Gammarus_oceanicus_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Gammarus_oceanicus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Gammarus_oceanicus_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Gammarus_oceanicus_Whole_body
#| fig-cap: "Crustaceans / Gammarus oceanicus / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Gammarus_oceanicus_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Gammarus_oceanicus_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Gammarus_oceanicus_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Gammarus_oceanicus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Gammarus_oceanicus_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Gammarus_oceanicus_Whole_body
#| fig-cap: "Crustaceans / Gammarus oceanicus / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Gammarus_oceanicus_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Gammarus_oceanicus_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Gammarus_oceanicus_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Gammarus_oceanicus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Gammarus_oceanicus_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Gammarus_oceanicus_Whole_body
#| fig-cap: "Crustaceans / Gammarus oceanicus / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Gammarus_oceanicus_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Gammarus_oceanicus_Whole_body_unit, ")")
)
```

## Hymenodora glacialis: Whole body

Crustaceans / Hymenodora glacialis / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Hymenodora_glacialis_Whole_body
grp_Crustaceans_Hymenodora_glacialis_Whole_body <- tar_read(outlier_biota_Crustaceans_Hymenodora_glacialis_Whole_body)
grp_Crustaceans_Hymenodora_glacialis_Whole_body$data <- grp_Crustaceans_Hymenodora_glacialis_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Hymenodora_glacialis_Whole_body_unit <- grp_Crustaceans_Hymenodora_glacialis_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Hymenodora_glacialis_Whole_body
#| fig-cap: "Crustaceans / Hymenodora glacialis / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Hymenodora_glacialis_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Hymenodora_glacialis_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Hymenodora_glacialis_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Hymenodora_glacialis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Hymenodora_glacialis_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Hymenodora_glacialis_Whole_body
#| fig-cap: "Crustaceans / Hymenodora glacialis / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Hymenodora_glacialis_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Hymenodora_glacialis_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Hymenodora_glacialis_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Hymenodora_glacialis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Hymenodora_glacialis_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Hymenodora_glacialis_Whole_body
#| fig-cap: "Crustaceans / Hymenodora glacialis / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Hymenodora_glacialis_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Hymenodora_glacialis_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Hymenodora_glacialis_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Hymenodora_glacialis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Hymenodora_glacialis_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Hymenodora_glacialis_Whole_body
#| fig-cap: "Crustaceans / Hymenodora glacialis / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Hymenodora_glacialis_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Hymenodora_glacialis_Whole_body_unit, ")")
)
```

## Isobactrus setosus: Whole body

Crustaceans / Isobactrus setosus / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Isobactrus_setosus_Whole_body
grp_Crustaceans_Isobactrus_setosus_Whole_body <- tar_read(outlier_biota_Crustaceans_Isobactrus_setosus_Whole_body)
grp_Crustaceans_Isobactrus_setosus_Whole_body$data <- grp_Crustaceans_Isobactrus_setosus_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Isobactrus_setosus_Whole_body_unit <- grp_Crustaceans_Isobactrus_setosus_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Isobactrus_setosus_Whole_body
#| fig-cap: "Crustaceans / Isobactrus setosus / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Isobactrus_setosus_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Isobactrus_setosus_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Isobactrus_setosus_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Isobactrus_setosus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Isobactrus_setosus_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Isobactrus_setosus_Whole_body
#| fig-cap: "Crustaceans / Isobactrus setosus / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Isobactrus_setosus_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Isobactrus_setosus_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Isobactrus_setosus_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Isobactrus_setosus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Isobactrus_setosus_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Isobactrus_setosus_Whole_body
#| fig-cap: "Crustaceans / Isobactrus setosus / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Isobactrus_setosus_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Isobactrus_setosus_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Isobactrus_setosus_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Isobactrus_setosus_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Isobactrus_setosus_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Isobactrus_setosus_Whole_body
#| fig-cap: "Crustaceans / Isobactrus setosus / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Isobactrus_setosus_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Isobactrus_setosus_Whole_body_unit, ")")
)
```

## Lithodidae: Total soft tissues

Crustaceans / Lithodidae / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Lithodidae_Total_soft_tissues
grp_Crustaceans_Lithodidae_Total_soft_tissues <- tar_read(outlier_biota_Crustaceans_Lithodidae_Total_soft_tissues)
grp_Crustaceans_Lithodidae_Total_soft_tissues$data <- grp_Crustaceans_Lithodidae_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Lithodidae_Total_soft_tissues_unit <- grp_Crustaceans_Lithodidae_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Lithodidae_Total_soft_tissues
#| fig-cap: "Crustaceans / Lithodidae / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Lithodidae_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Lithodidae_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Lithodidae_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Lithodidae_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Lithodidae_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Lithodidae_Total_soft_tissues
#| fig-cap: "Crustaceans / Lithodidae / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Lithodidae_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Lithodidae_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Lithodidae_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Lithodidae_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Lithodidae_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Lithodidae_Total_soft_tissues
#| fig-cap: "Crustaceans / Lithodidae / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Lithodidae_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Lithodidae_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Lithodidae_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Lithodidae_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Lithodidae_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Lithodidae_Total_soft_tissues
#| fig-cap: "Crustaceans / Lithodidae / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Lithodidae_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Crustaceans_Lithodidae_Total_soft_tissues_unit, ")")
)
```

## Meganyctiphanes norvegica: Whole body

Crustaceans / Meganyctiphanes norvegica / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Meganyctiphanes_norvegica_Whole_body
grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body <- tar_read(outlier_biota_Crustaceans_Meganyctiphanes_norvegica_Whole_body)
grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body$data <- grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body_unit <- grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Meganyctiphanes_norvegica_Whole_body
#| fig-cap: "Crustaceans / Meganyctiphanes norvegica / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Meganyctiphanes_norvegica_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Meganyctiphanes_norvegica_Whole_body
#| fig-cap: "Crustaceans / Meganyctiphanes norvegica / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Meganyctiphanes_norvegica_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Meganyctiphanes_norvegica_Whole_body
#| fig-cap: "Crustaceans / Meganyctiphanes norvegica / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Meganyctiphanes_norvegica_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Meganyctiphanes_norvegica_Whole_body
#| fig-cap: "Crustaceans / Meganyctiphanes norvegica / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Meganyctiphanes_norvegica_Whole_body_unit, ")")
)
```

## Metridia longa: Whole body

Crustaceans / Metridia longa / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Metridia_longa_Whole_body
grp_Crustaceans_Metridia_longa_Whole_body <- tar_read(outlier_biota_Crustaceans_Metridia_longa_Whole_body)
grp_Crustaceans_Metridia_longa_Whole_body$data <- grp_Crustaceans_Metridia_longa_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Metridia_longa_Whole_body_unit <- grp_Crustaceans_Metridia_longa_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Metridia_longa_Whole_body
#| fig-cap: "Crustaceans / Metridia longa / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Metridia_longa_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Metridia_longa_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Metridia_longa_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Metridia_longa_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Metridia_longa_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Metridia_longa_Whole_body
#| fig-cap: "Crustaceans / Metridia longa / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Metridia_longa_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Metridia_longa_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Metridia_longa_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Metridia_longa_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Metridia_longa_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Metridia_longa_Whole_body
#| fig-cap: "Crustaceans / Metridia longa / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Metridia_longa_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Metridia_longa_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Metridia_longa_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Metridia_longa_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Metridia_longa_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Metridia_longa_Whole_body
#| fig-cap: "Crustaceans / Metridia longa / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Metridia_longa_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Metridia_longa_Whole_body_unit, ")")
)
```

## Mysis relicta: Total soft tissues

Crustaceans / Mysis relicta / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Mysis_relicta_Total_soft_tissues
grp_Crustaceans_Mysis_relicta_Total_soft_tissues <- tar_read(outlier_biota_Crustaceans_Mysis_relicta_Total_soft_tissues)
grp_Crustaceans_Mysis_relicta_Total_soft_tissues$data <- grp_Crustaceans_Mysis_relicta_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Mysis_relicta_Total_soft_tissues_unit <- grp_Crustaceans_Mysis_relicta_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Mysis_relicta_Total_soft_tissues
#| fig-cap: "Crustaceans / Mysis relicta / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Mysis_relicta_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Mysis_relicta_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Mysis_relicta_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Mysis_relicta_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Mysis_relicta_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Mysis_relicta_Total_soft_tissues
#| fig-cap: "Crustaceans / Mysis relicta / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Mysis_relicta_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Mysis_relicta_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Mysis_relicta_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Mysis_relicta_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Mysis_relicta_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Mysis_relicta_Total_soft_tissues
#| fig-cap: "Crustaceans / Mysis relicta / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Mysis_relicta_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Mysis_relicta_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Mysis_relicta_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Mysis_relicta_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Mysis_relicta_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Mysis_relicta_Total_soft_tissues
#| fig-cap: "Crustaceans / Mysis relicta / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Mysis_relicta_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Crustaceans_Mysis_relicta_Total_soft_tissues_unit, ")")
)
```

## Nephrops norvegicus: Muscle

Crustaceans / Nephrops norvegicus / Muscle. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Nephrops_norvegicus_Muscle
grp_Crustaceans_Nephrops_norvegicus_Muscle <- tar_read(outlier_biota_Crustaceans_Nephrops_norvegicus_Muscle)
grp_Crustaceans_Nephrops_norvegicus_Muscle$data <- grp_Crustaceans_Nephrops_norvegicus_Muscle$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Nephrops_norvegicus_Muscle_unit <- grp_Crustaceans_Nephrops_norvegicus_Muscle$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Nephrops_norvegicus_Muscle
#| fig-cap: "Crustaceans / Nephrops norvegicus / Muscle: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Nephrops_norvegicus_Muscle$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Nephrops_norvegicus_Muscle$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Nephrops_norvegicus_Muscle_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Nephrops_norvegicus_Muscle
outlier_flag_table(flag_by_facet(grp_Crustaceans_Nephrops_norvegicus_Muscle$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Nephrops_norvegicus_Muscle
#| fig-cap: "Crustaceans / Nephrops norvegicus / Muscle: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Nephrops_norvegicus_Muscle$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Nephrops_norvegicus_Muscle$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Nephrops_norvegicus_Muscle_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Nephrops_norvegicus_Muscle
outlier_flag_table(flag_by_facet(grp_Crustaceans_Nephrops_norvegicus_Muscle$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Nephrops_norvegicus_Muscle
#| fig-cap: "Crustaceans / Nephrops norvegicus / Muscle: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Nephrops_norvegicus_Muscle$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Nephrops_norvegicus_Muscle$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Nephrops_norvegicus_Muscle_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Nephrops_norvegicus_Muscle
outlier_flag_table(flag_by_facet(grp_Crustaceans_Nephrops_norvegicus_Muscle$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Nephrops_norvegicus_Muscle
#| fig-cap: "Crustaceans / Nephrops norvegicus / Muscle: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Nephrops_norvegicus_Muscle,
  x_label = paste0("Measured value (", grp_Crustaceans_Nephrops_norvegicus_Muscle_unit, ")")
)
```

## Nucella lapillus: Total soft tissues

Crustaceans / Nucella lapillus / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Nucella_lapillus_Total_soft_tissues
grp_Crustaceans_Nucella_lapillus_Total_soft_tissues <- tar_read(outlier_biota_Crustaceans_Nucella_lapillus_Total_soft_tissues)
grp_Crustaceans_Nucella_lapillus_Total_soft_tissues$data <- grp_Crustaceans_Nucella_lapillus_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Nucella_lapillus_Total_soft_tissues_unit <- grp_Crustaceans_Nucella_lapillus_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Nucella_lapillus_Total_soft_tissues
#| fig-cap: "Crustaceans / Nucella lapillus / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Nucella_lapillus_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Nucella_lapillus_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Nucella_lapillus_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Nucella_lapillus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Nucella_lapillus_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Nucella_lapillus_Total_soft_tissues
#| fig-cap: "Crustaceans / Nucella lapillus / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Nucella_lapillus_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Nucella_lapillus_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Nucella_lapillus_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Nucella_lapillus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Nucella_lapillus_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Nucella_lapillus_Total_soft_tissues
#| fig-cap: "Crustaceans / Nucella lapillus / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Nucella_lapillus_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Nucella_lapillus_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Nucella_lapillus_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Nucella_lapillus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Nucella_lapillus_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Nucella_lapillus_Total_soft_tissues
#| fig-cap: "Crustaceans / Nucella lapillus / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Nucella_lapillus_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Crustaceans_Nucella_lapillus_Total_soft_tissues_unit, ")")
)
```

## Palaemon adspersus: Total soft tissues

Crustaceans / Palaemon adspersus / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Palaemon_adspersus_Total_soft_tissues
grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues <- tar_read(outlier_biota_Crustaceans_Palaemon_adspersus_Total_soft_tissues)
grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues$data <- grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues_unit <- grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Palaemon_adspersus_Total_soft_tissues
#| fig-cap: "Crustaceans / Palaemon adspersus / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Palaemon_adspersus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Palaemon_adspersus_Total_soft_tissues
#| fig-cap: "Crustaceans / Palaemon adspersus / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Palaemon_adspersus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Palaemon_adspersus_Total_soft_tissues
#| fig-cap: "Crustaceans / Palaemon adspersus / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Palaemon_adspersus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Palaemon_adspersus_Total_soft_tissues
#| fig-cap: "Crustaceans / Palaemon adspersus / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Crustaceans_Palaemon_adspersus_Total_soft_tissues_unit, ")")
)
```

## Pandalus borealis: Muscle

Crustaceans / Pandalus borealis / Muscle. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Pandalus_borealis_Muscle
grp_Crustaceans_Pandalus_borealis_Muscle <- tar_read(outlier_biota_Crustaceans_Pandalus_borealis_Muscle)
grp_Crustaceans_Pandalus_borealis_Muscle$data <- grp_Crustaceans_Pandalus_borealis_Muscle$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Pandalus_borealis_Muscle_unit <- grp_Crustaceans_Pandalus_borealis_Muscle$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Pandalus_borealis_Muscle
#| fig-cap: "Crustaceans / Pandalus borealis / Muscle: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Pandalus_borealis_Muscle$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Pandalus_borealis_Muscle$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Pandalus_borealis_Muscle_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Pandalus_borealis_Muscle
outlier_flag_table(flag_by_facet(grp_Crustaceans_Pandalus_borealis_Muscle$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Pandalus_borealis_Muscle
#| fig-cap: "Crustaceans / Pandalus borealis / Muscle: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Pandalus_borealis_Muscle$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Pandalus_borealis_Muscle$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Pandalus_borealis_Muscle_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Pandalus_borealis_Muscle
outlier_flag_table(flag_by_facet(grp_Crustaceans_Pandalus_borealis_Muscle$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Pandalus_borealis_Muscle
#| fig-cap: "Crustaceans / Pandalus borealis / Muscle: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Pandalus_borealis_Muscle$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Pandalus_borealis_Muscle$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Pandalus_borealis_Muscle_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Pandalus_borealis_Muscle
outlier_flag_table(flag_by_facet(grp_Crustaceans_Pandalus_borealis_Muscle$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Pandalus_borealis_Muscle
#| fig-cap: "Crustaceans / Pandalus borealis / Muscle: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Pandalus_borealis_Muscle,
  x_label = paste0("Measured value (", grp_Crustaceans_Pandalus_borealis_Muscle_unit, ")")
)
```

## Pandalus borealis: Total soft tissues

Crustaceans / Pandalus borealis / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Pandalus_borealis_Total_soft_tissues
grp_Crustaceans_Pandalus_borealis_Total_soft_tissues <- tar_read(outlier_biota_Crustaceans_Pandalus_borealis_Total_soft_tissues)
grp_Crustaceans_Pandalus_borealis_Total_soft_tissues$data <- grp_Crustaceans_Pandalus_borealis_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Pandalus_borealis_Total_soft_tissues_unit <- grp_Crustaceans_Pandalus_borealis_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Pandalus_borealis_Total_soft_tissues
#| fig-cap: "Crustaceans / Pandalus borealis / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Pandalus_borealis_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Pandalus_borealis_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Pandalus_borealis_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Pandalus_borealis_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Pandalus_borealis_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Pandalus_borealis_Total_soft_tissues
#| fig-cap: "Crustaceans / Pandalus borealis / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Pandalus_borealis_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Pandalus_borealis_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Pandalus_borealis_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Pandalus_borealis_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Pandalus_borealis_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Pandalus_borealis_Total_soft_tissues
#| fig-cap: "Crustaceans / Pandalus borealis / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Pandalus_borealis_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Pandalus_borealis_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Pandalus_borealis_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Pandalus_borealis_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Crustaceans_Pandalus_borealis_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Pandalus_borealis_Total_soft_tissues
#| fig-cap: "Crustaceans / Pandalus borealis / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Pandalus_borealis_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Crustaceans_Pandalus_borealis_Total_soft_tissues_unit, ")")
)
```

## Pandalus borealis: Whole body

Crustaceans / Pandalus borealis / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Pandalus_borealis_Whole_body
grp_Crustaceans_Pandalus_borealis_Whole_body <- tar_read(outlier_biota_Crustaceans_Pandalus_borealis_Whole_body)
grp_Crustaceans_Pandalus_borealis_Whole_body$data <- grp_Crustaceans_Pandalus_borealis_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Pandalus_borealis_Whole_body_unit <- grp_Crustaceans_Pandalus_borealis_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Pandalus_borealis_Whole_body
#| fig-cap: "Crustaceans / Pandalus borealis / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Pandalus_borealis_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Pandalus_borealis_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Pandalus_borealis_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Pandalus_borealis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Pandalus_borealis_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Pandalus_borealis_Whole_body
#| fig-cap: "Crustaceans / Pandalus borealis / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Pandalus_borealis_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Pandalus_borealis_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Pandalus_borealis_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Pandalus_borealis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Pandalus_borealis_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Pandalus_borealis_Whole_body
#| fig-cap: "Crustaceans / Pandalus borealis / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Pandalus_borealis_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Pandalus_borealis_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Pandalus_borealis_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Pandalus_borealis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Pandalus_borealis_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Pandalus_borealis_Whole_body
#| fig-cap: "Crustaceans / Pandalus borealis / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Pandalus_borealis_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Pandalus_borealis_Whole_body_unit, ")")
)
```

## Sabinea sarsi: Whole body

Crustaceans / Sabinea sarsi / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Sabinea_sarsi_Whole_body
grp_Crustaceans_Sabinea_sarsi_Whole_body <- tar_read(outlier_biota_Crustaceans_Sabinea_sarsi_Whole_body)
grp_Crustaceans_Sabinea_sarsi_Whole_body$data <- grp_Crustaceans_Sabinea_sarsi_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Sabinea_sarsi_Whole_body_unit <- grp_Crustaceans_Sabinea_sarsi_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Sabinea_sarsi_Whole_body
#| fig-cap: "Crustaceans / Sabinea sarsi / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Sabinea_sarsi_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Sabinea_sarsi_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Sabinea_sarsi_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Sabinea_sarsi_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Sabinea_sarsi_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Sabinea_sarsi_Whole_body
#| fig-cap: "Crustaceans / Sabinea sarsi / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Sabinea_sarsi_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Sabinea_sarsi_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Sabinea_sarsi_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Sabinea_sarsi_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Sabinea_sarsi_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Sabinea_sarsi_Whole_body
#| fig-cap: "Crustaceans / Sabinea sarsi / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Sabinea_sarsi_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Sabinea_sarsi_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Sabinea_sarsi_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Sabinea_sarsi_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Sabinea_sarsi_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Sabinea_sarsi_Whole_body
#| fig-cap: "Crustaceans / Sabinea sarsi / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Sabinea_sarsi_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Sabinea_sarsi_Whole_body_unit, ")")
)
```

## Themisto abyssorum: Whole body

Crustaceans / Themisto abyssorum / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Themisto_abyssorum_Whole_body
grp_Crustaceans_Themisto_abyssorum_Whole_body <- tar_read(outlier_biota_Crustaceans_Themisto_abyssorum_Whole_body)
grp_Crustaceans_Themisto_abyssorum_Whole_body$data <- grp_Crustaceans_Themisto_abyssorum_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Themisto_abyssorum_Whole_body_unit <- grp_Crustaceans_Themisto_abyssorum_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Themisto_abyssorum_Whole_body
#| fig-cap: "Crustaceans / Themisto abyssorum / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Themisto_abyssorum_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Themisto_abyssorum_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Themisto_abyssorum_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Themisto_abyssorum_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Themisto_abyssorum_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Themisto_abyssorum_Whole_body
#| fig-cap: "Crustaceans / Themisto abyssorum / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Themisto_abyssorum_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Themisto_abyssorum_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Themisto_abyssorum_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Themisto_abyssorum_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Themisto_abyssorum_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Themisto_abyssorum_Whole_body
#| fig-cap: "Crustaceans / Themisto abyssorum / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Themisto_abyssorum_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Themisto_abyssorum_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Themisto_abyssorum_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Themisto_abyssorum_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Themisto_abyssorum_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Themisto_abyssorum_Whole_body
#| fig-cap: "Crustaceans / Themisto abyssorum / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Themisto_abyssorum_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Themisto_abyssorum_Whole_body_unit, ")")
)
```

## Themisto libellula: Whole body

Crustaceans / Themisto libellula / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Themisto_libellula_Whole_body
grp_Crustaceans_Themisto_libellula_Whole_body <- tar_read(outlier_biota_Crustaceans_Themisto_libellula_Whole_body)
grp_Crustaceans_Themisto_libellula_Whole_body$data <- grp_Crustaceans_Themisto_libellula_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Themisto_libellula_Whole_body_unit <- grp_Crustaceans_Themisto_libellula_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Themisto_libellula_Whole_body
#| fig-cap: "Crustaceans / Themisto libellula / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Themisto_libellula_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Themisto_libellula_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Themisto_libellula_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Themisto_libellula_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Themisto_libellula_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Themisto_libellula_Whole_body
#| fig-cap: "Crustaceans / Themisto libellula / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Themisto_libellula_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Themisto_libellula_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Themisto_libellula_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Themisto_libellula_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Themisto_libellula_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Themisto_libellula_Whole_body
#| fig-cap: "Crustaceans / Themisto libellula / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Themisto_libellula_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Themisto_libellula_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Themisto_libellula_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Themisto_libellula_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Themisto_libellula_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Themisto_libellula_Whole_body
#| fig-cap: "Crustaceans / Themisto libellula / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Themisto_libellula_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Themisto_libellula_Whole_body_unit, ")")
)
```

## Thysanoessa inermis: Whole body

Crustaceans / Thysanoessa inermis / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Crustaceans_Thysanoessa_inermis_Whole_body
grp_Crustaceans_Thysanoessa_inermis_Whole_body <- tar_read(outlier_biota_Crustaceans_Thysanoessa_inermis_Whole_body)
grp_Crustaceans_Thysanoessa_inermis_Whole_body$data <- grp_Crustaceans_Thysanoessa_inermis_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Crustaceans_Thysanoessa_inermis_Whole_body_unit <- grp_Crustaceans_Thysanoessa_inermis_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Crustaceans_Thysanoessa_inermis_Whole_body
#| fig-cap: "Crustaceans / Thysanoessa inermis / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Thysanoessa_inermis_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Crustaceans_Thysanoessa_inermis_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Crustaceans_Thysanoessa_inermis_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Crustaceans_Thysanoessa_inermis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Thysanoessa_inermis_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Crustaceans_Thysanoessa_inermis_Whole_body
#| fig-cap: "Crustaceans / Thysanoessa inermis / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Thysanoessa_inermis_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Crustaceans_Thysanoessa_inermis_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Crustaceans_Thysanoessa_inermis_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Crustaceans_Thysanoessa_inermis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Thysanoessa_inermis_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Crustaceans_Thysanoessa_inermis_Whole_body
#| fig-cap: "Crustaceans / Thysanoessa inermis / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Crustaceans_Thysanoessa_inermis_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Crustaceans_Thysanoessa_inermis_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Crustaceans_Thysanoessa_inermis_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Crustaceans_Thysanoessa_inermis_Whole_body
outlier_flag_table(flag_by_facet(grp_Crustaceans_Thysanoessa_inermis_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Crustaceans_Thysanoessa_inermis_Whole_body
#| fig-cap: "Crustaceans / Thysanoessa inermis / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Crustaceans_Thysanoessa_inermis_Whole_body,
  x_label = paste0("Measured value (", grp_Crustaceans_Thysanoessa_inermis_Whole_body_unit, ")")
)
```
