---
title: "Fish"
format:
  html: 
    fig-width: 16
    fig-height: 20
  docx:
    reference-doc: ../custom-reference-doc.docx    
    fig-width: 16
    fig-dpi: 300                     # high res
    fig-format: png                  # png is fine; svg not well supported in docx
    link-citations: true             # TODO: Not working... why?
    link-bibliography: true

execute:
  echo: false
  warning: false                     # suppresses warnings even when echo is off
  message: false                     # suppresses package load messages (library() etc.)
  error: false                       # prevents errors from stopping the render

knitr:
  opts_chunk:
    dpi: 300                         # belt-and-suspenders for knitr devices
    dev: "ragg_png"                  # use ragg for better font rendering

filters:
  - flextable-qmd                    # needed, or the flextables render crazy 
  - at: post-render
    path: ../_extensions/flextable-qmd/unwrap-float.lua
---

```{r}
#| label: setup
# == Libraries ----
library(tidyverse)
library(targets)
library(hexbin)
library(ggrepel)
library(viridis)
library(patchwork)
library(ggnewscale)
library(igraph)
library(sf)
library(gridExtra)
library(grid)
library(knitr)
library(ggtext)
library(flextable)
library(glue)

opts_chunk$set(dev = "svglite")

pkgload::load_all(quiet = TRUE)

set_flextable_defaults(
  font.size = 10,
  font.family = "Aptos",
  digits = 2,
  theme_fun = theme_vanilla,
  padding = 1
)

ggplot_base_size <- 20
theme_set(theme_minimal(base_size = ggplot_base_size, base_family = "Aptos"))

# Reusable plot layers
common_dotsize <- 3
common_binwidth <- 0.01
common_group_height <- 0.5 # height of a single boxplot "row"

outlier_boxplot <- geom_boxplot(
  outliers = FALSE,
  whisker.colour = "darkgrey",
  staple.colour = "darkgrey",
  box.colour = "darkgrey"
)

outlier_plot_theme <- theme(
  plot.title = element_markdown(),
  axis.text.y.left = element_markdown(),
  panel.background = element_blank(),
  panel.grid.major = element_line(colour = "grey90"),
  panel.grid.minor = element_line(colour = "grey95"),
  plot.background = element_blank(),
  legend.background = element_blank(),
  legend.key = element_blank(),
  strip.background = element_blank(),
  axis.ticks = element_blank()
)

# == Shared map data ----
world_map <- map_data("world")

# == Study area bbox ----
# xmin/xmax: longitude; ymin/ymax: latitude
STUDY_BBOX <- c(xmin = -65, xmax = 40, ymin = 55, ymax = 75)

# Helper: append coord_sf with fixed study area to a ggplot object.
# ratio = 2 approximates equal-area appearance at these latitudes.
study_coords <- function(p, ratio = 2) {
  coord_sf(
    xlim = c(STUDY_BBOX[["xmin"]], STUDY_BBOX[["xmax"]]),
    ylim = c(STUDY_BBOX[["ymin"]], STUDY_BBOX[["ymax"]]),
    expand = FALSE
  )
}

# == Outlier filter ----
# Z-score threshold for excluding extreme values.
# Applied locally per matrix using MEASURED_VALUE_STANDARD.
# MEASURED_VALUE is retained in the data for traceability only; never used for analysis.
Z_THRESHOLD <- 3

is_outlier <- function(x, threshold = Z_THRESHOLD) {
  abs((x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)) > threshold
}

# == Site comment cleaning patterns ----
# NOTE: \p{Lu} requires Unicode-aware regex; stringr::regex() handles this correctly.
remove_patterns <- c(
  # URLs / emails
  "(?i)https?://\\S+|www\\.\\S+|\\S+@\\S+",
  # Parenthetical / bracket content
  "\\s*\\([^)]*\\)|\\s*\\[[^]]*\\]",
  # Report codes e.g. 01.01-385, 2021/09
  "\\b\\d{1,4}(?:[\\.\\-/]\\d{1,4})+\\b",
  # Station / sample codes e.g. L1, L2, Stasjon 3
  "\\bL\\d+\\b|\\bStasjon\\s*\\d+\\b|\\bStation\\s*\\d+\\b",
  # Short numeric tokens (excludes 4-digit years)
  "\\b(?!\\d{4}\\b)\\d{1,3}\\b",
  # Report words (Norwegian / English)
  "(?i)rapport(?:\\s+nr\\.?|nr\\.?|nummer|:)?.*?\\b|(?i)report(?:\\s+no\\.?|\\s+nr\\.?|\\s+nr)?",
  "(?i)rapporter?\\b|(?i)reference(?:\\s+site)?\\b|\\bref\\.?\\b",
  # Vannmiljø / resipient noise
  "(?i)vannmilj[oø]\\b|\\bvm\\b|(?i)resipient(?:unders[kq]økelse|unde)?",
  "(?i)vm\\s*emission\\s*|(?i)miljøgifter\\b|(?i)miljøgifter\\s+i",
  # Contractor / org names
  "(?i)akvaplan-?niva|niva\\b|nina\\b|ramb[ao]ll\\b|norconsult\\b|asplan\\s+viak\\b|ngi\\b|ng\\&i\\b|APN|AS",
  "\\b\\p{Lu}[\\p{L}''\\-]+(?:\\s+\\p{Lu}[\\p{L}''\\-]+)*\\s+AS\\b",
  # Page / folio refs
  "\\bpg?\\.?\\s*\\d+\\b|\\bp\\.\\s*\\d+\\b",
  # Unit / measurement fragments
  "\\bmg/kg\\b|\\bmg/kg\\s*dry\\b|\\bmg/kg\\s*wet\\b|\\bSD\\b|\\bStandard\\s+Deviat(?:ion|ions)\\b",
  # Stray punctuation
  "[,;:\\|]{2,}|--+|\\.\\.+",
  # Literal noise strings
  "Original Comment:",
  "Emission :",
  "rsøkelse",
  "Vannmiljø Station",
  "Station"
)

clean_regex <- str_c(remove_patterns, collapse = "|")
```

As for algae, but for fish. There are a lot more or these, and there's a lot more tissue differentiation.

# Fish

```{r}
#| label: tbl-fish-variation-summary
fish <- tar_read(load_literature_pqt) |>
  filter(SPECIES_GROUP == "Fish")

fish |>
  group_by(
    SAMPLE_SPECIES,
    REFERENCE_ID,
    SITE_GEOGRAPHIC_FEATURE,
    SAMPLE_TISSUE
  ) |>
  reframe(n = n()) |>
  arrange(desc(n)) |>
  flextable() |>
  autofit() |>
  set_header_labels(
    SAMPLE_SPECIES = "Species",
    REFERENCE_ID = "Reference",
    SITE_GEOGRAPHIC_FEATURE = "Geographic Feature",
    SAMPLE_TISSUE = "Tissue",
    n = "n"
  )
```

We have a great deal of fish data. Broadly speaking, we can group it by three factors: - Tissue (liver, gill, whole body, muscle, and in one case bile) - Site type (water ways, water bodies, coastal waters, and rarely, open ocean) - Species

It would also be useful to know the rough size, trophic level, lifestyle, etc. of these fish. We will attend to that later, though.

# Distribution of Copper Concentrations

No removal of outliers at this stage. We do need to check for unit types, though.

Two important questions:

1.  What units are used?
2.  Are they correlated with anything?

```{r}
#| label: tbl_fish_units
fish_units_exploration <- fish |>
  reframe(
    .by = c(
      'MEASURED_UNIT_STANDARD',
      "SAMPLE_SPECIES",
      "SAMPLE_TISSUE",
      "SITE_GEOGRAPHIC_FEATURE"
    ),
    mean = mean(MEASURED_VALUE_STANDARD),
    sd = sd(MEASURED_VALUE_STANDARD),
    n = sum(MEASURED_N)
  ) |>
  arrange(desc(n))

fish_units_exploration |>
  flextable() |>
  colformat_double(j = c("mean", "sd"), digits = 2) |>
  autofit() |>
  set_header_labels(
    MEASURED_UNIT_STANDARD = "Unit",
    SAMPLE_SPECIES = "Species",
    SAMPLE_TISSUE = "Tissue",
    SITE_GEOGRAPHIC_FEATURE = "Geographic Feature",
    mean = "Mean",
    sd = "SD",
    n = "n"
  )
```

For a subset of species/tissues/locations we have both dry and wet weights and can calculate some (rough) conversion factors.

```{r}
#| label: tbl_fish_units_pairwise
fish_units_lookup <- fish_units_exploration |>
  group_by(SAMPLE_SPECIES, SAMPLE_TISSUE, SITE_GEOGRAPHIC_FEATURE) |>
  filter(n() >= 2) |>
  pivot_wider(names_from = MEASURED_UNIT_STANDARD, values_from = mean) |>
  reframe(
    wet_over_dry = max(`mg/kg (wet)`, na.rm = TRUE) /
      max(`mg/kg (dry)`, na.rm = TRUE),
    n = sum(n)
  ) |>
  add_row(
    SAMPLE_SPECIES = "Generic",
    SAMPLE_TISSUE = "Generic",
    SITE_GEOGRAPHIC_FEATURE = "Generic",
    wet_over_dry = 5,
    n = NA
  )

fish_units_lookup |>
  flextable() |>
  colformat_double(j = c("wet_over_dry"), digits = 2) |>
  autofit() |>
  set_header_labels(
    SAMPLE_SPECIES = "Species",
    SAMPLE_TISSUE = "Tissue",
    SITE_GEOGRAPHIC_FEATURE = "Geographic Feature",
    wet_over_dry = "Wet/Dry Ratio",
    n = "n"
  )
```

Roughly 10% of our samples (537) are dry, the remainder (3965) are wet weight. According to @cressonUnderestimationChemicalContamination2017, a conversion factor of 1:5 for dry:wet weight is typically used, but individual species vary considerably from this. Under the circumstances, we will use the calculated conversion factors for *G. morhua*, *Melanogrammus aeglefinus*, and *Pollachius pollachius*, and the default figure of 5 (despite its obvious flaws) for the rest.

```{r}
#| label: fig-fish-pointrange
threshold_data <- generate_copper_thresholds() |>
  filter(SAMPLE_SPECIES == "Gadus morhua", SAMPLE_TISSUE == "Liver") # only one background threshold; for G. morhua livers

fish_dry_normalised <- fish |>
  mutate(
    # manual, since we have so few values
    SPECIES_DRY_WET_RATIO = case_when(
      SAMPLE_SPECIES == "Gadus morhua" & SAMPLE_TISSUE == "Liver" ~ 10.9,
      SAMPLE_SPECIES == "Melanogrammus aeglefinus" &
        SAMPLE_TISSUE == "Liver" ~ 1.11,
      SAMPLE_SPECIES == "Pollachius pollachius" &
        SAMPLE_TISSUE == "Liver" ~ 1.34,
      .default = 5
    ),
    # when unit is dry, multiply by the relevant ratio, otherwise leave untouched
    MEASURED_VALUE_STANDARD_NORMALISED = case_when(
      MEASURED_UNIT_STANDARD == "mg/kg (dry)" ~ MEASURED_VALUE_STANDARD *
        SPECIES_DRY_WET_RATIO,
      .default = MEASURED_VALUE_STANDARD
    ),
    UNCERTAINTY_UPPER_STANDARD_NORMALISED = case_when(
      MEASURED_UNIT_STANDARD == "mg/kg (dry)" ~ UNCERTAINTY_UPPER_STANDARD *
        SPECIES_DRY_WET_RATIO,
      .default = UNCERTAINTY_UPPER_STANDARD
    ),
    UNCERTAINTY_LOWER_STANDARD_NORMALISED = case_when(
      MEASURED_UNIT_STANDARD == "mg/kg (dry)" ~ UNCERTAINTY_LOWER_STANDARD *
        SPECIES_DRY_WET_RATIO,
      .default = SPECIES_DRY_WET_RATIO
    ),
    MEASURED_UNIT_STANDARD_NORMALISED = "mg/kg (wet)"
  )

nrow(fish)
nrow(fish_dry_normalised)

fish_pointrange <- fish_dry_normalised |>
  mutate(
    SITE_COMMENT = SITE_COMMENT |> str_remove("^.*?\\."),
    SITE_NAME = SITE_NAME |> str_remove("Vannmiljø Station")
  ) |>
  ggplot(aes(
    y = SITE_NAME,
    x = MEASURED_VALUE_STANDARD_NORMALISED,
    colour = MEASURED_VALUE_STANDARD_NORMALISED
  )) +
  scale_colour_viridis_c() +
  geom_pointrange(aes(
    xmin = UNCERTAINTY_LOWER_STANDARD_NORMALISED,
    xmax = UNCERTAINTY_UPPER_STANDARD_NORMALISED
  )) +
  new_scale_color() +
  geom_vline(
    data = threshold_data,
    aes(xintercept = THRESHOLD_VALUE, color = TITLE_SHORT),
    linetype = "dashed",
    linewidth = 0.8,
    alpha = 0.6
  ) +
  geom_text(
    aes(
      label = str_wrap(SITE_COMMENT),
      x = UNCERTAINTY_UPPER_STANDARD_NORMALISED + 0.5
    ),
    colour = "black",
    hjust = 0
  ) +
  labs(
    y = "Site Code",
    x = "Measured Value (standardised to mg/kg (wet))"
  ) +
  outlier_plot_theme +
  theme(legend.position = "none")


fish_pointrange
```

We have some improbably high values in some cases, all from Vannmiljø. Let's look at these by species/tissue also:

```{r}
#| label: fig-fish-outliers-boxplot
fish_boxplot <- fish_dry_normalised |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD_NORMALISED, y = SAMPLE_SPECIES)) +
  # geom_box uses > mean( + abs(IQR * 1.5)) as the cut-off
  # TODO: This or z-score?
  geom_boxplot(outlier.colour = "red") +
  scale_x_log10(
    breaks = 10^(-1:5),
    labels = scales::label_log(),
    minor_breaks = outer(1:9, 10^(-1:4)) |> as.vector()
  ) +
  facet_wrap(
    facets = vars(SAMPLE_TISSUE),
    ncol = 1,
    scales = "free_y",
    space = "free_y"
  ) +
  labs(
    x = "Measured Value (standardised to mg/kg (wet))"
  ) +
  outlier_plot_theme

fish_boxplot
```

# TODO: Best approach for outlier removal?

```{r}
#| label: fig-fish-outliers
fish_outliers <- fish_dry_normalised |>
  filter(is_outlier(MEASURED_VALUE_STANDARD_NORMALISED)) # Z > 3

fish_outliers |>
  mutate(
    SITE_COMMENT = SITE_COMMENT |> str_remove("^.*?\\."),
    SITE_NAME = SITE_NAME |> str_remove("Vannmiljø Station")
  ) |>
  ggplot(aes(
    y = SITE_NAME,
    x = MEASURED_VALUE_STANDARD,
    colour = MEASURED_VALUE_STANDARD
  )) +
  scale_colour_viridis_c() +
  geom_pointrange(aes(
    xmin = UNCERTAINTY_LOWER,
    xmax = UNCERTAINTY_UPPER,
    shape = SAMPLE_TISSUE
  )) +
  new_scale_color() +
  geom_vline(
    data = threshold_data,
    aes(xintercept = THRESHOLD_VALUE, color = TITLE_SHORT),
    linetype = "dashed",
    linewidth = 0.8,
    alpha = 0.6
  ) +
  geom_text(
    aes(
      label = str_wrap(str_glue("{SAMPLE_SPECIES}, {SAMPLE_TISSUE}")),
      x = 40000
    ),
    colour = "black",
    hjust = 0
  ) +
  labs(y = "Site Code", x = "Measured Value (mg/kg (wet))") +
  facet_wrap(
    facets = vars(MEASURED_UNIT_STANDARD),
    ncol = 1,
    scales = "free_y",
    space = "free_y"
  ) +
  outlier_plot_theme

fish_filtered <- fish_dry_normalised |>
  filter(!is_outlier(MEASURED_VALUE_STANDARD))

nrow(fish_filtered)
```

10,000 mg of copper is 10 g copper per kg of liver. The most likely explanation here is that values were mistakenly entered with the wrong unit. However, since only a small proportion (17 rows) of data points show such irregularly high values, the most pragmatic option is to simply exclude them.

# Lumping and Splitting

We have a lot of fish data, and a lot of different species and tissues. However, data are significantly biased towards a small number of species.

```{r}
#| label: tbl-fish-species-summary
fish_species_n <- fish_filtered |>
  group_by(
    SAMPLE_SPECIES
  ) |>
  reframe(n = n()) |>
  arrange(desc(n))

fish_species_n_sum <- sum(fish_species_n$n)

fish_species_n <- fish_species_n |>
  mutate(prop = n / fish_species_n_sum)
```

From a trophic perspective we'd like to have species that represent different parts of the food web (although we wouldn't expect this to be important for copper). In practical terms, 72% of our dataset is *Gadus morhua* (Atlantic cod), while *Salmo trutta* (Brown trout), *Ctenolabrus rupestris* (Goldsinny-wrasse), and *Coregonus lavaretus* (Common whitefish) represent 3-5% of the total each. We collect some information on each species to determine their relevance to our case.

```{r}
#| label: tbl-fish-species-info
#| tbl-cap: "Most common fish species in data."
tibble(
  species = c(
    "Gadus morhua (Atlantic cod)",
    "Salmo trutta (Brown or Sea trout)",
    "Ctenolabrus rupestris (Goldsinny-wrasse)",
    "Coregonus lavaretus (Common whitefish)"
  ),
  n = c("3146 (72%)", "204 (5%)", "192 (4%)", "139 (3%)"),
  distribution = c(
    "High occurrence above Arctic circle. Found from shoreline to edge of continental slope.",
    "Found across Norwegian & Icelandic coasts, including above Arctic circle. Freshwater/coastal lifecycle.",
    "Present in Norwegian territorial waters, including above Arctic circle.",
    "Freshwater sites across Norway, including above Arctic circle."
  ),
  trophic = c(
    "Predators of small fish and inverts, consumed by larger fish, aquatic mammals, squid.",
    "Larger (\u226572 cm) predator of small fish and inverts. Consumed by some larger fish and seals.",
    "Smaller (\u226418 cm) predator of invertebrates.",
    "Medium (~40 cm) predator of planktonic and benthic crustaceans."
  ),
  notes = c(
    "IUCN Red List Vulnerable. Important and overfished commercial species.",
    "IUCN Least Concern. Widely eaten.",
    "IUCN Least Concern. Used as a cleaner fish in salmon culture.",
    "IUCN Vulnerable. Some ambiguity over species identification."
  ),
  source = c(
    "marinespecies.org; fishbase.org",
    "marinespecies.org; fishbase.org",
    "marinespecies.org; fishbase.org",
    "artsdatabanken.no; fishbase.org"
  )
) |>
  flextable() |>
  italic(j = "species") |>
  autofit() |>
  set_header_labels(
    species = "Species",
    n = "n (% of total)",
    distribution = "Distribution",
    trophic = "Trophic Level",
    notes = "Notes",
    source = "Source"
  )
```

Now, with that said - is there a significant difference between these species?

```{r}
#| label: fish-species-variation
fish_filtered |>
  filter(
    SAMPLE_SPECIES %in%
      c(
        "Gadus morhua",
        "Salmo trutta",
        "Coregonus lavaretus",
        "Ctenolabrus rupestris"
      )
  ) |>
  ggplot(
    mapping = aes(x = MEASURED_VALUE_STANDARD_NORMALISED, y = SAMPLE_SPECIES)
  ) +
  geom_point() +
  outlier_boxplot +
  scale_x_log10(
    breaks = 10^(-1:5),
    labels = scales::label_log(),
    minor_breaks = outer(1:9, 10^(-1:4)) |> as.vector()
  ) +
  facet_wrap(
    facets = vars(SAMPLE_TISSUE),
    ncol = 1,
    scales = "free_y",
    space = "free_y"
  ) +
  labs(
    x = "Measured Value (mg/kg (wet))",
    y = "Species"
  ) +
  outlier_plot_theme
```