---
title: "Notebook XX - Sørfjorden Case Study"
date: last-modified
lightbox: true
format: html
execute:
  freeze: false   # parked 2026-07-29, see _quarto.yml render list
---

```{r}
#| label: setup
#| include: false
#| warning: false

library(targets)
library(here)
library(tidyverse)
library(flextable)
pkgload::load_all(quiet = TRUE) # sample_groups_flextable() etc.

here::i_am("index.qmd")
```

Sørfjorden ("The South Fjord") is the southerly arm of Hardangerfjorden on Norway's South-West coast [@thorsnaesSorfjordenHardanger2024]. In the early 1900s a number of heavy industrial plants began operation on the fjord's shores, and it became unusually contaminated with heavy metals [@ruusMonitoringEnvironmentalQuality2010]. Copper is known to have been emitted to Sørfjord as a byproduct of zinc production in the area, and unusually high copper concentrations have been detected in the tissues of fish caught in the area [@ruusMonitoringEnvironmentalQuality2010]

A003,AEP-003 - Sørfjorden,"Bounding box 59.77-60.37 N, 6.25-6.85 E, measured against load_literature_pqt the same way as A002. 760 measurements, 3 references (VannmiljøCopper2010-2025, 2003CoteurFieldContaminationOf, 2005GillanStructureOfSediment).",59.77,60.37,6.25,6.85,1900,2100,

```{r}
#| label: fig-soerfjorden-density
#| fig-cap: Sørfjorden bounding box (59.77-60.37N, 6.25-6.85E) with sampling density (sum of MEASURED_N per site), hexbinned. Adapted from scripts/build_study_area_figure.R panel a).
#| fig-width: 5
#| fig-height: 9
library(sf)

sf::sf_use_s2(FALSE)

soerfjorden_bbox <- st_bbox(
  c(xmin = 6.25, xmax = 6.85, ymin = 59.77, ymax = 60.37),
  crs = st_crs(4326)
)

wgs84_geo <- tar_read(wgs84_geography)
literature_merged_data <- tar_read(load_literature_pqt)

# Fjord-scale graticule: the project default (30deg lon / 10deg lat, built for
# the full Arctic study area in build_study_area_figure.R) draws zero lines
# inside a 0.6deg bbox like this one.
# wgs84_geo$graticule <- create_graticule(
#   lon_seq = seq(0, 20, 0.2),
#   lat_seq = seq(55, 65, 0.2)
# )

site_points <- literature_merged_data |>
  filter(
    !is.na(LATITUDE),
    !is.na(LONGITUDE),
    !is.na(MEASURED_N),
    LATITUDE >= soerfjorden_bbox["ymin"],
    LATITUDE <= soerfjorden_bbox["ymax"],
    LONGITUDE >= soerfjorden_bbox["xmin"],
    LONGITUDE <= soerfjorden_bbox["xmax"]
  ) |>
  summarise(
    total_n = sum(MEASURED_N, na.rm = TRUE),
    .by = c(SITE_CODE, LATITUDE, LONGITUDE)
  ) |>
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326, remove = FALSE)

panel_density <- ggplot() +
  # geom_sf(
  #   data = wgs84_geo$marine_polys,
  #   aes(fill = I(ocean_color)),
  #   color = "white",
  #   linewidth = 0.15
  # ) +
  geom_sf(
    data = wgs84_geo$countries,
    # aes(
    #   fill = I(ifelse(
    #     highlight_name,
    #     country_colours["highlight"],
    #     country_colours["default"]
    #   ))
    # ),
    color = "white",
    linewidth = 0.25
  ) +
  # geom_sf(
  #   data = wgs84_geo$graticule,
  #   linewidth = 0.15,
  #   color = "white",
  #   alpha = 0.45
  # ) +
  ggnewscale::new_scale_fill() +
  stat_bin_hex(
    data = site_points,
    aes(
      x = LONGITUDE,
      y = LATITUDE,
      weight = total_n,
      fill = after_stat(count)
    ),
    bins = 30,
    linewidth = 0,
    alpha = 0.75,
    inherit.aes = FALSE
  ) +
  scale_fill_viridis_c(
    name = "Samples (sum n)",
    trans = "log10",
    labels = scales::comma
  ) +
  # Equirectangular, true-scale at 60N: same trick as the hexbin cells in
  # build_study_area_figure.R panel a), keeps them regular hexagons rather
  # than stretched ones at this latitude (cos(60) = 0.5).
  coord_sf(
    crs = st_crs("+proj=eqc +lat_ts=60"),
    xlim = c(soerfjorden_bbox["xmin"] - 1, soerfjorden_bbox["xmax"] + 1),
    ylim = c(soerfjorden_bbox["ymin"] - 1, soerfjorden_bbox["ymax"] + 1),
    default_crs = st_crs(4326),
    expand = TRUE
  ) +
  theme_void() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(2, 2, 2, 2),
    legend.position = "bottom",
    legend.key.height = unit(0.3, "cm"),
    legend.key.width = unit(0.8, "cm"),
    plot.title = element_text(
      face = "bold",
      size = 11,
      hjust = 0.01,
      margin = margin(b = 2)
    )
  ) +
  labs(title = "Sørfjorden: study area & sampling density") |>
    suppressWarnings()

# Written to figures/ (not .images-raw/ like the parent script) since this
# notebook's output is a real, reviewable artefact rather than a one-off
# manuscript-figure draft. See CLAUDE.md 2.3.1: Sam can't see this inline, so
# the path is the thing that matters.
ggsave(
  "figures/soerfjorden-density.png",
  panel_density,
  width = 5,
  height = 9,
  dpi = 300,
  bg = "white"
) |>
  suppressWarnings()

panel_density
```

```{r}
# Table of groups filtered to bounding box
```

<!-- Embed AEP here -->