---
title: "Notebook XX - Aquaculture Copper Emissions (Hammerfest)"
subtitle: "Matching the Miljodirektoratet aquaculture Cu-to-water export to licensed farm locations"
date: last-modified
format: html
execute:
  freeze: true   # parked / temporary, not in the _quarto.yml render list
  echo: true     # working notebook: the name-matching logic is the point
  warning: false
  message: false
---

## What this is

`data/raw/Copper-Aquaculture-Emissions.xlsx` turned up as a per-facility,
per-year record of **copper discharged to water** by Norwegian aquaculture sites
(Miljodirektoratet / Norske utslipp style export, sheet `Utslipp`). It is
national, 2002-2023, one row per facility x operator x year, every value in kg
and every value carrying basis code **`B` (*Beregnet*, i.e. modelled)** rather
than measured. Copper from a marine farm is overwhelmingly net-pen antifouling
coating, so a modelled figure here is expected.

The file has **no location number** (`loknr`), only a free-text `Anleggsnavn` of
the form `"<site>, matfisk (<operator>)"`. To place the emissions we match that
name against the Fiskeridirektoratet aquaculture register
(`inst/extdata/shapefiles/Aquaculture_Locations/`), which does carry `loknr`,
coordinates, species and licence status.

**Scope: Hammerfest kommune only** (post-2020 merged kommune, so the old
Kvalsund sites are included; the file has already recoded them to `Hammerfest`).

```{r}
#| label: setup
library(readxl)
library(sf)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(flextable)
library(here)

here::i_am("README.md")
pkgload::load_all(here::here(), quiet = TRUE)

ft <- function(d, cap = NULL) {
  flextable(d) |>
    theme_vanilla() |>
    bold(part = "header") |>
    fontsize(size = 9, part = "all") |>
    padding(padding = 2, part = "all") |>
    set_caption(cap) |>
    set_table_properties(layout = "autofit", width = 1)
}

# Site-name key: lower-case, drop the ", matfisk (...)" / " akva (...)" tail,
# squish whitespace. Diacritics are kept -- both sources are Norwegian and
# agree on them.
norm_site <- function(x) {
  x |>
    str_to_lower() |>
    str_replace_all("\\s*[,(].*$", "") |>
    str_replace_all("\\s+akva$", "") |>
    str_replace_all("[[:space:]]+", " ") |>
    str_trim()
}
```

## Load and key both sources

```{r}
#| label: load-emissions
emissions <- read_excel(
  here::here("data/raw/Copper-Aquaculture-Emissions.xlsx"),
  sheet = "Utslipp",
  skip = 1,
  .name_repair = "unique_quiet"
) |>
  rename(
    anlegg = 1,
    fylke = 2,
    kommune = 3,
    year = 4,
    cu_kg = 5,
    basis = 6,
    unit = 7
  ) |>
  mutate(
    year = as.integer(year),
    cu_kg = as.numeric(cu_kg),
    site_key = norm_site(anlegg),
    operator = str_match(anlegg, "\\(([^)]*)\\)\\s*$")[, 2]
  )

em_ha <- emissions |>
  filter(str_to_upper(kommune) == "HAMMERFEST") |>
  group_by(site_key) |>
  summarise(
    operators = paste(sort(unique(operator)), collapse = " | "),
    year_from = min(year),
    year_to = max(year),
    n_year_rows = n(),
    cu_total_kg = round(sum(cu_kg, na.rm = TRUE), 1),
    cu_per_year = round(mean(cu_kg, na.rm = TRUE), 2),
    .groups = "drop"
  )
```

```{r}
#| label: load-register
register <- sf::read_sf(here::here(
  "inst/extdata/shapefiles/Aquaculture_Locations/Fiskeridir_Akvakulturregistere.shp"
)) |>
  sf::st_drop_geometry() |>
  as_tibble() |>
  filter(str_to_upper(kommune) == "HAMMERFEST") |>
  mutate(site_key = norm_site(navn)) |>
  group_by(site_key) |>
  summarise(
    loknr = paste(sort(unique(loknr)), collapse = "/"),
    navn = navn[1],
    status = paste(sort(unique(status_lok)), collapse = "/"),
    lon = mean(lon),
    lat = mean(lat),
    species = paste(sort(unique(til_arter)), collapse = " | "),
    capacity = paste0(kapasitet_[1], " ", kapasitet1[1]),
    .groups = "drop"
  )

# Combined Hammerfest-area AEP bounding box, read live from the manifest.
aep_box <- readr::read_csv(
  here::here("data/clean/aep/aep_manifest.csv"),
  show_col_types = FALSE
) |>
  summarise(
    lon_min = min(lon_min),
    lon_max = max(lon_max),
    lat_min = min(lat_min),
    lat_max = max(lat_max)
  )
```

## Match

```{r}
#| label: match
matched <- em_ha |>
  left_join(register, by = "site_key") |>
  mutate(
    in_aep_box = !is.na(lat) &
      lon >= aep_box$lon_min &
      lon <= aep_box$lon_max &
      lat >= aep_box$lat_min &
      lat <= aep_box$lat_max
  )

n_sites <- nrow(matched)
n_matched <- sum(!is.na(matched$loknr))
n_in_box <- sum(matched$in_aep_box, na.rm = TRUE)
```

`r n_matched` of `r n_sites` Hammerfest-kommune emission sites match a
currently-licensed location by name (1:1 -- no two current Hammerfest licences
share a normalised name). `r n_in_box` of the matched sites fall inside the
combined AEP bounding box (`r sprintf("%.2f-%.2f E, %.2f-%.2f N", aep_box$lon_min, aep_box$lon_max, aep_box$lat_min, aep_box$lat_max)`);
the rest are elsewhere in the kommune (Sorroysund, Revsbotn, the Kvalsund side).

### Matched: emission site to register `loknr`

```{r}
#| label: tbl-matched
#| tbl-cap: "Hammerfest-kommune aquaculture sites reporting copper to water, matched to the Fiskeridirektoratet register. Cu is the sum / mean of annual modelled discharge over the reporting years shown. Coordinates, `loknr` and species come from the register."
matched |>
  filter(!is.na(loknr)) |>
  transmute(
    Site = str_to_title(site_key),
    loknr,
    Lon = round(lon, 4),
    Lat = round(lat, 4),
    `In AEP box` = if_else(in_aep_box, "yes", "no"),
    `Reg. status` = status,
    Years = paste0(year_from, "-", year_to),
    `n yr` = n_year_rows,
    `Cu total (kg)` = cu_total_kg,
    `Cu (kg/yr)` = cu_per_year,
    Species = species
  ) |>
  arrange(desc(`Cu total (kg)`)) |>
  ft()
```

### Emissions but no current licence

Historic sites -- decommissioned or licence surrendered before the register
snapshot, so no coordinates are available from it. Operator strings are the
pre-2015 holders (Polarlaks, Stolt Sea Farm, early Cermaq).

```{r}
#| label: tbl-historic
#| tbl-cap: "Hammerfest-kommune sites with reported copper discharge but no current aquaculture licence."
matched |>
  filter(is.na(loknr)) |>
  transmute(
    Site = str_to_title(site_key),
    Operators = operators,
    Years = paste0(year_from, "-", year_to),
    `n yr` = n_year_rows,
    `Cu total (kg)` = cu_total_kg
  ) |>
  arrange(desc(`Cu total (kg)`)) |>
  ft()
```

### Licensed sites in the AEP box with no emissions record

```{r}
#| label: tbl-nomatch-reverse
#| tbl-cap: "Currently-licensed aquaculture locations inside the AEP bounding box that do not appear in the copper emissions file."
register |>
  filter(
    !is.na(lon),
    lon >= aep_box$lon_min,
    lon <= aep_box$lon_max,
    lat >= aep_box$lat_min,
    lat <= aep_box$lat_max,
    !site_key %in% em_ha$site_key
  ) |>
  transmute(
    loknr,
    Site = navn,
    Lon = round(lon, 4),
    Lat = round(lat, 4),
    Species = species,
    Capacity = capacity
  ) |>
  ft()
```

## Totals

```{r}
#| label: totals
tibble::tribble(
  ~Scope                                  , ~`Cu to water 2002-2023 (kg)`                                           ,
  "All Hammerfest-kommune emission sites" , round(sum(matched$cu_total_kg, na.rm = TRUE), 0)                        ,
  "Matched to a current licence"          , round(sum(matched$cu_total_kg[!is.na(matched$loknr)], na.rm = TRUE), 0) ,
  "Inside the AEP bounding box"           , round(sum(matched$cu_total_kg[matched$in_aep_box], na.rm = TRUE), 0)    ,
  "Historic sites (no current licence)"   , round(sum(matched$cu_total_kg[is.na(matched$loknr)], na.rm = TRUE), 0)
) |>
  ft()
```

```{r}
#| label: fig-cu-by-site
#| fig-cap: "Total modelled copper to water 2002-2023 per Hammerfest-kommune aquaculture site, split by whether the site matches a current licence. Written to images/dev/ for review outside the rendered notebook."
#| fig-width: 8
#| fig-height: 6
plot_df <- matched |>
  mutate(
    site_lab = str_to_title(site_key),
    group = if_else(is.na(loknr), "historic (no licence)", "current licence")
  )

p <- ggplot(
  plot_df,
  aes(
    x = cu_total_kg,
    y = reorder(site_lab, cu_total_kg),
    fill = group
  )
) +
  geom_col() +
  scale_fill_manual(
    values = c(
      "current licence" = "#2B5F75",
      "historic (no licence)" = "#B5876B"
    )
  ) +
  labs(
    x = "Cu to water 2002-2023 (kg, modelled)",
    y = NULL,
    fill = NULL,
    title = "Aquaculture copper discharge, Hammerfest kommune"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom", panel.grid.major.y = element_blank())

dir.create(here::here("images/dev"), showWarnings = FALSE, recursive = TRUE)
ggsave(
  here::here("images/dev/aquaculture-cu-by-site.png"),
  p,
  width = 8,
  height = 6,
  dpi = 150,
  bg = "white"
)
p
```

## Caveats

- **Name match only.** The emissions export carries no `loknr`; the join is on
  a normalised site name within Hammerfest kommune. It is 1:1 here, but a future
  register snapshot with two same-named Hammerfest sites would break that.
- **Coordinates come entirely from the register side.** The seven historic
  sites have none; they could be geolocated from Norske utslipp facility pages
  or an older Fiskeridirektoratet extract.
- **All discharge values are modelled** (`B`), consistent with copper being
  back-calculated from antifouling use. Relevant to the Evidence score for any
  aquaculture KTR edge.
- **Operator strings change across years** with company renames (Polarlaks ->
  Cermaq -> "Cermaq Norway Salmon AS" etc.); sites are collapsed on name, not
  operator.
- Not wired into `_targets.R` or the `_quarto.yml` render list. Build by hand
  with `quarto render docs/NBXX-aquaculture.qmd`.
