---
title: "Distributions: Aquatic > Aquatic Freshwater"
format: html
---

::: callout-note
This notebook was scaffolded by `scripts/generate_distribution_notebooks.R`
from the outlier-analysis targets factory (`_targets.R`,
`R/fct_outlier_groups.R`, `R/fct_outlier_detection.R`,
`R/fct_outlier_report.R`). The plots/tables below are generated, but
analysis and interpretation are added by hand, section by section.
Re-running the generator only appends sections for groups not yet present
here -- it never rewrites or removes anything you have written.

See [NBXX-Outliers.qmd](NBXX-Outliers.qmd) for the fully hand-authored
*G. morhua* liver case study this workflow was built from, including
year/site/campaign breakdowns and the reasoning behind the outlier flags.
:::

```{r}
#| label: setup
library(tidyverse)
library(targets)
library(patchwork)
library(flextable)
pkgload::load_all(quiet = TRUE)

set_flextable_defaults(
  font.size = 10,
  font.family = "Arial",
  digits = 2,
  theme_fun = theme_vanilla,
  padding = 1
)

ggplot_base_size <- 20
theme_set(theme_minimal(base_size = ggplot_base_size, base_family = "Aptos"))
```

## Freshwater

Aquatic / Freshwater. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Aquatic_Freshwater
grp_Aquatic_Freshwater <- tar_read(outlier_compartment_Aquatic_Freshwater)
grp_Aquatic_Freshwater$data <- grp_Aquatic_Freshwater$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Aquatic_Freshwater_unit <- grp_Aquatic_Freshwater$summary$unit
```

```{r}
#| label: fig-outlier-year-Aquatic_Freshwater
#| fig-cap: "Aquatic / Freshwater: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Freshwater$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Aquatic_Freshwater$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Aquatic_Freshwater_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Aquatic_Freshwater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Freshwater$data,
  "SAMPLING_YEAR",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-site-Aquatic_Freshwater
#| fig-cap: "Aquatic / Freshwater: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Freshwater$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Aquatic_Freshwater$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Aquatic_Freshwater_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Aquatic_Freshwater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Freshwater$data,
  "SITE_CODE",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-campaign-Aquatic_Freshwater
#| fig-cap: "Aquatic / Freshwater: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Freshwater$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Aquatic_Freshwater$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Aquatic_Freshwater_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Aquatic_Freshwater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Freshwater$data,
  "CAMPAIGN_NAME_SHORT",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-treatment-Aquatic_Freshwater
#| fig-cap: "Aquatic / Freshwater: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Aquatic_Freshwater,
  x_label = paste0("Measured value (", grp_Aquatic_Freshwater_unit, ")")
)
```
