---
title: "Notebook XX - Repparfjorden Case Study"
author: "Sam Welch"
date: last-modified
lightbox: true
---

I've been asked to put together a quick overview of the Repparfjorden, its history and likely relevant sources. So here goes.

Repparfjorden is a fjord in Hammerfest, Finnmark, at the far north of Norway. It has historically been a Submarine Tailing Disposal site, and may be again despite popular resistance. Nussir ASA wants to start mining copper near Repparfjorden, and this would be the logical place to put it (https://en.wikipedia.org/wiki/Repparfjorden).

```{r}
#| label: setup
#| warning: false
#| output: false
library(dplyr)
library(leaflet)
library(sf)
library(bslib)
library(eDataDRF)
library(targets)
library(devtools)
library(here)
library(readxl)
library(htmltools)
library(readxl)

suppressMessages(
  load_all(quiet = TRUE)
) # TODO: Reduce render time by sourcing only what we specifically need?
```

According to Wikipedia the coordinates of the fjord are (70.5, 24.17) (https://geohack.toolforge.org/geohack.php?pagename=Repparfjorden&params=70.5_N_24.17_E_type:waterbody_region:NO_dim:50000). We'll use this as a starting point for our map. It's apparently 14km long, so to get an appropriate amount of its surroundings we'll project it out by 30 km. (actually we won't yet).

# Define Study Area ----

```{r}
#| label: define-study-area

rf_coords <- st_sfc(st_point(c(70.50, 24.17)), crs = 4326)

rf_bbox <- st_bbox(
  c(xmin = 23.87, xmax = 24.47, ymin = 70.3, ymax = 70.8),
  crs = st_crs(4326)
)
```

# Literature Data ----

```{r}
#| label: load-literature-data

literature_merged_data <- tar_read(load_literature_pqt)
tar_load(copper_toxicity_thresholds)

literature_sf <- literature_merged_data |>
  filter(
    !is.na(LATITUDE),
    !is.na(LONGITUDE),
    !is.na(MEASURED_OR_IMPUTED_VALUE_STANDARD)
  ) |>
  st_as_sf(
    coords = c("LONGITUDE", "LATITUDE"),
    crs = st_crs("WGS84"),
    remove = FALSE
  ) |>
  st_crop(rf_bbox)
```

# Mining Sites ----

```{r}
#| label: load-mining-sites

mining_sites_sf <- mine_tailings_coords() |>
  st_as_sf(
    coords = c("LONGITUDE", "LATITUDE"),
    crs = st_crs("WGS84"),
    remove = FALSE
  )
```

# Aquaculture Sites ----

```{r}
#| label: load-aquaculture

aquaculture_sf <- sf::read_sf(here(
  "inst/extdata/shapefiles/Aquaculture_Locations/Fiskeridir_Akvakulturregistere.shx"
)) |>
  filter(fylke == "FINNMARK")
```

# Industrial Emissions ----

```{r}
#| label: load-industrial-emissions

copper_industrial_emissions <- read_excel(
  here::here(
    "data",
    "raw",
    "norske_utslipp",
    "norske_utslipp_copper_land_industries_individ.xlsx"
  ),
  skip = 2
) |>
  rename(
    facility_name = Anleggsnavn,
    county = Fylke,
    municipality = Kommune,
    nace_code = `NACE Kode`,
    year = År,
    annual_emission_air = `Årlig utslipp til luft`,
    basis_air = `Grunnlagsverdi for luft`,
    annual_emission_water = `Årlig utslipp til vann`,
    basis_water = `Grunnlagsverdi for vann`,
    unit = Enhet,
    facility_id = AnleggId,
    org_number = `Org.nr.`
  ) |>
  mutate(
    missing_pattern = case_when(
      is.na(annual_emission_air) &
        is.na(annual_emission_water) ~ "Both missing",
      is.na(annual_emission_air) ~ "Air missing",
      is.na(annual_emission_water) ~ "Water missing",
      TRUE ~ "Complete data"
    )
  )
```

# Interactive Map ----

```{r}
#| label: repparfjord-map

# Build labels for literature data ----
# NOTE: Adjust column names below to match your actual data structure
literature_labels <- lapply(seq_len(nrow(literature_sf)), function(i) {
  row <- literature_sf[i, ]
  htmltools::HTML(sprintf(
    "<strong>Literature Data</strong><br/>
    Value: %s<br/>
    Lat: %.4f, Lon: %.4f",
    row$MEASURED_OR_IMPUTED_VALUE_STANDARD,
    row$LATITUDE,
    row$LONGITUDE
  ))
})

# Build labels for mining sites ----
# NOTE: Adjust column names to match mine_tailings_coords() output
mining_labels <- lapply(seq_len(nrow(mining_sites_sf)), function(i) {
  row <- mining_sites_sf[i, ]
  htmltools::HTML(sprintf(
    "<strong>Mining/Tailings Site</strong><br/>
    Name: %s<br/>
    (%.4f, %.4f)<br/>
    Ore: %s",
    row$SITE_NAME,
    row$LATITUDE,
    row$LONGITUDE,
    row$ORE_TYPE
  ))
})

# Build labels for aquaculture sites ----
# NOTE: Adjust column names to match your shapefile attributes
aquaculture_labels <- lapply(seq_len(nrow(aquaculture_sf)), function(i) {
  row <- aquaculture_sf[i, ]
  htmltools::HTML(sprintf(
    "<strong>Aquaculture Site</strong><br/>
    Name: %s<br/>
    Kommune: %s<br/>
    Capacity: %s %s",
    if ("navn" %in% names(row)) row$navn else "Unknown",
    if ("kommune" %in% names(row)) row$kommune else "Unknown",
    row$kapasitet_,
    row$kapasitet1
  ))
})

# Render map ----
leaflet() |>
  addProviderTiles(providers$CartoDB.Positron) |>
  # Literature data points

  addCircleMarkers(
    data = literature_sf,
    radius = 6,
    color = "orange",
    fillOpacity = 0.6,
    stroke = TRUE,
    label = literature_labels,
    group = "Literature Data"
  ) |>
  # Mining/tailings sites

  addCircleMarkers(
    data = mining_sites_sf,
    radius = 20,
    color = "grey",
    fillOpacity = 0.6,
    stroke = TRUE,
    label = mining_labels,
    group = "Mining Sites"
  ) |>
  # Aquaculture sites
  addCircleMarkers(
    data = aquaculture_sf,
    radius = 10,
    color = "skyblue",
    fillOpacity = 0.6,
    stroke = TRUE,
    label = aquaculture_labels,
    group = "Aquaculture"
  ) |>
  # Layer control
  addLayersControl(
    overlayGroups = c("Literature Data", "Mining Sites", "Aquaculture"),
    options = layersControlOptions(collapsed = FALSE)
  )
```


```{r}
#| label: literature_data

literature_sf |>
  ggplot2::ggplot(aes(
    x = ENVIRON_COMPARTMENT_SUB,
    y = REFERENCE_ID,
    fill = MEASURED_VALUE_STANDARD
  )) +
  geom_tile()

# Hmmm.

# Not an incredible amount of data. Where's the Repparfjorden studies?
literature_sf |>
  group_by(ENVIRON_COMPARTMENT_SUB, SAMPLE_SPECIES, REFERENCE_ID) |>
  reframe(n = n(), mean = mean(MEASURED_VALUE_STANDARD))
```