---
title: "Notebook XX - Sørfjorden Case Study"
date: last-modified
lightbox: true
format: html
execute:
  freeze: false   # parked 2026-07-29, see _quarto.yml render list
---

```{r}
#| label: setup
#| include: false
#| warning: false

library(targets)
library(here)
library(tidyverse)
library(flextable)
pkgload::load_all(quiet = TRUE) # sample_groups_flextable() etc.

here::i_am("index.qmd")
```

Sørfjorden ("The South Fjord") is the southerly arm of Hardangerfjorden on Norway's South-West coast [@thorsnaesSorfjordenHardanger2024]. In the early 1900s a number of heavy industrial plants began operation on the fjord's shores, and it became unusually contaminated with heavy metals [@ruusMonitoringEnvironmentalQuality2010]. Copper is known to have been emitted to Sørfjord as a byproduct of zinc production in the area, and unusually high copper concentrations have been detected in the tissues of fish caught in the area [@ruusMonitoringEnvironmentalQuality2010]

```{r}
#| label: fig-soerfjorden-overview
#| fig-cap: map goes here

# map goes here

```

```{r}
# Table of groups filtered to bounding box
```

<!-- Embed AEP here -->