---
title: "Copper REACH Data for Norway"
execute:
  freeze: true   # parked 2026-07-29, see _quarto.yml render list
  echo: false
---

```{r}
suppressMessages(
  {
    library(tibble)
    library(dplyr)
    library(ggplot2)
    library(forcats)
    library(devtools)
    library(sf)
    library(summarytools)
    library(DT)
    library(readxl)
    library(visNetwork)

    load_all(quiet = TRUE)
  }
)

library(here)
i_am("README.md")

options(
  DT.options = list(
    pageLength = 25,
    language = list(search = "Filter:"),
    dom = "ftp" # filter, table, pagination — drops the length menu
  )
)

# Helper: render a table appropriately for the current output format ----
# Formats numeric columns to `digits` decimal places before display.
# Uses DT::datatable() for HTML, knitr::kable() for everything else (docx, pdf, etc.)
render_table <- function(df, caption = NULL, digits = 2, ...) {
  # Format numeric columns to specified DPs ----
  df_formatted <- df |>
    dplyr::mutate(
      dplyr::across(where(is.numeric), \(x) round(x, digits))
    )

  if (knitr::is_html_output()) {
    DT::datatable(df_formatted, caption = caption, ...)
  } else {
    knitr::kable(df_formatted, caption = caption)
  }
}
```

```{r}
# Note: covers declared chemicals only (deklarasjonspliktige kjemikalier).
#       "Netto mengde" = imported + produced - exported per declaration.

# ==============================================================================
# 2. Read data sheets ----
# ==============================================================================

path_reach <- here("inst/extdata/emissions/REACH_copper_prtd.xlsx")

# -- By sector (hovednæring) --------------------------------------------------
# Read via read_reach_sector_years() (R/fct_reach.R), not inline: this
# notebook and _targets.R's node_cards target both need the same
# sector-name translation and "Other" -> Unclassified handling (that logic's
# own TODO, "a mess - blank columns and negative numbers", is documented at
# read_reach_sector_years()), and two copies of it is how they drift.
# tar_read(), not a bare function call: this chunk's tar_read() calls ARE its
# dependency declaration (CLAUDE.md 4.4.2), so the notebook rebuilds when the
# REACH source file does.
reach_by_sector_en <- tar_read(reach_sector_years) |>
  mutate(
    NACE_supercategory = recode_values(
      isic_nace_section,
      c("A", "B") ~ "Primary",
      c("C", "F", "L") ~ "Secondary",
      c("E", "H", "L", "S") ~ "Tertiary",
      default = "Unclassified"
    )
  )

# -- By product type ----------------------------------------------------------
reach_by_product <- read_excel(
  path_reach,
  sheet = "kobber sum på produkttype"
) |>
  rename(
    year = AmountYear,
    netto_tonn = `Netto mengde i tonn`,
    product_type = Tekst
  ) |>
  mutate(
    year = as.integer(year),
    netto_tonn = as.numeric(netto_tonn),
    product_type = trimws(product_type)
  )

# The Norwegian-to-English NACE sector translation formerly lived here as a
# tribble; it's now reach_nace_sectors() in R/fct_reach.R (same source,
# https://ec.europa.eu/eurostat/documents/3859598/5902521/KS-RA-07-015-EN.PDF.pdf
# p143), applied inside read_reach_sector_years() before it ever reaches this
# notebook.

reach_by_sector_en_summarised <- reach_by_sector_en |>
  # mutate(z_score = scale(net_kg)[, 1], .by = sector_en) |>
  # ungroup() |>
  # filter(abs(z_score) < 2) |> # use z = 2 as a cutoff
  reframe(
    mean_net_kg = mean(net_kg),
    sd_net_kg = sd(net_kg),
    n_years_reported = n(),
    nace_section = range(isic_nace_section),
    .by = "sector_en"
  ) |>
  distinct() |>
  arrange(desc(mean_net_kg))

# TODO: Product category translations via LLM. Could find a better source
product_category_translations <- tibble::tribble(
  ~category_no                                                          , ~category_en                                       , ~confidence ,
  "BRENSELTILSETNINGER"                                                 , "Fuel additives"                                   , "high"      ,
  "Absorpsjons/ og adsorpsjonsmaterialer"                               , "Absorption and adsorption materials"              , "high"      ,
  "ANDRE SMØREMIDLER"                                                   , "Other lubricants"                                 , "high"      ,
  "ARMERINGSMIDLER"                                                     , "Reinforcing agents"                               , "medium"    , # could be "armoring agents" in some contexts
  "Bindemidler"                                                         , "Binders"                                          , "high"      ,
  "Biocider"                                                            , "Biocides"                                         , "high"      ,
  "Borekjemikalier inkl råolje/gass"                                    , "Drilling chemicals incl. crude oil/gas"           , "high"      ,
  "BOREOLJER"                                                           , "Drilling oils"                                    , "high"      ,
  "BUNNFELLINGSHINDRENDE MIDLER, GENERELT"                              , "Anti-precipitation agents, general"               , "medium"    , # could be "anti-settling" or "anti-sedimentation"
  "Fyllingsmidler"                                                      , "Fillers"                                          , "high"      ,
  "GJØDNING, GENERELT"                                                  , "Fertilisers, general"                             , "high"      ,
  "Impregnering"                                                        , "Impregnation agents"                              , "high"      ,
  "INSEKTSMIDDEL, INSEKTMIDLER OG ANDRE MIDLER MOT SKADEDYR PÅ PLANTER" , "Insecticides and other plant pest control agents" , "high"      ,
  "Konstruksjonsmaterialer"                                             , "Construction materials"                           , "high"      ,
  "Lim"                                                                 , "Adhesives"                                        , "high"      ,
  "Maling"                                                              , "Paint"                                            , "high"      ,
  "PH-REGULERENDE MIDLER, GENERELT"                                     , "pH-regulating agents, general"                    , "high"      ,
  "Prosessregulerendemidler"                                            , "Process control agents"                           , "medium"    , # or "process regulating agents"
  "Rengjøring"                                                          , "Cleaning agents"                                  , "high"      ,
  "Rustbeskyttelse"                                                     , "Corrosion protection"                             , "high"      ,
  "SALT TIL GALVANISKE BAD"                                             , "Salts for electroplating baths"                   , "high"      ,
  "Sprengstoff"                                                         , "Explosives"                                       , "high"      ,
  "STØPEMASSER, GENERELT"                                               , "Casting compounds, general"                       , "medium"    , # could be "moulding compounds"
  "SYNTESERÅVARER OG MELLOMPRODUKTER"                                   , "Synthesis raw materials and intermediates"        , "high"      ,
  "Trykkfarger"                                                         , "Printing inks"                                    , "high"      ,
  "BLEKERE TIL FOTOGRAFISK FILM"                                        , "Bleaching agents for photographic film"           , "high"      ,
  "BRUNERINGSSALTER"                                                    , "Browning salts"                                   , "medium"    , # technical term for metal darkening/bluing salts
  "Glasur, emalje"                                                      , "Glaze, enamel"                                    , "high"      ,
  "Herdere"                                                             , "Hardeners/curing agents"                          , "high"      ,
  "Loddemidler"                                                         , "Soldering agents/fluxes"                          , "medium"    , # loddemidler is broad; could be just "solders"
  "BILPLEIEMIDLER, GENERELT"                                            , "Car care products, general"                       , "high"      ,
  "FLUSSMIDLER (SVEISING)"                                              , "Fluxes (welding)"                                 , "high"      ,
  "PIGMENT TIL GLASURER, EMALJER OG GLASS"                              , "Pigments for glazes, enamels and glass"           , "high"      ,
  "Metalloverflatebehandlingsmidler"                                    , "Metal surface treatment agents"                   , "high"      ,
  "ANTIOKSIDANTER (ANTIOZONANTER)"                                      , "Antioxidants (antiozonants)"                      , "high"      ,
  "Poler og pleieMIDLER"                                                , "Polishes and care products"                       , "medium" # odd capitalisation in source; "poler" = polishes
)

reach_by_product_translated <- reach_by_product |>
  left_join(
    product_category_translations,
    by = join_by(product_type == category_no)
  ) |>
  mutate(net_kg = netto_tonn * 1000)

reach_by_product_summarised <- reach_by_product_translated |>
  reframe(
    mean_net_kg = mean(net_kg),
    sd_net_kg = sd(net_kg),
    n_years_reported = n(),
    .by = "category_en"
  ) |>
  distinct() |>
  arrange(desc(mean_net_kg))

reach_by_product_summarised_lumped <- reach_by_product_translated |>
  # need to remove negative rows for fct_limp
  filter(netto_tonn > 0) |>
  mutate(
    category_en = fct_lump_n(
      category_en,
      n = 7,
      w = netto_tonn,
      other_level = "Other"
    )
  ) |>
  reframe(
    mean_net_kg = mean(net_kg),
    sd_net_kg = sd(net_kg),
    n_years_reported = n(),
    .by = "category_en"
  ) |>
  distinct() |>
  arrange(desc(mean_net_kg))
```

REACH PR data for copper was obtained by contacting Miljødirektoratet (personal communications, 2nd May 2025).

> Hei\
> Din henvendelse REQ0014380 - Copper Pollution/Use Statistics er oppdatert med følgende: Beklager sent svar.\
> \
> Legger ved tall for utvalgte kobberstoffer i produktregisteret. Stoffene er fra liste som vi har brukt i forbindelse med statistikken for kobber i notimpregnering i akvakultur. Du kjenner kanskje tallene fra HI rapporter etc. Det er godt mulig det er flere kobberstoffer i registeret. Har du forslag til stoffer vi kunne sjekke, eller bare søke kjapt å se om det er noe? (altså andre enn de i listen) Det er også mulig å dele opp bransjer og typer mye mer. Du kan se mer om produkttypene våre og bransjekodene våre på hjemmesidene. Jeg brukte nå en hovedgruppering for at du kunne få en bilde av hva vi har av disse kobberstoffene. Mer spesifikt for typer og bransjer kan du spørre om. Det er da også mulig at det blir uoff tall særlig om du også spesifiserer spørsmål om stoffene. Lagde også en med mengder per stoff. Tallene per stoff kan være at noen er unntatt offentlighet. Vil du se fordelingen per stoff så kan jeg sende deg en kryptert fil. Da trenger jeg telefonnummeret så jeg kan sende deg passord.(så sender jeg en mail med en separat fil)
>
> Hvis du svarer på denne e-posten, blir svaret automatisk lagt til i saken. Hilsen Miljødirektoratet

Attached was a spreadsheet consisting of five sheets:

- Forklaring (Explanation)
- Figurer (Figures)
- Sum HovedgruppeAndvendelse (Sum Main Group... something)
- kobber sum på produkttype (Copper sum of product type)
- Ark4 (an empty sheet)

The most salient details:

- The search was run on copper and copper-containing substances (see table below)
- Net quantities are calculated from the weight % of chemicals
- The net quantity is the imported quantity, plus produced quantity, minus exported quantity for each declaration

Although REACH declarations are mandatory, it has to be assumed that the level of detail and quality is, overall, low. Nevertheless, this information can give us a rough idea of the relative importance of uses of copper in Norway.

We probably don't have permission to republish the raw data, but I believe analysis is fine. Will check beforehand.

```{r}
#| label: tbl-copper-cas
#| 
reach_cas <- tribble(
  ~cas_nr      , ~kjemisk_navn             , ~name_en                      ,
  "1111-67-7"  , "kobberthiocyanat"        , "Copper thiocyanate"          ,
  "1317-38-0"  , "Kobber(II)oksid"         , "Copper(II) oxide"            ,
  "1317-39-1"  , "Kobber(I)oksid"          , "Copper(I) oxide"             ,
  "1338-02-9"  , "Kobbernaftenat"          , "Copper naphthenate"          ,
  "7440-50-8"  , "kobber pulver"           , "Copper powder"               ,
  "7447-39-4"  , "Kobberklorid"            , "Copper chloride"             ,
  "7758-89-6"  , "Kobber(I)klorid"         , "Copper(I) chloride"          ,
  "7758-98-7"  , "Kobber(II)sulfat"        , "Copper(II) sulfate"          ,
  "12069-69-1" , "Kobberhydroksidkarbonat" , "Copper hydroxycarbonate"     ,
  "14915-37-8" , "kobberpyrition"          , "Copper pyrithione"           ,
  "54253-62-2" , "Kobber(II)metansulfonat" , "Copper(II) methanesulfonate"
)

reach_cas |>
  render_table()
```

# By Sector

Looking at sales by sector is an obvious place to start. Product details are more complicated (but potentially asl more useful).

Sectors are based on NACE codes. We translate names to English based on the NACE standard, and put categories into supercategories (primary, secondary, tertiary/service) for easier plotting.

```{r}
#| label: fig-reach-by-sector
#| fig-cap: "Net copper quantities reported to Produktregisteret by industry sector. Net = (imported + produced) − exported quantity. Facets ordered by NACE supercategory."
#| fig-width: 10
#| fig-height: 12
#| warning: false
reach_by_sector_en |>
  group_by(sector_en) |>
  mutate(mean_net = mean(net_kg)) |>
  ungroup() |>
  mutate(
    sector_en = factor(
      sector_en,
      levels = unique(sector_en[order(NACE_supercategory, -mean_net)])
    )
  ) |>
  ggplot(
    aes(
      x = year,
      y = net_kg,
      group = sector_en,
      fill = NACE_supercategory
    )
  ) +
  geom_col() +
  geom_hline(yintercept = 0) +
  scale_fill_brewer(palette = "Set2", name = "NACE supercategory") +
  scale_y_continuous(labels = scales::label_comma()) +
  facet_wrap(vars(sector_en), scales = "free_y", ncol = 2) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(
    title = "Net Copper Reported by Industry Sector",
    subtitle = "Net = (imported quantity + produced quantity) − exported quantity",
    x = "Year",
    y = "Net Copper (kg)"
  )
```

This data is not particularly trustworthy. Spikes in net quantity may represent improper data entry or simply normal year-on-year staggered purchasing behaviour. Lacking the ability to follow up on it to a satisfactory level of detail, it seems the safest option is to:

a)  take the mean value over the period
b)  assign it the lowest level of quality

This leaves us with the following dataset (@tbl-reach-by-sector):

```{r}
#| label: tbl-reach-by-sector
reach_by_sector_en_summarised |> render_table()
```

## Sector groupings used for the AEP source nodes

Eleven NACE sectors is more resolution than the underlying data can support at
`n_years_reported` as low as 3-6, so before these become AEP source nodes two
lumps are applied:

- **Construction and real estate**: Construction (NACE F) + Real estate
  activities (NACE L).
- **Other services and administration**: Other service activities (S),
  Unclassified, Arts/entertainment/recreation (R), Professional/scientific/
  technical (M), Electricity/gas/steam/air conditioning supply (D),
  Administrative and support service activities (N).

The other six sectors (Manufacturing, Agriculture/forestry/fishing, Mining and
quarrying, Wholesale and retail trade, Transportation and storage, Water
supply/sewerage/waste) stand alone. This gives eight groups, each a source node
in `data/clean/aep/aep_nodes.csv` (`N004`-`N011`), scored with the LoE
assessment below and using the mean net quantity per year as `external_value`.

The lumping itself (which raw sector rolls into which node) is
`reach_sector_lump_map()`/`reach_node_id_by_sector()` in `R/fct_reach.R`, not
computed here: `_targets.R`'s `node_cards` target needs the identical
lumping to draw each node's per-year series on its card, and two copies is how
they drift.

```{r}
#| label: tbl-reach-sector-lumping

reach_by_sector_lumped <- tar_read(reach_node_summary_data)

reach_by_sector_lumped |> render_table()
```

# Products

We might as well also look at products. The raw data is in Norwegian and not immediately very informative. First we inspect the data for any obvious outliers or suspect data.

These are less obviously problematic than those per industry sector, although some categories drop to 0 year-on-year with more alacrity than we might expect.

```{r}
#| label: fig-reach-by-product
#| fig-cap: "Net copper quantities reported to Produktregisteret by product type. Net = (imported + produced) − exported quantity."
#| fig-width: 10
#| fig-height: 20
#| warning: false
reach_by_product_translated |>
  group_by(category_en) |>
  ggplot(
    aes(
      x = year,
      y = net_kg,
      fill = category_en
    )
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::label_comma()) +
  facet_wrap(vars(category_en), scales = "free_y", ncol = 2) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(
    title = "Net Copper Reported by Product Category",
    subtitle = "Net = (imported quantity + produced quantity) − exported quantity",
    x = "Year",
    y = "Net Copper (kg)"
  )
```

We translate and average across the reported periods.

```{r}
#| label: tbl-reach-by-product
reach_by_product_summarised |> render_table()
```

This isn't massively informative either because we don't have any useful grouping categories, but it's a start. Basically synthesis raw materials, biocides, and insecticides are the most important categories by far. Which is good to know.

We can lump stuff together and then graph:

```{r}
reach_by_product_summarised_lumped |>
  mutate(category_en = fct_reorder(category_en, mean_net_kg)) |>
  ggplot(aes(x = category_en, y = mean_net_kg, fill = category_en)) +
  geom_col() +
  scale_fill_manual(
    values = c(
      colorRampPalette(RColorBrewer::brewer.pal(8, "Dark2"))(7),
      "grey70" # this sucks
    ),
    guide = "none"
  ) +
  scale_y_continuous(labels = scales::label_comma(), trans = "log10") +
  coord_flip() +
  theme_minimal(base_size = 12) +
  labs(
    x = NULL,
    y = "Mean net copper (kg)"
  )
```

We lack the ability to easily break down this information into compartment-specific emissions or directly map it to the emissions data provided by [Norske Utslipp](NBXX-norske-utslipp.html). However, we can still include it in the AEP as a line of evidence.

# Line of Evidence Assessment

REACH self-reported net values are considered weak evidence of copper occurrence in our LoE framework. However, they provide an irreplaceable quantification of flows of copper early in the supply chain.

| Criterion | Score | Justification |
|---------------------------|------------------|---------------------------|
| Essentiality | 1 | This information does not provide any direct evidence of emissions to the environment. |
| Theoretical Plausibility | 3 | The mechanistic underpinnings of industrial copper use are well-understood. |
| Empirical Support | 1 | Data are self-reported and impossible to verify. |
| Quantitative Understanding | 1 | Relevant models of copper use and manufacture may exist, but were outside of the scope of this study. |

: REACH WoE Assessment

# Resulting Source Nodes

The eight lumped sectors above are written by hand into
`data/clean/aep/aep_nodes.csv` as source nodes `N004`-`N011`
(`node_type = "external"`, `external_value` = mean net kg/year,
`external_sd`/`external_n` from the same table), carrying the LoE scores above.
They feed `N012` (coastal benthic sediment) and `N013` (coastal water column)
in `data/clean/aep/aep_edges.csv`, all `putative` pending a literature check
(PLAN.md P4.2), and are members of the national AEP (`A001`).

Their cards are drawn by the same `node_cards` target every other
node's card comes from (via `write_aep_node_cards()`'s `external_series`
argument, fed by `reach_external_series_data`), not written separately by this
notebook (Sam, 2026-08-11: "why create an entirely separate dir structure for
source cards?" -- there is no good reason to, once the pipeline itself knows
how to draw the bars). That also means these bars now appear on the real
`figures/aep_A001.png`, not just here.

```{r}
#| label: reach-node-cards
#| include: false

tr <- function(expr) tryCatch(expr, error = function(e) NULL)

reach_node_ids <- unname(reach_node_id_by_sector())

reach_card_paths <- tr(tar_read(node_cards))
if (!is.null(reach_card_paths)) {
  reach_card_paths <- reach_card_paths[
    basename(dirname(reach_card_paths)) == "A001" &
      tools::file_path_sans_ext(basename(reach_card_paths)) %in% reach_node_ids
  ]
  # Largest mean net quantity first, not file order.
  reach_card_paths <- reach_card_paths[
    match(reach_node_ids, tools::file_path_sans_ext(basename(reach_card_paths)))
  ]
}
```

From `images/node_cards/A001/`, largest mean net quantity first:

```{r}
#| echo: false
#| results: asis
if (is.null(reach_card_paths) || length(reach_card_paths) == 0) {
  cat(
    "`node_cards` is not in the target store yet. Run `tar_make()`.\n"
  )
} else {
  cat('::: {layout-ncol="4"}\n')
  for (p in reach_card_paths) {
    # node_cards's paths come from here_rel() in _targets.R, i.e.
    # already relative to the PROJECT ROOT, not absolute. here::here(p)
    # re-anchors that to an absolute path off the real project root
    # (found via the .here/README.md anchor) regardless of the R working
    # directory this chunk runs in, which Quarto sets to this .qmd's own
    # directory (docs/) -- fs::path_rel() on the bare relative string
    # produced a link with no "../" at all, silently 404ing at
    # docs/images/... instead of the real ../images/... (caught by
    # actually rendering this notebook and checking, not by inspection).
    rel <- fs::path_rel(here::here(p), here("docs"))
    cat(sprintf("![](%s)\n\n", rel))
  }
  cat(":::\n")
}
```