---
title: "Distributions: Biota Compartment"
format:
  html:
    fig-width: 16
    fig-height: 20
  docx:
    fig-width: 16
    fig-height: 20

execute:
  echo: false

knitr:
  opts_chunk:
    dpi: 300
    dev: "ragg_png"

filters:
  - flextable-qmd
  - at: post-render
    path: ../_extensions/flextable-qmd/unwrap-float.lua
---

::: callout-note
This notebook was scaffolded by `scripts/generate_distribution_notebooks.R`
from the outlier-analysis targets factory (`_targets.R`,
`R/fct_outlier_groups.R`, `R/fct_outlier_detection.R`,
`R/fct_outlier_report.R`). The plots/tables below are generated, but
analysis and interpretation are added by hand, section by section.
Re-running the generator only appends sections for groups not yet present
here -- it never rewrites or removes anything you have written.

See [NBXX-Outliers.qmd](NBXX-Outliers.qmd) for the fully hand-authored
*G. morhua* liver case study this workflow was built from, including
year/site/campaign breakdowns and the reasoning behind the outlier flags.
:::

```{r}
#| label: setup
library(tidyverse)
library(targets)
library(patchwork)
library(flextable)
pkgload::load_all(quiet = TRUE)

set_flextable_defaults(
  font.size = 10,
  font.family = "Arial",
  digits = 2,
  theme_fun = theme_vanilla,
  padding = 1
)

ggplot_base_size <- 20
theme_set(theme_minimal(base_size = ggplot_base_size, base_family = "Aptos"))
```

## Biota, Aquatic

Biota / Biota, Aquatic. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Biota_Biota_Aquatic
grp_Biota_Biota_Aquatic <- tar_read(outlier_compartment_Biota_Biota_Aquatic)
grp_Biota_Biota_Aquatic$data <- grp_Biota_Biota_Aquatic$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Biota_Biota_Aquatic_unit <- grp_Biota_Biota_Aquatic$summary$unit
```

```{r}
#| label: fig-outlier-year-Biota_Biota_Aquatic
#| fig-cap: "Biota / Biota, Aquatic: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Biota_Biota_Aquatic$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Biota_Biota_Aquatic$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Biota_Biota_Aquatic_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Biota_Biota_Aquatic
outlier_flag_table(flag_by_facet(grp_Biota_Biota_Aquatic$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Biota_Biota_Aquatic
#| fig-cap: "Biota / Biota, Aquatic: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Biota_Biota_Aquatic$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Biota_Biota_Aquatic$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Biota_Biota_Aquatic_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Biota_Biota_Aquatic
outlier_flag_table(flag_by_facet(grp_Biota_Biota_Aquatic$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Biota_Biota_Aquatic
#| fig-cap: "Biota / Biota, Aquatic: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Biota_Biota_Aquatic$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Biota_Biota_Aquatic$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Biota_Biota_Aquatic_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Biota_Biota_Aquatic
outlier_flag_table(flag_by_facet(grp_Biota_Biota_Aquatic$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Biota_Biota_Aquatic
#| fig-cap: "Biota / Biota, Aquatic: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Biota_Biota_Aquatic,
  x_label = paste0("Measured value (", grp_Biota_Biota_Aquatic_unit, ")")
)
```

## Biota, Terrestrial

Biota / Biota, Terrestrial. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Biota_Biota_Terrestrial
grp_Biota_Biota_Terrestrial <- tar_read(outlier_compartment_Biota_Biota_Terrestrial)
grp_Biota_Biota_Terrestrial$data <- grp_Biota_Biota_Terrestrial$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Biota_Biota_Terrestrial_unit <- grp_Biota_Biota_Terrestrial$summary$unit
```

```{r}
#| label: fig-outlier-year-Biota_Biota_Terrestrial
#| fig-cap: "Biota / Biota, Terrestrial: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Biota_Biota_Terrestrial$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Biota_Biota_Terrestrial$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Biota_Biota_Terrestrial_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Biota_Biota_Terrestrial
outlier_flag_table(flag_by_facet(grp_Biota_Biota_Terrestrial$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Biota_Biota_Terrestrial
#| fig-cap: "Biota / Biota, Terrestrial: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Biota_Biota_Terrestrial$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Biota_Biota_Terrestrial$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Biota_Biota_Terrestrial_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Biota_Biota_Terrestrial
outlier_flag_table(flag_by_facet(grp_Biota_Biota_Terrestrial$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Biota_Biota_Terrestrial
#| fig-cap: "Biota / Biota, Terrestrial: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Biota_Biota_Terrestrial$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Biota_Biota_Terrestrial$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Biota_Biota_Terrestrial_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Biota_Biota_Terrestrial
outlier_flag_table(flag_by_facet(grp_Biota_Biota_Terrestrial$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Biota_Biota_Terrestrial
#| fig-cap: "Biota / Biota, Terrestrial: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Biota_Biota_Terrestrial,
  x_label = paste0("Measured value (", grp_Biota_Biota_Terrestrial_unit, ")")
)
```
