---
title: "Distributions: Aquatic Compartment"
format:
  html:
    fig-width: 16
    fig-height: 20
  docx:
    fig-width: 16
    fig-height: 20

execute:
  echo: false

knitr:
  opts_chunk:
    dpi: 300
    dev: "ragg_png"

filters:
  - flextable-qmd
  - at: post-render
    path: ../_extensions/flextable-qmd/unwrap-float.lua
---

::: callout-note
This notebook was scaffolded by `scripts/generate_distribution_notebooks.R`
from the outlier-analysis targets factory (`_targets.R`,
`R/fct_outlier_groups.R`, `R/fct_outlier_detection.R`,
`R/fct_outlier_report.R`). The plots/tables below are generated, but
analysis and interpretation are added by hand, section by section.
Re-running the generator only appends sections for groups not yet present
here -- it never rewrites or removes anything you have written.

See [NBXX-Outliers.qmd](NBXX-Outliers.qmd) for the fully hand-authored
*G. morhua* liver case study this workflow was built from, including
year/site/campaign breakdowns and the reasoning behind the outlier flags.
:::

```{r}
#| label: setup
library(tidyverse)
library(targets)
library(patchwork)
library(flextable)
pkgload::load_all(quiet = TRUE)

set_flextable_defaults(
  font.size = 10,
  font.family = "Arial",
  digits = 2,
  theme_fun = theme_vanilla,
  padding = 1
)

ggplot_base_size <- 20
theme_set(theme_minimal(base_size = ggplot_base_size, base_family = "Aptos"))
```

## Aquatic Sediment

Aquatic / Aquatic Sediment. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Aquatic_Aquatic_Sediment
grp_Aquatic_Aquatic_Sediment <- tar_read(
  outlier_compartment_Aquatic_Aquatic_Sediment
)
grp_Aquatic_Aquatic_Sediment$data <- grp_Aquatic_Aquatic_Sediment$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Aquatic_Aquatic_Sediment_unit <- grp_Aquatic_Aquatic_Sediment$summary$unit
```

```{r}
#| label: fig-outlier-year-Aquatic_Aquatic_Sediment
#| fig-cap: "Aquatic / Aquatic Sediment: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Aquatic_Sediment$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Aquatic_Aquatic_Sediment$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Aquatic_Aquatic_Sediment_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Aquatic_Aquatic_Sediment
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Aquatic_Sediment$data,
  "SAMPLING_YEAR",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-site-Aquatic_Aquatic_Sediment
#| fig-cap: "Aquatic / Aquatic Sediment: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Aquatic_Sediment$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Aquatic_Aquatic_Sediment$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Aquatic_Aquatic_Sediment_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Aquatic_Aquatic_Sediment
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Aquatic_Sediment$data,
  "SITE_CODE",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-campaign-Aquatic_Aquatic_Sediment
#| fig-cap: "Aquatic / Aquatic Sediment: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Aquatic_Sediment$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Aquatic_Aquatic_Sediment$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Aquatic_Aquatic_Sediment_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Aquatic_Aquatic_Sediment
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Aquatic_Sediment$data,
  "CAMPAIGN_NAME_SHORT",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-treatment-Aquatic_Aquatic_Sediment
#| fig-cap: "Aquatic / Aquatic Sediment: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Aquatic_Aquatic_Sediment,
  x_label = paste0("Measured value (", grp_Aquatic_Aquatic_Sediment_unit, ")")
)
```

## Brackish/Transitional Water

Aquatic / Brackish/Transitional Water. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Aquatic_Brackish_Transitional_Water
grp_Aquatic_Brackish_Transitional_Water <- tar_read(
  outlier_compartment_Aquatic_Brackish_Transitional_Water
)
grp_Aquatic_Brackish_Transitional_Water$data <- grp_Aquatic_Brackish_Transitional_Water$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Aquatic_Brackish_Transitional_Water_unit <- grp_Aquatic_Brackish_Transitional_Water$summary$unit
```

```{r}
#| label: fig-outlier-year-Aquatic_Brackish_Transitional_Water
#| fig-cap: "Aquatic / Brackish/Transitional Water: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Brackish_Transitional_Water$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Aquatic_Brackish_Transitional_Water$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0(
    "Measured value (",
    grp_Aquatic_Brackish_Transitional_Water_unit,
    ")"
  ),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Aquatic_Brackish_Transitional_Water
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Brackish_Transitional_Water$data,
  "SAMPLING_YEAR",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-site-Aquatic_Brackish_Transitional_Water
#| fig-cap: "Aquatic / Brackish/Transitional Water: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Brackish_Transitional_Water$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Aquatic_Brackish_Transitional_Water$data,
  facet_col = "SITE_CODE",
  x_label = paste0(
    "Measured value (",
    grp_Aquatic_Brackish_Transitional_Water_unit,
    ")"
  ),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Aquatic_Brackish_Transitional_Water
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Brackish_Transitional_Water$data,
  "SITE_CODE",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-campaign-Aquatic_Brackish_Transitional_Water
#| fig-cap: "Aquatic / Brackish/Transitional Water: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Brackish_Transitional_Water$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Aquatic_Brackish_Transitional_Water$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0(
    "Measured value (",
    grp_Aquatic_Brackish_Transitional_Water_unit,
    ")"
  ),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Aquatic_Brackish_Transitional_Water
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Brackish_Transitional_Water$data,
  "CAMPAIGN_NAME_SHORT",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-treatment-Aquatic_Brackish_Transitional_Water
#| fig-cap: "Aquatic / Brackish/Transitional Water: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Aquatic_Brackish_Transitional_Water,
  x_label = paste0(
    "Measured value (",
    grp_Aquatic_Brackish_Transitional_Water_unit,
    ")"
  )
)
```

## Freshwater

Aquatic / Freshwater. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Aquatic_Freshwater
grp_Aquatic_Freshwater <- tar_read(outlier_compartment_Aquatic_Freshwater)
grp_Aquatic_Freshwater$data <- grp_Aquatic_Freshwater$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Aquatic_Freshwater_unit <- grp_Aquatic_Freshwater$summary$unit
```

```{r}
#| label: fig-outlier-year-Aquatic_Freshwater
#| fig-cap: "Aquatic / Freshwater: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Freshwater$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Aquatic_Freshwater$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Aquatic_Freshwater_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Aquatic_Freshwater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Freshwater$data,
  "SAMPLING_YEAR",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-site-Aquatic_Freshwater
#| fig-cap: "Aquatic / Freshwater: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Freshwater$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Aquatic_Freshwater$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Aquatic_Freshwater_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Aquatic_Freshwater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Freshwater$data,
  "SITE_CODE",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-campaign-Aquatic_Freshwater
#| fig-cap: "Aquatic / Freshwater: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Freshwater$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Aquatic_Freshwater$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Aquatic_Freshwater_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Aquatic_Freshwater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Freshwater$data,
  "CAMPAIGN_NAME_SHORT",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-treatment-Aquatic_Freshwater
#| fig-cap: "Aquatic / Freshwater: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Aquatic_Freshwater,
  x_label = paste0("Measured value (", grp_Aquatic_Freshwater_unit, ")")
)
```

## Groundwater

Aquatic / Groundwater. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Aquatic_Groundwater
grp_Aquatic_Groundwater <- tar_read(outlier_compartment_Aquatic_Groundwater)
grp_Aquatic_Groundwater$data <- grp_Aquatic_Groundwater$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Aquatic_Groundwater_unit <- grp_Aquatic_Groundwater$summary$unit
```

```{r}
#| label: fig-outlier-year-Aquatic_Groundwater
#| fig-cap: "Aquatic / Groundwater: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Groundwater$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Aquatic_Groundwater$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Aquatic_Groundwater_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Aquatic_Groundwater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Groundwater$data,
  "SAMPLING_YEAR",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-site-Aquatic_Groundwater
#| fig-cap: "Aquatic / Groundwater: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Groundwater$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Aquatic_Groundwater$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Aquatic_Groundwater_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Aquatic_Groundwater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Groundwater$data,
  "SITE_CODE",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-campaign-Aquatic_Groundwater
#| fig-cap: "Aquatic / Groundwater: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Groundwater$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Aquatic_Groundwater$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Aquatic_Groundwater_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Aquatic_Groundwater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Groundwater$data,
  "CAMPAIGN_NAME_SHORT",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-treatment-Aquatic_Groundwater
#| fig-cap: "Aquatic / Groundwater: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Aquatic_Groundwater,
  x_label = paste0("Measured value (", grp_Aquatic_Groundwater_unit, ")")
)
```

## Leachate

Aquatic / Leachate. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Aquatic_Leachate
grp_Aquatic_Leachate <- tar_read(outlier_compartment_Aquatic_Leachate)
grp_Aquatic_Leachate$data <- grp_Aquatic_Leachate$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Aquatic_Leachate_unit <- grp_Aquatic_Leachate$summary$unit
```

```{r}
#| label: fig-outlier-year-Aquatic_Leachate
#| fig-cap: "Aquatic / Leachate: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Leachate$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Aquatic_Leachate$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Aquatic_Leachate_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Aquatic_Leachate
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Leachate$data,
  "SAMPLING_YEAR",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-site-Aquatic_Leachate
#| fig-cap: "Aquatic / Leachate: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Leachate$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Aquatic_Leachate$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Aquatic_Leachate_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Aquatic_Leachate
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Leachate$data,
  "SITE_CODE",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-campaign-Aquatic_Leachate
#| fig-cap: "Aquatic / Leachate: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Leachate$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Aquatic_Leachate$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Aquatic_Leachate_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Aquatic_Leachate
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Leachate$data,
  "CAMPAIGN_NAME_SHORT",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-treatment-Aquatic_Leachate
#| fig-cap: "Aquatic / Leachate: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Aquatic_Leachate,
  x_label = paste0("Measured value (", grp_Aquatic_Leachate_unit, ")")
)
```

## Marine/Salt Water

Aquatic / Marine/Salt Water. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Aquatic_Marine_Salt_Water
grp_Aquatic_Marine_Salt_Water <- tar_read(
  outlier_compartment_Aquatic_Marine_Salt_Water
)
grp_Aquatic_Marine_Salt_Water$data <- grp_Aquatic_Marine_Salt_Water$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Aquatic_Marine_Salt_Water_unit <- grp_Aquatic_Marine_Salt_Water$summary$unit
```

```{r}
#| label: fig-outlier-year-Aquatic_Marine_Salt_Water
#| fig-cap: "Aquatic / Marine/Salt Water: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Marine_Salt_Water$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Aquatic_Marine_Salt_Water$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Aquatic_Marine_Salt_Water_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Aquatic_Marine_Salt_Water
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Marine_Salt_Water$data,
  "SAMPLING_YEAR",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-site-Aquatic_Marine_Salt_Water
#| fig-cap: "Aquatic / Marine/Salt Water: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Marine_Salt_Water$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Aquatic_Marine_Salt_Water$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Aquatic_Marine_Salt_Water_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Aquatic_Marine_Salt_Water
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Marine_Salt_Water$data,
  "SITE_CODE",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-campaign-Aquatic_Marine_Salt_Water
#| fig-cap: "Aquatic / Marine/Salt Water: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Marine_Salt_Water$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Aquatic_Marine_Salt_Water$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Aquatic_Marine_Salt_Water_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Aquatic_Marine_Salt_Water
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Marine_Salt_Water$data,
  "CAMPAIGN_NAME_SHORT",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-treatment-Aquatic_Marine_Salt_Water
#| fig-cap: "Aquatic / Marine/Salt Water: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Aquatic_Marine_Salt_Water,
  x_label = paste0("Measured value (", grp_Aquatic_Marine_Salt_Water_unit, ")")
)
```

## Porewater

Aquatic / Porewater. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Aquatic_Porewater
grp_Aquatic_Porewater <- tar_read(outlier_compartment_Aquatic_Porewater)
grp_Aquatic_Porewater$data <- grp_Aquatic_Porewater$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Aquatic_Porewater_unit <- grp_Aquatic_Porewater$summary$unit
```

```{r}
#| label: fig-outlier-year-Aquatic_Porewater
#| fig-cap: "Aquatic / Porewater: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Porewater$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Aquatic_Porewater$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Aquatic_Porewater_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Aquatic_Porewater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Porewater$data,
  "SAMPLING_YEAR",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-site-Aquatic_Porewater
#| fig-cap: "Aquatic / Porewater: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Porewater$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Aquatic_Porewater$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Aquatic_Porewater_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Aquatic_Porewater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Porewater$data,
  "SITE_CODE",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-campaign-Aquatic_Porewater
#| fig-cap: "Aquatic / Porewater: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Porewater$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Aquatic_Porewater$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Aquatic_Porewater_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Aquatic_Porewater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Porewater$data,
  "CAMPAIGN_NAME_SHORT",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-treatment-Aquatic_Porewater
#| fig-cap: "Aquatic / Porewater: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Aquatic_Porewater,
  x_label = paste0("Measured value (", grp_Aquatic_Porewater_unit, ")")
)
```

## Sludge

Aquatic / Sludge. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Aquatic_Sludge
grp_Aquatic_Sludge <- tar_read(outlier_compartment_Aquatic_Sludge)
grp_Aquatic_Sludge$data <- grp_Aquatic_Sludge$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Aquatic_Sludge_unit <- grp_Aquatic_Sludge$summary$unit
```

```{r}
#| label: fig-outlier-year-Aquatic_Sludge
#| fig-cap: "Aquatic / Sludge: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Sludge$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Aquatic_Sludge$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Aquatic_Sludge_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Aquatic_Sludge
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Sludge$data,
  "SAMPLING_YEAR",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-site-Aquatic_Sludge
#| fig-cap: "Aquatic / Sludge: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Sludge$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Aquatic_Sludge$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Aquatic_Sludge_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Aquatic_Sludge
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Sludge$data,
  "SITE_CODE",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-campaign-Aquatic_Sludge
#| fig-cap: "Aquatic / Sludge: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Sludge$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Aquatic_Sludge$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Aquatic_Sludge_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Aquatic_Sludge
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Sludge$data,
  "CAMPAIGN_NAME_SHORT",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-treatment-Aquatic_Sludge
#| fig-cap: "Aquatic / Sludge: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Aquatic_Sludge,
  x_label = paste0("Measured value (", grp_Aquatic_Sludge_unit, ")")
)
```

## Stormwater

Aquatic / Stormwater. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Aquatic_Stormwater
grp_Aquatic_Stormwater <- tar_read(outlier_compartment_Aquatic_Stormwater)
grp_Aquatic_Stormwater$data <- grp_Aquatic_Stormwater$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Aquatic_Stormwater_unit <- grp_Aquatic_Stormwater$summary$unit
```

```{r}
#| label: fig-outlier-year-Aquatic_Stormwater
#| fig-cap: "Aquatic / Stormwater: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Stormwater$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Aquatic_Stormwater$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Aquatic_Stormwater_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Aquatic_Stormwater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Stormwater$data,
  "SAMPLING_YEAR",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-site-Aquatic_Stormwater
#| fig-cap: "Aquatic / Stormwater: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Stormwater$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Aquatic_Stormwater$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Aquatic_Stormwater_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Aquatic_Stormwater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Stormwater$data,
  "SITE_CODE",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-campaign-Aquatic_Stormwater
#| fig-cap: "Aquatic / Stormwater: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Stormwater$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Aquatic_Stormwater$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Aquatic_Stormwater_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Aquatic_Stormwater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Stormwater$data,
  "CAMPAIGN_NAME_SHORT",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-treatment-Aquatic_Stormwater
#| fig-cap: "Aquatic / Stormwater: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Aquatic_Stormwater,
  x_label = paste0("Measured value (", grp_Aquatic_Stormwater_unit, ")")
)
```

## Wastewater

Aquatic / Wastewater. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Aquatic_Wastewater
grp_Aquatic_Wastewater <- tar_read(outlier_compartment_Aquatic_Wastewater)
grp_Aquatic_Wastewater$data <- grp_Aquatic_Wastewater$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Aquatic_Wastewater_unit <- grp_Aquatic_Wastewater$summary$unit
```

```{r}
#| label: fig-outlier-year-Aquatic_Wastewater
#| fig-cap: "Aquatic / Wastewater: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Wastewater$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Aquatic_Wastewater$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Aquatic_Wastewater_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Aquatic_Wastewater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Wastewater$data,
  "SAMPLING_YEAR",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-site-Aquatic_Wastewater
#| fig-cap: "Aquatic / Wastewater: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Wastewater$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Aquatic_Wastewater$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Aquatic_Wastewater_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Aquatic_Wastewater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Wastewater$data,
  "SITE_CODE",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-campaign-Aquatic_Wastewater
#| fig-cap: "Aquatic / Wastewater: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Wastewater$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Aquatic_Wastewater$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Aquatic_Wastewater_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Aquatic_Wastewater
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Wastewater$data,
  "CAMPAIGN_NAME_SHORT",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-treatment-Aquatic_Wastewater
#| fig-cap: "Aquatic / Wastewater: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Aquatic_Wastewater,
  x_label = paste0("Measured value (", grp_Aquatic_Wastewater_unit, ")")
)
```
