---
title: "Distributions: Birds"
format:
  html:
    fig-width: 16
    fig-height: 20
  docx:
    fig-width: 16
    fig-height: 20

execute:
  echo: false

knitr:
  opts_chunk:
    dpi: 300
    dev: "ragg_png"

filters:
  - flextable-qmd
  - at: post-render
    path: ../_extensions/flextable-qmd/unwrap-float.lua
---

::: callout-note
This notebook was scaffolded by `scripts/generate_distribution_notebooks.R`
from the outlier-analysis targets factory (`_targets.R`,
`R/fct_outlier_groups.R`, `R/fct_outlier_detection.R`,
`R/fct_outlier_report.R`). The plots/tables below are generated, but
analysis and interpretation are added by hand, section by section.
Re-running the generator only appends sections for groups not yet present
here -- it never rewrites or removes anything you have written.

See [NBXX-Outliers.qmd](NBXX-Outliers.qmd) for the fully hand-authored
*G. morhua* liver case study this workflow was built from, including
year/site/campaign breakdowns and the reasoning behind the outlier flags.
:::

```{r}
#| label: setup
library(tidyverse)
library(targets)
library(patchwork)
library(flextable)
pkgload::load_all(quiet = TRUE)

set_flextable_defaults(
  font.size = 10,
  font.family = "Arial",
  digits = 2,
  theme_fun = theme_vanilla,
  padding = 1
)

ggplot_base_size <- 20
theme_set(theme_minimal(base_size = ggplot_base_size, base_family = "Aptos"))
```

## Gulosus aristotelis: Egg

Birds / Gulosus aristotelis / Egg. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Birds_Gulosus_aristotelis_Egg
grp_Birds_Gulosus_aristotelis_Egg <- tar_read(outlier_biota_Birds_Gulosus_aristotelis_Egg)
grp_Birds_Gulosus_aristotelis_Egg$data <- grp_Birds_Gulosus_aristotelis_Egg$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Birds_Gulosus_aristotelis_Egg_unit <- grp_Birds_Gulosus_aristotelis_Egg$summary$unit
```

```{r}
#| label: fig-outlier-year-Birds_Gulosus_aristotelis_Egg
#| fig-cap: "Birds / Gulosus aristotelis / Egg: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Birds_Gulosus_aristotelis_Egg$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Birds_Gulosus_aristotelis_Egg$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Birds_Gulosus_aristotelis_Egg_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Birds_Gulosus_aristotelis_Egg
outlier_flag_table(flag_by_facet(grp_Birds_Gulosus_aristotelis_Egg$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Birds_Gulosus_aristotelis_Egg
#| fig-cap: "Birds / Gulosus aristotelis / Egg: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Birds_Gulosus_aristotelis_Egg$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Birds_Gulosus_aristotelis_Egg$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Birds_Gulosus_aristotelis_Egg_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Birds_Gulosus_aristotelis_Egg
outlier_flag_table(flag_by_facet(grp_Birds_Gulosus_aristotelis_Egg$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Birds_Gulosus_aristotelis_Egg
#| fig-cap: "Birds / Gulosus aristotelis / Egg: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Birds_Gulosus_aristotelis_Egg$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Birds_Gulosus_aristotelis_Egg$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Birds_Gulosus_aristotelis_Egg_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Birds_Gulosus_aristotelis_Egg
outlier_flag_table(flag_by_facet(grp_Birds_Gulosus_aristotelis_Egg$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Birds_Gulosus_aristotelis_Egg
#| fig-cap: "Birds / Gulosus aristotelis / Egg: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Birds_Gulosus_aristotelis_Egg,
  x_label = paste0("Measured value (", grp_Birds_Gulosus_aristotelis_Egg_unit, ")")
)
```

## Larus argentatus: Egg

Birds / Larus argentatus / Egg. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Birds_Larus_argentatus_Egg
grp_Birds_Larus_argentatus_Egg <- tar_read(outlier_biota_Birds_Larus_argentatus_Egg)
grp_Birds_Larus_argentatus_Egg$data <- grp_Birds_Larus_argentatus_Egg$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Birds_Larus_argentatus_Egg_unit <- grp_Birds_Larus_argentatus_Egg$summary$unit
```

```{r}
#| label: fig-outlier-year-Birds_Larus_argentatus_Egg
#| fig-cap: "Birds / Larus argentatus / Egg: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Birds_Larus_argentatus_Egg$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Birds_Larus_argentatus_Egg$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Birds_Larus_argentatus_Egg_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Birds_Larus_argentatus_Egg
outlier_flag_table(flag_by_facet(grp_Birds_Larus_argentatus_Egg$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Birds_Larus_argentatus_Egg
#| fig-cap: "Birds / Larus argentatus / Egg: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Birds_Larus_argentatus_Egg$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Birds_Larus_argentatus_Egg$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Birds_Larus_argentatus_Egg_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Birds_Larus_argentatus_Egg
outlier_flag_table(flag_by_facet(grp_Birds_Larus_argentatus_Egg$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Birds_Larus_argentatus_Egg
#| fig-cap: "Birds / Larus argentatus / Egg: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Birds_Larus_argentatus_Egg$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Birds_Larus_argentatus_Egg$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Birds_Larus_argentatus_Egg_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Birds_Larus_argentatus_Egg
outlier_flag_table(flag_by_facet(grp_Birds_Larus_argentatus_Egg$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Birds_Larus_argentatus_Egg
#| fig-cap: "Birds / Larus argentatus / Egg: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Birds_Larus_argentatus_Egg,
  x_label = paste0("Measured value (", grp_Birds_Larus_argentatus_Egg_unit, ")")
)
```

## Rissa tridactyla: Liver

Birds / Rissa tridactyla / Liver. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Birds_Rissa_tridactyla_Liver
grp_Birds_Rissa_tridactyla_Liver <- tar_read(outlier_biota_Birds_Rissa_tridactyla_Liver)
grp_Birds_Rissa_tridactyla_Liver$data <- grp_Birds_Rissa_tridactyla_Liver$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Birds_Rissa_tridactyla_Liver_unit <- grp_Birds_Rissa_tridactyla_Liver$summary$unit
```

```{r}
#| label: fig-outlier-year-Birds_Rissa_tridactyla_Liver
#| fig-cap: "Birds / Rissa tridactyla / Liver: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Birds_Rissa_tridactyla_Liver$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Birds_Rissa_tridactyla_Liver$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Birds_Rissa_tridactyla_Liver_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Birds_Rissa_tridactyla_Liver
outlier_flag_table(flag_by_facet(grp_Birds_Rissa_tridactyla_Liver$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Birds_Rissa_tridactyla_Liver
#| fig-cap: "Birds / Rissa tridactyla / Liver: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Birds_Rissa_tridactyla_Liver$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Birds_Rissa_tridactyla_Liver$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Birds_Rissa_tridactyla_Liver_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Birds_Rissa_tridactyla_Liver
outlier_flag_table(flag_by_facet(grp_Birds_Rissa_tridactyla_Liver$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Birds_Rissa_tridactyla_Liver
#| fig-cap: "Birds / Rissa tridactyla / Liver: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Birds_Rissa_tridactyla_Liver$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Birds_Rissa_tridactyla_Liver$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Birds_Rissa_tridactyla_Liver_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Birds_Rissa_tridactyla_Liver
outlier_flag_table(flag_by_facet(grp_Birds_Rissa_tridactyla_Liver$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Birds_Rissa_tridactyla_Liver
#| fig-cap: "Birds / Rissa tridactyla / Liver: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Birds_Rissa_tridactyla_Liver,
  x_label = paste0("Measured value (", grp_Birds_Rissa_tridactyla_Liver_unit, ")")
)
```

## Rissa tridactyla: Muscle

Birds / Rissa tridactyla / Muscle. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Birds_Rissa_tridactyla_Muscle
grp_Birds_Rissa_tridactyla_Muscle <- tar_read(outlier_biota_Birds_Rissa_tridactyla_Muscle)
grp_Birds_Rissa_tridactyla_Muscle$data <- grp_Birds_Rissa_tridactyla_Muscle$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Birds_Rissa_tridactyla_Muscle_unit <- grp_Birds_Rissa_tridactyla_Muscle$summary$unit
```

```{r}
#| label: fig-outlier-year-Birds_Rissa_tridactyla_Muscle
#| fig-cap: "Birds / Rissa tridactyla / Muscle: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Birds_Rissa_tridactyla_Muscle$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Birds_Rissa_tridactyla_Muscle$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Birds_Rissa_tridactyla_Muscle_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Birds_Rissa_tridactyla_Muscle
outlier_flag_table(flag_by_facet(grp_Birds_Rissa_tridactyla_Muscle$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Birds_Rissa_tridactyla_Muscle
#| fig-cap: "Birds / Rissa tridactyla / Muscle: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Birds_Rissa_tridactyla_Muscle$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Birds_Rissa_tridactyla_Muscle$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Birds_Rissa_tridactyla_Muscle_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Birds_Rissa_tridactyla_Muscle
outlier_flag_table(flag_by_facet(grp_Birds_Rissa_tridactyla_Muscle$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Birds_Rissa_tridactyla_Muscle
#| fig-cap: "Birds / Rissa tridactyla / Muscle: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Birds_Rissa_tridactyla_Muscle$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Birds_Rissa_tridactyla_Muscle$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Birds_Rissa_tridactyla_Muscle_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Birds_Rissa_tridactyla_Muscle
outlier_flag_table(flag_by_facet(grp_Birds_Rissa_tridactyla_Muscle$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Birds_Rissa_tridactyla_Muscle
#| fig-cap: "Birds / Rissa tridactyla / Muscle: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Birds_Rissa_tridactyla_Muscle,
  x_label = paste0("Measured value (", grp_Birds_Rissa_tridactyla_Muscle_unit, ")")
)
```

## Somateria mollissima: Blood

Birds / Somateria mollissima / Blood. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Birds_Somateria_mollissima_Blood
grp_Birds_Somateria_mollissima_Blood <- tar_read(outlier_biota_Birds_Somateria_mollissima_Blood)
grp_Birds_Somateria_mollissima_Blood$data <- grp_Birds_Somateria_mollissima_Blood$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Birds_Somateria_mollissima_Blood_unit <- grp_Birds_Somateria_mollissima_Blood$summary$unit
```

```{r}
#| label: fig-outlier-year-Birds_Somateria_mollissima_Blood
#| fig-cap: "Birds / Somateria mollissima / Blood: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Birds_Somateria_mollissima_Blood$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Birds_Somateria_mollissima_Blood$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Birds_Somateria_mollissima_Blood_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Birds_Somateria_mollissima_Blood
outlier_flag_table(flag_by_facet(grp_Birds_Somateria_mollissima_Blood$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Birds_Somateria_mollissima_Blood
#| fig-cap: "Birds / Somateria mollissima / Blood: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Birds_Somateria_mollissima_Blood$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Birds_Somateria_mollissima_Blood$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Birds_Somateria_mollissima_Blood_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Birds_Somateria_mollissima_Blood
outlier_flag_table(flag_by_facet(grp_Birds_Somateria_mollissima_Blood$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Birds_Somateria_mollissima_Blood
#| fig-cap: "Birds / Somateria mollissima / Blood: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Birds_Somateria_mollissima_Blood$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Birds_Somateria_mollissima_Blood$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Birds_Somateria_mollissima_Blood_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Birds_Somateria_mollissima_Blood
outlier_flag_table(flag_by_facet(grp_Birds_Somateria_mollissima_Blood$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Birds_Somateria_mollissima_Blood
#| fig-cap: "Birds / Somateria mollissima / Blood: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Birds_Somateria_mollissima_Blood,
  x_label = paste0("Measured value (", grp_Birds_Somateria_mollissima_Blood_unit, ")")
)
```

## Somateria mollissima: Egg

Birds / Somateria mollissima / Egg. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Birds_Somateria_mollissima_Egg
grp_Birds_Somateria_mollissima_Egg <- tar_read(outlier_biota_Birds_Somateria_mollissima_Egg)
grp_Birds_Somateria_mollissima_Egg$data <- grp_Birds_Somateria_mollissima_Egg$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Birds_Somateria_mollissima_Egg_unit <- grp_Birds_Somateria_mollissima_Egg$summary$unit
```

```{r}
#| label: fig-outlier-year-Birds_Somateria_mollissima_Egg
#| fig-cap: "Birds / Somateria mollissima / Egg: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Birds_Somateria_mollissima_Egg$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Birds_Somateria_mollissima_Egg$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Birds_Somateria_mollissima_Egg_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Birds_Somateria_mollissima_Egg
outlier_flag_table(flag_by_facet(grp_Birds_Somateria_mollissima_Egg$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Birds_Somateria_mollissima_Egg
#| fig-cap: "Birds / Somateria mollissima / Egg: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Birds_Somateria_mollissima_Egg$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Birds_Somateria_mollissima_Egg$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Birds_Somateria_mollissima_Egg_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Birds_Somateria_mollissima_Egg
outlier_flag_table(flag_by_facet(grp_Birds_Somateria_mollissima_Egg$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Birds_Somateria_mollissima_Egg
#| fig-cap: "Birds / Somateria mollissima / Egg: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Birds_Somateria_mollissima_Egg$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Birds_Somateria_mollissima_Egg$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Birds_Somateria_mollissima_Egg_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Birds_Somateria_mollissima_Egg
outlier_flag_table(flag_by_facet(grp_Birds_Somateria_mollissima_Egg$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Birds_Somateria_mollissima_Egg
#| fig-cap: "Birds / Somateria mollissima / Egg: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Birds_Somateria_mollissima_Egg,
  x_label = paste0("Measured value (", grp_Birds_Somateria_mollissima_Egg_unit, ")")
)
```

## Turdus pilaris: Egg

Birds / Turdus pilaris / Egg. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Birds_Turdus_pilaris_Egg
grp_Birds_Turdus_pilaris_Egg <- tar_read(outlier_biota_Birds_Turdus_pilaris_Egg)
grp_Birds_Turdus_pilaris_Egg$data <- grp_Birds_Turdus_pilaris_Egg$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Birds_Turdus_pilaris_Egg_unit <- grp_Birds_Turdus_pilaris_Egg$summary$unit
```

```{r}
#| label: fig-outlier-year-Birds_Turdus_pilaris_Egg
#| fig-cap: "Birds / Turdus pilaris / Egg: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Birds_Turdus_pilaris_Egg$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Birds_Turdus_pilaris_Egg$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Birds_Turdus_pilaris_Egg_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Birds_Turdus_pilaris_Egg
outlier_flag_table(flag_by_facet(grp_Birds_Turdus_pilaris_Egg$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Birds_Turdus_pilaris_Egg
#| fig-cap: "Birds / Turdus pilaris / Egg: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Birds_Turdus_pilaris_Egg$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Birds_Turdus_pilaris_Egg$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Birds_Turdus_pilaris_Egg_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Birds_Turdus_pilaris_Egg
outlier_flag_table(flag_by_facet(grp_Birds_Turdus_pilaris_Egg$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Birds_Turdus_pilaris_Egg
#| fig-cap: "Birds / Turdus pilaris / Egg: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Birds_Turdus_pilaris_Egg$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Birds_Turdus_pilaris_Egg$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Birds_Turdus_pilaris_Egg_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Birds_Turdus_pilaris_Egg
outlier_flag_table(flag_by_facet(grp_Birds_Turdus_pilaris_Egg$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Birds_Turdus_pilaris_Egg
#| fig-cap: "Birds / Turdus pilaris / Egg: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Birds_Turdus_pilaris_Egg,
  x_label = paste0("Measured value (", grp_Birds_Turdus_pilaris_Egg_unit, ")")
)
```
