---
title: "Distributions: Invertebrates"
format: html
---

::: callout-note
This notebook was scaffolded by `scripts/generate_distribution_notebooks.R`
from the outlier-analysis targets factory (`_targets.R`,
`R/fct_outlier_groups.R`, `R/fct_outlier_detection.R`,
`R/fct_outlier_report.R`). The plots/tables below are generated, but
analysis and interpretation are added by hand, section by section.
Re-running the generator only appends sections for groups not yet present
here -- it never rewrites or removes anything you have written.

See [NBXX-Outliers.qmd](NBXX-Outliers.qmd) for the fully hand-authored
*G. morhua* liver case study this workflow was built from, including
year/site/campaign breakdowns and the reasoning behind the outlier flags.
:::

```{r}
#| label: setup
library(tidyverse)
library(targets)
library(patchwork)
library(flextable)
pkgload::load_all(quiet = TRUE)

set_flextable_defaults(
  font.size = 10,
  font.family = "Arial",
  digits = 2,
  theme_fun = theme_vanilla,
  padding = 1
)

ggplot_base_size <- 20
theme_set(theme_minimal(base_size = ggplot_base_size, base_family = "Aptos"))
```

## Astacus astacus: Total soft tissues

Invertebrates / Astacus astacus / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Astacus_astacus_Total_soft_tissues
grp_Invertebrates_Astacus_astacus_Total_soft_tissues <- tar_read(outlier_biota_Invertebrates_Astacus_astacus_Total_soft_tissues)
grp_Invertebrates_Astacus_astacus_Total_soft_tissues$data <- grp_Invertebrates_Astacus_astacus_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Astacus_astacus_Total_soft_tissues_unit <- grp_Invertebrates_Astacus_astacus_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Astacus_astacus_Total_soft_tissues
#| fig-cap: "Invertebrates / Astacus astacus / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Astacus_astacus_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Astacus_astacus_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Astacus_astacus_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Astacus_astacus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Invertebrates_Astacus_astacus_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Astacus_astacus_Total_soft_tissues
#| fig-cap: "Invertebrates / Astacus astacus / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Astacus_astacus_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Astacus_astacus_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Astacus_astacus_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Astacus_astacus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Invertebrates_Astacus_astacus_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Astacus_astacus_Total_soft_tissues
#| fig-cap: "Invertebrates / Astacus astacus / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Astacus_astacus_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Astacus_astacus_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Astacus_astacus_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Astacus_astacus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Invertebrates_Astacus_astacus_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Astacus_astacus_Total_soft_tissues
#| fig-cap: "Invertebrates / Astacus astacus / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Astacus_astacus_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Invertebrates_Astacus_astacus_Total_soft_tissues_unit, ")")
)
```

## Asterias rubens: Body wall

Invertebrates / Asterias rubens / Body wall. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Asterias_rubens_Body_wall
grp_Invertebrates_Asterias_rubens_Body_wall <- tar_read(outlier_biota_Invertebrates_Asterias_rubens_Body_wall)
grp_Invertebrates_Asterias_rubens_Body_wall$data <- grp_Invertebrates_Asterias_rubens_Body_wall$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Asterias_rubens_Body_wall_unit <- grp_Invertebrates_Asterias_rubens_Body_wall$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Asterias_rubens_Body_wall
#| fig-cap: "Invertebrates / Asterias rubens / Body wall: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Asterias_rubens_Body_wall$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Asterias_rubens_Body_wall$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Asterias_rubens_Body_wall_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Asterias_rubens_Body_wall
outlier_flag_table(flag_by_facet(grp_Invertebrates_Asterias_rubens_Body_wall$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Asterias_rubens_Body_wall
#| fig-cap: "Invertebrates / Asterias rubens / Body wall: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Asterias_rubens_Body_wall$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Asterias_rubens_Body_wall$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Asterias_rubens_Body_wall_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Asterias_rubens_Body_wall
outlier_flag_table(flag_by_facet(grp_Invertebrates_Asterias_rubens_Body_wall$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Asterias_rubens_Body_wall
#| fig-cap: "Invertebrates / Asterias rubens / Body wall: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Asterias_rubens_Body_wall$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Asterias_rubens_Body_wall$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Asterias_rubens_Body_wall_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Asterias_rubens_Body_wall
outlier_flag_table(flag_by_facet(grp_Invertebrates_Asterias_rubens_Body_wall$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Asterias_rubens_Body_wall
#| fig-cap: "Invertebrates / Asterias rubens / Body wall: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Asterias_rubens_Body_wall,
  x_label = paste0("Measured value (", grp_Invertebrates_Asterias_rubens_Body_wall_unit, ")")
)
```

## Asterias rubens: Pyloric caeca

Invertebrates / Asterias rubens / Pyloric caeca. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Asterias_rubens_Pyloric_caeca
grp_Invertebrates_Asterias_rubens_Pyloric_caeca <- tar_read(outlier_biota_Invertebrates_Asterias_rubens_Pyloric_caeca)
grp_Invertebrates_Asterias_rubens_Pyloric_caeca$data <- grp_Invertebrates_Asterias_rubens_Pyloric_caeca$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Asterias_rubens_Pyloric_caeca_unit <- grp_Invertebrates_Asterias_rubens_Pyloric_caeca$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Asterias_rubens_Pyloric_caeca
#| fig-cap: "Invertebrates / Asterias rubens / Pyloric caeca: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Asterias_rubens_Pyloric_caeca$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Asterias_rubens_Pyloric_caeca$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Asterias_rubens_Pyloric_caeca_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Asterias_rubens_Pyloric_caeca
outlier_flag_table(flag_by_facet(grp_Invertebrates_Asterias_rubens_Pyloric_caeca$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Asterias_rubens_Pyloric_caeca
#| fig-cap: "Invertebrates / Asterias rubens / Pyloric caeca: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Asterias_rubens_Pyloric_caeca$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Asterias_rubens_Pyloric_caeca$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Asterias_rubens_Pyloric_caeca_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Asterias_rubens_Pyloric_caeca
outlier_flag_table(flag_by_facet(grp_Invertebrates_Asterias_rubens_Pyloric_caeca$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Asterias_rubens_Pyloric_caeca
#| fig-cap: "Invertebrates / Asterias rubens / Pyloric caeca: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Asterias_rubens_Pyloric_caeca$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Asterias_rubens_Pyloric_caeca$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Asterias_rubens_Pyloric_caeca_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Asterias_rubens_Pyloric_caeca
outlier_flag_table(flag_by_facet(grp_Invertebrates_Asterias_rubens_Pyloric_caeca$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Asterias_rubens_Pyloric_caeca
#| fig-cap: "Invertebrates / Asterias rubens / Pyloric caeca: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Asterias_rubens_Pyloric_caeca,
  x_label = paste0("Measured value (", grp_Invertebrates_Asterias_rubens_Pyloric_caeca_unit, ")")
)
```

## Chironomidae: Whole body

Invertebrates / Chironomidae / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Chironomidae_Whole_body
grp_Invertebrates_Chironomidae_Whole_body <- tar_read(outlier_biota_Invertebrates_Chironomidae_Whole_body)
grp_Invertebrates_Chironomidae_Whole_body$data <- grp_Invertebrates_Chironomidae_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Chironomidae_Whole_body_unit <- grp_Invertebrates_Chironomidae_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Chironomidae_Whole_body
#| fig-cap: "Invertebrates / Chironomidae / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Chironomidae_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Chironomidae_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Chironomidae_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Chironomidae_Whole_body
outlier_flag_table(flag_by_facet(grp_Invertebrates_Chironomidae_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Chironomidae_Whole_body
#| fig-cap: "Invertebrates / Chironomidae / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Chironomidae_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Chironomidae_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Chironomidae_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Chironomidae_Whole_body
outlier_flag_table(flag_by_facet(grp_Invertebrates_Chironomidae_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Chironomidae_Whole_body
#| fig-cap: "Invertebrates / Chironomidae / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Chironomidae_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Chironomidae_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Chironomidae_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Chironomidae_Whole_body
outlier_flag_table(flag_by_facet(grp_Invertebrates_Chironomidae_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Chironomidae_Whole_body
#| fig-cap: "Invertebrates / Chironomidae / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Chironomidae_Whole_body,
  x_label = paste0("Measured value (", grp_Invertebrates_Chironomidae_Whole_body_unit, ")")
)
```

## Crossaster papposus: Disc skeleton

Invertebrates / Crossaster papposus / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Crossaster_papposus_Disc_skeleton
grp_Invertebrates_Crossaster_papposus_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Crossaster_papposus_Disc_skeleton)
grp_Invertebrates_Crossaster_papposus_Disc_skeleton$data <- grp_Invertebrates_Crossaster_papposus_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Crossaster_papposus_Disc_skeleton_unit <- grp_Invertebrates_Crossaster_papposus_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Crossaster_papposus_Disc_skeleton
#| fig-cap: "Invertebrates / Crossaster papposus / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Crossaster_papposus_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Crossaster_papposus_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Crossaster_papposus_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Crossaster_papposus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Crossaster_papposus_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Crossaster_papposus_Disc_skeleton
#| fig-cap: "Invertebrates / Crossaster papposus / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Crossaster_papposus_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Crossaster_papposus_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Crossaster_papposus_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Crossaster_papposus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Crossaster_papposus_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Crossaster_papposus_Disc_skeleton
#| fig-cap: "Invertebrates / Crossaster papposus / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Crossaster_papposus_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Crossaster_papposus_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Crossaster_papposus_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Crossaster_papposus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Crossaster_papposus_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Crossaster_papposus_Disc_skeleton
#| fig-cap: "Invertebrates / Crossaster papposus / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Crossaster_papposus_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Crossaster_papposus_Disc_skeleton_unit, ")")
)
```

## Ctenodiscus crispatus: Disc skeleton

Invertebrates / Ctenodiscus crispatus / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Ctenodiscus_crispatus_Disc_skeleton
grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton)
grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton$data <- grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton_unit <- grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Ctenodiscus_crispatus_Disc_skeleton
#| fig-cap: "Invertebrates / Ctenodiscus crispatus / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Ctenodiscus_crispatus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Ctenodiscus_crispatus_Disc_skeleton
#| fig-cap: "Invertebrates / Ctenodiscus crispatus / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Ctenodiscus_crispatus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Ctenodiscus_crispatus_Disc_skeleton
#| fig-cap: "Invertebrates / Ctenodiscus crispatus / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Ctenodiscus_crispatus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Ctenodiscus_crispatus_Disc_skeleton
#| fig-cap: "Invertebrates / Ctenodiscus crispatus / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Ctenodiscus_crispatus_Disc_skeleton_unit, ")")
)
```

## Gorgonocephalus arcticus: Disc skeleton

Invertebrates / Gorgonocephalus arcticus / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton
grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton)
grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton$data <- grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton_unit <- grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton
#| fig-cap: "Invertebrates / Gorgonocephalus arcticus / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton
#| fig-cap: "Invertebrates / Gorgonocephalus arcticus / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton
#| fig-cap: "Invertebrates / Gorgonocephalus arcticus / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton
#| fig-cap: "Invertebrates / Gorgonocephalus arcticus / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Gorgonocephalus_arcticus_Disc_skeleton_unit, ")")
)
```

## Gorgonocephalus lamarckii: Disc skeleton

Invertebrates / Gorgonocephalus lamarckii / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton
grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton)
grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton$data <- grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton_unit <- grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton
#| fig-cap: "Invertebrates / Gorgonocephalus lamarckii / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton
#| fig-cap: "Invertebrates / Gorgonocephalus lamarckii / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton
#| fig-cap: "Invertebrates / Gorgonocephalus lamarckii / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton
#| fig-cap: "Invertebrates / Gorgonocephalus lamarckii / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Gorgonocephalus_lamarckii_Disc_skeleton_unit, ")")
)
```

## Heliometra glacialis: Disc skeleton

Invertebrates / Heliometra glacialis / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Heliometra_glacialis_Disc_skeleton
grp_Invertebrates_Heliometra_glacialis_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Heliometra_glacialis_Disc_skeleton)
grp_Invertebrates_Heliometra_glacialis_Disc_skeleton$data <- grp_Invertebrates_Heliometra_glacialis_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Heliometra_glacialis_Disc_skeleton_unit <- grp_Invertebrates_Heliometra_glacialis_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Heliometra_glacialis_Disc_skeleton
#| fig-cap: "Invertebrates / Heliometra glacialis / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Heliometra_glacialis_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Heliometra_glacialis_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Heliometra_glacialis_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Heliometra_glacialis_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Heliometra_glacialis_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Heliometra_glacialis_Disc_skeleton
#| fig-cap: "Invertebrates / Heliometra glacialis / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Heliometra_glacialis_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Heliometra_glacialis_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Heliometra_glacialis_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Heliometra_glacialis_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Heliometra_glacialis_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Heliometra_glacialis_Disc_skeleton
#| fig-cap: "Invertebrates / Heliometra glacialis / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Heliometra_glacialis_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Heliometra_glacialis_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Heliometra_glacialis_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Heliometra_glacialis_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Heliometra_glacialis_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Heliometra_glacialis_Disc_skeleton
#| fig-cap: "Invertebrates / Heliometra glacialis / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Heliometra_glacialis_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Heliometra_glacialis_Disc_skeleton_unit, ")")
)
```

## Hippasteria phrygiana: Disc skeleton

Invertebrates / Hippasteria phrygiana / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Hippasteria_phrygiana_Disc_skeleton
grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Hippasteria_phrygiana_Disc_skeleton)
grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton$data <- grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton_unit <- grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Hippasteria_phrygiana_Disc_skeleton
#| fig-cap: "Invertebrates / Hippasteria phrygiana / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Hippasteria_phrygiana_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Hippasteria_phrygiana_Disc_skeleton
#| fig-cap: "Invertebrates / Hippasteria phrygiana / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Hippasteria_phrygiana_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Hippasteria_phrygiana_Disc_skeleton
#| fig-cap: "Invertebrates / Hippasteria phrygiana / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Hippasteria_phrygiana_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Hippasteria_phrygiana_Disc_skeleton
#| fig-cap: "Invertebrates / Hippasteria phrygiana / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Hippasteria_phrygiana_Disc_skeleton_unit, ")")
)
```

## Icasterias panopla: Disc skeleton

Invertebrates / Icasterias panopla / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Icasterias_panopla_Disc_skeleton
grp_Invertebrates_Icasterias_panopla_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Icasterias_panopla_Disc_skeleton)
grp_Invertebrates_Icasterias_panopla_Disc_skeleton$data <- grp_Invertebrates_Icasterias_panopla_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Icasterias_panopla_Disc_skeleton_unit <- grp_Invertebrates_Icasterias_panopla_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Icasterias_panopla_Disc_skeleton
#| fig-cap: "Invertebrates / Icasterias panopla / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Icasterias_panopla_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Icasterias_panopla_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Icasterias_panopla_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Icasterias_panopla_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Icasterias_panopla_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Icasterias_panopla_Disc_skeleton
#| fig-cap: "Invertebrates / Icasterias panopla / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Icasterias_panopla_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Icasterias_panopla_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Icasterias_panopla_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Icasterias_panopla_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Icasterias_panopla_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Icasterias_panopla_Disc_skeleton
#| fig-cap: "Invertebrates / Icasterias panopla / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Icasterias_panopla_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Icasterias_panopla_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Icasterias_panopla_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Icasterias_panopla_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Icasterias_panopla_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Icasterias_panopla_Disc_skeleton
#| fig-cap: "Invertebrates / Icasterias panopla / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Icasterias_panopla_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Icasterias_panopla_Disc_skeleton_unit, ")")
)
```

## Insecta: Whole body

Invertebrates / Insecta / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Insecta_Whole_body
grp_Invertebrates_Insecta_Whole_body <- tar_read(outlier_biota_Invertebrates_Insecta_Whole_body)
grp_Invertebrates_Insecta_Whole_body$data <- grp_Invertebrates_Insecta_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Insecta_Whole_body_unit <- grp_Invertebrates_Insecta_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Insecta_Whole_body
#| fig-cap: "Invertebrates / Insecta / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Insecta_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Insecta_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Insecta_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Insecta_Whole_body
outlier_flag_table(flag_by_facet(grp_Invertebrates_Insecta_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Insecta_Whole_body
#| fig-cap: "Invertebrates / Insecta / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Insecta_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Insecta_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Insecta_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Insecta_Whole_body
outlier_flag_table(flag_by_facet(grp_Invertebrates_Insecta_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Insecta_Whole_body
#| fig-cap: "Invertebrates / Insecta / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Insecta_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Insecta_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Insecta_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Insecta_Whole_body
outlier_flag_table(flag_by_facet(grp_Invertebrates_Insecta_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Insecta_Whole_body
#| fig-cap: "Invertebrates / Insecta / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Insecta_Whole_body,
  x_label = paste0("Measured value (", grp_Invertebrates_Insecta_Whole_body_unit, ")")
)
```

## Ophiopleura borealis: Disc skeleton

Invertebrates / Ophiopleura borealis / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Ophiopleura_borealis_Disc_skeleton
grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Ophiopleura_borealis_Disc_skeleton)
grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton$data <- grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton_unit <- grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Ophiopleura_borealis_Disc_skeleton
#| fig-cap: "Invertebrates / Ophiopleura borealis / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Ophiopleura_borealis_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Ophiopleura_borealis_Disc_skeleton
#| fig-cap: "Invertebrates / Ophiopleura borealis / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Ophiopleura_borealis_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Ophiopleura_borealis_Disc_skeleton
#| fig-cap: "Invertebrates / Ophiopleura borealis / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Ophiopleura_borealis_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Ophiopleura_borealis_Disc_skeleton
#| fig-cap: "Invertebrates / Ophiopleura borealis / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Ophiopleura_borealis_Disc_skeleton_unit, ")")
)
```

## Ophiura sarsii: Disc skeleton

Invertebrates / Ophiura sarsii / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Ophiura_sarsii_Disc_skeleton
grp_Invertebrates_Ophiura_sarsii_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Ophiura_sarsii_Disc_skeleton)
grp_Invertebrates_Ophiura_sarsii_Disc_skeleton$data <- grp_Invertebrates_Ophiura_sarsii_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Ophiura_sarsii_Disc_skeleton_unit <- grp_Invertebrates_Ophiura_sarsii_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Ophiura_sarsii_Disc_skeleton
#| fig-cap: "Invertebrates / Ophiura sarsii / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Ophiura_sarsii_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Ophiura_sarsii_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Ophiura_sarsii_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Ophiura_sarsii_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Ophiura_sarsii_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Ophiura_sarsii_Disc_skeleton
#| fig-cap: "Invertebrates / Ophiura sarsii / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Ophiura_sarsii_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Ophiura_sarsii_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Ophiura_sarsii_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Ophiura_sarsii_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Ophiura_sarsii_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Ophiura_sarsii_Disc_skeleton
#| fig-cap: "Invertebrates / Ophiura sarsii / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Ophiura_sarsii_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Ophiura_sarsii_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Ophiura_sarsii_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Ophiura_sarsii_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Ophiura_sarsii_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Ophiura_sarsii_Disc_skeleton
#| fig-cap: "Invertebrates / Ophiura sarsii / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Ophiura_sarsii_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Ophiura_sarsii_Disc_skeleton_unit, ")")
)
```

## Pontaster tenuispinus: Disc skeleton

Invertebrates / Pontaster tenuispinus / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Pontaster_tenuispinus_Disc_skeleton
grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Pontaster_tenuispinus_Disc_skeleton)
grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton$data <- grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton_unit <- grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Pontaster_tenuispinus_Disc_skeleton
#| fig-cap: "Invertebrates / Pontaster tenuispinus / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Pontaster_tenuispinus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Pontaster_tenuispinus_Disc_skeleton
#| fig-cap: "Invertebrates / Pontaster tenuispinus / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Pontaster_tenuispinus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Pontaster_tenuispinus_Disc_skeleton
#| fig-cap: "Invertebrates / Pontaster tenuispinus / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Pontaster_tenuispinus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Pontaster_tenuispinus_Disc_skeleton
#| fig-cap: "Invertebrates / Pontaster tenuispinus / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Pontaster_tenuispinus_Disc_skeleton_unit, ")")
)
```

## Pteraster obscurus: Disc skeleton

Invertebrates / Pteraster obscurus / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Pteraster_obscurus_Disc_skeleton
grp_Invertebrates_Pteraster_obscurus_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Pteraster_obscurus_Disc_skeleton)
grp_Invertebrates_Pteraster_obscurus_Disc_skeleton$data <- grp_Invertebrates_Pteraster_obscurus_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Pteraster_obscurus_Disc_skeleton_unit <- grp_Invertebrates_Pteraster_obscurus_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Pteraster_obscurus_Disc_skeleton
#| fig-cap: "Invertebrates / Pteraster obscurus / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Pteraster_obscurus_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Pteraster_obscurus_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Pteraster_obscurus_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Pteraster_obscurus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Pteraster_obscurus_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Pteraster_obscurus_Disc_skeleton
#| fig-cap: "Invertebrates / Pteraster obscurus / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Pteraster_obscurus_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Pteraster_obscurus_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Pteraster_obscurus_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Pteraster_obscurus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Pteraster_obscurus_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Pteraster_obscurus_Disc_skeleton
#| fig-cap: "Invertebrates / Pteraster obscurus / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Pteraster_obscurus_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Pteraster_obscurus_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Pteraster_obscurus_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Pteraster_obscurus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Pteraster_obscurus_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Pteraster_obscurus_Disc_skeleton
#| fig-cap: "Invertebrates / Pteraster obscurus / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Pteraster_obscurus_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Pteraster_obscurus_Disc_skeleton_unit, ")")
)
```

## Pteraster pulvillus: Disc skeleton

Invertebrates / Pteraster pulvillus / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Pteraster_pulvillus_Disc_skeleton
grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Pteraster_pulvillus_Disc_skeleton)
grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton$data <- grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton_unit <- grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Pteraster_pulvillus_Disc_skeleton
#| fig-cap: "Invertebrates / Pteraster pulvillus / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Pteraster_pulvillus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Pteraster_pulvillus_Disc_skeleton
#| fig-cap: "Invertebrates / Pteraster pulvillus / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Pteraster_pulvillus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Pteraster_pulvillus_Disc_skeleton
#| fig-cap: "Invertebrates / Pteraster pulvillus / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Pteraster_pulvillus_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Pteraster_pulvillus_Disc_skeleton
#| fig-cap: "Invertebrates / Pteraster pulvillus / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Pteraster_pulvillus_Disc_skeleton_unit, ")")
)
```

## Strongylocentrotus droebachiensis: Gonad

Invertebrates / Strongylocentrotus droebachiensis / Gonad. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Strongylocentrotus_droebachiensis_Gonad
grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad <- tar_read(outlier_biota_Invertebrates_Strongylocentrotus_droebachiensis_Gonad)
grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad$data <- grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad_unit <- grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Strongylocentrotus_droebachiensis_Gonad
#| fig-cap: "Invertebrates / Strongylocentrotus droebachiensis / Gonad: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Strongylocentrotus_droebachiensis_Gonad
outlier_flag_table(flag_by_facet(grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Strongylocentrotus_droebachiensis_Gonad
#| fig-cap: "Invertebrates / Strongylocentrotus droebachiensis / Gonad: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Strongylocentrotus_droebachiensis_Gonad
outlier_flag_table(flag_by_facet(grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Strongylocentrotus_droebachiensis_Gonad
#| fig-cap: "Invertebrates / Strongylocentrotus droebachiensis / Gonad: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Strongylocentrotus_droebachiensis_Gonad
outlier_flag_table(flag_by_facet(grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Strongylocentrotus_droebachiensis_Gonad
#| fig-cap: "Invertebrates / Strongylocentrotus droebachiensis / Gonad: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad,
  x_label = paste0("Measured value (", grp_Invertebrates_Strongylocentrotus_droebachiensis_Gonad_unit, ")")
)
```

## Strongylocentrotus droebachiensis: Total soft tissues minus gonads

Invertebrates / Strongylocentrotus droebachiensis / Total soft tissues minus gonads. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads
grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads <- tar_read(outlier_biota_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads)
grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads$data <- grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads_unit <- grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads
#| fig-cap: "Invertebrates / Strongylocentrotus droebachiensis / Total soft tissues minus gonads: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads
outlier_flag_table(flag_by_facet(grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads
#| fig-cap: "Invertebrates / Strongylocentrotus droebachiensis / Total soft tissues minus gonads: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads
outlier_flag_table(flag_by_facet(grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads
#| fig-cap: "Invertebrates / Strongylocentrotus droebachiensis / Total soft tissues minus gonads: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads
outlier_flag_table(flag_by_facet(grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads
#| fig-cap: "Invertebrates / Strongylocentrotus droebachiensis / Total soft tissues minus gonads: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads,
  x_label = paste0("Measured value (", grp_Invertebrates_Strongylocentrotus_droebachiensis_Total_soft_tissues_minus_gonads_unit, ")")
)
```

## Strongylocentrotus pallidus: Echinoid corona

Invertebrates / Strongylocentrotus pallidus / Echinoid corona. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona
grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona <- tar_read(outlier_biota_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona)
grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona$data <- grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona_unit <- grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona
#| fig-cap: "Invertebrates / Strongylocentrotus pallidus / Echinoid corona: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona
outlier_flag_table(flag_by_facet(grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona
#| fig-cap: "Invertebrates / Strongylocentrotus pallidus / Echinoid corona: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona
outlier_flag_table(flag_by_facet(grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona
#| fig-cap: "Invertebrates / Strongylocentrotus pallidus / Echinoid corona: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona
outlier_flag_table(flag_by_facet(grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona
#| fig-cap: "Invertebrates / Strongylocentrotus pallidus / Echinoid corona: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona,
  x_label = paste0("Measured value (", grp_Invertebrates_Strongylocentrotus_pallidus_Echinoid_corona_unit, ")")
)
```

## Urasterias lincki: Disc skeleton

Invertebrates / Urasterias lincki / Disc skeleton. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Invertebrates_Urasterias_lincki_Disc_skeleton
grp_Invertebrates_Urasterias_lincki_Disc_skeleton <- tar_read(outlier_biota_Invertebrates_Urasterias_lincki_Disc_skeleton)
grp_Invertebrates_Urasterias_lincki_Disc_skeleton$data <- grp_Invertebrates_Urasterias_lincki_Disc_skeleton$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Invertebrates_Urasterias_lincki_Disc_skeleton_unit <- grp_Invertebrates_Urasterias_lincki_Disc_skeleton$summary$unit
```

```{r}
#| label: fig-outlier-year-Invertebrates_Urasterias_lincki_Disc_skeleton
#| fig-cap: "Invertebrates / Urasterias lincki / Disc skeleton: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Urasterias_lincki_Disc_skeleton$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Invertebrates_Urasterias_lincki_Disc_skeleton$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Invertebrates_Urasterias_lincki_Disc_skeleton_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Invertebrates_Urasterias_lincki_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Urasterias_lincki_Disc_skeleton$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Invertebrates_Urasterias_lincki_Disc_skeleton
#| fig-cap: "Invertebrates / Urasterias lincki / Disc skeleton: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Urasterias_lincki_Disc_skeleton$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Invertebrates_Urasterias_lincki_Disc_skeleton$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Invertebrates_Urasterias_lincki_Disc_skeleton_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Invertebrates_Urasterias_lincki_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Urasterias_lincki_Disc_skeleton$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Invertebrates_Urasterias_lincki_Disc_skeleton
#| fig-cap: "Invertebrates / Urasterias lincki / Disc skeleton: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Invertebrates_Urasterias_lincki_Disc_skeleton$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Invertebrates_Urasterias_lincki_Disc_skeleton$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Invertebrates_Urasterias_lincki_Disc_skeleton_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Invertebrates_Urasterias_lincki_Disc_skeleton
outlier_flag_table(flag_by_facet(grp_Invertebrates_Urasterias_lincki_Disc_skeleton$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Invertebrates_Urasterias_lincki_Disc_skeleton
#| fig-cap: "Invertebrates / Urasterias lincki / Disc skeleton: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Invertebrates_Urasterias_lincki_Disc_skeleton,
  x_label = paste0("Measured value (", grp_Invertebrates_Urasterias_lincki_Disc_skeleton_unit, ")")
)
```
