---
title: "Sample Groups"
# Formatting lives in docs/_metadata.yml, which Quarto applies to every notebook
# in this directory and to nothing above it. page-layout, lightbox, toc-depth
# and embed-resources are all set there.
---

This is a top-but-one level notebook containing all sample groups.

```{r}
#| label: setup
#| include: false
#| warning: false

library(targets)
library(here)
library(tidyverse)
library(flextable)
pkgload::load_all(quiet = TRUE) # sample_groups_flextable() etc.

here::i_am("index.qmd")
```


```{r}
#| label: tbl-groups-summary
#| tbl-caption: tbc

# Reshaping lives in the sample_groups_table target (R/fct_sample_groups_table.R),
# so this notebook and index.qmd cannot drift apart. All 245 groups shown here;
# the manuscript shows only the large ones.
#
# link_sections is passed the anchors that exist as headings further down this
# page, so those group names become links and the rest stay plain text. The
# manuscript omits the argument, because it holds none of these sections.
tar_read(sample_groups_table) |>
  sample_groups_flextable(
    link_sections = tar_read(triage_pilot_groups)$heading_slug
  )
```

# Triage plots

One row of five views per group, generated by `R/fct_group_triage.R` and
written to `triage/` as PNGs by the `triage_pilot_plots` target. Click any
plot to enlarge it.

Currently **25 groups**, sampled from the `r nrow(tar_read(sample_groups_table))`-group
table and restricted to those with `n >= 100`. The sample is seeded, so it stays
the same between runs. Widen it by raising `n_sample` in the
`triage_pilot_groups` target, or set it to `Inf` for all of them.

Headings nest by the group key: compartment, sub-compartment, then taxonomy for
biota, with geographic feature and sub-feature sharing the leaf. Within each
level, siblings are ordered by total measurements, so the heaviest branches come
first. Units are **not** a heading level: where a group holds both dry and wet
weight measurements, each gets its own plot row under the shared heading.

::: callout-note
## Why some plots show points and others show bins

Every view switches presentation on the number of **rows** available, at a
threshold of 30 (`triage_use_points()`). Above it, values are summarised into
2D bins or hexes; below it, they are drawn as individual points. This is
deliberate: a hex map built from four observations is mostly empty cells and
conveys less than the four points themselves, while forty thousand points
overplot into a solid block and take an age to draw.

So a group can legitimately show hexes in one panel and points in another, and
two groups on this page can present the same view differently. **Aquatic /
Freshwater / River, stream, canal** is the clearest example: it clears
`n >= 100` on the summary table but has only 4 underlying rows, so its spatial
panel falls back to points.

That gap is worth watching. `n` in the summary table is `sum(MEASURED_N)`, a
count of *measurements*, whereas the plots draw one mark per *row*. A group
that aggregates many measurements into few rows will look far sparser here
than its `n` suggests. `n_rows` is reported in each heading below so the
discrepancy is visible rather than surprising.
:::

::: callout-warning
## The threshold lines are sanity checks, not an assessment

Coloured reference lines are Norwegian M-608 classification boundaries, PROREF
background values, and ICES BACs. Their **class is named on the secondary axis**,
and the axis title names the source. Colour and line style both follow the
Miljødirektoratet class number, running blue and dotted for Background (I)
through to red and solid for Very Poor (V), so severity reads off the line
itself. Copper has no Class III in either freshwater or sediment: M-608 runs
Class II straight into Class IV, and the styling is keyed on the class number so
Poor stays orange despite the gap.

We hold thresholds for four situations (freshwater, coastal water, marine
sediment, and two biota species) and apply them to every compartment, species and
tissue that shares a unit. So:

- water classes are for *dissolved* copper, most measurements here are total;
- the sediment classes are marine, applied also to lake and river sediment;
- the vertebrate comparator is *Gadus morhua* **liver**, which concentrates
  copper well above muscle, so muscle groups will look clean against it;
- the invertebrate comparator is whole *Mytilus edulis* soft tissue.

Units are matched, never converted, so dry-weight groups get dry-weight
thresholds or none at all. Nothing here is a risk assessment; the lines exist to
make an error of three orders of magnitude obvious, and they will not catch a
wrong value that happens to look plausible. See the header of
`R/fct_threshold_match.R` for the full argument.

Panel (b) draws the lines but omits the class axis: on a vertical axis the
numerals collide. Noted as `PLAN.md` P1.1g.
:::

```{r}
#| label: fig-triage-plots
#| output: asis

groups <- tar_read(triage_pilot_groups)
plot_paths <- tar_read(triage_pilot_plots)

# Filename suffix -> caption. Keys match the list names in
# write_triage_plots_for_group(), which also drive the a/b/c/d/e file ordering.
captions <- c(
  a_density = "a) Overall distribution (all units)",
  b_date = "b) Concentration by date",
  c_campaign = "c) Distribution by campaign",
  d_site_type = "d) Distribution by site type (all geographies)",
  e_spatial = "e) Spatial distribution"
)

# Columns forming the heading tree, outermost first. Unit is absent by design:
# see triage_heading_cols(). The last two share the leaf heading because
# markdown stops at six levels and the biota chain is seven deep.
heading_cols <- triage_heading_cols()
leaf_cols <- tail(heading_cols, 2)
level_cols <- head(heading_cols, -2)

# groups arrives already sorted by sort_triage_groups(): hierarchical, with
# siblings ranked by total measurements. Tracking the previous row's values is
# therefore enough to know which headings to (re)open.
previous <- rep(NA_character_, length(heading_cols))
names(previous) <- heading_cols
previous_slug <- ""

for (i in seq_len(nrow(groups))) {
  grp <- groups[i, , drop = FALSE]
  current <- vapply(
    heading_cols,
    function(col) as.character(grp[[col]][1]),
    character(1)
  )

  # Unit variants of one group share a heading and an anchor, so the second and
  # subsequent variants emit no heading at all and go straight to their plot
  # row. Emitting it again would duplicate both the visible heading and its HTML
  # id, and duplicate ids break the link from the summary table.
  new_heading <- !identical(grp$heading_slug[1], previous_slug)

  if (new_heading) {
    # A group whose taxonomy columns are NA (everything non-biota) simply has
    # fewer levels, so the tree is naturally shallower there rather than
    # carrying empty "NA" headings.
    present <- level_cols[!is.na(current[level_cols])]
    leaf_parts <- current[leaf_cols]
    leaf <- paste(
      ifelse(is.na(leaf_parts), "Unknown", leaf_parts),
      collapse = " › "
    )

    # Reopen from the shallowest level that changed. Without this, a sibling at
    # depth 2 would sit under the previous row's depth-3 heading. Where nothing
    # changed, no ancestor headings are emitted and the leaf correctly appears
    # as a sibling under the headings already open.
    changed <- which(
      current[present] != previous[present] | is.na(previous[present])
    )
    from <- if (length(changed) == 0) length(present) + 1L else min(changed)

    for (d in seq_along(present)) {
      if (d < from) next
      cat("\n\n", strrep("#", d), " ", current[present[d]], "\n", sep = "")
    }

    # The leaf carries an explicit id so the summary table can link to it. The
    # anchor is shared by every unit variant of this group, which is why
    # heading_anchor() does not go through slugify_name()'s make.unique().
    cat(
      "\n\n",
      strrep("#", length(present) + 1L),
      " ",
      leaf,
      " {#",
      grp$heading_slug[1],
      "}\n\n",
      sep = ""
    )

    previous <- current
    previous_slug <- grp$heading_slug[1]
  }

  # Built by group_summary_line() rather than assembled here, so the flag wording
  # comes from the same place the summary table's highlighting does and the two
  # cannot disagree about what is flagged.
  cat(group_summary_line(grp), "\n\n", sep = "")

  # Build the expected filename per view rather than pattern-matching the
  # path list. Prefix matching broke when one slug was a string prefix of
  # another, which make.unique() suffixes used to produce.
  # Paths are project-root-relative; this notebook renders from docs/
  cat("::: {layout-ncol=5}\n\n")
  for (key in names(captions)) {
    p <- file.path("triage", paste0(grp$group_slug[1], "_", key, ".png"))
    # Checked against the target's own output rather than file.exists(): that
    # keeps this chunk's tar_read(triage_pilot_plots) dependency live, so the
    # render target rebuilds after the plots. Dropping that read once caused
    # the notebook to render before the PNGs were written.
    if (!p %in% plot_paths) {
      cat("*", captions[[key]], " not available.*\n\n", sep = "")
      next
    }
    cat(
      "![",
      captions[[key]],
      "](",
      file.path("..", p),
      ")",
      "{group=\"",
      grp$group_slug[1],
      "\"}\n\n",
      sep = ""
    )
  }
  cat(":::\n\n")
}
```
