---
title: "Distributions: Aquatic > Aquatic Sediment"
format: html
---

::: callout-note
This notebook was scaffolded by `scripts/generate_distribution_notebooks.R`
from the outlier-analysis targets factory (`_targets.R`,
`R/fct_outlier_groups.R`, `R/fct_outlier_detection.R`,
`R/fct_outlier_report.R`). The plots/tables below are generated, but
analysis and interpretation are added by hand, section by section.
Re-running the generator only appends sections for groups not yet present
here -- it never rewrites or removes anything you have written.

See [NBXX-Outliers.qmd](NBXX-Outliers.qmd) for the fully hand-authored
*G. morhua* liver case study this workflow was built from, including
year/site/campaign breakdowns and the reasoning behind the outlier flags.
:::

```{r}
#| label: setup
library(tidyverse)
library(targets)
library(patchwork)
library(flextable)
pkgload::load_all(quiet = TRUE)

set_flextable_defaults(
  font.size = 10,
  font.family = "Arial",
  digits = 2,
  theme_fun = theme_vanilla,
  padding = 1
)

ggplot_base_size <- 20
theme_set(theme_minimal(base_size = ggplot_base_size, base_family = "Aptos"))
```

# Overview

```{r}
#| label: read-Aquatic_Aquatic_Sediment
grp_Aquatic_Aquatic_Sediment <- tar_read(
  outlier_compartment_Aquatic_Aquatic_Sediment
)
grp_Aquatic_Aquatic_Sediment$data <- grp_Aquatic_Aquatic_Sediment$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE))) |>
  # 24 rows from Vm_2010_2025 campaigns have MEASURED_VALUE == 0 with no LOD/LOQ
  # or censoring flag. True zero Cu in marine sediment is implausible; these are
  # almost certainly "not detected" entries stored as 0 by the Vannmiljø source.
  # Treated as NA pending confirmation of detection limits from the upstream source.
  mutate(MEASURED_VALUE_STANDARD = na_if(MEASURED_VALUE_STANDARD, 0))
grp_Aquatic_Aquatic_Sediment_unit <- grp_Aquatic_Aquatic_Sediment$summary$unit
```

- Aquatic sediment might be expected to vary considerably - sediment is a known sink for copper pollution, and it isn't subject to active regulation as in e.g. biota tissues
- It's a commonly sampled compartment/medium, so we can expect a lot of data points and need to adjust our plots accordingly

<!-- TODO: Stop trying to use patchwork here and just do an HTML graph gallery for now -->

```{r}
#| label: aquatic-sediment-overview
#| fig-width: 14
#| fig-height: 18

unit_label <- paste0("Measured value (", grp_Aquatic_Aquatic_Sediment_unit, ")")

# Overall distribution
p1 <- grp_Aquatic_Aquatic_Sediment$data |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD, colour = MEASURED_UNIT_STANDARD)) +
  geom_density() +
  geom_rug(alpha = 0.1) +
  scale_x_log10() +
  labs(x = unit_label, y = "Density", title = "a) Overall Density") +
  coord_cartesian(clip = "off") +
  theme(
    axis.text.y = element_text(
      hjust = 0,
      margin = margin(r = -2, unit = "lines")
    )
  )

# Relationship between sampling date and value (time flows left to right)
# data covers ~ 1998 - 2023, so 25 years
p2 <- grp_Aquatic_Aquatic_Sediment$data |>
  ggplot(aes(x = SAMPLING_DATE, y = MEASURED_VALUE_STANDARD)) +
  geom_bin2d(bins = 125) +
  scale_fill_viridis_b(option = "plasma") +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x) +
  scale_y_log10() +
  labs(
    x = "Sampling date",
    y = unit_label,
    title = "b) Concentration by Date"
  )

# Per-campaign distributions as 2D bin heatmap (n >= 10), campaigns ordered by median
p3 <- grp_Aquatic_Aquatic_Sediment$data |>
  add_count(CAMPAIGN_NAME_SHORT) |>
  filter(n >= 10) |>
  mutate(
    campaign_name_pretty = case_when(
      str_detect(CAMPAIGN_NAME_SHORT, "Vm_2010_2025") ~ str_remove_all(
        CAMPAIGN_NAME_SHORT,
        "Vm_2010_2025 \\("
      ) |>
        str_remove_all("\\)$") # TODO: Prettier regex.
    ),
    campaign_name_pretty = fct_reorder(
      campaign_name_pretty,
      MEASURED_VALUE_STANDARD,
      median,
      .na_rm = TRUE
    ) |>
      fct_relabel(str_wrap, width = 15)
  ) |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD, y = campaign_name_pretty)) +
  geom_bin2d(bins = 40) +
  scale_x_log10() +
  scale_fill_viridis_b(name = "Count") +
  coord_cartesian(clip = "off") +
  labs(
    x = unit_label,
    title = "c) Distribution by Campaign",
    subtitle = "(n \u2265 10)"
  ) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_text(
      hjust = 0,
      size = rel(0.6),
      margin = margin(r = -4, unit = "lines")
    )
  )

# Spatial pattern: median measured value per hex cell
world_map <- map_data("world")

aas_spatial <- grp_Aquatic_Aquatic_Sediment$data |>
  filter(!is.na(LONGITUDE), !is.na(LATITUDE)) |>
  sf::st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326) |>
  dplyr::mutate(
    x = sf::st_coordinates(geometry)[, 1],
    y = sf::st_coordinates(geometry)[, 2]
  )

p4 <- ggplot() +
  geom_polygon(
    data = world_map,
    aes(x = long, y = lat, group = group),
    fill = "lightgray",
    colour = "white"
  ) +
  stat_summary_hex(
    data = aas_spatial,
    aes(x = x, y = y, z = MEASURED_VALUE_STANDARD),
    fun = "median",
    bins = 60,
    alpha = 0.75
  ) +
  scale_fill_viridis_b(
    name = paste0("(", grp_Aquatic_Aquatic_Sediment_unit, ")"),
    trans = "log10",
    n.breaks = 6,
    option = "rocket"
  ) +
  coord_fixed(
    ratio = 2,
    xlim = c(get_study_area_bbox()[[1]], get_study_area_bbox()[[3]]),
    ylim = c(50, get_study_area_bbox()[[4]])
  ) +
  labs(
    x = "",
    y = "",
    title = "e) Spatial Distribution",
    subtitle = "median concentration"
  ) +
  theme(legend.position = "right")

# Distribution by site type (geographic feature sub-type, faceted by feature)
p5 <- grp_Aquatic_Aquatic_Sediment$data |>
  filter(
    !is.na(SITE_GEOGRAPHIC_FEATURE),
    !is.na(SITE_GEOGRAPHIC_FEATURE_SUB)
  ) |>
  mutate(
    SITE_GEOGRAPHIC_FEATURE_SUB = fct_reorder(
      SITE_GEOGRAPHIC_FEATURE_SUB,
      MEASURED_VALUE_STANDARD,
      median,
      .na_rm = TRUE
    ) |>
      fct_relabel(str_wrap, width = 15)
  ) |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD, y = SITE_GEOGRAPHIC_FEATURE_SUB)) +
  geom_bin2d(bins = 40) +
  scale_x_log10() +
  scale_fill_viridis_b(name = "Count") +
  scale_y_discrete(position = "right") +
  facet_wrap(
    ~SITE_GEOGRAPHIC_FEATURE,
    scales = "free_y",
    ncol = 1,
    space = "free_y",
    labeller = label_wrap_gen(width = 15)
  ) +
  labs(
    x = unit_label,
    y = NULL,
    title = "d) Distribution by Site Type"
  ) +
  theme(axis.text.y = element_text(size = rel(0.6)))

pw <- (p1 | p2) / (p3 | p5) / (p4 | plot_spacer())
pw + plot_layout(heights = c(1, 2, 1))
```

## Aquatic Sediment

Aquatic / Aquatic Sediment. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.



```{r}
#| label: fig-outlier-year-Aquatic_Aquatic_Sediment
#| fig-cap: "Aquatic / Aquatic Sediment: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Aquatic_Sediment$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Aquatic_Aquatic_Sediment$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Aquatic_Aquatic_Sediment_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Aquatic_Aquatic_Sediment
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Aquatic_Sediment$data,
  "SAMPLING_YEAR",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-site-Aquatic_Aquatic_Sediment
#| fig-cap: "Aquatic / Aquatic Sediment: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Aquatic_Sediment$data$SITE_CODE, base = 4)

# TODO: This plot is very slow and probably
plot_outlier_distribution(
  grp_Aquatic_Aquatic_Sediment$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Aquatic_Aquatic_Sediment_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Aquatic_Aquatic_Sediment
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Aquatic_Sediment$data,
  "SITE_CODE",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-campaign-Aquatic_Aquatic_Sediment
#| fig-cap: "Aquatic / Aquatic Sediment: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Aquatic_Aquatic_Sediment$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Aquatic_Aquatic_Sediment$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Aquatic_Aquatic_Sediment_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Aquatic_Aquatic_Sediment
outlier_flag_table(flag_by_facet(
  grp_Aquatic_Aquatic_Sediment$data,
  "CAMPAIGN_NAME_SHORT",
  min_n = 10
))
```

```{r}
#| label: fig-outlier-treatment-Aquatic_Aquatic_Sediment
#| fig-cap: "Aquatic / Aquatic Sediment: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Aquatic_Aquatic_Sediment,
  x_label = paste0("Measured value (", grp_Aquatic_Aquatic_Sediment_unit, ")")
)
```
