---
title: "Outlier Detection"
format:
  html:
    fig-width: 16
    fig-height: 20
  docx:
    fig-width: 16
    fig-height: 20

execute:
  freeze: true   # parked 2026-07-29, see _quarto.yml render list
  error: false                       # prevents errors from stopping the render
---

Some of our datasets include a lot of outliers. For example, let's take our fish data for *Gadus morhua* (excluding any measurements in dry weight for the sake of argument, and using liver measurements because they're the most data-rich).

::: callout-tip
Via JMO: Hi Sam, based on my experience with QC for EEA: For me the first step would be to distinguish between records identifies as outliers vs. suspicious values. "Outliers" can be identified by various algorithms, like you have described. "Suspicious values" (or some other label) are those that you suspect are wrong, based on additional judgement (reasoning, evidence, confirmation, etc). Those can be flagged as such an then excluded from the assessment, in my view. (Along with the reason, documentation etc).\
\
Via SME: [USGS stuff](https://niva365.sharepoint.com/sites/int_Codinganddataanalysis/Shared%20Documents/Forms/AllItems.aspx?viewid=15d63d96%2D2899%2D45aa%2D9f98%2D7599ea00eef1&csf=1&ovuser=a7c4c1dd%2Dd500%2D4316%2D8296%2D3dd9569b49d1%2Csam%2Ewelch%40niva%2Eno&TeamsCID=8bfb9e0f%2D9a26%2D4407%2D9b27%2D96330e615667&OR=Teams%2DHL&CT=1782712848580&clickparams=eyJBcHBOYW1lIjoiVGVhbXMtRGVza3RvcCIsIkFwcFZlcnNpb24iOiI0OS8yNjA1MjkwNjExOCJ9&CID=404222a2%2D604c%2Df000%2Df645%2D84ef1eecfeb6&cidOR=SPO&FolderCTID=0x012000BFC3DE0BD26664418BC8FF1602CD54E0&id=%2Fsites%2Fint%5FCodinganddataanalysis%2FShared%20Documents%2FGeneral%2FHelsel%5FCensoredDataAnalysis%2FStatisticalMethodsin%20WaterResources%5FHelseltm4a3%2Epdf&parent=%2Fsites%2Fint%5FCodinganddataanalysis%2FShared%20Documents%2FGeneral%2FHelsel%5FCensoredDataAnalysis) (but not very useful)
:::

```{r}
#| label: setup
library(tidyverse)
library(targets)
library(patchwork)
library(ggnewscale)
library(knitr)
library(ggtext)
library(diptest)
library(flextable)
library(glue)
pkgload::load_all(quiet = TRUE) # prettify_campaign_name() etc.

# PARKED 2026-07-29. The outlier tar_map factory this notebook read from was
# removed (PLAN.md P0.2) and its objects pruned, so the tar_read() below no
# longer resolves. Rewiring it onto literature_analysis_ready is deferred; the
# prose and the reasoning about outlier flags are still the reference for how
# a group gets scrutinised by hand.
#
# This WAS one branch of the outlier-analysis targets factory (see
# R/fct_outlier_detection.R for the underlying functions, which survive).
# It already contains group-wise RMZ/IQR flags, a dip test, and
# Winsorized value columns computed on the whole Gadus morhua / Liver group,
# restricted to its dominant unit (mg/kg wet weight).
cod_liver <- tar_read(outlier_biota_Fish_Gadus_morhua_Liver)

gadus_morhua <- cod_liver$data |>
  mutate(
    SAMPLING_YEAR = factor(year(SAMPLING_DATE)),
    # Was a case_when() with no .default, so all 28 non-Vannmiljø campaign
    # names became NA. The `.default =` below it was being passed to mutate(),
    # not to case_when(), so it also created a spurious `.default` column.
    # Both fixed by using the shared helper.
    campaign_name_pretty = prettify_campaign_name(CAMPAIGN_NAME_SHORT)
  )

str_wrap_br <- function(string, width) {
  str_wrap(string) |> str_replace_all("//n", "<br>")
}

ggplot_base_size <- 20
theme_set(theme_minimal(base_size = ggplot_base_size, base_family = "Aptos"))

# Reusable plot layers
common_dotsize <- 3
common_binwidth <- 0.01
common_group_height <- 0.5 # set the height of a single boxplot "row"

outlier_boxplot <- geom_boxplot(
  outliers = FALSE,
  whisker.colour = "darkgrey",
  staple.colour = "darkgrey",
  box.colour = "darkgrey"
)

outlier_dots <- geom_dotplot(
  aes(fill = dot_fill),
  dotsize = common_dotsize,
  binwidth = common_binwidth,
  stackdir = "centerwhole",
  binpositions = "all",
  stroke = 0
)

outlier_fill_scale <- list(
  scale_fill_manual(
    values = c(
      "neither" = "black",
      "IQR" = "#0077BB",
      "RMZ" = "#cc6811",
      "both" = "#d51bcf",
      "not tested" = "grey70"
    ),
    labels = c(
      "neither" = "Neither",
      "IQR" = "IQR only",
      "RMZ" = "RMZ only",
      "both" = "Both",
      # matches outlier_min_n in _targets.R
      "not tested" = "n < 10 (not tested)"
    ),
    name = "Outlier flags"
  ),
  guides(fill = guide_legend(override.aes = list(colour = NA, size = 5)))
)

outlier_plot_theme <- theme(
  plot.title = element_markdown(),
  axis.text.y.left = element_markdown(),
  panel.background = element_blank(),
  panel.grid.major = element_line(colour = "grey90"),
  panel.grid.minor = element_line(colour = "grey95"),
  plot.background = element_blank(),
  legend.background = element_blank(),
  legend.key = element_blank(),
  strip.background = element_blank(),
  axis.ticks = element_blank()
)
```

A simple density plot of measured value distributions immediately shows an extremely long right tail, a signal that some measurements are many orders of magnitude higher than the mean. Even when we simply filter out the highest values, the right tail remains extremely long.

```{r}
#| label: fig-distribution
#| fig-width: 16
#| fig-height: 16
#| fig-cap: "Density distributions of copper concentrations in *G. morhua* livers when values above a certain threshold are removed."
no_filter <- gadus_morhua |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD)) +
  geom_density(colour = "#D55E00") +
  labs(
    title = "Values not filtered",
    y = "Density",
    x = "mg/kg Cu (wet weight)"
  )

filter_5000 <- gadus_morhua |>
  filter(MEASURED_VALUE_STANDARD <= 5000) |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD)) +
  geom_density(colour = "#D55E00") +
  labs(
    title = "Values >5000 mg/kg removed",
    y = "Density",
    x = "mg/kg Cu (wet weight)"
  )

filter_1000 <- gadus_morhua |>
  filter(MEASURED_VALUE_STANDARD <= 1000) |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD)) +
  geom_density(colour = "#D55E00") +
  labs(
    title = "Values >1000 mg/kg removed",
    y = "Density",
    x = "mg/kg Cu (wet weight)"
  )

filter_500 <- gadus_morhua |>
  filter(MEASURED_VALUE_STANDARD <= 500) |>
  ggplot(aes(x = MEASURED_VALUE_STANDARD)) +
  geom_density(colour = "#D55E00") +
  labs(
    title = "Values >500 mg/kg removed",
    y = "Density",
    x = "mg/kg Cu (wet weight)"
  )

no_filter /
  filter_5000 /
  filter_1000 /
  filter_500 +
  plot_layout(axes = "keep")
```

Our initial plot (@fig-distribution) peaks around 0 but extends beyond the 15,000 mg/kg wet weight (that is, 15g of copper per kg of wet liver). The most extreme of these values sufficiently improbable that we can remove, and as we do so our plot begins to resemble a log-normal distribution. However, where do we draw the line? Furthermore, we want our workflow to be as reusable and non-specific as possible.

Expert judgement is generally sufficient to make this call, but we cannot draw on this expert judgement for each conceivable case of species, tissue, site type, etc. Vannmiljø data dominate our overall dataset, and this includes both sites monitored specifically because they are uncontaminated (e.g. reference rivers), as well as those monitored because they are known to be highly contaminated (e.g. sites affected by mining).

In general, we would expect natural, "real" outliers to represent one or more of the following scenarios:

- A campaign has selected specific sites known to have high contamination
- An acute contamination event has occurred
- A specific site has highly localised contamination

Conversely, "artifact" outliers may be a product of:

- Analytical methodological defects or contamination
- Mistakes in data entry
- Mistakes in data processing or manipulation

Directly detecting the former is likely to be much easier than the latter. We can start by examining the distributions of concentrations by year, site, and campaign.

# Methods

TODO: Dot-binning method defaults:

- Wilkinson, L. (1999) Dot plots. The American Statistician, 53(3), 276-281.
- via https://ggplot2.tidyverse.org/reference/geom_dotplot.html

## Outlier detection

Detection and labelling of outliers is a important part of quality controlling data, especially as we often lack the hollistic understanding of the many studies making up our dataset to confidently attribute an outlier to real variation or methodlogical error. A number of statistical techniques exist to flag outliers, and the choice of both which is most appropriate, and where to set any thresholds for outlier-ness is to an extent subjective.

As there is no single consensus test for outlier labelling (especially when the number of potential outliers is unknown), we use two common, simple approaches to flag potential outliers in each group.

### Tukey fences (IQR method)

The Tukey fence method (@schwertmanIdentifyingOutliersSequential2007, originally @tukeyExploratoryDataAnalysis1977 but latter not available online) flags values that fall more than $1.5 \times IQR$ beyond the first or third quartile of the log-transformed values, where $IQR = Q_3 - Q_1$. Specifically, a value $x_i$ is flagged if:

$$\log_{10}(x_i) < Q_1 - 1.5 \cdot IQR \quad \text{or} \quad \log_{10}(x_i) > Q_3 + 1.5 \cdot IQR$$

We apply fences to log-transformed values because concentrations are approximately log-normally distributed.

Tukey recommends using 1.5 or 3 times the IQR as thresholds depending on the desired degree of conservatism. We elect to use 1.5, as although this may increase the false positive rate, it is also more sensitive to outliers when the distribution is largely log-normal but with a long right tail.

### Robust Modified Z-Score

The Robust Modified Z-Score (RMZ) is a median-based analogue of the standard Z-score that is more resistant to the influence of extreme values [@iglewiczHowDetectHandle1993; @leysDetectingOutliersNot2013], at the cost of being less appropriate for truly normal distrubtions.

For a value $x_i$ in a group with median $\tilde{x}$, it is defined as:

$$\text{RMZ}_i = \frac{x_i - \tilde{x}}{\text{MAD}}$$

where $\text{MAD} = \text{median}(|x_i - \tilde{x}|)$ is the median absolute deviation. In R, `mad()` already scales MAD by a consistency factor ($\approx 1 / 0.6745$) to make it comparable to the standard deviation under normality

@iglewiczHowDetectHandle1993 recommends a threshold of $|\text{RMZ}| > 3.5$ for outliers, and we have no reason to depart from this recommendation.

```{r}
#| label: robust-modified-z-score
robust_modified_z_score <- function(x) {
  x_median <- median(x, na.rm = TRUE)
  # mad() in R already includes the 1/0.6745 consistency factor
  x_mad <- mad(x, na.rm = TRUE)
  (x - x_median) / x_mad
}
```

In the graphs below, we flag points by which flags they receive: none (black), IQR only (blue), RMZ only (red), or both (purple).

### Test for unimodality

To identify groups whose distribution may be multimodal, which can indicate a mixture of genuinely distinct populations rather than a single skewed distribution, we apply Hartigan's Dip Test via the R package `diptest` [@maechlerDiptestHartigansDip2025]. This provides a statistical adjunct to visual inspection of distributions useful for flagging. Its precise description can be found in @hartiganDipTestUnimodality1985, but in essence it compares the difference between the cumulative density function of a normal distribution and the sample distribution; with a larger "dip" between the former and the latter indicating the "wobbly" cumulative distribution indicative of a multimodal distribution.

The test returns a p-value for the null hypothesis of unimodality; small p-values suggest the distribution is not unimodal. We use $p < 0.05$ as a standard threshold. In the graphs below, p values below the threshold are highlighted in red.

# Results

## Outliers by year

@fig-outliers-year shows the distribution of individual measurements grouped by sampling year, with colours indicating the number of outlier flags received.

```{r}
#| label: outliers-year-flag
gadus_morhua_by_year <- gadus_morhua |>
  group_by(SAMPLING_YEAR) |>
  mutate(
    RMZ = robust_modified_z_score(MEASURED_VALUE_STANDARD),
    log_val = log10(MEASURED_VALUE_STANDARD),
    Q1 = quantile(log_val, 0.25),
    Q3 = quantile(log_val, 0.75),
    IQR = IQR(log_val) * 1.5,
    outlier_RMZ = abs(RMZ) > 3.5,
    outlier_IQR = log_val < (Q1 - IQR) | log_val > (Q3 + IQR),
    dot_fill = factor(case_when(
      outlier_RMZ & outlier_IQR ~ "both",
      outlier_RMZ ~ "RMZ",
      outlier_IQR ~ "IQR",
      .default = "neither",
    )),
    year_and_n = glue("**{SAMPLING_YEAR}**<br>(n = {n()})")
  )

# TODO: Replace with standardised Z-dip with intepretation and avoid rendering stuff where n is too small
dip_results_year <- gadus_morhua_by_year |>
  group_by(SAMPLING_YEAR) |>
  reframe(
    dip_p = dip.test(MEASURED_VALUE_STANDARD)$p.value,
    draw_dip_p = dip.test(MEASURED_VALUE_STANDARD)$nobs > 4
  ) |>
  mutate(
    multimodal = dip_p < 0.05
  )

n_rows_year <- n_distinct(gadus_morhua_by_year$SAMPLING_YEAR) # used for consistent plot heights
```

```{r}
#| label: fig-outliers-year
#| fig-cap: "Measured copper concentration in *G. morhua* livers, published on Vannmiljø, by reported sampling year. Outlying points flagged by Tukey fences (IQR × 1.5) and Robust Modified Z-Score (threshold 3.5)."
#| fig-height: !expr (n_rows_year * common_group_height + 3)
main_plot <- gadus_morhua_by_year |>
  ggplot(aes(
    y = year_and_n,
    x = MEASURED_VALUE_STANDARD
  )) +
  outlier_boxplot +
  outlier_dots +
  outlier_fill_scale +
  scale_x_log10(
    breaks = 10^(-1:5),
    labels = scales::label_log(),
    minor_breaks = outer(1:9, 10^(-1:4)) |> as.vector()
  ) +
  labs(
    y = "Year",
    x = "Measured Value (mg/kg (wet))"
  ) +
  outlier_plot_theme

dip_panel <- dip_results_year |>
  ggplot(aes(
    y = SAMPLING_YEAR,
    x = 0,
    # draw only if n > 4 (functional min for dip test)
    label = ifelse(draw_dip_p, format(round(dip_p, 3), nsmall = 3), ""),
    colour = multimodal
  )) +
  geom_text(size = ggplot_base_size / .pt) +
  scale_colour_manual(values = c("FALSE" = "grey40", "TRUE" = "#CC3311")) +
  labs(x = "p(not unimodal)", y = NULL) +
  theme_void(base_size = ggplot_base_size, base_family = "Aptos") +
  theme(
    legend.position = "none",
    axis.title.x = element_text()
  )

main_plot + dip_panel + plot_layout(widths = c(6, 1), guides = "collect")
```

@tbl-outliers-year-comments lists the flagged measurements alongside their associated campaign and sample metadata.

```{r}
#| label: tbl-outliers-year-comments
#| tbl-cap: "Flagged outlier measurements in *G. morhua* liver copper concentrations, grouped by sampling year, with associated campaign and sample metadata."
gadus_morhua_by_year |>
  filter(dot_fill != "neither") |>
  group_by(MEASUREMENT_COMMENT, campaign_name_pretty) |>
  reframe(
    n = n(),
    mean = mean(MEASURED_VALUE_STANDARD)
  ) |>
  arrange(desc(mean)) |>
  select(campaign_name_pretty, mean, n, MEASUREMENT_COMMENT) |>
  flextable() |>
  autofit() |>
  set_table_properties(width = 1, layout = "autofit") |>
  colformat_double(j = c("mean")) |>
  set_header_labels(
    campaign_name_pretty = "Campaign",
    n = "n",
    mean = "Mean (mg/kg wet weight, 2 d.p)",
    MEASUREMENT_COMMENT = "Sample Comment"
  )
```

## Outliers by site

@fig-outliers-site shows the same measurements grouped by monitoring site, faceted by geographic feature type.

```{r}
#| label: outliers-site-flag
gadus_morhua_by_site <- gadus_morhua |>
  mutate(SITE_CODE = str_remove(SITE_CODE, "Vannmiljø_")) |>
  group_by(SITE_CODE) |>
  mutate(
    RMZ = robust_modified_z_score(MEASURED_VALUE_STANDARD),
    log_val = log10(MEASURED_VALUE_STANDARD),
    Q1 = quantile(log_val, 0.25),
    Q3 = quantile(log_val, 0.75),
    IQR = IQR(log_val) * 1.5,
    outlier_RMZ = abs(RMZ) > 3.5,
    outlier_IQR = log_val < (Q1 - IQR) | log_val > (Q3 + IQR),
    dot_fill = factor(case_when(
      outlier_RMZ & outlier_IQR ~ "both",
      outlier_RMZ ~ "RMZ",
      outlier_IQR ~ "IQR",
      .default = "neither"
    )),
    site_and_n = glue("**{SITE_CODE}**  (n = {n()})")
  )

# TODO: Warning here. What does it mean?
dip_results_site <- gadus_morhua_by_site |>
  group_by(SITE_CODE, SITE_GEOGRAPHIC_FEATURE) |>
  reframe(
    dip_p = dip.test(MEASURED_VALUE_STANDARD)$p.value,
    draw_dip_p = dip.test(MEASURED_VALUE_STANDARD)$nobs > 4
  ) |>
  mutate(multimodal = dip_p < 0.05)

n_rows_site <- n_distinct(gadus_morhua_by_site$SITE_CODE)
```

```{r}
#| label: fig-outliers-site
#| fig-cap: "Measured copper concentration in *G. morhua* livers, published on Vannmiljø, by reported site code. Outlying points flagged by Tukey fences (IQR × 1.5) and Robust Modified Z-Score (threshold 3.5)."
#| fig-height: !expr (n_rows_site * common_group_height + 4)
# sites height is biggger, to account for the gab between the two facets
main_plot_site <- gadus_morhua_by_site |>
  ggplot(aes(
    y = site_and_n,
    x = MEASURED_VALUE_STANDARD
  )) +
  outlier_boxplot +
  outlier_dots +
  outlier_fill_scale +
  scale_x_log10(
    breaks = 10^(-1:5),
    labels = scales::label_log(),
    minor_breaks = outer(1:9, 10^(-1:4)) |> as.vector()
  ) +
  facet_wrap(
    facets = vars(SITE_GEOGRAPHIC_FEATURE),
    ncol = 1,
    scales = "free_y",
    space = "free_y"
  ) +
  labs(
    y = "Site Code",
    x = "Measured Value (mg/kg (wet))"
  ) +
  outlier_plot_theme

# TODO: Tweak height of text elements, as they become misaligned over taller graphs
dip_panel_site <- dip_results_site |>
  ggplot(aes(
    y = SITE_CODE,
    x = 0,
    label = ifelse(draw_dip_p, format(round(dip_p, 3), nsmall = 3), ""),
    colour = multimodal
  )) +
  geom_text(size = ggplot_base_size / .pt) +
  facet_wrap(
    facets = vars(SITE_GEOGRAPHIC_FEATURE),
    ncol = 1,
    scales = "free_y",
    space = "free_y"
  ) +
  scale_colour_manual(values = c("FALSE" = "grey40", "TRUE" = "#CC3311")) +
  labs(x = "p(not unimodal)", y = NULL) +
  theme_void(base_size = ggplot_base_size, base_family = "Aptos") +
  theme(
    legend.position = "none",
    strip.text = element_blank()
  )

main_plot_site +
  dip_panel_site +
  plot_layout(widths = c(6, 1), guides = "collect")
```

@tbl-outliers-site-comments summarises the flagged measurements by site.

```{r}
#| label: tbl-outliers-site-comments
#| tbl-cap: "Flagged outlier measurements in *G. morhua* liver copper concentrations, grouped by monitoring site, with associated campaign and sample metadata."
gadus_morhua_by_site |>
  filter(dot_fill != "neither") |>
  group_by(MEASUREMENT_COMMENT, campaign_name_pretty) |>
  reframe(n = n(), mean = mean(MEASURED_VALUE_STANDARD)) |>
  arrange(desc(mean)) |>
  select(campaign_name_pretty, mean, n, MEASUREMENT_COMMENT) |>
  flextable() |>
  autofit() |>
  set_table_properties(width = 1, layout = "autofit") |>
  colformat_double(j = c("mean")) |>
  set_header_labels(
    campaign_name_pretty = "Campaign",
    n = "n",
    mean = "Mean (mg/kg wet weight, 2 d.p)",
    MEASUREMENT_COMMENT = "Sample Comment"
  )
```

## Outliers by campaign

@fig-outliers-campaign shows distributions grouped by monitoring campaign.

```{r}
#| label: outliers-campaign-flag
gadus_morhua_by_campaign <- gadus_morhua |>
  mutate(
    CAMPAIGN_NAME = str_remove(
      CAMPAIGN_NAME,
      "Vannmiljø Copper Monitoring 2010-2025 \\("
    ) |>
      str_remove("\\)$")
  ) |>
  group_by(CAMPAIGN_NAME) |>
  mutate(
    RMZ = robust_modified_z_score(MEASURED_VALUE_STANDARD),
    log_val = log10(MEASURED_VALUE_STANDARD),
    Q1 = quantile(log_val, 0.25),
    Q3 = quantile(log_val, 0.75),
    IQR = IQR(log_val) * 1.5,
    outlier_RMZ = abs(RMZ) > 3.5,
    outlier_IQR = log_val < (Q1 - IQR) | log_val > (Q3 + IQR),
    dot_fill = factor(case_when(
      outlier_RMZ & outlier_IQR ~ "both",
      outlier_RMZ ~ "RMZ",
      outlier_IQR ~ "IQR",
      .default = "neither"
    )),
    campaign_and_n = glue(
      "**{str_wrap_br(CAMPAIGN_NAME, width = 1)}**<br>(n = {n()})"
    ) |>
      as.character()
  )

# More errors here - sample size too small, I think
dip_results_campaign <- gadus_morhua_by_campaign |>
  group_by(CAMPAIGN_NAME, campaign_and_n) |>
  reframe(
    dip_p = dip.test(MEASURED_VALUE_STANDARD)$p.value,
    draw_dip_p = dip.test(MEASURED_VALUE_STANDARD)$nobs > 4
  ) |>
  mutate(multimodal = dip_p < 0.05)

n_rows_campaign <- n_distinct(gadus_morhua_by_campaign$CAMPAIGN_NAME)
```

```{r}
#| label: fig-outliers-campaign
#| fig-cap: "Measured copper concentration in *G. morhua* livers, published on Vannmiljø, by monitoring campaign. Outlying points flagged by Tukey fences (IQR × 1.5) and Robust Modified Z-Score (threshold 3.5)."
#| fig-height: !expr (n_rows_campaign * common_group_height + 3)
main_plot_campaign <- gadus_morhua_by_campaign |>
  ggplot(aes(
    y = campaign_and_n,
    x = MEASURED_VALUE_STANDARD
  )) +
  outlier_boxplot +
  outlier_dots +
  outlier_fill_scale +
  scale_x_log10(
    breaks = 10^(-1:5),
    labels = scales::label_log(),
    minor_breaks = outer(1:9, 10^(-1:4)) |> as.vector()
  ) +
  labs(
    y = "Campaign",
    x = "Measured Value (mg/kg (wet))"
  ) +
  outlier_plot_theme

dip_panel_campaign <- dip_results_campaign |>
  ggplot(aes(
    y = campaign_and_n,
    x = 0,
    label = ifelse(draw_dip_p, format(round(dip_p, 3), nsmall = 3), ""),
    colour = multimodal
  )) +
  geom_text(size = ggplot_base_size / .pt) +
  scale_colour_manual(values = c("FALSE" = "grey40", "TRUE" = "#CC3311")) +
  labs(x = "p(not unimodal)", y = NULL) +
  theme_void(base_size = ggplot_base_size, base_family = "Aptos") +
  theme(
    legend.position = "none",
    axis.title.x = element_text()
  )

main_plot_campaign +
  dip_panel_campaign +
  plot_layout(widths = c(6, 1), guides = "collect")
```

@tbl-outliers-campaign-comments summarises the flagged measurements.

```{r}
#| label: tbl-outliers-campaign-comments
#| tbl-cap: "Flagged outlier measurements in *G. morhua* liver copper concentrations, grouped by monitoring campaign, with associated campaign and sample metadata."
gadus_morhua_by_campaign |>
  filter(dot_fill != "neither") |>
  group_by(MEASUREMENT_COMMENT, campaign_name_pretty) |>
  reframe(n = n(), mean = mean(MEASURED_VALUE_STANDARD)) |>
  arrange(desc(mean)) |>
  select(campaign_name_pretty, mean, n, MEASUREMENT_COMMENT) |>
  flextable() |>
  autofit() |>
  # TODO: Needed as table overflows due to the Polluted Seabed row. (note that you need this aaand the filters at the top)
  set_table_properties(width = 1, layout = "autofit") |>
  colformat_double(j = c("mean")) |>
  set_header_labels(
    campaign_name_pretty = "Campaign",
    n = "n",
    mean = "Mean (mg/kg wet weight, 2 d.p)",
    MEASUREMENT_COMMENT = "Sample Comment"
  )
```

## Handling Outliers

Winsorization - reducing extreme values to a given percentile or trimming - eliminating them entirely - are the most common approaches for dealing with outliers. High quality data treatment remains, as above, dependent on a thorough understanding of the causes of outliers, and is not practically scalable given the quanaity and provenance of the data in this analysis.

Under the circumstances, we elected to Winsorize all double-flagged outliers to 98%, that is, reduce any values beyond the 1st and 99th-percentiles to these percentile values. Although this approach is somewhat crude, it is appropriate in this test case where large outliers are most likely a consequence of errors in data entry. In addition, robust summary statistics (median) will be used in addition to the less robust mean for interpretation of data.

Note that other datasets (other organisms, matrices, etc.) will also be inspected for outliers, and should available data suggest a different explanation for outliers in such cases, a variant strategy may be applied.

For review

Note: I haven't actually Winsorized the data yet. From what I've read it's not a great approach compared with more robust statistics, but I want to hurry up and get this done so I can go back to actually assembling the AEPs. I propose the following:

- In this case, I feel comfortable stating that we could both trim or Windsorise our massive outliers without affecting the quality of our analysis, because they're almost certainly transcription errors
- It doesn't look like there's any real difference between year/campaign/site groups. It's probably safe to merge them, but it would be good to have a test to back this up. Welch's ANOVA? We may need to Winsorise first
- Once we have the single distribution for copper in fish livers, make sure we report it as a distribution (with mean, median, etc.) in the AEPs (maybe we can even show the whole distribution, Bayesian network-style) rather than just a mean
- Repeat this exercise for all the most important compartments/tissues/species

```{r}
#| fig-cap: "Measured copper concentration in *G. morhua* livers, by sampling site feature. Points beyond the 1st and 99th percentiles per feature have been Windsorized to these values. Boxplots (Quartiles), and mean ± standard deviation (turquoise) also plotted."
#| label: fig-winsorization-and-grouping
#| fig-height: !expr (3 * common_group_height + 8)
gadus_morhua |>
  mutate(
    p01 = quantile(MEASURED_VALUE_STANDARD, 0.01),
    p99 = quantile(MEASURED_VALUE_STANDARD, 0.99),
    MEASURED_VALUE_STANDARD_windsor = case_when(
      MEASURED_VALUE_STANDARD < p01 ~ p01,
      MEASURED_VALUE_STANDARD > p99 ~ p99,
      .default = MEASURED_VALUE_STANDARD
    ),
    dot_fill = factor(case_when(
      MEASURED_VALUE_STANDARD < p01 |
        MEASURED_VALUE_STANDARD > p99 ~ "Winsorized",
      .default = "Original"
    )),
    geo_and_n = glue(
      "**{SITE_GEOGRAPHIC_FEATURE}**",
      "<br>n = {n()}<br>{n_distinct(CAMPAIGN_NAME_SHORT)} campaigns<br>{n_distinct(REFERENCE_ID)} refs",
      "<br>{format(min(SAMPLING_DATE), '%Y')}\u2013{format(max(SAMPLING_DATE), '%Y')}",
      "<br>{str_flatten_comma(unique(COUNTRY_ISO))}",
      "<br>{str_flatten_comma(unique(OCEAN_IHO))}",
      .sep = ""
    ),
    .by = "SITE_GEOGRAPHIC_FEATURE"
  ) |>
  ggplot(aes(
    x = MEASURED_VALUE_STANDARD_windsor,
    y = geo_and_n
  )) +
  outlier_boxplot +
  geom_dotplot(
    aes(fill = dot_fill),
    dotsize = common_dotsize / 2,
    binwidth = common_binwidth,
    stackdir = "centerwhole",
    binpositions = "all",
    stroke = 0
  ) +
  stat_summary(
    fun = mean,
    fun.min = function(x) mean(x) - sd(x),
    fun.max = function(x) mean(x) + sd(x),
    geom = "pointrange",
    colour = "#1baf59",
    size = 0.8,
    linewidth = 1
  ) +
  scale_fill_manual(
    values = c("Original" = "black", "Winsorized" = "#0072B2"),
    name = "Point status"
  ) +
  guides(fill = guide_legend(override.aes = list(colour = NA, size = 5))) +
  scale_x_log10(
    breaks = 10^(-1:5),
    labels = scales::label_log(),
    minor_breaks = outer(1:9, 10^(-1:4)) |> as.vector()
  ) +
  labs(
    x = "Cu concentration (mg/kg wet weight)",
    y = "Geographic feature type"
  ) +
  outlier_plot_theme
```

And as a table:

```{r}
#| label: tbl-winsorization-summary
#| tbl-cap: "Summary statistics for Cu concentrations in *G. morhua* livers, grouped by geographic feature type, prior to Winsorisation."
gadus_morhua |>
  reframe(
    sample_n_sum = sum(MEASURED_N),
    median = median(MEASURED_VALUE_STANDARD),
    mean = mean(MEASURED_VALUE_STANDARD),
    unit = unique(MEASURED_UNIT_STANDARD),
    sd = sd(MEASURED_VALUE_STANDARD),
    date_range = str_flatten(range(year(SAMPLING_DATE)), collapse = " \u2013 "),
    distribution_country = str_flatten_comma(unique(COUNTRY_ISO)),
    distribution_ocean = str_flatten_comma(unique(OCEAN_IHO)),
    sources = str_flatten_comma(unique(campaign_name_pretty)),
    .by = "SITE_GEOGRAPHIC_FEATURE"
  ) |>
  flextable() |>
  colformat_double(j = c("median", "mean", "sd"), digits = 2) |>
  autofit() |>
  set_table_properties(width = 1, layout = "autofit") |>
  set_header_labels(
    SITE_GEOGRAPHIC_FEATURE = "Geographic Feature",
    sample_n_sum = "n",
    median = "Median",
    mean = "Mean",
    sd = "SD",
    date_range = "Date Range",
    distribution_country = "Countries",
    distribution_ocean = "Ocean Regions",
    unit = "Unit",
    sources = "Campaigns"
  )
```

Going forwards, I think it'll make more sense to concentration distributions as violin plots, rather than dotplots, or else the sheer sample size of cod livers will drown things that come as smaller samples.

Although this will also take some graphical tweaking.

```{r}
#| fig-cap: "Measured copper concentration in *G. morhua* livers, by sampling site feature. Points beyond the 1st and 99th percentiles per feature have been Windsorized to these values. Boxplots (Quartiles), and mean ± standard deviation (turquoise) also plotted."
#| label: fig-winsorization-and-grouping-violin
#| fig-height: !expr (3 * common_group_height + 8)
gadus_morhua |>
  mutate(
    p01 = quantile(MEASURED_VALUE_STANDARD, 0.01),
    p99 = quantile(MEASURED_VALUE_STANDARD, 0.99),
    MEASURED_VALUE_STANDARD_windsor = case_when(
      MEASURED_VALUE_STANDARD < p01 ~ p01,
      MEASURED_VALUE_STANDARD > p99 ~ p99,
      .default = MEASURED_VALUE_STANDARD
    ),
    dot_fill = factor(case_when(
      MEASURED_VALUE_STANDARD < p01 |
        MEASURED_VALUE_STANDARD > p99 ~ "Winsorized",
      .default = "Original"
    )),
    geo_and_n = glue(
      "**{SITE_GEOGRAPHIC_FEATURE}**",
      "<br>n = {n()}<br>{n_distinct(CAMPAIGN_NAME_SHORT)} campaigns<br>{n_distinct(REFERENCE_ID)} refs",
      "<br>{format(min(SAMPLING_DATE), '%Y')}\u2013{format(max(SAMPLING_DATE), '%Y')}",
      "<br>{str_flatten_comma(unique(COUNTRY_ISO))}",
      "<br>{str_flatten_comma(unique(OCEAN_IHO))}",
      .sep = ""
    ),
    .by = "SITE_GEOGRAPHIC_FEATURE"
  ) |>
  ggplot(aes(
    x = MEASURED_VALUE_STANDARD_windsor,
    y = geo_and_n
  )) +
  geom_violin(
    scale = "count",
    draw_quantiles = c(0.25, 0.5, 0.75),
    aes(fill = SITE_GEOGRAPHIC_FEATURE),
    colour = NA,
    width = 2,
    alpha = 0.75
  ) +
  scale_fill_brewer(palette = "BuGn") +
  # TODO: Set alpha based for lines only
  geom_boxplot(fill = NA, width = 0.05, alpha = 0.75) +
  guides(fill = guide_legend(override.aes = list(colour = NA, size = 5))) +
  scale_x_log10(
    breaks = 10^(-1:5),
    labels = scales::label_log(),
    minor_breaks = outer(1:9, 10^(-1:4)) |> as.vector()
  ) +
  labs(
    x = "Cu concentration (mg/kg wet weight)",
    y = "Geographic feature type"
  ) +
  outlier_plot_theme +
  theme(legend.position = "none")
# TODO: Add a map?
```

# TODO: How much does the Windzorization actually change the summary stats - compared to Trimming?

Looking at the whole population, the conclusions are:

- Unprocessed is fine in terms of the main body of the distribution, but ends up with unacceptably long tails
- Winsorized 90% is clearly ridiculous
- 98% looks fine
- Trimming double-flagged is probably also fine (TODO: What about single-flagged?)

```{r}
# Global flags and Winsorization bounds are already computed on the whole
# group by the outlier-analysis factory (outlier_group_analysis()) -- see
# `cod_liver`, read at the top of this document. We just relabel here for
# the comparison plot below.
base <- gadus_morhua |>
  mutate(double_flagged = dot_fill == "both")

n_double <- sum(base$double_flagged)
n_total <- nrow(base)

# Assemble long-format comparison dataset
plot_data <- bind_rows(
  base |> transmute(value = MEASURED_VALUE_STANDARD, treatment = "Unprocessed"),
  base |> transmute(value = value_winsor_98, treatment = "Winsorized (98%)"),
  base |> transmute(value = value_winsor_90, treatment = "Winsorized (90%)"),
  base |>
    filter(!double_flagged) |>
    transmute(
      value = MEASURED_VALUE_STANDARD,
      treatment = glue(
        "Trimmed (double-flagged,\nn = {n_double} removed)"
      ) |>
        as.character()
    )
) |>
  mutate(
    treatment = factor(
      treatment,
      levels = c(
        "Unprocessed",
        "Winsorized (98%)",
        "Winsorized (90%)",
        # TODO: Fix this, add element_markdown below
        glue("Trimmed (double-flagged,\nn = {n_double} removed)") |>
          as.character()
      )
    )
  )

plot_data |>
  # TODO: Unprocessed at the top, add numerical values for summary stats - how robust are they?
  # Better colour scheme?
  ggplot(aes(x = value, y = treatment, fill = treatment)) +
  geom_violin(
    scale = "width",
    colour = NA,
    alpha = 0.65,
    width = 0.8
  ) +
  geom_boxplot(
    fill = NA,
    width = 0.08,
    outliers = FALSE,
    whisker.colour = "grey40",
    staple.colour = "grey40",
    box.colour = "grey40"
  ) +
  scale_x_log10(
    breaks = 10^(-1:5),
    labels = scales::label_log(),
    minor_breaks = outer(1:9, 10^(-1:4)) |> as.vector()
  ) +
  scale_fill_brewer(palette = "BuGn") +
  labs(
    x = "Cu concentration (mg/kg wet weight)",
    y = NULL,
    title = "Effect of outlier treatment on Cu distribution in *G. morhua* liver"
  ) +
  outlier_plot_theme +
  theme(
    legend.position = "none",
    plot.title = element_markdown()
  )
```

# References