---
title: "Distributions: Molluscs"
format:
  html:
    fig-width: 16
    fig-height: 20
  docx:
    fig-width: 16
    fig-height: 20

execute:
  echo: false

knitr:
  opts_chunk:
    dpi: 300
    dev: "ragg_png"

filters:
  - flextable-qmd
  - at: post-render
    path: ../_extensions/flextable-qmd/unwrap-float.lua
---

::: callout-note
This notebook was scaffolded by `scripts/generate_distribution_notebooks.R`
from the outlier-analysis targets factory (`_targets.R`,
`R/fct_outlier_groups.R`, `R/fct_outlier_detection.R`,
`R/fct_outlier_report.R`). The plots/tables below are generated, but
analysis and interpretation are added by hand, section by section.
Re-running the generator only appends sections for groups not yet present
here -- it never rewrites or removes anything you have written.

See [NBXX-Outliers.qmd](NBXX-Outliers.qmd) for the fully hand-authored
*G. morhua* liver case study this workflow was built from, including
year/site/campaign breakdowns and the reasoning behind the outlier flags.
:::

```{r}
#| label: setup
library(tidyverse)
library(targets)
library(patchwork)
library(flextable)
pkgload::load_all(quiet = TRUE)

set_flextable_defaults(
  font.size = 10,
  font.family = "Arial",
  digits = 2,
  theme_fun = theme_vanilla,
  padding = 1
)

ggplot_base_size <- 20
theme_set(theme_minimal(base_size = ggplot_base_size, base_family = "Aptos"))
```

## Acesta excavata: Total soft tissues

Molluscs / Acesta excavata / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Acesta_excavata_Total_soft_tissues
grp_Molluscs_Acesta_excavata_Total_soft_tissues <- tar_read(outlier_biota_Molluscs_Acesta_excavata_Total_soft_tissues)
grp_Molluscs_Acesta_excavata_Total_soft_tissues$data <- grp_Molluscs_Acesta_excavata_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Acesta_excavata_Total_soft_tissues_unit <- grp_Molluscs_Acesta_excavata_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Acesta_excavata_Total_soft_tissues
#| fig-cap: "Molluscs / Acesta excavata / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Acesta_excavata_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Acesta_excavata_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Acesta_excavata_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Acesta_excavata_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Acesta_excavata_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Acesta_excavata_Total_soft_tissues
#| fig-cap: "Molluscs / Acesta excavata / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Acesta_excavata_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Acesta_excavata_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Acesta_excavata_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Acesta_excavata_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Acesta_excavata_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Acesta_excavata_Total_soft_tissues
#| fig-cap: "Molluscs / Acesta excavata / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Acesta_excavata_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Acesta_excavata_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Acesta_excavata_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Acesta_excavata_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Acesta_excavata_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Acesta_excavata_Total_soft_tissues
#| fig-cap: "Molluscs / Acesta excavata / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Acesta_excavata_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Molluscs_Acesta_excavata_Total_soft_tissues_unit, ")")
)
```

## Astarte borealis: Total soft tissues

Molluscs / Astarte borealis / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Astarte_borealis_Total_soft_tissues
grp_Molluscs_Astarte_borealis_Total_soft_tissues <- tar_read(outlier_biota_Molluscs_Astarte_borealis_Total_soft_tissues)
grp_Molluscs_Astarte_borealis_Total_soft_tissues$data <- grp_Molluscs_Astarte_borealis_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Astarte_borealis_Total_soft_tissues_unit <- grp_Molluscs_Astarte_borealis_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Astarte_borealis_Total_soft_tissues
#| fig-cap: "Molluscs / Astarte borealis / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Astarte_borealis_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Astarte_borealis_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Astarte_borealis_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Astarte_borealis_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Astarte_borealis_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Astarte_borealis_Total_soft_tissues
#| fig-cap: "Molluscs / Astarte borealis / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Astarte_borealis_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Astarte_borealis_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Astarte_borealis_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Astarte_borealis_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Astarte_borealis_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Astarte_borealis_Total_soft_tissues
#| fig-cap: "Molluscs / Astarte borealis / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Astarte_borealis_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Astarte_borealis_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Astarte_borealis_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Astarte_borealis_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Astarte_borealis_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Astarte_borealis_Total_soft_tissues
#| fig-cap: "Molluscs / Astarte borealis / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Astarte_borealis_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Molluscs_Astarte_borealis_Total_soft_tissues_unit, ")")
)
```

## Buccinum undatum: Whole body

Molluscs / Buccinum undatum / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Buccinum_undatum_Whole_body
grp_Molluscs_Buccinum_undatum_Whole_body <- tar_read(outlier_biota_Molluscs_Buccinum_undatum_Whole_body)
grp_Molluscs_Buccinum_undatum_Whole_body$data <- grp_Molluscs_Buccinum_undatum_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Buccinum_undatum_Whole_body_unit <- grp_Molluscs_Buccinum_undatum_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Buccinum_undatum_Whole_body
#| fig-cap: "Molluscs / Buccinum undatum / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Buccinum_undatum_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Buccinum_undatum_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Buccinum_undatum_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Buccinum_undatum_Whole_body
outlier_flag_table(flag_by_facet(grp_Molluscs_Buccinum_undatum_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Buccinum_undatum_Whole_body
#| fig-cap: "Molluscs / Buccinum undatum / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Buccinum_undatum_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Buccinum_undatum_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Buccinum_undatum_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Buccinum_undatum_Whole_body
outlier_flag_table(flag_by_facet(grp_Molluscs_Buccinum_undatum_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Buccinum_undatum_Whole_body
#| fig-cap: "Molluscs / Buccinum undatum / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Buccinum_undatum_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Buccinum_undatum_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Buccinum_undatum_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Buccinum_undatum_Whole_body
outlier_flag_table(flag_by_facet(grp_Molluscs_Buccinum_undatum_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Buccinum_undatum_Whole_body
#| fig-cap: "Molluscs / Buccinum undatum / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Buccinum_undatum_Whole_body,
  x_label = paste0("Measured value (", grp_Molluscs_Buccinum_undatum_Whole_body_unit, ")")
)
```

## Erginus: Total soft tissues

Molluscs / Erginus / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Erginus_Total_soft_tissues
grp_Molluscs_Erginus_Total_soft_tissues <- tar_read(outlier_biota_Molluscs_Erginus_Total_soft_tissues)
grp_Molluscs_Erginus_Total_soft_tissues$data <- grp_Molluscs_Erginus_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Erginus_Total_soft_tissues_unit <- grp_Molluscs_Erginus_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Erginus_Total_soft_tissues
#| fig-cap: "Molluscs / Erginus / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Erginus_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Erginus_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Erginus_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Erginus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Erginus_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Erginus_Total_soft_tissues
#| fig-cap: "Molluscs / Erginus / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Erginus_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Erginus_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Erginus_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Erginus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Erginus_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Erginus_Total_soft_tissues
#| fig-cap: "Molluscs / Erginus / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Erginus_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Erginus_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Erginus_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Erginus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Erginus_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Erginus_Total_soft_tissues
#| fig-cap: "Molluscs / Erginus / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Erginus_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Molluscs_Erginus_Total_soft_tissues_unit, ")")
)
```

## Gastropoda: Whole body

Molluscs / Gastropoda / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Gastropoda_Whole_body
grp_Molluscs_Gastropoda_Whole_body <- tar_read(outlier_biota_Molluscs_Gastropoda_Whole_body)
grp_Molluscs_Gastropoda_Whole_body$data <- grp_Molluscs_Gastropoda_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Gastropoda_Whole_body_unit <- grp_Molluscs_Gastropoda_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Gastropoda_Whole_body
#| fig-cap: "Molluscs / Gastropoda / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Gastropoda_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Gastropoda_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Gastropoda_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Gastropoda_Whole_body
outlier_flag_table(flag_by_facet(grp_Molluscs_Gastropoda_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Gastropoda_Whole_body
#| fig-cap: "Molluscs / Gastropoda / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Gastropoda_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Gastropoda_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Gastropoda_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Gastropoda_Whole_body
outlier_flag_table(flag_by_facet(grp_Molluscs_Gastropoda_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Gastropoda_Whole_body
#| fig-cap: "Molluscs / Gastropoda / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Gastropoda_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Gastropoda_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Gastropoda_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Gastropoda_Whole_body
outlier_flag_table(flag_by_facet(grp_Molluscs_Gastropoda_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Gastropoda_Whole_body
#| fig-cap: "Molluscs / Gastropoda / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Gastropoda_Whole_body,
  x_label = paste0("Measured value (", grp_Molluscs_Gastropoda_Whole_body_unit, ")")
)
```

## Lepetella: Total soft tissues

Molluscs / Lepetella / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Lepetella_Total_soft_tissues
grp_Molluscs_Lepetella_Total_soft_tissues <- tar_read(outlier_biota_Molluscs_Lepetella_Total_soft_tissues)
grp_Molluscs_Lepetella_Total_soft_tissues$data <- grp_Molluscs_Lepetella_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Lepetella_Total_soft_tissues_unit <- grp_Molluscs_Lepetella_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Lepetella_Total_soft_tissues
#| fig-cap: "Molluscs / Lepetella / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Lepetella_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Lepetella_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Lepetella_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Lepetella_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Lepetella_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Lepetella_Total_soft_tissues
#| fig-cap: "Molluscs / Lepetella / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Lepetella_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Lepetella_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Lepetella_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Lepetella_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Lepetella_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Lepetella_Total_soft_tissues
#| fig-cap: "Molluscs / Lepetella / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Lepetella_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Lepetella_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Lepetella_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Lepetella_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Lepetella_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Lepetella_Total_soft_tissues
#| fig-cap: "Molluscs / Lepetella / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Lepetella_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Molluscs_Lepetella_Total_soft_tissues_unit, ")")
)
```

## Littorina: Total soft tissues

Molluscs / Littorina / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Littorina_Total_soft_tissues
grp_Molluscs_Littorina_Total_soft_tissues <- tar_read(outlier_biota_Molluscs_Littorina_Total_soft_tissues)
grp_Molluscs_Littorina_Total_soft_tissues$data <- grp_Molluscs_Littorina_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Littorina_Total_soft_tissues_unit <- grp_Molluscs_Littorina_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Littorina_Total_soft_tissues
#| fig-cap: "Molluscs / Littorina / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Littorina_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Littorina_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Littorina_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Littorina_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Littorina_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Littorina_Total_soft_tissues
#| fig-cap: "Molluscs / Littorina / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Littorina_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Littorina_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Littorina_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Littorina_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Littorina_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Littorina_Total_soft_tissues
#| fig-cap: "Molluscs / Littorina / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Littorina_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Littorina_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Littorina_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Littorina_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Littorina_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Littorina_Total_soft_tissues
#| fig-cap: "Molluscs / Littorina / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Littorina_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Molluscs_Littorina_Total_soft_tissues_unit, ")")
)
```

## Littorina littorea: Total soft tissues

Molluscs / Littorina littorea / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Littorina_littorea_Total_soft_tissues
grp_Molluscs_Littorina_littorea_Total_soft_tissues <- tar_read(outlier_biota_Molluscs_Littorina_littorea_Total_soft_tissues)
grp_Molluscs_Littorina_littorea_Total_soft_tissues$data <- grp_Molluscs_Littorina_littorea_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Littorina_littorea_Total_soft_tissues_unit <- grp_Molluscs_Littorina_littorea_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Littorina_littorea_Total_soft_tissues
#| fig-cap: "Molluscs / Littorina littorea / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Littorina_littorea_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Littorina_littorea_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Littorina_littorea_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Littorina_littorea_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Littorina_littorea_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Littorina_littorea_Total_soft_tissues
#| fig-cap: "Molluscs / Littorina littorea / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Littorina_littorea_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Littorina_littorea_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Littorina_littorea_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Littorina_littorea_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Littorina_littorea_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Littorina_littorea_Total_soft_tissues
#| fig-cap: "Molluscs / Littorina littorea / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Littorina_littorea_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Littorina_littorea_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Littorina_littorea_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Littorina_littorea_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Littorina_littorea_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Littorina_littorea_Total_soft_tissues
#| fig-cap: "Molluscs / Littorina littorea / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Littorina_littorea_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Molluscs_Littorina_littorea_Total_soft_tissues_unit, ")")
)
```

## Littorina rudis: Total soft tissues

Molluscs / Littorina rudis / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Littorina_rudis_Total_soft_tissues
grp_Molluscs_Littorina_rudis_Total_soft_tissues <- tar_read(outlier_biota_Molluscs_Littorina_rudis_Total_soft_tissues)
grp_Molluscs_Littorina_rudis_Total_soft_tissues$data <- grp_Molluscs_Littorina_rudis_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Littorina_rudis_Total_soft_tissues_unit <- grp_Molluscs_Littorina_rudis_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Littorina_rudis_Total_soft_tissues
#| fig-cap: "Molluscs / Littorina rudis / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Littorina_rudis_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Littorina_rudis_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Littorina_rudis_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Littorina_rudis_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Littorina_rudis_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Littorina_rudis_Total_soft_tissues
#| fig-cap: "Molluscs / Littorina rudis / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Littorina_rudis_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Littorina_rudis_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Littorina_rudis_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Littorina_rudis_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Littorina_rudis_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Littorina_rudis_Total_soft_tissues
#| fig-cap: "Molluscs / Littorina rudis / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Littorina_rudis_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Littorina_rudis_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Littorina_rudis_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Littorina_rudis_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Littorina_rudis_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Littorina_rudis_Total_soft_tissues
#| fig-cap: "Molluscs / Littorina rudis / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Littorina_rudis_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Molluscs_Littorina_rudis_Total_soft_tissues_unit, ")")
)
```

## Modiolus modiolus: Total soft tissues

Molluscs / Modiolus modiolus / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Modiolus_modiolus_Total_soft_tissues
grp_Molluscs_Modiolus_modiolus_Total_soft_tissues <- tar_read(outlier_biota_Molluscs_Modiolus_modiolus_Total_soft_tissues)
grp_Molluscs_Modiolus_modiolus_Total_soft_tissues$data <- grp_Molluscs_Modiolus_modiolus_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Modiolus_modiolus_Total_soft_tissues_unit <- grp_Molluscs_Modiolus_modiolus_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Modiolus_modiolus_Total_soft_tissues
#| fig-cap: "Molluscs / Modiolus modiolus / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Modiolus_modiolus_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Modiolus_modiolus_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Modiolus_modiolus_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Modiolus_modiolus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Modiolus_modiolus_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Modiolus_modiolus_Total_soft_tissues
#| fig-cap: "Molluscs / Modiolus modiolus / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Modiolus_modiolus_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Modiolus_modiolus_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Modiolus_modiolus_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Modiolus_modiolus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Modiolus_modiolus_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Modiolus_modiolus_Total_soft_tissues
#| fig-cap: "Molluscs / Modiolus modiolus / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Modiolus_modiolus_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Modiolus_modiolus_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Modiolus_modiolus_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Modiolus_modiolus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Modiolus_modiolus_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Modiolus_modiolus_Total_soft_tissues
#| fig-cap: "Molluscs / Modiolus modiolus / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Modiolus_modiolus_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Molluscs_Modiolus_modiolus_Total_soft_tissues_unit, ")")
)
```

## Mya arenaria: Total soft tissues

Molluscs / Mya arenaria / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Mya_arenaria_Total_soft_tissues
grp_Molluscs_Mya_arenaria_Total_soft_tissues <- tar_read(outlier_biota_Molluscs_Mya_arenaria_Total_soft_tissues)
grp_Molluscs_Mya_arenaria_Total_soft_tissues$data <- grp_Molluscs_Mya_arenaria_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Mya_arenaria_Total_soft_tissues_unit <- grp_Molluscs_Mya_arenaria_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Mya_arenaria_Total_soft_tissues
#| fig-cap: "Molluscs / Mya arenaria / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Mya_arenaria_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Mya_arenaria_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Mya_arenaria_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Mya_arenaria_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Mya_arenaria_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Mya_arenaria_Total_soft_tissues
#| fig-cap: "Molluscs / Mya arenaria / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Mya_arenaria_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Mya_arenaria_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Mya_arenaria_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Mya_arenaria_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Mya_arenaria_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Mya_arenaria_Total_soft_tissues
#| fig-cap: "Molluscs / Mya arenaria / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Mya_arenaria_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Mya_arenaria_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Mya_arenaria_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Mya_arenaria_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Mya_arenaria_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Mya_arenaria_Total_soft_tissues
#| fig-cap: "Molluscs / Mya arenaria / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Mya_arenaria_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Molluscs_Mya_arenaria_Total_soft_tissues_unit, ")")
)
```

## Mytilus edulis: Total soft tissues

Molluscs / Mytilus edulis / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Mytilus_edulis_Total_soft_tissues
grp_Molluscs_Mytilus_edulis_Total_soft_tissues <- tar_read(outlier_biota_Molluscs_Mytilus_edulis_Total_soft_tissues)
grp_Molluscs_Mytilus_edulis_Total_soft_tissues$data <- grp_Molluscs_Mytilus_edulis_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Mytilus_edulis_Total_soft_tissues_unit <- grp_Molluscs_Mytilus_edulis_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Mytilus_edulis_Total_soft_tissues
#| fig-cap: "Molluscs / Mytilus edulis / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Mytilus_edulis_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Mytilus_edulis_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Mytilus_edulis_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Mytilus_edulis_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Mytilus_edulis_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Mytilus_edulis_Total_soft_tissues
#| fig-cap: "Molluscs / Mytilus edulis / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Mytilus_edulis_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Mytilus_edulis_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Mytilus_edulis_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Mytilus_edulis_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Mytilus_edulis_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Mytilus_edulis_Total_soft_tissues
#| fig-cap: "Molluscs / Mytilus edulis / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Mytilus_edulis_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Mytilus_edulis_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Mytilus_edulis_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Mytilus_edulis_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Mytilus_edulis_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Mytilus_edulis_Total_soft_tissues
#| fig-cap: "Molluscs / Mytilus edulis / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Mytilus_edulis_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Molluscs_Mytilus_edulis_Total_soft_tissues_unit, ")")
)
```

## Mytilus edulis: Whole body

Molluscs / Mytilus edulis / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Mytilus_edulis_Whole_body
grp_Molluscs_Mytilus_edulis_Whole_body <- tar_read(outlier_biota_Molluscs_Mytilus_edulis_Whole_body)
grp_Molluscs_Mytilus_edulis_Whole_body$data <- grp_Molluscs_Mytilus_edulis_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Mytilus_edulis_Whole_body_unit <- grp_Molluscs_Mytilus_edulis_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Mytilus_edulis_Whole_body
#| fig-cap: "Molluscs / Mytilus edulis / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Mytilus_edulis_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Mytilus_edulis_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Mytilus_edulis_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Mytilus_edulis_Whole_body
outlier_flag_table(flag_by_facet(grp_Molluscs_Mytilus_edulis_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Mytilus_edulis_Whole_body
#| fig-cap: "Molluscs / Mytilus edulis / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Mytilus_edulis_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Mytilus_edulis_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Mytilus_edulis_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Mytilus_edulis_Whole_body
outlier_flag_table(flag_by_facet(grp_Molluscs_Mytilus_edulis_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Mytilus_edulis_Whole_body
#| fig-cap: "Molluscs / Mytilus edulis / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Mytilus_edulis_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Mytilus_edulis_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Mytilus_edulis_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Mytilus_edulis_Whole_body
outlier_flag_table(flag_by_facet(grp_Molluscs_Mytilus_edulis_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Mytilus_edulis_Whole_body
#| fig-cap: "Molluscs / Mytilus edulis / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Mytilus_edulis_Whole_body,
  x_label = paste0("Measured value (", grp_Molluscs_Mytilus_edulis_Whole_body_unit, ")")
)
```

## Nucella lapillus: Total soft tissues

Molluscs / Nucella lapillus / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Nucella_lapillus_Total_soft_tissues
grp_Molluscs_Nucella_lapillus_Total_soft_tissues <- tar_read(outlier_biota_Molluscs_Nucella_lapillus_Total_soft_tissues)
grp_Molluscs_Nucella_lapillus_Total_soft_tissues$data <- grp_Molluscs_Nucella_lapillus_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Nucella_lapillus_Total_soft_tissues_unit <- grp_Molluscs_Nucella_lapillus_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Nucella_lapillus_Total_soft_tissues
#| fig-cap: "Molluscs / Nucella lapillus / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Nucella_lapillus_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Nucella_lapillus_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Nucella_lapillus_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Nucella_lapillus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Nucella_lapillus_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Nucella_lapillus_Total_soft_tissues
#| fig-cap: "Molluscs / Nucella lapillus / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Nucella_lapillus_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Nucella_lapillus_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Nucella_lapillus_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Nucella_lapillus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Nucella_lapillus_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Nucella_lapillus_Total_soft_tissues
#| fig-cap: "Molluscs / Nucella lapillus / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Nucella_lapillus_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Nucella_lapillus_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Nucella_lapillus_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Nucella_lapillus_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Nucella_lapillus_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Nucella_lapillus_Total_soft_tissues
#| fig-cap: "Molluscs / Nucella lapillus / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Nucella_lapillus_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Molluscs_Nucella_lapillus_Total_soft_tissues_unit, ")")
)
```

## Patella vulgata: Total soft tissues

Molluscs / Patella vulgata / Total soft tissues. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Patella_vulgata_Total_soft_tissues
grp_Molluscs_Patella_vulgata_Total_soft_tissues <- tar_read(outlier_biota_Molluscs_Patella_vulgata_Total_soft_tissues)
grp_Molluscs_Patella_vulgata_Total_soft_tissues$data <- grp_Molluscs_Patella_vulgata_Total_soft_tissues$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Patella_vulgata_Total_soft_tissues_unit <- grp_Molluscs_Patella_vulgata_Total_soft_tissues$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Patella_vulgata_Total_soft_tissues
#| fig-cap: "Molluscs / Patella vulgata / Total soft tissues: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Patella_vulgata_Total_soft_tissues$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Patella_vulgata_Total_soft_tissues$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Patella_vulgata_Total_soft_tissues_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Patella_vulgata_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Patella_vulgata_Total_soft_tissues$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Patella_vulgata_Total_soft_tissues
#| fig-cap: "Molluscs / Patella vulgata / Total soft tissues: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Patella_vulgata_Total_soft_tissues$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Patella_vulgata_Total_soft_tissues$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Patella_vulgata_Total_soft_tissues_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Patella_vulgata_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Patella_vulgata_Total_soft_tissues$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Patella_vulgata_Total_soft_tissues
#| fig-cap: "Molluscs / Patella vulgata / Total soft tissues: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Patella_vulgata_Total_soft_tissues$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Patella_vulgata_Total_soft_tissues$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Patella_vulgata_Total_soft_tissues_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Patella_vulgata_Total_soft_tissues
outlier_flag_table(flag_by_facet(grp_Molluscs_Patella_vulgata_Total_soft_tissues$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Patella_vulgata_Total_soft_tissues
#| fig-cap: "Molluscs / Patella vulgata / Total soft tissues: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Patella_vulgata_Total_soft_tissues,
  x_label = paste0("Measured value (", grp_Molluscs_Patella_vulgata_Total_soft_tissues_unit, ")")
)
```

## Patella vulgata: Whole body

Molluscs / Patella vulgata / Whole body. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Molluscs_Patella_vulgata_Whole_body
grp_Molluscs_Patella_vulgata_Whole_body <- tar_read(outlier_biota_Molluscs_Patella_vulgata_Whole_body)
grp_Molluscs_Patella_vulgata_Whole_body$data <- grp_Molluscs_Patella_vulgata_Whole_body$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Molluscs_Patella_vulgata_Whole_body_unit <- grp_Molluscs_Patella_vulgata_Whole_body$summary$unit
```

```{r}
#| label: fig-outlier-year-Molluscs_Patella_vulgata_Whole_body
#| fig-cap: "Molluscs / Patella vulgata / Whole body: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Patella_vulgata_Whole_body$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Molluscs_Patella_vulgata_Whole_body$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Molluscs_Patella_vulgata_Whole_body_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Molluscs_Patella_vulgata_Whole_body
outlier_flag_table(flag_by_facet(grp_Molluscs_Patella_vulgata_Whole_body$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Molluscs_Patella_vulgata_Whole_body
#| fig-cap: "Molluscs / Patella vulgata / Whole body: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Patella_vulgata_Whole_body$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Molluscs_Patella_vulgata_Whole_body$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Molluscs_Patella_vulgata_Whole_body_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Molluscs_Patella_vulgata_Whole_body
outlier_flag_table(flag_by_facet(grp_Molluscs_Patella_vulgata_Whole_body$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Molluscs_Patella_vulgata_Whole_body
#| fig-cap: "Molluscs / Patella vulgata / Whole body: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Molluscs_Patella_vulgata_Whole_body$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Molluscs_Patella_vulgata_Whole_body$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Molluscs_Patella_vulgata_Whole_body_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Molluscs_Patella_vulgata_Whole_body
outlier_flag_table(flag_by_facet(grp_Molluscs_Patella_vulgata_Whole_body$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Molluscs_Patella_vulgata_Whole_body
#| fig-cap: "Molluscs / Patella vulgata / Whole body: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Molluscs_Patella_vulgata_Whole_body,
  x_label = paste0("Measured value (", grp_Molluscs_Patella_vulgata_Whole_body_unit, ")")
)
```
