---
title: "Distributions: Mammals"
format:
  html:
    fig-width: 16
    fig-height: 20
  docx:
    fig-width: 16
    fig-height: 20

execute:
  echo: false

knitr:
  opts_chunk:
    dpi: 300
    dev: "ragg_png"

filters:
  - flextable-qmd
  - at: post-render
    path: ../_extensions/flextable-qmd/unwrap-float.lua
---

::: callout-note
This notebook was scaffolded by `scripts/generate_distribution_notebooks.R`
from the outlier-analysis targets factory (`_targets.R`,
`R/fct_outlier_groups.R`, `R/fct_outlier_detection.R`,
`R/fct_outlier_report.R`). The plots/tables below are generated, but
analysis and interpretation are added by hand, section by section.
Re-running the generator only appends sections for groups not yet present
here -- it never rewrites or removes anything you have written.

See [NBXX-Outliers.qmd](NBXX-Outliers.qmd) for the fully hand-authored
*G. morhua* liver case study this workflow was built from, including
year/site/campaign breakdowns and the reasoning behind the outlier flags.
:::

```{r}
#| label: setup
library(tidyverse)
library(targets)
library(patchwork)
library(flextable)
pkgload::load_all(quiet = TRUE)

set_flextable_defaults(
  font.size = 10,
  font.family = "Arial",
  digits = 2,
  theme_fun = theme_vanilla,
  padding = 1
)

ggplot_base_size <- 20
theme_set(theme_minimal(base_size = ggplot_base_size, base_family = "Aptos"))
```

## Balaenoptera physalus: Fat/Adipose

Mammals / Balaenoptera physalus / Fat/Adipose. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Balaenoptera_physalus_Fat_Adipose
grp_Mammals_Balaenoptera_physalus_Fat_Adipose <- tar_read(outlier_biota_Mammals_Balaenoptera_physalus_Fat_Adipose)
grp_Mammals_Balaenoptera_physalus_Fat_Adipose$data <- grp_Mammals_Balaenoptera_physalus_Fat_Adipose$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Balaenoptera_physalus_Fat_Adipose_unit <- grp_Mammals_Balaenoptera_physalus_Fat_Adipose$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Balaenoptera_physalus_Fat_Adipose
#| fig-cap: "Mammals / Balaenoptera physalus / Fat/Adipose: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Balaenoptera_physalus_Fat_Adipose$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Balaenoptera_physalus_Fat_Adipose$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Balaenoptera_physalus_Fat_Adipose_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Balaenoptera_physalus_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Balaenoptera_physalus_Fat_Adipose$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Balaenoptera_physalus_Fat_Adipose
#| fig-cap: "Mammals / Balaenoptera physalus / Fat/Adipose: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Balaenoptera_physalus_Fat_Adipose$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Balaenoptera_physalus_Fat_Adipose$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Balaenoptera_physalus_Fat_Adipose_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Balaenoptera_physalus_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Balaenoptera_physalus_Fat_Adipose$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Balaenoptera_physalus_Fat_Adipose
#| fig-cap: "Mammals / Balaenoptera physalus / Fat/Adipose: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Balaenoptera_physalus_Fat_Adipose$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Balaenoptera_physalus_Fat_Adipose$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Balaenoptera_physalus_Fat_Adipose_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Balaenoptera_physalus_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Balaenoptera_physalus_Fat_Adipose$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Balaenoptera_physalus_Fat_Adipose
#| fig-cap: "Mammals / Balaenoptera physalus / Fat/Adipose: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Balaenoptera_physalus_Fat_Adipose,
  x_label = paste0("Measured value (", grp_Mammals_Balaenoptera_physalus_Fat_Adipose_unit, ")")
)
```

## Capreolus capreolus: Liver

Mammals / Capreolus capreolus / Liver. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Capreolus_capreolus_Liver
grp_Mammals_Capreolus_capreolus_Liver <- tar_read(outlier_biota_Mammals_Capreolus_capreolus_Liver)
grp_Mammals_Capreolus_capreolus_Liver$data <- grp_Mammals_Capreolus_capreolus_Liver$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Capreolus_capreolus_Liver_unit <- grp_Mammals_Capreolus_capreolus_Liver$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Capreolus_capreolus_Liver
#| fig-cap: "Mammals / Capreolus capreolus / Liver: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Capreolus_capreolus_Liver$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Capreolus_capreolus_Liver$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Capreolus_capreolus_Liver_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Capreolus_capreolus_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Capreolus_capreolus_Liver$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Capreolus_capreolus_Liver
#| fig-cap: "Mammals / Capreolus capreolus / Liver: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Capreolus_capreolus_Liver$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Capreolus_capreolus_Liver$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Capreolus_capreolus_Liver_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Capreolus_capreolus_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Capreolus_capreolus_Liver$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Capreolus_capreolus_Liver
#| fig-cap: "Mammals / Capreolus capreolus / Liver: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Capreolus_capreolus_Liver$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Capreolus_capreolus_Liver$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Capreolus_capreolus_Liver_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Capreolus_capreolus_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Capreolus_capreolus_Liver$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Capreolus_capreolus_Liver
#| fig-cap: "Mammals / Capreolus capreolus / Liver: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Capreolus_capreolus_Liver,
  x_label = paste0("Measured value (", grp_Mammals_Capreolus_capreolus_Liver_unit, ")")
)
```

## Cystophora cristata: Kidney

Mammals / Cystophora cristata / Kidney. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Cystophora_cristata_Kidney
grp_Mammals_Cystophora_cristata_Kidney <- tar_read(outlier_biota_Mammals_Cystophora_cristata_Kidney)
grp_Mammals_Cystophora_cristata_Kidney$data <- grp_Mammals_Cystophora_cristata_Kidney$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Cystophora_cristata_Kidney_unit <- grp_Mammals_Cystophora_cristata_Kidney$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Cystophora_cristata_Kidney
#| fig-cap: "Mammals / Cystophora cristata / Kidney: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Cystophora_cristata_Kidney$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Cystophora_cristata_Kidney$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Cystophora_cristata_Kidney_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Cystophora_cristata_Kidney
outlier_flag_table(flag_by_facet(grp_Mammals_Cystophora_cristata_Kidney$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Cystophora_cristata_Kidney
#| fig-cap: "Mammals / Cystophora cristata / Kidney: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Cystophora_cristata_Kidney$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Cystophora_cristata_Kidney$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Cystophora_cristata_Kidney_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Cystophora_cristata_Kidney
outlier_flag_table(flag_by_facet(grp_Mammals_Cystophora_cristata_Kidney$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Cystophora_cristata_Kidney
#| fig-cap: "Mammals / Cystophora cristata / Kidney: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Cystophora_cristata_Kidney$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Cystophora_cristata_Kidney$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Cystophora_cristata_Kidney_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Cystophora_cristata_Kidney
outlier_flag_table(flag_by_facet(grp_Mammals_Cystophora_cristata_Kidney$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Cystophora_cristata_Kidney
#| fig-cap: "Mammals / Cystophora cristata / Kidney: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Cystophora_cristata_Kidney,
  x_label = paste0("Measured value (", grp_Mammals_Cystophora_cristata_Kidney_unit, ")")
)
```

## Cystophora cristata: Liver

Mammals / Cystophora cristata / Liver. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Cystophora_cristata_Liver
grp_Mammals_Cystophora_cristata_Liver <- tar_read(outlier_biota_Mammals_Cystophora_cristata_Liver)
grp_Mammals_Cystophora_cristata_Liver$data <- grp_Mammals_Cystophora_cristata_Liver$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Cystophora_cristata_Liver_unit <- grp_Mammals_Cystophora_cristata_Liver$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Cystophora_cristata_Liver
#| fig-cap: "Mammals / Cystophora cristata / Liver: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Cystophora_cristata_Liver$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Cystophora_cristata_Liver$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Cystophora_cristata_Liver_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Cystophora_cristata_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Cystophora_cristata_Liver$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Cystophora_cristata_Liver
#| fig-cap: "Mammals / Cystophora cristata / Liver: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Cystophora_cristata_Liver$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Cystophora_cristata_Liver$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Cystophora_cristata_Liver_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Cystophora_cristata_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Cystophora_cristata_Liver$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Cystophora_cristata_Liver
#| fig-cap: "Mammals / Cystophora cristata / Liver: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Cystophora_cristata_Liver$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Cystophora_cristata_Liver$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Cystophora_cristata_Liver_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Cystophora_cristata_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Cystophora_cristata_Liver$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Cystophora_cristata_Liver
#| fig-cap: "Mammals / Cystophora cristata / Liver: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Cystophora_cristata_Liver,
  x_label = paste0("Measured value (", grp_Mammals_Cystophora_cristata_Liver_unit, ")")
)
```

## Cystophora cristata: Muscle

Mammals / Cystophora cristata / Muscle. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Cystophora_cristata_Muscle
grp_Mammals_Cystophora_cristata_Muscle <- tar_read(outlier_biota_Mammals_Cystophora_cristata_Muscle)
grp_Mammals_Cystophora_cristata_Muscle$data <- grp_Mammals_Cystophora_cristata_Muscle$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Cystophora_cristata_Muscle_unit <- grp_Mammals_Cystophora_cristata_Muscle$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Cystophora_cristata_Muscle
#| fig-cap: "Mammals / Cystophora cristata / Muscle: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Cystophora_cristata_Muscle$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Cystophora_cristata_Muscle$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Cystophora_cristata_Muscle_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Cystophora_cristata_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Cystophora_cristata_Muscle$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Cystophora_cristata_Muscle
#| fig-cap: "Mammals / Cystophora cristata / Muscle: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Cystophora_cristata_Muscle$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Cystophora_cristata_Muscle$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Cystophora_cristata_Muscle_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Cystophora_cristata_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Cystophora_cristata_Muscle$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Cystophora_cristata_Muscle
#| fig-cap: "Mammals / Cystophora cristata / Muscle: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Cystophora_cristata_Muscle$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Cystophora_cristata_Muscle$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Cystophora_cristata_Muscle_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Cystophora_cristata_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Cystophora_cristata_Muscle$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Cystophora_cristata_Muscle
#| fig-cap: "Mammals / Cystophora cristata / Muscle: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Cystophora_cristata_Muscle,
  x_label = paste0("Measured value (", grp_Mammals_Cystophora_cristata_Muscle_unit, ")")
)
```

## Lagenorhynchus albirostris: Muscle

Mammals / Lagenorhynchus albirostris / Muscle. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Lagenorhynchus_albirostris_Muscle
grp_Mammals_Lagenorhynchus_albirostris_Muscle <- tar_read(outlier_biota_Mammals_Lagenorhynchus_albirostris_Muscle)
grp_Mammals_Lagenorhynchus_albirostris_Muscle$data <- grp_Mammals_Lagenorhynchus_albirostris_Muscle$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Lagenorhynchus_albirostris_Muscle_unit <- grp_Mammals_Lagenorhynchus_albirostris_Muscle$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Lagenorhynchus_albirostris_Muscle
#| fig-cap: "Mammals / Lagenorhynchus albirostris / Muscle: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Lagenorhynchus_albirostris_Muscle$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Lagenorhynchus_albirostris_Muscle$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Lagenorhynchus_albirostris_Muscle_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Lagenorhynchus_albirostris_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Lagenorhynchus_albirostris_Muscle$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Lagenorhynchus_albirostris_Muscle
#| fig-cap: "Mammals / Lagenorhynchus albirostris / Muscle: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Lagenorhynchus_albirostris_Muscle$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Lagenorhynchus_albirostris_Muscle$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Lagenorhynchus_albirostris_Muscle_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Lagenorhynchus_albirostris_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Lagenorhynchus_albirostris_Muscle$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Lagenorhynchus_albirostris_Muscle
#| fig-cap: "Mammals / Lagenorhynchus albirostris / Muscle: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Lagenorhynchus_albirostris_Muscle$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Lagenorhynchus_albirostris_Muscle$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Lagenorhynchus_albirostris_Muscle_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Lagenorhynchus_albirostris_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Lagenorhynchus_albirostris_Muscle$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Lagenorhynchus_albirostris_Muscle
#| fig-cap: "Mammals / Lagenorhynchus albirostris / Muscle: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Lagenorhynchus_albirostris_Muscle,
  x_label = paste0("Measured value (", grp_Mammals_Lagenorhynchus_albirostris_Muscle_unit, ")")
)
```

## Megaptera novaeangliae: Fat/Adipose

Mammals / Megaptera novaeangliae / Fat/Adipose. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Megaptera_novaeangliae_Fat_Adipose
grp_Mammals_Megaptera_novaeangliae_Fat_Adipose <- tar_read(outlier_biota_Mammals_Megaptera_novaeangliae_Fat_Adipose)
grp_Mammals_Megaptera_novaeangliae_Fat_Adipose$data <- grp_Mammals_Megaptera_novaeangliae_Fat_Adipose$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Megaptera_novaeangliae_Fat_Adipose_unit <- grp_Mammals_Megaptera_novaeangliae_Fat_Adipose$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Megaptera_novaeangliae_Fat_Adipose
#| fig-cap: "Mammals / Megaptera novaeangliae / Fat/Adipose: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Megaptera_novaeangliae_Fat_Adipose$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Megaptera_novaeangliae_Fat_Adipose$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Megaptera_novaeangliae_Fat_Adipose_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Megaptera_novaeangliae_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Megaptera_novaeangliae_Fat_Adipose$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Megaptera_novaeangliae_Fat_Adipose
#| fig-cap: "Mammals / Megaptera novaeangliae / Fat/Adipose: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Megaptera_novaeangliae_Fat_Adipose$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Megaptera_novaeangliae_Fat_Adipose$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Megaptera_novaeangliae_Fat_Adipose_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Megaptera_novaeangliae_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Megaptera_novaeangliae_Fat_Adipose$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Megaptera_novaeangliae_Fat_Adipose
#| fig-cap: "Mammals / Megaptera novaeangliae / Fat/Adipose: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Megaptera_novaeangliae_Fat_Adipose$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Megaptera_novaeangliae_Fat_Adipose$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Megaptera_novaeangliae_Fat_Adipose_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Megaptera_novaeangliae_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Megaptera_novaeangliae_Fat_Adipose$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Megaptera_novaeangliae_Fat_Adipose
#| fig-cap: "Mammals / Megaptera novaeangliae / Fat/Adipose: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Megaptera_novaeangliae_Fat_Adipose,
  x_label = paste0("Measured value (", grp_Mammals_Megaptera_novaeangliae_Fat_Adipose_unit, ")")
)
```

## Odobenus rosmarus divergens: Heart

Mammals / Odobenus rosmarus divergens / Heart. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Odobenus_rosmarus_divergens_Heart
grp_Mammals_Odobenus_rosmarus_divergens_Heart <- tar_read(outlier_biota_Mammals_Odobenus_rosmarus_divergens_Heart)
grp_Mammals_Odobenus_rosmarus_divergens_Heart$data <- grp_Mammals_Odobenus_rosmarus_divergens_Heart$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Odobenus_rosmarus_divergens_Heart_unit <- grp_Mammals_Odobenus_rosmarus_divergens_Heart$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Odobenus_rosmarus_divergens_Heart
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Heart: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Heart$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Heart$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Heart_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Odobenus_rosmarus_divergens_Heart
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Heart$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Odobenus_rosmarus_divergens_Heart
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Heart: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Heart$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Heart$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Heart_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Odobenus_rosmarus_divergens_Heart
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Heart$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Odobenus_rosmarus_divergens_Heart
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Heart: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Heart$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Heart$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Heart_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Odobenus_rosmarus_divergens_Heart
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Heart$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Odobenus_rosmarus_divergens_Heart
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Heart: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Odobenus_rosmarus_divergens_Heart,
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Heart_unit, ")")
)
```

## Odobenus rosmarus divergens: Kidney

Mammals / Odobenus rosmarus divergens / Kidney. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Odobenus_rosmarus_divergens_Kidney
grp_Mammals_Odobenus_rosmarus_divergens_Kidney <- tar_read(outlier_biota_Mammals_Odobenus_rosmarus_divergens_Kidney)
grp_Mammals_Odobenus_rosmarus_divergens_Kidney$data <- grp_Mammals_Odobenus_rosmarus_divergens_Kidney$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Odobenus_rosmarus_divergens_Kidney_unit <- grp_Mammals_Odobenus_rosmarus_divergens_Kidney$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Odobenus_rosmarus_divergens_Kidney
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Kidney: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Kidney$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Kidney$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Kidney_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Odobenus_rosmarus_divergens_Kidney
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Kidney$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Odobenus_rosmarus_divergens_Kidney
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Kidney: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Kidney$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Kidney$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Kidney_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Odobenus_rosmarus_divergens_Kidney
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Kidney$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Odobenus_rosmarus_divergens_Kidney
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Kidney: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Kidney$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Kidney$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Kidney_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Odobenus_rosmarus_divergens_Kidney
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Kidney$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Odobenus_rosmarus_divergens_Kidney
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Kidney: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Odobenus_rosmarus_divergens_Kidney,
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Kidney_unit, ")")
)
```

## Odobenus rosmarus divergens: Liver

Mammals / Odobenus rosmarus divergens / Liver. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Odobenus_rosmarus_divergens_Liver
grp_Mammals_Odobenus_rosmarus_divergens_Liver <- tar_read(outlier_biota_Mammals_Odobenus_rosmarus_divergens_Liver)
grp_Mammals_Odobenus_rosmarus_divergens_Liver$data <- grp_Mammals_Odobenus_rosmarus_divergens_Liver$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Odobenus_rosmarus_divergens_Liver_unit <- grp_Mammals_Odobenus_rosmarus_divergens_Liver$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Odobenus_rosmarus_divergens_Liver
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Liver: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Liver$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Liver$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Liver_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Odobenus_rosmarus_divergens_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Liver$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Odobenus_rosmarus_divergens_Liver
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Liver: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Liver$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Liver$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Liver_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Odobenus_rosmarus_divergens_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Liver$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Odobenus_rosmarus_divergens_Liver
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Liver: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Liver$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Liver$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Liver_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Odobenus_rosmarus_divergens_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Liver$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Odobenus_rosmarus_divergens_Liver
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Liver: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Odobenus_rosmarus_divergens_Liver,
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Liver_unit, ")")
)
```

## Odobenus rosmarus divergens: Lung

Mammals / Odobenus rosmarus divergens / Lung. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Odobenus_rosmarus_divergens_Lung
grp_Mammals_Odobenus_rosmarus_divergens_Lung <- tar_read(outlier_biota_Mammals_Odobenus_rosmarus_divergens_Lung)
grp_Mammals_Odobenus_rosmarus_divergens_Lung$data <- grp_Mammals_Odobenus_rosmarus_divergens_Lung$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Odobenus_rosmarus_divergens_Lung_unit <- grp_Mammals_Odobenus_rosmarus_divergens_Lung$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Odobenus_rosmarus_divergens_Lung
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Lung: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Lung$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Lung$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Lung_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Odobenus_rosmarus_divergens_Lung
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Lung$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Odobenus_rosmarus_divergens_Lung
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Lung: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Lung$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Lung$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Lung_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Odobenus_rosmarus_divergens_Lung
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Lung$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Odobenus_rosmarus_divergens_Lung
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Lung: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Lung$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Lung$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Lung_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Odobenus_rosmarus_divergens_Lung
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Lung$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Odobenus_rosmarus_divergens_Lung
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Lung: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Odobenus_rosmarus_divergens_Lung,
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Lung_unit, ")")
)
```

## Odobenus rosmarus divergens: Muscle

Mammals / Odobenus rosmarus divergens / Muscle. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Odobenus_rosmarus_divergens_Muscle
grp_Mammals_Odobenus_rosmarus_divergens_Muscle <- tar_read(outlier_biota_Mammals_Odobenus_rosmarus_divergens_Muscle)
grp_Mammals_Odobenus_rosmarus_divergens_Muscle$data <- grp_Mammals_Odobenus_rosmarus_divergens_Muscle$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Odobenus_rosmarus_divergens_Muscle_unit <- grp_Mammals_Odobenus_rosmarus_divergens_Muscle$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Odobenus_rosmarus_divergens_Muscle
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Muscle: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Muscle$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Muscle$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Muscle_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Odobenus_rosmarus_divergens_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Muscle$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Odobenus_rosmarus_divergens_Muscle
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Muscle: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Muscle$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Muscle$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Muscle_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Odobenus_rosmarus_divergens_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Muscle$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Odobenus_rosmarus_divergens_Muscle
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Muscle: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Muscle$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Muscle$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Muscle_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Odobenus_rosmarus_divergens_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Muscle$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Odobenus_rosmarus_divergens_Muscle
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Muscle: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Odobenus_rosmarus_divergens_Muscle,
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Muscle_unit, ")")
)
```

## Odobenus rosmarus divergens: Skin

Mammals / Odobenus rosmarus divergens / Skin. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Odobenus_rosmarus_divergens_Skin
grp_Mammals_Odobenus_rosmarus_divergens_Skin <- tar_read(outlier_biota_Mammals_Odobenus_rosmarus_divergens_Skin)
grp_Mammals_Odobenus_rosmarus_divergens_Skin$data <- grp_Mammals_Odobenus_rosmarus_divergens_Skin$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Odobenus_rosmarus_divergens_Skin_unit <- grp_Mammals_Odobenus_rosmarus_divergens_Skin$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Odobenus_rosmarus_divergens_Skin
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Skin: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Skin$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Skin$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Skin_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Odobenus_rosmarus_divergens_Skin
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Skin$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Odobenus_rosmarus_divergens_Skin
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Skin: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Skin$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Skin$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Skin_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Odobenus_rosmarus_divergens_Skin
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Skin$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Odobenus_rosmarus_divergens_Skin
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Skin: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_divergens_Skin$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_divergens_Skin$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Skin_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Odobenus_rosmarus_divergens_Skin
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_divergens_Skin$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Odobenus_rosmarus_divergens_Skin
#| fig-cap: "Mammals / Odobenus rosmarus divergens / Skin: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Odobenus_rosmarus_divergens_Skin,
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_divergens_Skin_unit, ")")
)
```

## Odobenus rosmarus rosmarus: Skin

Mammals / Odobenus rosmarus rosmarus / Skin. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Odobenus_rosmarus_rosmarus_Skin
grp_Mammals_Odobenus_rosmarus_rosmarus_Skin <- tar_read(outlier_biota_Mammals_Odobenus_rosmarus_rosmarus_Skin)
grp_Mammals_Odobenus_rosmarus_rosmarus_Skin$data <- grp_Mammals_Odobenus_rosmarus_rosmarus_Skin$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Odobenus_rosmarus_rosmarus_Skin_unit <- grp_Mammals_Odobenus_rosmarus_rosmarus_Skin$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Odobenus_rosmarus_rosmarus_Skin
#| fig-cap: "Mammals / Odobenus rosmarus rosmarus / Skin: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_rosmarus_Skin$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_rosmarus_Skin$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_rosmarus_Skin_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Odobenus_rosmarus_rosmarus_Skin
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_rosmarus_Skin$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Odobenus_rosmarus_rosmarus_Skin
#| fig-cap: "Mammals / Odobenus rosmarus rosmarus / Skin: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_rosmarus_Skin$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_rosmarus_Skin$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_rosmarus_Skin_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Odobenus_rosmarus_rosmarus_Skin
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_rosmarus_Skin$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Odobenus_rosmarus_rosmarus_Skin
#| fig-cap: "Mammals / Odobenus rosmarus rosmarus / Skin: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Odobenus_rosmarus_rosmarus_Skin$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Odobenus_rosmarus_rosmarus_Skin$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_rosmarus_Skin_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Odobenus_rosmarus_rosmarus_Skin
outlier_flag_table(flag_by_facet(grp_Mammals_Odobenus_rosmarus_rosmarus_Skin$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Odobenus_rosmarus_rosmarus_Skin
#| fig-cap: "Mammals / Odobenus rosmarus rosmarus / Skin: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Odobenus_rosmarus_rosmarus_Skin,
  x_label = paste0("Measured value (", grp_Mammals_Odobenus_rosmarus_rosmarus_Skin_unit, ")")
)
```

## Orcinus orca: Fat/Adipose

Mammals / Orcinus orca / Fat/Adipose. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Orcinus_orca_Fat_Adipose
grp_Mammals_Orcinus_orca_Fat_Adipose <- tar_read(outlier_biota_Mammals_Orcinus_orca_Fat_Adipose)
grp_Mammals_Orcinus_orca_Fat_Adipose$data <- grp_Mammals_Orcinus_orca_Fat_Adipose$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Orcinus_orca_Fat_Adipose_unit <- grp_Mammals_Orcinus_orca_Fat_Adipose$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Orcinus_orca_Fat_Adipose
#| fig-cap: "Mammals / Orcinus orca / Fat/Adipose: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Orcinus_orca_Fat_Adipose$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Orcinus_orca_Fat_Adipose$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Orcinus_orca_Fat_Adipose_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Orcinus_orca_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Orcinus_orca_Fat_Adipose$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Orcinus_orca_Fat_Adipose
#| fig-cap: "Mammals / Orcinus orca / Fat/Adipose: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Orcinus_orca_Fat_Adipose$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Orcinus_orca_Fat_Adipose$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Orcinus_orca_Fat_Adipose_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Orcinus_orca_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Orcinus_orca_Fat_Adipose$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Orcinus_orca_Fat_Adipose
#| fig-cap: "Mammals / Orcinus orca / Fat/Adipose: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Orcinus_orca_Fat_Adipose$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Orcinus_orca_Fat_Adipose$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Orcinus_orca_Fat_Adipose_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Orcinus_orca_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Orcinus_orca_Fat_Adipose$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Orcinus_orca_Fat_Adipose
#| fig-cap: "Mammals / Orcinus orca / Fat/Adipose: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Orcinus_orca_Fat_Adipose,
  x_label = paste0("Measured value (", grp_Mammals_Orcinus_orca_Fat_Adipose_unit, ")")
)
```

## Pagophilus groenlandicus: Kidney

Mammals / Pagophilus groenlandicus / Kidney. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Pagophilus_groenlandicus_Kidney
grp_Mammals_Pagophilus_groenlandicus_Kidney <- tar_read(outlier_biota_Mammals_Pagophilus_groenlandicus_Kidney)
grp_Mammals_Pagophilus_groenlandicus_Kidney$data <- grp_Mammals_Pagophilus_groenlandicus_Kidney$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Pagophilus_groenlandicus_Kidney_unit <- grp_Mammals_Pagophilus_groenlandicus_Kidney$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Pagophilus_groenlandicus_Kidney
#| fig-cap: "Mammals / Pagophilus groenlandicus / Kidney: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Pagophilus_groenlandicus_Kidney$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Pagophilus_groenlandicus_Kidney$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Pagophilus_groenlandicus_Kidney_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Pagophilus_groenlandicus_Kidney
outlier_flag_table(flag_by_facet(grp_Mammals_Pagophilus_groenlandicus_Kidney$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Pagophilus_groenlandicus_Kidney
#| fig-cap: "Mammals / Pagophilus groenlandicus / Kidney: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Pagophilus_groenlandicus_Kidney$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Pagophilus_groenlandicus_Kidney$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Pagophilus_groenlandicus_Kidney_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Pagophilus_groenlandicus_Kidney
outlier_flag_table(flag_by_facet(grp_Mammals_Pagophilus_groenlandicus_Kidney$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Pagophilus_groenlandicus_Kidney
#| fig-cap: "Mammals / Pagophilus groenlandicus / Kidney: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Pagophilus_groenlandicus_Kidney$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Pagophilus_groenlandicus_Kidney$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Pagophilus_groenlandicus_Kidney_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Pagophilus_groenlandicus_Kidney
outlier_flag_table(flag_by_facet(grp_Mammals_Pagophilus_groenlandicus_Kidney$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Pagophilus_groenlandicus_Kidney
#| fig-cap: "Mammals / Pagophilus groenlandicus / Kidney: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Pagophilus_groenlandicus_Kidney,
  x_label = paste0("Measured value (", grp_Mammals_Pagophilus_groenlandicus_Kidney_unit, ")")
)
```

## Pagophilus groenlandicus: Liver

Mammals / Pagophilus groenlandicus / Liver. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Pagophilus_groenlandicus_Liver
grp_Mammals_Pagophilus_groenlandicus_Liver <- tar_read(outlier_biota_Mammals_Pagophilus_groenlandicus_Liver)
grp_Mammals_Pagophilus_groenlandicus_Liver$data <- grp_Mammals_Pagophilus_groenlandicus_Liver$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Pagophilus_groenlandicus_Liver_unit <- grp_Mammals_Pagophilus_groenlandicus_Liver$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Pagophilus_groenlandicus_Liver
#| fig-cap: "Mammals / Pagophilus groenlandicus / Liver: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Pagophilus_groenlandicus_Liver$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Pagophilus_groenlandicus_Liver$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Pagophilus_groenlandicus_Liver_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Pagophilus_groenlandicus_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Pagophilus_groenlandicus_Liver$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Pagophilus_groenlandicus_Liver
#| fig-cap: "Mammals / Pagophilus groenlandicus / Liver: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Pagophilus_groenlandicus_Liver$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Pagophilus_groenlandicus_Liver$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Pagophilus_groenlandicus_Liver_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Pagophilus_groenlandicus_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Pagophilus_groenlandicus_Liver$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Pagophilus_groenlandicus_Liver
#| fig-cap: "Mammals / Pagophilus groenlandicus / Liver: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Pagophilus_groenlandicus_Liver$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Pagophilus_groenlandicus_Liver$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Pagophilus_groenlandicus_Liver_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Pagophilus_groenlandicus_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Pagophilus_groenlandicus_Liver$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Pagophilus_groenlandicus_Liver
#| fig-cap: "Mammals / Pagophilus groenlandicus / Liver: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Pagophilus_groenlandicus_Liver,
  x_label = paste0("Measured value (", grp_Mammals_Pagophilus_groenlandicus_Liver_unit, ")")
)
```

## Pagophilus groenlandicus: Muscle

Mammals / Pagophilus groenlandicus / Muscle. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Pagophilus_groenlandicus_Muscle
grp_Mammals_Pagophilus_groenlandicus_Muscle <- tar_read(outlier_biota_Mammals_Pagophilus_groenlandicus_Muscle)
grp_Mammals_Pagophilus_groenlandicus_Muscle$data <- grp_Mammals_Pagophilus_groenlandicus_Muscle$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Pagophilus_groenlandicus_Muscle_unit <- grp_Mammals_Pagophilus_groenlandicus_Muscle$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Pagophilus_groenlandicus_Muscle
#| fig-cap: "Mammals / Pagophilus groenlandicus / Muscle: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Pagophilus_groenlandicus_Muscle$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Pagophilus_groenlandicus_Muscle$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Pagophilus_groenlandicus_Muscle_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Pagophilus_groenlandicus_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Pagophilus_groenlandicus_Muscle$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Pagophilus_groenlandicus_Muscle
#| fig-cap: "Mammals / Pagophilus groenlandicus / Muscle: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Pagophilus_groenlandicus_Muscle$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Pagophilus_groenlandicus_Muscle$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Pagophilus_groenlandicus_Muscle_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Pagophilus_groenlandicus_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Pagophilus_groenlandicus_Muscle$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Pagophilus_groenlandicus_Muscle
#| fig-cap: "Mammals / Pagophilus groenlandicus / Muscle: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Pagophilus_groenlandicus_Muscle$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Pagophilus_groenlandicus_Muscle$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Pagophilus_groenlandicus_Muscle_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Pagophilus_groenlandicus_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Pagophilus_groenlandicus_Muscle$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Pagophilus_groenlandicus_Muscle
#| fig-cap: "Mammals / Pagophilus groenlandicus / Muscle: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Pagophilus_groenlandicus_Muscle,
  x_label = paste0("Measured value (", grp_Mammals_Pagophilus_groenlandicus_Muscle_unit, ")")
)
```

## Phoca groenlandica: Kidney

Mammals / Phoca groenlandica / Kidney. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Phoca_groenlandica_Kidney
grp_Mammals_Phoca_groenlandica_Kidney <- tar_read(outlier_biota_Mammals_Phoca_groenlandica_Kidney)
grp_Mammals_Phoca_groenlandica_Kidney$data <- grp_Mammals_Phoca_groenlandica_Kidney$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Phoca_groenlandica_Kidney_unit <- grp_Mammals_Phoca_groenlandica_Kidney$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Phoca_groenlandica_Kidney
#| fig-cap: "Mammals / Phoca groenlandica / Kidney: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phoca_groenlandica_Kidney$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Phoca_groenlandica_Kidney$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Phoca_groenlandica_Kidney_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Phoca_groenlandica_Kidney
outlier_flag_table(flag_by_facet(grp_Mammals_Phoca_groenlandica_Kidney$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Phoca_groenlandica_Kidney
#| fig-cap: "Mammals / Phoca groenlandica / Kidney: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phoca_groenlandica_Kidney$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Phoca_groenlandica_Kidney$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Phoca_groenlandica_Kidney_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Phoca_groenlandica_Kidney
outlier_flag_table(flag_by_facet(grp_Mammals_Phoca_groenlandica_Kidney$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Phoca_groenlandica_Kidney
#| fig-cap: "Mammals / Phoca groenlandica / Kidney: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phoca_groenlandica_Kidney$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Phoca_groenlandica_Kidney$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Phoca_groenlandica_Kidney_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Phoca_groenlandica_Kidney
outlier_flag_table(flag_by_facet(grp_Mammals_Phoca_groenlandica_Kidney$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Phoca_groenlandica_Kidney
#| fig-cap: "Mammals / Phoca groenlandica / Kidney: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Phoca_groenlandica_Kidney,
  x_label = paste0("Measured value (", grp_Mammals_Phoca_groenlandica_Kidney_unit, ")")
)
```

## Phoca groenlandica: Liver

Mammals / Phoca groenlandica / Liver. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Phoca_groenlandica_Liver
grp_Mammals_Phoca_groenlandica_Liver <- tar_read(outlier_biota_Mammals_Phoca_groenlandica_Liver)
grp_Mammals_Phoca_groenlandica_Liver$data <- grp_Mammals_Phoca_groenlandica_Liver$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Phoca_groenlandica_Liver_unit <- grp_Mammals_Phoca_groenlandica_Liver$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Phoca_groenlandica_Liver
#| fig-cap: "Mammals / Phoca groenlandica / Liver: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phoca_groenlandica_Liver$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Phoca_groenlandica_Liver$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Phoca_groenlandica_Liver_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Phoca_groenlandica_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Phoca_groenlandica_Liver$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Phoca_groenlandica_Liver
#| fig-cap: "Mammals / Phoca groenlandica / Liver: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phoca_groenlandica_Liver$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Phoca_groenlandica_Liver$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Phoca_groenlandica_Liver_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Phoca_groenlandica_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Phoca_groenlandica_Liver$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Phoca_groenlandica_Liver
#| fig-cap: "Mammals / Phoca groenlandica / Liver: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phoca_groenlandica_Liver$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Phoca_groenlandica_Liver$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Phoca_groenlandica_Liver_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Phoca_groenlandica_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Phoca_groenlandica_Liver$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Phoca_groenlandica_Liver
#| fig-cap: "Mammals / Phoca groenlandica / Liver: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Phoca_groenlandica_Liver,
  x_label = paste0("Measured value (", grp_Mammals_Phoca_groenlandica_Liver_unit, ")")
)
```

## Phoca vitulina: Liver

Mammals / Phoca vitulina / Liver. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Phoca_vitulina_Liver
grp_Mammals_Phoca_vitulina_Liver <- tar_read(outlier_biota_Mammals_Phoca_vitulina_Liver)
grp_Mammals_Phoca_vitulina_Liver$data <- grp_Mammals_Phoca_vitulina_Liver$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Phoca_vitulina_Liver_unit <- grp_Mammals_Phoca_vitulina_Liver$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Phoca_vitulina_Liver
#| fig-cap: "Mammals / Phoca vitulina / Liver: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phoca_vitulina_Liver$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Phoca_vitulina_Liver$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Phoca_vitulina_Liver_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Phoca_vitulina_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Phoca_vitulina_Liver$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Phoca_vitulina_Liver
#| fig-cap: "Mammals / Phoca vitulina / Liver: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phoca_vitulina_Liver$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Phoca_vitulina_Liver$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Phoca_vitulina_Liver_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Phoca_vitulina_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Phoca_vitulina_Liver$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Phoca_vitulina_Liver
#| fig-cap: "Mammals / Phoca vitulina / Liver: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phoca_vitulina_Liver$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Phoca_vitulina_Liver$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Phoca_vitulina_Liver_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Phoca_vitulina_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Phoca_vitulina_Liver$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Phoca_vitulina_Liver
#| fig-cap: "Mammals / Phoca vitulina / Liver: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Phoca_vitulina_Liver,
  x_label = paste0("Measured value (", grp_Mammals_Phoca_vitulina_Liver_unit, ")")
)
```

## Phocoena phocoena: Fat/Adipose

Mammals / Phocoena phocoena / Fat/Adipose. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Phocoena_phocoena_Fat_Adipose
grp_Mammals_Phocoena_phocoena_Fat_Adipose <- tar_read(outlier_biota_Mammals_Phocoena_phocoena_Fat_Adipose)
grp_Mammals_Phocoena_phocoena_Fat_Adipose$data <- grp_Mammals_Phocoena_phocoena_Fat_Adipose$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Phocoena_phocoena_Fat_Adipose_unit <- grp_Mammals_Phocoena_phocoena_Fat_Adipose$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Phocoena_phocoena_Fat_Adipose
#| fig-cap: "Mammals / Phocoena phocoena / Fat/Adipose: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phocoena_phocoena_Fat_Adipose$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Phocoena_phocoena_Fat_Adipose$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Phocoena_phocoena_Fat_Adipose_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Phocoena_phocoena_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Phocoena_phocoena_Fat_Adipose$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Phocoena_phocoena_Fat_Adipose
#| fig-cap: "Mammals / Phocoena phocoena / Fat/Adipose: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phocoena_phocoena_Fat_Adipose$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Phocoena_phocoena_Fat_Adipose$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Phocoena_phocoena_Fat_Adipose_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Phocoena_phocoena_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Phocoena_phocoena_Fat_Adipose$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Phocoena_phocoena_Fat_Adipose
#| fig-cap: "Mammals / Phocoena phocoena / Fat/Adipose: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phocoena_phocoena_Fat_Adipose$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Phocoena_phocoena_Fat_Adipose$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Phocoena_phocoena_Fat_Adipose_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Phocoena_phocoena_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Phocoena_phocoena_Fat_Adipose$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Phocoena_phocoena_Fat_Adipose
#| fig-cap: "Mammals / Phocoena phocoena / Fat/Adipose: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Phocoena_phocoena_Fat_Adipose,
  x_label = paste0("Measured value (", grp_Mammals_Phocoena_phocoena_Fat_Adipose_unit, ")")
)
```

## Phocoena phocoena: Liver

Mammals / Phocoena phocoena / Liver. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Phocoena_phocoena_Liver
grp_Mammals_Phocoena_phocoena_Liver <- tar_read(outlier_biota_Mammals_Phocoena_phocoena_Liver)
grp_Mammals_Phocoena_phocoena_Liver$data <- grp_Mammals_Phocoena_phocoena_Liver$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Phocoena_phocoena_Liver_unit <- grp_Mammals_Phocoena_phocoena_Liver$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Phocoena_phocoena_Liver
#| fig-cap: "Mammals / Phocoena phocoena / Liver: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phocoena_phocoena_Liver$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Phocoena_phocoena_Liver$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Phocoena_phocoena_Liver_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Phocoena_phocoena_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Phocoena_phocoena_Liver$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Phocoena_phocoena_Liver
#| fig-cap: "Mammals / Phocoena phocoena / Liver: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phocoena_phocoena_Liver$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Phocoena_phocoena_Liver$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Phocoena_phocoena_Liver_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Phocoena_phocoena_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Phocoena_phocoena_Liver$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Phocoena_phocoena_Liver
#| fig-cap: "Mammals / Phocoena phocoena / Liver: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phocoena_phocoena_Liver$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Phocoena_phocoena_Liver$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Phocoena_phocoena_Liver_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Phocoena_phocoena_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Phocoena_phocoena_Liver$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Phocoena_phocoena_Liver
#| fig-cap: "Mammals / Phocoena phocoena / Liver: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Phocoena_phocoena_Liver,
  x_label = paste0("Measured value (", grp_Mammals_Phocoena_phocoena_Liver_unit, ")")
)
```

## Phocoena phocoena: Muscle

Mammals / Phocoena phocoena / Muscle. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Phocoena_phocoena_Muscle
grp_Mammals_Phocoena_phocoena_Muscle <- tar_read(outlier_biota_Mammals_Phocoena_phocoena_Muscle)
grp_Mammals_Phocoena_phocoena_Muscle$data <- grp_Mammals_Phocoena_phocoena_Muscle$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Phocoena_phocoena_Muscle_unit <- grp_Mammals_Phocoena_phocoena_Muscle$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Phocoena_phocoena_Muscle
#| fig-cap: "Mammals / Phocoena phocoena / Muscle: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phocoena_phocoena_Muscle$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Phocoena_phocoena_Muscle$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Phocoena_phocoena_Muscle_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Phocoena_phocoena_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Phocoena_phocoena_Muscle$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Phocoena_phocoena_Muscle
#| fig-cap: "Mammals / Phocoena phocoena / Muscle: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phocoena_phocoena_Muscle$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Phocoena_phocoena_Muscle$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Phocoena_phocoena_Muscle_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Phocoena_phocoena_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Phocoena_phocoena_Muscle$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Phocoena_phocoena_Muscle
#| fig-cap: "Mammals / Phocoena phocoena / Muscle: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Phocoena_phocoena_Muscle$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Phocoena_phocoena_Muscle$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Phocoena_phocoena_Muscle_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Phocoena_phocoena_Muscle
outlier_flag_table(flag_by_facet(grp_Mammals_Phocoena_phocoena_Muscle$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Phocoena_phocoena_Muscle
#| fig-cap: "Mammals / Phocoena phocoena / Muscle: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Phocoena_phocoena_Muscle,
  x_label = paste0("Measured value (", grp_Mammals_Phocoena_phocoena_Muscle_unit, ")")
)
```

## Physeter macrocephalus: Fat/Adipose

Mammals / Physeter macrocephalus / Fat/Adipose. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Physeter_macrocephalus_Fat_Adipose
grp_Mammals_Physeter_macrocephalus_Fat_Adipose <- tar_read(outlier_biota_Mammals_Physeter_macrocephalus_Fat_Adipose)
grp_Mammals_Physeter_macrocephalus_Fat_Adipose$data <- grp_Mammals_Physeter_macrocephalus_Fat_Adipose$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Physeter_macrocephalus_Fat_Adipose_unit <- grp_Mammals_Physeter_macrocephalus_Fat_Adipose$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Physeter_macrocephalus_Fat_Adipose
#| fig-cap: "Mammals / Physeter macrocephalus / Fat/Adipose: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Physeter_macrocephalus_Fat_Adipose$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Physeter_macrocephalus_Fat_Adipose$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Physeter_macrocephalus_Fat_Adipose_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Physeter_macrocephalus_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Physeter_macrocephalus_Fat_Adipose$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Physeter_macrocephalus_Fat_Adipose
#| fig-cap: "Mammals / Physeter macrocephalus / Fat/Adipose: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Physeter_macrocephalus_Fat_Adipose$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Physeter_macrocephalus_Fat_Adipose$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Physeter_macrocephalus_Fat_Adipose_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Physeter_macrocephalus_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Physeter_macrocephalus_Fat_Adipose$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Physeter_macrocephalus_Fat_Adipose
#| fig-cap: "Mammals / Physeter macrocephalus / Fat/Adipose: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Physeter_macrocephalus_Fat_Adipose$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Physeter_macrocephalus_Fat_Adipose$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Physeter_macrocephalus_Fat_Adipose_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Physeter_macrocephalus_Fat_Adipose
outlier_flag_table(flag_by_facet(grp_Mammals_Physeter_macrocephalus_Fat_Adipose$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Physeter_macrocephalus_Fat_Adipose
#| fig-cap: "Mammals / Physeter macrocephalus / Fat/Adipose: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Physeter_macrocephalus_Fat_Adipose,
  x_label = paste0("Measured value (", grp_Mammals_Physeter_macrocephalus_Fat_Adipose_unit, ")")
)
```

## Ursus maritimus: Liver

Mammals / Ursus maritimus / Liver. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Ursus_maritimus_Liver
grp_Mammals_Ursus_maritimus_Liver <- tar_read(outlier_biota_Mammals_Ursus_maritimus_Liver)
grp_Mammals_Ursus_maritimus_Liver$data <- grp_Mammals_Ursus_maritimus_Liver$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Ursus_maritimus_Liver_unit <- grp_Mammals_Ursus_maritimus_Liver$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Ursus_maritimus_Liver
#| fig-cap: "Mammals / Ursus maritimus / Liver: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Ursus_maritimus_Liver$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Ursus_maritimus_Liver$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Ursus_maritimus_Liver_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Ursus_maritimus_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Ursus_maritimus_Liver$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Ursus_maritimus_Liver
#| fig-cap: "Mammals / Ursus maritimus / Liver: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Ursus_maritimus_Liver$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Ursus_maritimus_Liver$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Ursus_maritimus_Liver_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Ursus_maritimus_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Ursus_maritimus_Liver$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Ursus_maritimus_Liver
#| fig-cap: "Mammals / Ursus maritimus / Liver: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Ursus_maritimus_Liver$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Ursus_maritimus_Liver$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Ursus_maritimus_Liver_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Ursus_maritimus_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Ursus_maritimus_Liver$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Ursus_maritimus_Liver
#| fig-cap: "Mammals / Ursus maritimus / Liver: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Ursus_maritimus_Liver,
  x_label = paste0("Measured value (", grp_Mammals_Ursus_maritimus_Liver_unit, ")")
)
```

## Vulpes vulpes: Liver

Mammals / Vulpes vulpes / Liver. Flags are computed separately within each year/site/campaign
group below (Tukey fences x RMZ, `n < 10` not tested); the final
figure compares Winsorization/trimming of the pooled distribution.

TODO: add your own analysis/interpretation of this group here.

```{r}
#| label: read-Mammals_Vulpes_vulpes_Liver
grp_Mammals_Vulpes_vulpes_Liver <- tar_read(outlier_biota_Mammals_Vulpes_vulpes_Liver)
grp_Mammals_Vulpes_vulpes_Liver$data <- grp_Mammals_Vulpes_vulpes_Liver$data |>
  mutate(SAMPLING_YEAR = factor(year(SAMPLING_DATE)))
grp_Mammals_Vulpes_vulpes_Liver_unit <- grp_Mammals_Vulpes_vulpes_Liver$summary$unit
```

```{r}
#| label: fig-outlier-year-Mammals_Vulpes_vulpes_Liver
#| fig-cap: "Mammals / Vulpes vulpes / Liver: measured value by sampling year."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Vulpes_vulpes_Liver$data$SAMPLING_YEAR)
plot_outlier_distribution(
  grp_Mammals_Vulpes_vulpes_Liver$data,
  facet_col = "SAMPLING_YEAR",
  x_label = paste0("Measured value (", grp_Mammals_Vulpes_vulpes_Liver_unit, ")"),
  y_label = "Year",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-year-Mammals_Vulpes_vulpes_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Vulpes_vulpes_Liver$data, "SAMPLING_YEAR", min_n = 10))
```

```{r}
#| label: fig-outlier-site-Mammals_Vulpes_vulpes_Liver
#| fig-cap: "Mammals / Vulpes vulpes / Liver: measured value by site."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Vulpes_vulpes_Liver$data$SITE_CODE, base = 4)
plot_outlier_distribution(
  grp_Mammals_Vulpes_vulpes_Liver$data,
  facet_col = "SITE_CODE",
  x_label = paste0("Measured value (", grp_Mammals_Vulpes_vulpes_Liver_unit, ")"),
  y_label = "Site",
  panel_facet = "SITE_GEOGRAPHIC_FEATURE",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-site-Mammals_Vulpes_vulpes_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Vulpes_vulpes_Liver$data, "SITE_CODE", min_n = 10))
```

```{r}
#| label: fig-outlier-campaign-Mammals_Vulpes_vulpes_Liver
#| fig-cap: "Mammals / Vulpes vulpes / Liver: measured value by campaign."
#| fig-height: !expr outlier_fig_height(grp_Mammals_Vulpes_vulpes_Liver$data$CAMPAIGN_NAME_SHORT)
plot_outlier_distribution(
  grp_Mammals_Vulpes_vulpes_Liver$data,
  facet_col = "CAMPAIGN_NAME_SHORT",
  x_label = paste0("Measured value (", grp_Mammals_Vulpes_vulpes_Liver_unit, ")"),
  y_label = "Campaign",
  min_n = 10
)
```

```{r}
#| label: tbl-outlier-campaign-Mammals_Vulpes_vulpes_Liver
outlier_flag_table(flag_by_facet(grp_Mammals_Vulpes_vulpes_Liver$data, "CAMPAIGN_NAME_SHORT", min_n = 10))
```

```{r}
#| label: fig-outlier-treatment-Mammals_Vulpes_vulpes_Liver
#| fig-cap: "Mammals / Vulpes vulpes / Liver: effect of outlier treatment on the pooled distribution."
plot_outlier_treatment_comparison(
  grp_Mammals_Vulpes_vulpes_Liver,
  x_label = paste0("Measured value (", grp_Mammals_Vulpes_vulpes_Liver_unit, ")")
)
```
